### 2013-09-16 Paper/grass.tex
Added Paper

%\documentclass[aps,twocolumn,floatfix,prl]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{slashbox}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{dashrule}

\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\graphicspath{ {images/} }

\newcommand{\ellcap}{{\ell_\text{c}}}
\newcommand{\Fh}{{\boldsymbol{F}_h}}
\newcommand{\Fv}{{\boldsymbol{F}_v}}
\newcommand{\Tv}{{\boldsymbol{T}}}
\newcommand{\bn}{{\boldsymbol{n}}}
\newcommand{\bt}{{\boldsymbol{t}}}
\newcommand{\bz}{{\boldsymbol{z}}}
\newcommand{\bx}{{\boldsymbol{x}}}
\newcommand{\by}{{\boldsymbol{y}}}
\newcommand{\bT}{{\boldsymbol{T}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

\begin{document}

\title{Waving and swaying}
\author{Ravi Singh}
\author{Shreyas Mandre}
\affiliation{Brown University, Providence RI 02912 USA}
\author{Mahesh Bandi\footnote{work performed while visiting Brown University}}
\affiliation{Okinawa Institute of Science and Technology, Okinawa, Japan}
\author{Amala Mahadevan}
\affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
\author{L. Mahadevan}
\affiliation{Harvard University, Cambridge MA 02138 USA}

\begin{abstract}
%We propose a coupled fluid-canopy model to study dynamics of canopy as well as flow field of fluid. The presence of 
%vegetation is incorporated as a continuous medium which is coupled to flow field via a drag term. We
%did stability analysis of mean flow having strong shear near tip of canopy via solving modified form of 
%Orr-Somerfield equation which has some extra term due to coupling of grass and flow field. Our calculation
%predicts existence of threshold velocity for synchronous waving to commence which have been experienced in different field observations

Flow through submerged aquatic vegetation generates coherent vortices on top of vegetation depending on flow strength and vegetation 
properties, generation of coherent vortices are understood to be manifestation of a hydrodynamic instability of fluid flow experiencing 
different amount of drag force in and out of vegetation. In this study we propose a mechanism based on coupled fluid-vegetation model where presence of vegetation is incorporated 
as a continuous medium coupled to flow field via a drag term. It is predicted that in case of submerged vegetation with high number density,
wave length of coherent vortices are %not governed by shear length scale $L_s = U(h)/U^{'}(h)$ near tip of vegetation 
%(where h is vegetation height, $U(z)$ is mean velocity as function of $z$ and $U'(z) =dU/dz$)
controlled by unvegetated water channel height. Mechanism also predicts existence of threshold velocity for coherent vortices to be observed.     
\end{abstract}
\maketitle

\section{Introduction}
Different kinds of motion shown by field canopy and sea grass due to wind and water flow respectively are very common 
phenomena which affects the hydrodynamic conditions. Altered hydrodynamic conditions can influence number of environmental 
variables and processes such as transport of heat, water, vapor, gas , pollens, sediments, nutrients, contaminants, dissolved oxygen 
and other different kind of nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature; 
wind damage to jungle and crop canopies. Driven by such applications number of scientific work regarding flow over canopies has been done over 
last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
of canopy scales.\newline
A key feature of flow through submerged vegetation or terrestrial canopies is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of this strong shear is manifested as
existence of inflection point near top of canopy in velocity profile. A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
etc.\newline
Previous studies of monami and homani have been done under various assumption such as existence of mixing layer profile near top of canopy; Viscosity playing minor role 
in generating cohrent eddies; Irrotationality of flow; Ignorance of drag in stability analysis. While assumption of viscosity playing minor role seems to be 
a good approximation in large Raynols number, assumption of irrotionality may not be good approximation and more importantly exclusion of drag of canopy in stability 
calculation of mean flow may not be right approximation given the fact that it is drag due to canopy which is responsible for existence of inflection point in mean velocity
profile\newline     
In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow predicted by model in two different scenario via solving 
modified orr-sommerfield equation, our calculation predicts threshold condition for waving, we we further confirm threshold using CFD calculation as well.

\section{The model}
We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow. Presence of grass modifies the flow of fluid via exerting drag, this drag force is
taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
shape of grass at that location, which can be obtained by assuming grass to be in quasi-equilibrium at any moment and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
\begin{equation}
 \mathbf{f}=-N\mathbf{f_{d}}
\end{equation}
where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density 
Drag force exerted by fluid on grass depends on its density $\rho$, flow velocity u, orientation and diameter of grass, which we model as 
\begin{equation}
 \mathbf{f_{d}}=C_{N}\rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
\end{equation}

\begin{equation}
\begin{split}
 \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
 \frac{\del M}{\del s}&-N = 0\\
 M &= B\kappa
\end{split}
\end{equation}
where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass and $\kappa$ is curvature 
of grass.  
\begin{center}
\begin{figure}[ht]
\begin{minipage}[b]{5cm}
%\begin{center}
\centering
 \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
\caption{Typical grass orientation}
\end{minipage}
\hspace{1.0cm}
\begin{minipage}[b]{5cm}
\centering
 \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
\caption{Forces and moment on a section of grass}
\end{minipage}
\end{figure}
\end{center}
For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
set of equation(3).
\section{Stability Analysis}
 In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow 
in presence of grass, we investigate stability of base state obtained through equation of motion. We consider small perturbations $u, v, p$  associated 
with base profile $U$ and $P$ respectively. Momentum and mass balance equation expanded in 1st order of perturbed variable yields.
\begin{equation}
\begin{split}
 \frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}(u_{xx}+u_{yy})-2C_{N}dN_{g}Uu\\
 \frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}(v_{xx}+v_{yy})\\
 \nabla\cdot \bu &= 0
\end{split}
\end{equation}
In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
of having been produced through equation of motion itself.\newline
We performed stability analysis on two different kind of base flow namely 1) flow driven by constant pressure gradient between two parallel plate  2) constant pressure gradient driven 
flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{c_{1}H^2}{4\mu} \]
where $c_{1}$ is constant pressure gradient, H is height at which fluid is present. With these scaling and using stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations 
can be simplified into a single equation.
\begin{equation}
\begin{split}
 R_{e}\left[\psi_{yyt}+\psi_{xxt}+U\left(\psi_{xyy}+\psi_{xxx}\right)-U_{yy}\psi_{x} \right] &=\left[\psi_{yyxx}+\psi_{yyyy}+\psi_{xxxx}+\psi_{xxyy} \right]\\
& + 2D_{drag}U\psi_{y}\delta(y-h)\\
& - 2D_{drag}U_{y}\psi_{y} -2D_{drag}U\psi_{yy}
\end{split}
\end{equation}
where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional drag
 seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get modified orr-sommerfield equation which is an eigenvalue problem
whose solution will provide frequency and wavelength of eddies traveling over canopy.
\begin{equation}
\begin{split}
\sigma \left(D^2-k^2\phi \right) &= \frac{1}{R_{e}}\left[D^4\phi -2k^{2}D^2\phi +k^{4}\phi \right]\\
				  & +ikU_{yy}\phi-ikU\left(D^2\phi-k^2\phi \right)\\
 & -\frac{2D_{drag}}{R_{e}}U_{y}\phi_{y}-\frac{2D_{drag}}{R_{e}}U\phi_{yy}+\frac{2D_{drag}}{R_{e}}U\phi_{y}\delta(y-h)
\end{split}
\end{equation}
critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
  \begin{figure}[htb!]
  \includegraphics[scale=0.35]{Instability2}
\caption{Critical Reynolds number with flow between parallel plate for different grass height as a function of grass number density}
\end{figure}

\begin{figure}[htb!]
  \includegraphics[scale=0.35]{Instability1_free.jpg}
\caption{Critical Reynolds number with no shear stress at top for different grass height as a function of grass number density}
\end{figure}

\end{document}

### 2013-09-17 Paper/grass.tex
Modified author ordering
Updated dashing abstract.

@@
 \author{Ravi Singh}
-\author{Shreyas Mandre}
 \affiliation{Brown University, Providence RI 02912 USA}
+\author{L. Mahadevan}
+\affiliation{Harvard University, Cambridge MA 02138 USA}
 \author{Mahesh Bandi\footnote{work performed while visiting Brown University}}
@@
 \affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
-\author{L. Mahadevan}
-\affiliation{Harvard University, Cambridge MA 02138 USA}
+\author{Shreyas Mandre}
+\affiliation{Brown University, Providence RI 02912 USA}
 
 \begin{abstract}
-%We propose a coupled fluid-canopy model to study dynamics of canopy as well as flow field of fluid. The presence of 
-%vegetation is incorporated as a continuous medium which is coupled to flow field via a drag term. We
-%did stability analysis of mean flow having strong shear near tip of canopy via solving modified form of 
-%Orr-Somerfield equation which has some extra term due to coupling of grass and flow field. Our calculation
-%predicts existence of threshold velocity for synchronous waving to commence which have been experienced in different field observations
-
-Flow through submerged aquatic vegetation generates coherent vortices on top of vegetation depending on flow strength and vegetation 
-properties, generation of coherent vortices are understood to be manifestation of a hydrodynamic instability of fluid flow experiencing 
-different amount of drag force in and out of vegetation. In this study we propose a mechanism based on coupled fluid-vegetation model where presence of vegetation is incorporated 
-as a continuous medium coupled to flow field via a drag term. It is predicted that in case of submerged vegetation with high number density,
-wave length of coherent vortices are %not governed by shear length scale $L_s = U(h)/U^{'}(h)$ near tip of vegetation 
-%(where h is vegetation height, $U(z)$ is mean velocity as function of $z$ and $U'(z) =dU/dz$)
-controlled by unvegetated water channel height. Mechanism also predicts existence of threshold velocity for coherent vortices to be observed.     
+The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile.
+We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer.
+We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability.
+Based on these results, we propose a mechanism for waving marine grass based on the shear instability of the flow above the canopy.
 \end{abstract}

### 2013-12-13 Paper/grass.tex
update_draft_dec13

@@
 The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
-We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer.
-We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability.
-Based on these results, we propose a mechanism for waving marine grass based on the shear instability of the flow above the canopy.
+We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, 
+normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
+the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
+marine grass based on the shear instability of the flow above the canopy.
 \end{abstract}
@@
 \section{Introduction}
-Different kinds of motion shown by field canopy and sea grass due to wind and water flow respectively are very common 
-phenomena which affects the hydrodynamic conditions. Altered hydrodynamic conditions can influence number of environmental 
-variables and processes such as transport of heat, water, vapor, gas , pollens, sediments, nutrients, contaminants, dissolved oxygen 
-and other different kind of nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature; 
-wind damage to jungle and crop canopies. Driven by such applications number of scientific work regarding flow over canopies has been done over 
-last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
-of canopy scales.\newline
-A key feature of flow through submerged vegetation or terrestrial canopies is the existence of strong shear near top of canopy due to
-different amount of drag experienced by fluid in and above the canopy, existence of this strong shear is manifested as
-existence of inflection point near top of canopy in velocity profile. A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
-process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
-may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
-Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
-etc.\newline
-Previous studies of monami and homani have been done under various assumption such as existence of mixing layer profile near top of canopy; Viscosity playing minor role 
-in generating cohrent eddies; Irrotationality of flow; Ignorance of drag in stability analysis. While assumption of viscosity playing minor role seems to be 
-a good approximation in large Raynols number, assumption of irrotionality may not be good approximation and more importantly exclusion of drag of canopy in stability 
-calculation of mean flow may not be right approximation given the fact that it is drag due to canopy which is responsible for existence of inflection point in mean velocity
-profile\newline     
+Rich set of dynamics are exhibited by sea grasses due to its interaction with water flow which can considerably affect the hydrodynamic conditions near the 
+sea grasses, resulting change in hydrodynamic conditions can influence number of environmental 
+variables and processes relative to unobstructed flow such as transport of heat, water, vapor, sediments, contaminants, dissolved oxygen 
+and various other nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature etc. 
+%Driven by such applications number of scientific work regarding flow over canopies has been done over 
+%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
+%of canopy scales.\newline
+Among many dynamics shown by flexible grasses due to their interaction with water flow, the response to steady current in form of large amplitude coherent oscillations known as
+Monami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude oscillation of terrestrial canopy in wind have also been observed named as Honami 
+(Inoue 1993). Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
+\newline
+A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
+different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
+existence of an inflection point near top of canopy in velocity profile. 
+%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
+%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
+%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
+%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
+%etc.\newline
+Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
+have been done under assumption that a pattern of coherent eddies propagating downstream is set by a hydrodynamic instability of shear layer 
+similar to that of free shear flow, response of sea grass to coherent eddies propagating downstream results in large amplitude synchronous oscillations known as monami. 
+Among some of drawback of previous theory includes
+(1) While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, atmospheric flow is unbounded in vertical direction 
+but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
+by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
+\newline     
 In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
-medium which is coupled to flow field via a drag term. We did linear stability analysis of flow predicted by model in two different scenario via solving 
-modified orr-sommerfield equation, our calculation predicts threshold condition for waving, we we further confirm threshold using CFD calculation as well.
+medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
+a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
+We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
+prediction from theory with existing data
 
 \section{The model}
-We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow. Presence of grass modifies the flow of fluid via exerting drag, this drag force is
+We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
 taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
-shape of grass at that location, which can be obtained by assuming grass to be in quasi-equilibrium at any moment and solving force balance and momentum balance for grass  
+shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
+% and solving force balance and momentum balance for grass  
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density 
-Drag force exerted by fluid on grass depends on its density $\rho$, flow velocity u, orientation and diameter of grass, which we model as 
+where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
+Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, orientation and diameter of grass which we model as 
 \begin{equation}
@@
 
-\begin{equation}
-\begin{split}
- \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
- \frac{\del M}{\del s}&-N = 0\\
- M &= B\kappa
-\end{split}
-\end{equation}
-where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass and $\kappa$ is curvature 
-of grass.  
-\begin{center}
-\begin{figure}[ht]
-\begin{minipage}[b]{5cm}
+%\begin{equation}
+%\begin{split}
+% \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
+% \frac{\del M}{\del s}&-N = 0\\
+% M &= B\kappa
+%\end{split}
+
+%\end{equation}
+where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
+
 %\begin{center}
-\centering
- \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
-\caption{Typical grass orientation}
-\end{minipage}
-\hspace{1.0cm}
-\begin{minipage}[b]{5cm}
-\centering
- \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
-\caption{Forces and moment on a section of grass}
-\end{minipage}
-\end{figure}
-\end{center}
-For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
-set of equation(3).
+%\begin{figure}[ht]
+%\begin{minipage}[b]{5cm}
+%\begin{center}
+%\centering
+% \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
+%\caption{Typical grass orientation}
+%\end{minipage}
+%\hspace{1.0cm}
+%\begin{minipage}[b]{5cm}
+%\centering
+% \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
+%\caption{Forces and moment on a section of grass}
+%\end{minipage}
+%\end{figure}
+%\end{center}
+%For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
+%set of equation(3).
 \section{Stability Analysis}
- In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow 
-in presence of grass, we investigate stability of base state obtained through equation of motion. We consider small perturbations $u, v, p$  associated 
+In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow, 
+we investigate stability of base flow profile which can be obtained by solving steady state solution of equation(1). We consider small perturbations $u, v, p$  associated 
 with base profile $U$ and $P$ respectively. Momentum and mass balance equation expanded in 1st order of perturbed variable yields.
@@
 \end{equation}
-In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
-It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
-profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
-of having been produced through equation of motion itself.\newline
-We performed stability analysis on two different kind of base flow namely 1) flow driven by constant pressure gradient between two parallel plate  2) constant pressure gradient driven 
+%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
+%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
+%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
+%of having been produced through equation of motion itself.\newline
+We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
 flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
@@
 where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional drag
- seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get modified orr-sommerfield equation which is an eigenvalue problem
-whose solution will provide frequency and wavelength of eddies traveling over canopy.
+which is zero outside the grass. Seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
+which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
 \begin{equation}

### 2013-12-17 Paper/grass.tex
update text file

@@
 \section{Introduction}
-Rich set of dynamics are exhibited by sea grasses due to its interaction with water flow which can considerably affect the hydrodynamic conditions near the 
-sea grasses, resulting change in hydrodynamic conditions can influence number of environmental 
-variables and processes relative to unobstructed flow such as transport of heat, water, vapor, sediments, contaminants, dissolved oxygen 
-and various other nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature etc. 
+Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
+grass. Resulting change in hydrodynamic conditions relative to unobstructed flow can influence number of environmental 
+variables and processes such as transport of sediments, contaminants, dissolved oxygen;
+plant growth and biomass production (known as thigmomorphogenesis) etc. 
 %Driven by such applications number of scientific work regarding flow over canopies has been done over 
@@
 %of canopy scales.\newline
-Among many dynamics shown by flexible grasses due to their interaction with water flow, the response to steady current in form of large amplitude coherent oscillations known as
-Monami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude oscillation of terrestrial canopy in wind have also been observed named as Honami 
-(Inoue 1993). Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
+Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
+mo-nami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude coherent oscillation of terrestrial canopy in wind have also been observed named as Honami 
+(Inoue 1993).
+% Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
 \newline
@@
 different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
-existence of an inflection point near top of canopy in velocity profile. 
+existence of an inflection point in mean velocity near top of canopy. In addition pattern of coherent eddies propagating downstream 
+along have also been observed.
 %A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
@@
 %Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
-%etc.\newline
+%etc.
+\newline
 Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
-have been done under assumption that a pattern of coherent eddies propagating downstream is set by a hydrodynamic instability of shear layer 
-similar to that of free shear flow, response of sea grass to coherent eddies propagating downstream results in large amplitude synchronous oscillations known as monami. 
-Among some of drawback of previous theory includes
-(1) While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, atmospheric flow is unbounded in vertical direction 
-but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
-by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
+have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
+similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami.
+Several aspect of existing theory remain unclear. The proposed hydrodynamic instability of shear layer layer does not take into account presence of grass whereas shear layer 
+was produced due to drag caused by presence vegetation. Since shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
+observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
+atmospheric flow is unbounded in vertical direction.
+% but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
+%by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
 \newline     

### 2013-12-17 Paper/grass.tex
commit2

@@
 different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
-existence of an inflection point in mean velocity near top of canopy. In addition pattern of coherent eddies propagating downstream 
-along have also been observed.
+presence of an inflection point in mean velocity near canopy top.
+% In addition pattern of coherent eddies propagating downstream 
+%along have also been observed.
 %A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
@@
 %etc.
-\newline
-Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
+%\newline
+Following the work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
 have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
-similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami.
-Several aspect of existing theory remain unclear. The proposed hydrodynamic instability of shear layer layer does not take into account presence of grass whereas shear layer 
-was produced due to drag caused by presence vegetation. Since shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
-observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
-atmospheric flow is unbounded in vertical direction.
+similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami. While shear layer
+model sucessfully predict frequency of passage of corent vortices for number of experimental observations,
+several aspect of existing theory remain unclear. First, the proposed hydrodynamic instability of shear layer does not take into account influence of
+drag due to grasses whreas shear layer itself is produced by drag force provided by vegetation. 
+%the proposed hydrodynamic instability mechanism of shear layer layer does not take into account presence of grass whereas existence of 
+%shear layer is due to drag provided by vegetation. 
+Second, classical free shear flow is known to be unstable above a very small Reynolds number $\geq 5 $. On the contrary experimental observations
+(Grizzle,1996) shows that an event of mo-nami is not observed untill flow speed exceed a certain threshold value. Although atmoshpheric flow through terrestrial canopy seems similar to 
+aquatic flow through vegetation, atmoshpheric flow is essentially unbounded, due to difference of vertical extent of flow in terrestrial and aquatic case a mechanism similar to 
+terrestrial case might not be adequate for aquatic scenario.
+%might not hold  
+%was produced due to drag caused by presence vegetation. 
+%Second, shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
+%observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
+%atmospheric flow is unbounded in vertical direction.
 % but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
@@
 \newline 
-In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
-medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
-a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
-We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
-prediction from theory with existing data
+ In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegeation. Our caluation of linear stability analysis of flow in presence of grasses 
+shows that drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer. We also show that an event of mo-nami is caused by instablity 
+of fluid flow above vegeation rather than that of shear layer near grass. We further show that inclusion of grass predicts a threshold criteria for waving characterizedd by
+Reynolds number. We confirm our analyis by comparing predictin from theory with field and experimental observations.  
+%In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
+%medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
+%a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
+%We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
+%prediction from theory with existing data
 
 \section{The model}
-We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
-taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
-shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
+We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continous drag filed which is taken as additional body force 
+in momentum equation of fluid. 
+%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
+%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
+%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
 % and solving force balance and momentum balance for grass  
@@
 \end{equation}
-\begin{equation}
- \mathbf{f}=-N\mathbf{f_{d}}
-\end{equation}
-where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
-Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, orientation and diameter of grass which we model as 
-\begin{equation}
- \mathbf{f_{d}}=C_{N}\rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
-\end{equation}
+%\begin{equation}
+% \mathbf{f}=-N\mathbf{f_{d}}
+%\end{equation}
+where $\mathbf{f}=-N\mathbf{f_d}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
+Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
+$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
+%\begin{equation}
+% \mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
+%\end{equation}\
 
@@
 %\end{equation}
-where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
+%where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
 
@@
 \section{Stability Analysis}
-In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow, 
-we investigate stability of base flow profile which can be obtained by solving steady state solution of equation(1). We consider small perturbations $u, v, p$  associated 
-with base profile $U$ and $P$ respectively. Momentum and mass balance equation expanded in 1st order of perturbed variable yields.
+In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of hydrodynamic instability associated with mean flow, 
+we investigate stability of base flow profile which obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 
+with base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
 \begin{equation}
@@
 %of having been produced through equation of motion itself.\newline
-We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
-flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
-   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{c_{1}H^2}{4\mu} \]
-where $c_{1}$ is constant pressure gradient, H is height at which fluid is present. With these scaling and using stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations 
-can be simplified into a single equation.
+%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
+%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
+Where drag coeffient $C_N$ is set zero out side vagetation. Using height $H$ of fluid column and associated advection time H/$U_{0}$, we use following scalaing to non-dimensinlize equation 
+of motion
+   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
+With these scaling along with use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.
 \begin{equation}

### 2013-12-18 Paper/grass.tex
Introduction_almost

@@
 \section{Introduction}
-Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
-grass. Resulting change in hydrodynamic conditions relative to unobstructed flow can influence number of environmental 
-variables and processes such as transport of sediments, contaminants, dissolved oxygen;
-plant growth and biomass production (known as thigmomorphogenesis) etc. 
+Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
+%Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
+%grass.
+Resulting changes in hydrodynamic conditions relative to unobstructed flow can influence number of environmental variables and processes such as 
+transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 
 %Driven by such applications number of scientific work regarding flow over canopies has been done over 
@@
 %of canopy scales.\newline
-Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
-mo-nami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude coherent oscillation of terrestrial canopy in wind have also been observed named as Honami 
-(Inoue 1993).
+The response of flexible grass canopies to steady current in form of large amplitude coherent oscillations known as mo-nami (Ackerman and Okubo, 1993) plays a central role
+in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 
+%Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
+%mo-nami (Ackerman and Okubo 1993) is focus of this paper. 
+Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955) have also been observed.
 % Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
 \newline
-A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
-different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
-presence of an inflection point in mean velocity near canopy top.
+Although atmosphere flow through terrestrial canopies seems similar to aquatic flow through vegetation, atmospheric flows are essentially unbounded vertically. Another major
+difference between atmospheric flow and aquatic flow is the considerable difference of stiffness of canopies, terrestrial canopies tends to be much more rigid than marine canopies.
+\newline   
+The previous explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
+different amounts of drag experienced by fluid in and above the canopy. Existence of strong shear is manifested by
+the presence of an inflection point in the mean velocity profile near the top of the canopy.
 % In addition pattern of coherent eddies propagating downstream 
@@
 %\newline
-Following the work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
-have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
-similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami. While shear layer
-model sucessfully predict frequency of passage of corent vortices for number of experimental observations,
-several aspect of existing theory remain unclear. First, the proposed hydrodynamic instability of shear layer does not take into account influence of
-drag due to grasses whreas shear layer itself is produced by drag force provided by vegetation. 
+%Following the work of Raupach (1996) on terrestrial canopies, 
+Existing explanation of monami have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
+similar to that of free shear flow known as Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses are exhibited by their large amplitude synchronous 
+oscillations.
+%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
+%are exhibited in form of large amplitude oscillations knows as mo-nami.
+\newline
+While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
+several aspect of existing theory remain unexplained. First, the proposed hydrodynamic instability of shear layer produced due to drag of 
+grass does not take into account influence of drag due to grasses itself. 
 %the proposed hydrodynamic instability mechanism of shear layer layer does not take into account presence of grass whereas existence of 
 %shear layer is due to drag provided by vegetation. 
-Second, classical free shear flow is known to be unstable above a very small Reynolds number $\geq 5 $. On the contrary experimental observations
-(Grizzle,1996) shows that an event of mo-nami is not observed untill flow speed exceed a certain threshold value. Although atmoshpheric flow through terrestrial canopy seems similar to 
-aquatic flow through vegetation, atmoshpheric flow is essentially unbounded, due to difference of vertical extent of flow in terrestrial and aquatic case a mechanism similar to 
-terrestrial case might not be adequate for aquatic scenario.
+Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) for monami is much
+higher $\sim 1000$
+%(Grizzle,1996) shows that an event of mo-nami is not observed until flow speed exceed a certain threshold value. Although atmospheric flow through terrestrial canopy seems similar to 
+%aquatic flow through vegetation, atmospheric flow is essentially unbounded, due to these difference of vertical extent of flow in two cases a mechanism similar to 
+%terrestrial case might not be adequate for aquatic scenario.
 %might not hold  
@@
 \newline 
- In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegeation. Our caluation of linear stability analysis of flow in presence of grasses 
-shows that drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer. We also show that an event of mo-nami is caused by instablity 
-of fluid flow above vegeation rather than that of shear layer near grass. We further show that inclusion of grass predicts a threshold criteria for waving characterizedd by
-Reynolds number. We confirm our analyis by comparing predictin from theory with field and experimental observations.  
+ In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grasses 
+shows that drag from the grass blades damps out the Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
+of fluid flow above vegetation coupled with the flow through canopy. We further show that inclusion of drag due to grass predicts a threshold condition for waving where threshold is
+characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.  
 %In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
@@
 \section{The model}
-We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continous drag filed which is taken as additional body force 
-in momentum equation of fluid. 
+We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continuous drag field 
+which is incorporated as an additional body force $\mathbf{f}=-N\mathbf{f_d}$ in fluid momentum equation. 
 %We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
@@
 %\end{equation}
-where $\mathbf{f}=-N\mathbf{f_d}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
-Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
+where $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. Drag force exerted by fluid on grass depends on
+fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
@@
 In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of hydrodynamic instability associated with mean flow, 
-we investigate stability of base flow profile which obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 
+we investigate stability of base flow profile which is obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 
 with base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
@@
    \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-With these scaling along with use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.
+With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.
 \begin{equation}
@@
 \end{equation}
-where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional drag
-which is zero outside the grass. Seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
+where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional.
+Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
 which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
@@
 \end{equation}
-critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
+Critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
   \begin{figure}[htb!]

### 2013-12-18 Paper/grass.tex
Model

@@
 \section{The model}
-We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continuous drag field 
-which is incorporated as an additional body force $\mathbf{f}=-N\mathbf{f_d}$ in fluid momentum equation. 
+We use a mean field model for the coupling between the flow and the canopy. Assuming strong oscillation in fluid flow is caused by instability of flow
+composed of regions experiencing higher drag with in grass and less outside while monami being response of flexible grass to oscillations of fluid flow. We expect stiffness 
+of grass to play minor role in development of flow instability. We approximate grasses with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N/m^2$. 
+Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N\mathbf{f_d}$ in the fluid momentum balance as
 %We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
@@
 %\end{equation}
-where $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. Drag force exerted by fluid on grass depends on
-fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
-$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
+where $\mathbf{f_{d}}$ is drag force per unit length of grass, $\rho$ the fluid density, $\mathbf{u}$ the velocity, $P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$
+the acceleration due to gravity. Drag force itself is modeled as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
+$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
+normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. We further expect $C_N$ to
+not change along the height of the grass and take $C_N$ to be constant for further analysis. 
 %\begin{equation}

### 2013-12-26 Paper/grass.tex
near_christmas

@@
-%\documentclass[aps,twocolumn,floatfix,prl]{revtex4-1}
+\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
 %\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
-\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
+%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
 \usepackage{fullpage}
@@
 composed of regions experiencing higher drag with in grass and less outside while monami being response of flexible grass to oscillations of fluid flow. We expect stiffness 
-of grass to play minor role in development of flow instability. We approximate grasses with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N/m^2$. 
-Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N\mathbf{f_d}$ in the fluid momentum balance as
+of grass to play minor role in development of flow instability. We approximate grasses with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N_g/m^2$. 
+Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
 %We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
@@
 \section{Stability Analysis}
-In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of hydrodynamic instability associated with mean flow, 
-we investigate stability of base flow profile which is obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 
-with base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
+We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
+Base state flow can be obtained by solving steady state form of constant pressure gradient driven flow associated with equation(1)
+$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. Width of shear layer associated with the base flow can be 
+obtained by balancing viscous force and vegetative drag in shear layer along with the balance of shear stress exerted by un-vegetated flow at the tip and total vegetative 
+drag in shear layer see figure(1). With these balance we can get the scale of shear layer $\delta \sim H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$
+where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
+\begin{figure}[htb!]
+  \includegraphics[]{fig1}
+\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (top center) }
+\end{figure}
+Knowing the scale of shear layer we can systematically investigate its effect on flow instability
+by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
+respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
 \begin{equation}
@@
 %flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
-Where drag coeffient $C_N$ is set zero out side vagetation. Using height $H$ of fluid column and associated advection time H/$U_{0}$, we use following scalaing to non-dimensinlize equation 
-of motion
-   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.
+Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.
+%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
+%of motion
+%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
+With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, momentum and mass balance can be combined 
+into a single equation.
 \begin{equation}
-\begin{split}
- R_{e}\left[\psi_{yyt}+\psi_{xxt}+U\left(\psi_{xyy}+\psi_{xxx}\right)-U_{yy}\psi_{x} \right] &=\left[\psi_{yyxx}+\psi_{yyyy}+\psi_{xxxx}+\psi_{xxyy} \right]\\
-& + 2D_{drag}U\psi_{y}\delta(y-h)\\
-& - 2D_{drag}U_{y}\psi_{y} -2D_{drag}U\psi_{yy}
-\end{split}
+Re\left[\del_t+\del_x U \right]\grad^2\phi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
+%\begin{split}
+% R_{e}\left[\psi_{yyt}+\psi_{xxt}+U\left(\psi_{xyy}+\psi_{xxx}\right)-U_{yy}\psi_{x} \right] &=\left[\psi_{yyxx}+\psi_{yyyy}+\psi_{xxxx}+\psi_{xxyy} \right]\\
+%& + 2D_{drag}U\psi_{y}\delta(y-h)\\
+%& - 2D_{drag}U_{y}\psi_{y} -2D_{drag}U\psi_{yy}
+%\end{split}
 \end{equation}
-where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional.
-Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
+where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
+non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
 which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
@@
 \begin{split}
-\sigma \left(D^2-k^2\phi \right) &= \frac{1}{R_{e}}\left[D^4\phi -2k^{2}D^2\phi +k^{4}\phi \right]\\
-				  & +ikU_{yy}\phi-ikU\left(D^2\phi-k^2\phi \right)\\
- & -\frac{2D_{drag}}{R_{e}}U_{y}\phi_{y}-\frac{2D_{drag}}{R_{e}}U\phi_{yy}+\frac{2D_{drag}}{R_{e}}U\phi_{y}\delta(y-h)
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi\\
+&-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)
+%& -\frac{2D_{drag}}{R_{e}}U_{y}\phi_{y}-\frac{2D_{drag}}{R_{e}}U\phi_{yy}+\frac{2D_{drag}}{R_{e}}U\phi_{y}\delta(y-h)
 \end{split}
@@
   \begin{figure}[htb!]
-  \includegraphics[scale=0.35]{Instability2}
+  \includegraphics[]{Critical_Re_vs_Ng_Grizzle}
 \caption{Critical Reynolds number with flow between parallel plate for different grass height as a function of grass number density}
@@
 
-\begin{figure}[htb!]
-  \includegraphics[scale=0.35]{Instability1_free.jpg}
-\caption{Critical Reynolds number with no shear stress at top for different grass height as a function of grass number density}
-\end{figure}
-
 \end{document}

### 2014-01-13 Paper/grass.tex
jan13

@@
 \usepackage{dashrule}
-
 \usepackage{amsmath}
@@
 \begin{document}
-
 \title{Waving and swaying}
-\author{Ravi Singh}
+\author{Ravi Singh, Shreyas Mandre}
 \affiliation{Brown University, Providence RI 02912 USA}
@@
 \affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
-\author{Shreyas Mandre}
-\affiliation{Brown University, Providence RI 02912 USA}
+%\author{Shreyas Mandre}
+%\affiliation{Brown University, Providence RI 02912 USA}
 
 \begin{abstract}
-The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
-We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, 
-normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
-the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
-marine grass based on the shear instability of the flow above the canopy.
+%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
+%We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, 
+%normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
+%the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
+%marine grass based on the shear instability of the flow above the canopy.
+The spontaneous waving of marine grass is though to be due to a Kelving-helmholtz instability resulting from strong shear near top of grass. On the contrary
+we find that strong velocity shear resulting from vertical discontinuity of drag due to presence of grass does not lead to Kelvin-Helmholtz instability which is 
+normally observed in free shear flows. We also show that wavelength of coherent vortices scales with the unvegetated water depth. Based on these results, 
+we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy. 
 \end{abstract}
@@
 Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-%Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
-%grass.
-Resulting changes in hydrodynamic conditions relative to unobstructed flow can influence number of environmental variables and processes such as 
+Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
 transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 
-%Driven by such applications number of scientific work regarding flow over canopies has been done over 
-%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
-%of canopy scales.\newline
 The response of flexible grass canopies to steady current in form of large amplitude coherent oscillations known as mo-nami (Ackerman and Okubo, 1993) plays a central role
 in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 
-%Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
-%mo-nami (Ackerman and Okubo 1993) is focus of this paper. 
-Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955) have also been observed.
-% Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
 \newline
+Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.
 Although atmosphere flow through terrestrial canopies seems similar to aquatic flow through vegetation, atmospheric flows are essentially unbounded vertically. Another major
-difference between atmospheric flow and aquatic flow is the considerable difference of stiffness of canopies, terrestrial canopies tends to be much more rigid than marine canopies.
+difference between atmospheric flow and aquatic flow is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than marine canopies.
 \newline   
-The previous explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
-different amounts of drag experienced by fluid in and above the canopy. Existence of strong shear is manifested by
-the presence of an inflection point in the mean velocity profile near the top of the canopy.
-% In addition pattern of coherent eddies propagating downstream 
-%along have also been observed.
-%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
-%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
-%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
-%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
-%etc.
-%\newline
-%Following the work of Raupach (1996) on terrestrial canopies, 
-Existing explanation of monami have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
-similar to that of free shear flow known as Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses are exhibited by their large amplitude synchronous 
-oscillations.
+Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
+different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices throguh a mechanism similar to Kelvin-Helmholtz
+instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
 %Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
@@
 grass does not take into account influence of drag due to grasses itself.
-%the proposed hydrodynamic instability mechanism of shear layer layer does not take into account presence of grass whereas existence of 
-%shear layer is due to drag provided by vegetation. 
+Linear stability analysis of base flow does not take into account influence of drag due to grass  
 Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) for monami is much
-higher $\sim 1000$
-%(Grizzle,1996) shows that an event of mo-nami is not observed until flow speed exceed a certain threshold value. Although atmospheric flow through terrestrial canopy seems similar to 
-%aquatic flow through vegetation, atmospheric flow is essentially unbounded, due to these difference of vertical extent of flow in two cases a mechanism similar to 
-%terrestrial case might not be adequate for aquatic scenario.
-%might not hold  
-%was produced due to drag caused by presence vegetation. 
-%Second, shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
-%observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
-%atmospheric flow is unbounded in vertical direction.
-% but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
-%by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
+higher $O(1000)$
 \newline 
@@
 characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.
-%In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
-%medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
-%a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
-%We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
-%prediction from theory with existing data
 
 \section{The model}
-We use a mean field model for the coupling between the flow and the canopy. Assuming strong oscillation in fluid flow is caused by instability of flow
-composed of regions experiencing higher drag with in grass and less outside while monami being response of flexible grass to oscillations of fluid flow. We expect stiffness 
-of grass to play minor role in development of flow instability. We approximate grasses with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N_g/m^2$. 
+We use a mean field model for the coupling between the flow and the canopy. The drag force per unit volume experienced by fluid flow is believed to be dependent on
+mean frontal area of grass while flexibility of grass playing only minor role in development of flow (Raupach-1996, Gishalberti-2002).   
+We approximate mean frontal area of grass field with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N_g/m^2$. 
 Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
@@
 %\end{equation}
-where $\mathbf{f_{d}}$ is drag force per unit length of grass, $\rho$ the fluid density, $\mathbf{u}$ the velocity, $P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $\rho$ the fluid density, $\mathbf{u}$ the velocity, $P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$
 the acceleration due to gravity. Drag force itself is modeled as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
@@
 normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. We further expect $C_N$ to
-not change along the height of the grass and take $C_N$ to be constant for further analysis. 
+not change along the height of the grass and take $C_N$ to be constant. 
 %\begin{equation}
@@
 We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
-Base state flow can be obtained by solving steady state form of constant pressure gradient driven flow associated with equation(1)
-$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. Width of shear layer associated with the base flow can be 
-obtained by balancing viscous force and vegetative drag in shear layer along with the balance of shear stress exerted by un-vegetated flow at the tip and total vegetative 
-drag in shear layer see figure(1). With these balance we can get the scale of shear layer $\delta \sim H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$
-where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
+Base state flow is obtained by solving steady state form of constant pressure gradient driven flow associated with equation(1)
+$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. Dominant balance of drag force with pressure gradient with in
+grass leads to a constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ while balance of viscous force with pressure gradient above grass provide a parabolic velocity profile.
+The base flow profile consist of two boundary layer one near the bottom of grass due to no slip condition and other near tip of canopy known as shear layer due to vertical discontinuity 
+of drag force. Width of boundary layer near bottom of grass can be obtained by balance of viscous force with drag and pressure gradient forces which gives boundary layer
+width $\sim \sqrt{\frac{\mu}{\rho C_NdN_g U_g}}$. Thickness of shear layer near the top of grass known as shear layer width is obtained by balance of viscous force 
+with drag along with matching of shear stress near tip exerted by flow in unvegetated region at canopy top which leads to shear layer 
+width $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$ where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
+
 \begin{figure}[htb!]
   \includegraphics[]{fig1}
-\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (top center) }
+\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
 \end{figure}
@@
 by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
-respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
+respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
 \begin{equation}
@@
 %   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, momentum and mass balance can be combined 
+With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
 into a single equation.
 \begin{equation}
-Re\left[\del_t+\del_x U \right]\grad^2\phi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
+Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
 %\begin{split}
@@
 \end{equation}
-Critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
+Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability. Fig(2) shows 
+this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
   \begin{figure}[htb!]

### 2014-01-21 Paper/grass.tex
Extra Figure

--- 2013-12-26 Paper/grass.tex
@@
 \affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
-\author{Shreyas Mandre}
-\affiliation{Brown University, Providence RI 02912 USA}
+%\author{Shreyas Mandre}
+%\affiliation{Brown University, Providence RI 02912 USA}
 
@@
 The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
-We find that this inflection point is located inside the grass canopy, and the drag from the grass blades damps out the Kelvin Helmholtz instability, 
+We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
 normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
@@
 marine grass based on the shear instability of the flow above the canopy.
+
+%The spontaneous waving of marine grass is though to be due to a Kelving-helmholtz instability resulting from strong shear near top of grass. On the contrary
+%we find that strong velocity shear resulting from vertical discontinuity of drag due to presence of grass does not lead to Kelvin-Helmholtz instability which is 
+%normally observed in free shear flows. We also show that wavelength of coherent vortices scales with the unvegetated water depth. Based on these results, 
+%we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy. 
 \end{abstract}
@@
 Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-%Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
-%grass.
-Resulting changes in hydrodynamic conditions relative to unobstructed flow can influence number of environmental variables and processes such as 
+Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
 transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 
-%Driven by such applications number of scientific work regarding flow over canopies has been done over 
-%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
-%of canopy scales.\newline
-The response of flexible grass canopies to steady current in form of large amplitude coherent oscillations known as mo-nami (Ackerman and Okubo, 1993) plays a central role
+The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami (Ackerman and Okubo, 1993), plays a central role
 in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 
-%Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
-%mo-nami (Ackerman and Okubo 1993) is focus of this paper. 
-Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955) have also been observed.
-% Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
 \newline
-Although atmosphere flow through terrestrial canopies seems similar to aquatic flow through vegetation, atmospheric flows are essentially unbounded vertically. Another major
-difference between atmospheric flow and aquatic flow is the considerable difference of stiffness of canopies, terrestrial canopies tends to be much more rigid than marine canopies.
+Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
+difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
+Due to these considerable differences between aquatic and marine flow the analysis in this paper is applicable only to marine canopies  
 \newline   
-The previous explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
-different amounts of drag experienced by fluid in and above the canopy. Existence of strong shear is manifested by
-the presence of an inflection point in the mean velocity profile near the top of the canopy.
-% In addition pattern of coherent eddies propagating downstream 
-%along have also been observed.
-%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
-%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
-%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
-%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
-%etc.
-%\newline
-%Following the work of Raupach (1996) on terrestrial canopies, 
-Existing explanation of monami have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
-similar to that of free shear flow known as Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses are exhibited by their large amplitude synchronous 
-oscillations.
+Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
+different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz
+instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
 %Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
@@
 While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
-several aspect of existing theory remain unexplained. First, the proposed hydrodynamic instability of shear layer produced due to drag of 
-grass does not take into account influence of drag due to grasses itself. 
-%the proposed hydrodynamic instability mechanism of shear layer layer does not take into account presence of grass whereas existence of 
-%shear layer is due to drag provided by vegetation. 
-Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) for monami is much
-higher $\sim 1000$
-%(Grizzle,1996) shows that an event of mo-nami is not observed until flow speed exceed a certain threshold value. Although atmospheric flow through terrestrial canopy seems similar to 
-%aquatic flow through vegetation, atmospheric flow is essentially unbounded, due to these difference of vertical extent of flow in two cases a mechanism similar to 
-%terrestrial case might not be adequate for aquatic scenario.
-%might not hold  
-%was produced due to drag caused by presence vegetation. 
-%Second, shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
-%observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
-%atmospheric flow is unbounded in vertical direction.
-% but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
-%by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
+several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent 
+theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) 
+for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of
+phenomenon.
 \newline 
 In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grasses 
-shows that drag from the grass blades damps out the Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
+shows that existence of strong shear near canopy top does not lead to Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
 of fluid flow above vegetation coupled with the flow through canopy. We further show that inclusion of drag due to grass predicts a threshold condition for waving where threshold is
 characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.  
-%In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
-%medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
-%a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
-%We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
-%prediction from theory with existing data
-
-\section{The model}
-We use a mean field model for the coupling between the flow and the canopy. Assuming strong oscillation in fluid flow is caused by instability of flow
-composed of regions experiencing higher drag with in grass and less outside while monami being response of flexible grass to oscillations of fluid flow. We expect stiffness 
-of grass to play minor role in development of flow instability. We approximate grasses with stiff cylindrical rod of diameter $d$ and height $h$ with number density $N_g/m^2$. 
+%\section{The model}
+%Monami is believed to be response to strong oscillations of flow in streamwise velocity associated with coherent vortices.
+\newline
+While Drag exerted by canopy is believed to play 
+a significant role in the development of coherent vortices leading to monami, flexibility of grass provide striking visualization of these structures without significantly affecting the
+dynamics of flow (Raupach 1996, Ghisalberti 2002). For this reason we replace grass with stiff structure to investigate dynamics leading to generation of 
+coherent vortices. We use a mean field model for the coupling between the flow and the canopy. 
+%The drag force per unit volume experienced by fluid flow is believed to be dependent on
+%mean frontal area of grass. We approximate mean frontal area of grass field of height $h$ with $dN_g/m^2$. Where $d$ is average width of grass and $N_g$ the grass number density per unit area 
 Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
@@
 %\end{equation}
-where $\mathbf{f_{d}}$ is drag force per unit length of grass, $\rho$ the fluid density, $\mathbf{u}$ the velocity, $P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$
-the acceleration due to gravity. Drag force itself is modeled as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
+$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity. Drag force itself is modeled 
+as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade 
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
 normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. We further expect $C_N$ to
-not change along the height of the grass and take $C_N$ to be constant for further analysis. 
+not change along the height of the grass and take $C_N$ to be constant. 
 %\begin{equation}
@@
 %\end{equation}\
-
 %\begin{equation}
@@
 %\end{split}
-
 %\end{equation}
 %where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
+%\section{Stability Analysis}
+\newline
+A solution of equation(1) gives uniform steady flow which we expect to be unstable leading to generate coherent eddies. To understand the instability characteristic associated with this
+steady flow termed as base flow we perform linear stability analysis.  
+%We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
+This base state flow is obtained by solving steady state form of constant pressure gradient driven flow of equation(1)
+$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface Fig(1). Flow consist of a region of approximately constant velocity
+$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient with in grass and a parabolic velocity profile in unvegetated region due to 
+balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top give rise to two boundary layer in the flow profile one 
+near bottom and near the tip of the grass which is conventionally known as shear layer.
+%eads to a 
+%constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ while balance of viscous force with pressure gradient above grass provide a parabolic velocity profile.
+%The base flow profile consist of two boundary layer one near the bottom of grass due to no slip condition and other near tip of canopy known as shear layer due to vertical discontinuity 
+%of drag force. Width of boundary layer near bottom of grass can be obtained by balance of viscous force with drag and pressure gradient forces which gives boundary layer
+%width $\sim \sqrt{\frac{\mu}{\rho C_NdN_g U_g}}$. 
+Thickness of shear layer near top of grass is obtained by balance of viscous force 
+with drag along with match of shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
+width $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$ where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
 
-%\begin{center}
-%\begin{figure}[ht]
-%\begin{minipage}[b]{5cm}
-%\begin{center}
-%\centering
-% \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
-%\caption{Typical grass orientation}
-%\end{minipage}
-%\hspace{1.0cm}
-%\begin{minipage}[b]{5cm}
-%\centering
-% \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
-%\caption{Forces and moment on a section of grass}
-%\end{minipage}
-%\end{figure}
-%\end{center}
-%For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
-%set of equation(3).
-\section{Stability Analysis}
-We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
-Base state flow can be obtained by solving steady state form of constant pressure gradient driven flow associated with equation(1)
-$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. Width of shear layer associated with the base flow can be 
-obtained by balancing viscous force and vegetative drag in shear layer along with the balance of shear stress exerted by un-vegetated flow at the tip and total vegetative 
-drag in shear layer see figure(1). With these balance we can get the scale of shear layer $\delta \sim H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$
-where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
 \begin{figure}[htb!]
   \includegraphics[]{fig1}
-\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (top center) }
+\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
 \end{figure}
@@
 by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
-respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.
+respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
 \begin{equation}
+%\scriptstyle{
 \begin{split}
- \frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}(u_{xx}+u_{yy})-2C_{N}dN_{g}Uu\\
- \frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}(v_{xx}+v_{yy})\\
- \nabla\cdot \bu &= 0
+ %u_t+Uu_x+vU_y &= \frac{1}{\rho}p_x+\frac{\mu}{\rho}\left(u_xx+u_yy \right) -2C_NdN_gUu\\
+\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\
+\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0\\
+ %\nabla\cdot \bu &= 0
 \end{split}
+%}
 \end{equation}
@@
 %   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, momentum and mass balance can be combined 
+With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
 into a single equation.
 \begin{equation}
-Re\left[\del_t+\del_x U \right]\grad^2\phi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
-%\begin{split}
-% R_{e}\left[\psi_{yyt}+\psi_{xxt}+U\left(\psi_{xyy}+\psi_{xxx}\right)-U_{yy}\psi_{x} \right] &=\left[\psi_{yyxx}+\psi_{yyyy}+\psi_{xxxx}+\psi_{xxyy} \right]\\
-%& + 2D_{drag}U\psi_{y}\delta(y-h)\\
-%& - 2D_{drag}U_{y}\psi_{y} -2D_{drag}U\psi_{yy}
-%\end{split}
+\scriptstyle{
+Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
+}
 \end{equation}
@@
 \begin{split}
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi\\
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
 &-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)
-%& -\frac{2D_{drag}}{R_{e}}U_{y}\phi_{y}-\frac{2D_{drag}}{R_{e}}U\phi_{yy}+\frac{2D_{drag}}{R_{e}}U\phi_{y}\delta(y-h)
 \end{split}
 \end{equation}
-Critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
+Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2))
+
+% shows 
+%this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
 \begin{figure}[htb!]
-  \includegraphics[]{Critical_Re_vs_Ng_Grizzle}
+\includegraphics[scale=0.8]{Critical_Re_vs_Ng_Grizzle}
 \caption{Critical Reynolds number with flow between parallel plate for different grass height as a function of grass number density}
+\end{figure}
+\begin{figure}[htb!]
+\includegraphics[]{K_vs_shear_width}
+\caption{Critical Wavenumber with for different grass height as a function of shear layer thickness}
+\end{figure}
+\begin{figure}[htb!]
+\includegraphics[scale=0.34]{Observed_vs_calculated}
+\caption{Comparison between experimently oberved frequency and predicted frequency from model}
 \end{figure}

### 2014-01-25 Paper/grass.tex

@@
 \includegraphics[]{K_vs_shear_width}
-\caption{Critical Wavenumber with for different grass height as a function of shear layer thickness}
+\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
 \end{figure}
@@
 \includegraphics[scale=0.34]{Observed_vs_calculated}
-\caption{Comparison between experimently oberved frequency and predicted frequency from model}
+\caption{Comparison between experimently oberved frequency and predicted frequency}
 \end{figure}

### 2014-01-27 Paper/grass.tex
before aspen

@@
 \begin{abstract}
-The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
-We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
-normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
-the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
-marine grass based on the shear instability of the flow above the canopy.
+%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
+%We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
+%normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
+%the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
+%marine grass based on the shear instability of the flow above the canopy.
 
@@
 %we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy.
+
+The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
+is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable either through a mechanism of Kelvin-Helmholtz instability associated with 
+inflection point or via shear instability of flow above the grass.
+
 \end{abstract}
@@
 \begin{figure}[htb!]
-  \includegraphics[]{fig1}
+  \includegraphics[scale=0.8]{fig1}
 \caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
@@
 \end{equation}
-Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2))
-
+Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2)).
+Variation of critical wave number with shear layer thickness for different grass height is shown in Fig-3.
 % shows 
@@
 \begin{figure}[htb!]
-\includegraphics[scale=0.8]{Critical_Re_vs_Ng_Grizzle}
-\caption{Critical Reynolds number with flow between parallel plate for different grass height as a function of grass number density}
+\includegraphics[]{Critical_Re_vs_delta}
+\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
 \end{figure}
+\newline
+The following features can be noted from Fig(3). For the certain grass height region of linear variation of wavenumber with inverse shear layer thickness suggest that the flow is subject 
+to Kelvin-Helmholtz instability above a threshold shear thickness. Region of constant wave number suggest that length scale of the instability is no longer governed by shear layer
+thickness in this this regime, rather it is governed by flow above the grass, which can be understood by asymptotic analysis in regime of small shear layer thickness and large Reynolds 
+number
 \begin{figure}[htb!]
@@
 \end{figure}
+%\newline
+Collapse of $\phi$ for different shear thickness in the regime of constant critical wavenumber $k$ suggest that we have obtained the asymptotic solution in limit of small shear layer 
+thickness. Noting that non-dimensional flow velocity $U \sim \delta^{3/2}$ inside the grass and $R_{e} \gg 1$, we can simplify the equation-(4) considerably both with in the grass and 
+above the grass canopy. Inside the grass equation simplifies into $\sigma\left(D^2-k^2\right)\phi = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$. Together with simplified
+force balance inside the grass canopy and constant value of $k\approx O(1)$, we can estimate that critical Reynolds number $R_e \sim \delta^{-3/2}$ for small shear layer thickness Fig(2) 
+\begin{figure}[htb!]
+ \includegraphics[]{Asymptotic_eig_set3}
+\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
+\end{figure}
+
 \begin{figure}[htb!]

### 2014-02-21 Paper/grass.tex
updatedFeb21

@@
+%\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
 \documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 %\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
-\usepackage{fullpage}
+%\usepackage{fullpage}
 \usepackage{amsmath}
@@
 \usepackage{graphicx}
-\usepackage{slashbox}
+%\usepackage{slashbox}
 \usepackage{color}
@@
 \usepackage{dashrule}
-\usepackage{amsmath}
-
 \usepackage{dcolumn}% Align table columns on decimal point
-
-\usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
+%\usepackage{dcolumn}% Align table columns on decimal point
 \usepackage{bm}% bold math
@@
 \usepackage{calrsfs}
+\usepackage{subfigure}
 \graphicspath{ {images/} }
@@
 The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
-is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable either through a mechanism of Kelvin-Helmholtz instability associated with 
-inflection point or via shear instability of flow above the grass.
-
+is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
+inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
 \end{abstract}
@@
 \newline
-Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.
+Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
 difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
-Due to these considerable differences between aquatic and marine flow the analysis in this paper is applicable only to marine canopies  
+Due to these considerable differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
 \newline   
 Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
-different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz
-instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
+different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
+Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
 %Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
@@
 While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
-several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent 
-theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) 
-for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of
-phenomenon.
+several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
+consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical 
+Reynolds number (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
+further investigation for better understanding of phenomenon.
 \newline 
-In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grasses 
-shows that existence of strong shear near canopy top does not lead to Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
-of fluid flow above vegetation coupled with the flow through canopy. We further show that inclusion of drag due to grass predicts a threshold condition for waving where threshold is
-characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.  
+In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
+shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz 
+instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies arising due to existence of flow boundary.
+We further show that inclusion of drag due to grass predicts a threshold condition for waving where 
+threshold is characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.
 %\section{The model}
@@
 \newline  
-While Drag exerted by canopy is believed to play 
-a significant role in the development of coherent vortices leading to monami, flexibility of grass provide striking visualization of these structures without significantly affecting the
-dynamics of flow (Raupach 1996, Ghisalberti 2002). For this reason we replace grass with stiff structure to investigate dynamics leading to generation of 
-coherent vortices. We use a mean field model for the coupling between the flow and the canopy. 
-%The drag force per unit volume experienced by fluid flow is believed to be dependent on
-%mean frontal area of grass. We approximate mean frontal area of grass field of height $h$ with $dN_g/m^2$. Where $d$ is average width of grass and $N_g$ the grass number density per unit area 
-Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
-%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
-%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
-%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
-% and solving force balance and momentum balance for grass  
+The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
+primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depends weakly on flexibility of grass
+unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
+very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
+Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices.
+We use a mean field model for the coupling between the flow and the canopy.
+Drag force exerted by the grass is approximated via a
+continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
 \begin{equation}
 \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
+\label{navier-stokes}
 \end{equation}
@@
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
-normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. We further expect $C_N$ to
-not change along the height of the grass and take $C_N$ to be constant. 
+normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field such as marine
+both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N \& N_g$ to be constants.
 %\begin{equation}
@@
 \newline
-A solution of equation(1) gives uniform steady flow which we expect to be unstable leading to generate coherent eddies. To understand the instability characteristic associated with this
-steady flow termed as base flow we perform linear stability analysis.  
+The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. steady state flow driven by   
 %We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
-This base state flow is obtained by solving steady state form of constant pressure gradient driven flow of equation(1)
-$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface Fig(1). Flow consist of a region of approximately constant velocity
-$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient with in grass and a parabolic velocity profile in unvegetated region due to 
-balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top give rise to two boundary layer in the flow profile one 
-near bottom and near the tip of the grass which is conventionally known as shear layer.
-%eads to a 
-%constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ while balance of viscous force with pressure gradient above grass provide a parabolic velocity profile.
-%The base flow profile consist of two boundary layer one near the bottom of grass due to no slip condition and other near tip of canopy known as shear layer due to vertical discontinuity 
-%of drag force. Width of boundary layer near bottom of grass can be obtained by balance of viscous force with drag and pressure gradient forces which gives boundary layer
-%width $\sim \sqrt{\frac{\mu}{\rho C_NdN_g U_g}}$. 
-Thickness of shear layer near top of grass is obtained by balance of viscous force 
+constant pressure gradient satisfies
+$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. A sample solution is shown in figure (1) and 
+compared with mean flow observed in lab scale experiment. Flow consist of a region of approximately constant velocity
+\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with in grass and a parabolic velocity 
+profile in unvegetated region due to balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top 
+give rise to two boundary layer in the flow profile one near bottom and other near the tip of the grass. The boundary layer near the canopy top is identified 
+as shear layer in previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
+In our model thickness of shear layer near top of grass is obtained by balance of viscous force 
 with drag along with match of shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
-width $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$ where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.
-
-\begin{figure}[htb!]
-  \includegraphics[scale=0.8]{fig1}
-\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
-\end{figure}
+width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity.
+\newline
 Knowing the scale of shear layer we can systematically investigate its effect on flow instability
@@
 respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
+\small
 \begin{equation}
-%\scriptstyle{
 \begin{split}
- %u_t+Uu_x+vU_y &= \frac{1}{\rho}p_x+\frac{\mu}{\rho}\left(u_xx+u_yy \right) -2C_NdN_gUu\\
 \frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\
@@
 \end{split}
-%}
 \end{equation}
-%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
-%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
-%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
-%of having been produced through equation of motion itself.\newline
-%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
-%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
+\normalsize
 Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.
-%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
-%of motion
-%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
 With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
@@
 \end{equation}
+\begin{figure}[htb]
+  \includegraphics[scale=0.8]{fig1}
+\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
+\end{figure}
+%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
+%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
+%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
+%of having been produced through equation of motion itself.\newline
+%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
+%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
+
+%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
+%of motion
+%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
 where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
-non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
+non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, we get generalized orr-sommerfield equation 
 which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
+\small
 \begin{equation}
@@
 \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
-&-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)
+&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}U\phi_y\right)
+\label{Orr-somerfield}
 \end{split}
 \end{equation}
+\normalsize
 Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2)).
-Variation of critical wave number with shear layer thickness for different grass height is shown in Fig-3.
-% shows 
-%this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
-\begin{figure}[htb!]
+Variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}
+\begin{figure}[htb]
 \includegraphics[]{Critical_Re_vs_delta}
@@
 \newline
-The following features can be noted from Fig(3). For the certain grass height region of linear variation of wavenumber with inverse shear layer thickness suggest that the flow is subject 
-to Kelvin-Helmholtz instability above a threshold shear thickness. Region of constant wave number suggest that length scale of the instability is no longer governed by shear layer
-thickness in this this regime, rather it is governed by flow above the grass, which can be understood by asymptotic analysis in regime of small shear layer thickness and large Reynolds 
-number
-\begin{figure}[htb!]
+Variation of critical wavenumber with shear layer thickness exhibits a region of linear variation of critical wavelength with shear layer thickness showing flow instability arising from
+a mechanism similar to Kelvin-Helmholtz as hypothesized earlier (Ghisalberti, Raupach). However the region of almost no variation of wavelength for a given submergence ratio with 
+shear layer thickness shows the existence of another regime where flow instability does not arise from Kelvin-Helmholtz instability of shear layer rather flow instability is dominated by
+the instability of flow in unvegetated region. The existence of a regime dominated by shear instability of flow above grass is further corroborated by the collapse of eigen-modes 
+for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the grass further predicts the
+asymptotic variation of critical Reynolds number $R_e \sim \delta^{-3/2}$ with shear layer thickness which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+\begin{figure}[htb]
 \includegraphics[]{K_vs_shear_width}
 \caption{Critical Wavenumber for different grass height as function of shear layer thickness}
+\label{K_vs_shear}
 \end{figure}
-%\newline
-Collapse of $\phi$ for different shear thickness in the regime of constant critical wavenumber $k$ suggest that we have obtained the asymptotic solution in limit of small shear layer 
-thickness. Noting that non-dimensional flow velocity $U \sim \delta^{3/2}$ inside the grass and $R_{e} \gg 1$, we can simplify the equation-(4) considerably both with in the grass and 
-above the grass canopy. Inside the grass equation simplifies into $\sigma\left(D^2-k^2\right)\phi = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$. Together with simplified
-force balance inside the grass canopy and constant value of $k\approx O(1)$, we can estimate that critical Reynolds number $R_e \sim \delta^{-3/2}$ for small shear layer thickness Fig(2) 
-\begin{figure}[htb!]
+Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $R_{e} \gg 1$, equation \eqref{Orr-somerfield}
+can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}{R_e} \ll 1$ for the regime of unvegetated flow instability
+ the equation within the grass simplifies into
+into $\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
+$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+\begin{figure}[htb]
  \includegraphics[]{Asymptotic_eig_set3}
 \caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
+\label{Asymptotic_mode}
 \end{figure}
-
-\begin{figure}[htb!]
+\begin{figure}[htb]
 \includegraphics[scale=0.34]{Observed_vs_calculated}
@@
 \end{figure}
-
+\begin{figure}[htb]
+\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
+\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
+\caption{Comparison between observed and calculated velocity profile}
+\end{figure}
+%\begin{figure}[htb]
+%\includegraphics[]{Vivoni_Fig3_7_hb_14match}
+%\caption{Comparison between observed and calculated velocity profile,hb=14}
+%\end{figure}
+%\begin{figure}[htb]
+%\includegraphics[]{Vivoni_Fig3_7_hb_17match}
+%\caption{Comparison between observed and calculated velocity profile,hb=17}
+%\end{figure}
 \end{document}

### 2014-02-27 Paper/grass.tex
UpdateFeb27

@@
-%\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
-\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
+\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
+%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
 %\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
@@
 \graphicspath{ {images/} }
-
 \newcommand{\ellcap}{{\ell_\text{c}}}
@@
 \end{figure}
-\newline
-Variation of critical wavenumber with shear layer thickness exhibits a region of linear variation of critical wavelength with shear layer thickness showing flow instability arising from
-a mechanism similar to Kelvin-Helmholtz as hypothesized earlier (Ghisalberti, Raupach). However the region of almost no variation of wavelength for a given submergence ratio with 
-shear layer thickness shows the existence of another regime where flow instability does not arise from Kelvin-Helmholtz instability of shear layer rather flow instability is dominated by
-the instability of flow in unvegetated region. The existence of a regime dominated by shear instability of flow above grass is further corroborated by the collapse of eigen-modes 
-for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the grass further predicts the
-asymptotic variation of critical Reynolds number $R_e \sim \delta^{-3/2}$ with shear layer thickness which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
 \begin{figure}[htb]
-\includegraphics[]{K_vs_shear_width}
+\includegraphics[]{K_vs_shear_width_noshear}
 \caption{Critical Wavenumber for different grass height as function of shear layer thickness}
@@
 \end{figure}
-Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $R_{e} \gg 1$, equation \eqref{Orr-somerfield}
-can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}{R_e} \ll 1$ for the regime of unvegetated flow instability
- the equation within the grass simplifies into
-into $\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
-$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-\begin{figure}[htb]
- \includegraphics[]{Asymptotic_eig_set3}
+Variation of critical wavenumber with shear layer thickness can be divided into three distinct region. First region, where critical wavelength does not increases linearly with the
+increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region named as shear layer region, where variation in critical
+wavelength is proportional to the variation in shear layer thickness, which can be understood to be arising from a mechanism similar to Kelin-Helmholtz instability observed in the free
+shear flow. Third region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
+submergence ratio and the shear layer thickness experienced by the canopy.
+\begin{figure}[]
+\includegraphics[]{Asymptotic_eig_set3.pdf}
 \caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
@@
 \end{figure}
+\newline
+The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $ <5H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
+Canopy with low submergence ratio can also be understood to be close to the canopy in unbounded flow $h/H\sim0$ such as terrestrial canopy with atmospheric flow. 
+The linear variation of critical wavelength with the shear layer thickness in this regime is also consistent with the observed wavelength of honami for terrestrial canopies.
+However the deviation of critical wavelength from the prediction based on of instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
+regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
+comprising flow in shear layer and flow above the canopies. Whereas constant value of critical wavelength for all the shear layer smaller than a critical values shows the instability 
+in the region-3 is arising from the shear instability of flow above the canopy. The existence of a regime dominated by shear instability of flow above grass is further corroborated 
+by the collapse of eigen-modes for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the 
+grass further predicts that for large drag $R_e \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $R_{e} \gg 1$. So equation \eqref{Orr-somerfield}
+can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}{R_e} \ll 1$,
+the equation within the grass simplifies into
+$\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
+$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. From the balance of coefficients of $\phi_{yy}$ inside the grass we can easily estimate that asymptotically 
+$R_e \sim (\frac{\delta}{H})^{-3/2}$.
 \begin{figure}[htb]
@@
 \caption{Comparison between experimently oberved frequency and predicted frequency}
+\label{Observed_calculated}
 \end{figure}
-\begin{figure}[htb]
-\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
-\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
-\caption{Comparison between observed and calculated velocity profile}
-\end{figure}
+\newline
+Comparison of frequency associated with mode of highest growth predicted by model with that of experimental observations of frequencies in the lab scale experiments is shown in
+Figure \ref{Observed_calculated}.
+The observed frequencies in the lab scale are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage. The agreement between the observed
+frequency and the predicted frequency along with shear layer width suggest that monami is associated with the instability of flow comprising shear layer flow and flow above the canopy.
+The deviation of predicted frequency from the observed one can be attributed to the various simplification we have made in our model. For real canopy the drag coefficients are known to
+decrease from bottom to tip of the grass blades. The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy. Although these variation
+are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
+observed frequency and the predicted frequency.   
+\newline
+Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monai. The occurrence of these three flow instability in a canopy
+depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to Kelvin-Helmholtz
+instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer with flow above the grass.
+Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold Reynolds number is $O(100)$, which is 
+close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the grass filament
+might lead to better agreement between the observed and predicted frequency of monami. 
+\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
+%\cite{Nepf02}\cite{Nepf04}\cite{Nepf08}\cite{Nepf08_2}\cite{Raupach11}\cite{Raupach94}
+%\cite{Delangre06}
+%\begin{figure}[htb]Q
+%\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
+%\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
+%\caption{Comparison between observed and calculated velocity profile}
+%\end{figure}
 %\begin{figure}[htb]
@@
 %\end{figure}
+\bibliography{Grass}{}
+\bibliographystyle{plain}
+%\nocite{*}
 \end{document}

### 2014-03-03 Paper/grass.tex
update March3

@@
 Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
-transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 
-The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami (Ackerman and Okubo, 1993), plays a central role
-in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 
+transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87}\cite{Nepf99}. 
+The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
+in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
 \newline
-Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.
+Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56}\cite{Raupach96}, have also been observed.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
 difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
-Due to these considerable differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
+Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
 \newline   
-Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
+Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02}\cite{Raupach96} (Ghisalberti 2002, Raupach 1996) due to
 different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
@@
 several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
-consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical 
-Reynolds number (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
+consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
+Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
 further investigation for better understanding of phenomenon.
 \newline 
-In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
+In this paper we extend the idea of Raupach et.al\cite{Raupach96} via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
 shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz 
-instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies arising due to existence of flow boundary.
-We further show that inclusion of drag due to grass predicts a threshold condition for waving where 
-threshold is characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.
+instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies and shear instability of flow comprising flow above the canopy 
+and flow in the shear layer. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory 
+with field and experimental observations.
 %\section{The model}
@@
 The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
-primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depends weakly on flexibility of grass
+primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
 unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
 very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
-Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices.
-We use a mean field model for the coupling between the flow and the canopy.
-Drag force exerted by the grass is approximated via a
+Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the 
+flow and the canopy. Drag force exerted by the grass is approximated via a 
 continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
@@
 \newline
-The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. steady state flow driven by   
+The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   
 %We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
@@
 compared with mean flow observed in lab scale experiment. Flow consist of a region of approximately constant velocity
-\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with in grass and a parabolic velocity 
+\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with-in grass and a parabolic velocity 
 profile in unvegetated region due to balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top 
 give rise to two boundary layer in the flow profile one near bottom and other near the tip of the grass. The boundary layer near the canopy top is identified 
-as shear layer in previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
+as a free shear layer in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
 In our model thickness of shear layer near top of grass is obtained by balance of viscous force 
-with drag along with match of shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
-width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity.
+with drag along with matching shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
+width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and
+$H$ is half channel width.
+\begin{figure}[htb]
+  \includegraphics[scale=0.8]{fig1}
+\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
+\end{figure}
 \newline
@@
 \normalsize
-Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0 = \frac{dP/dxH^2}{\mu0}$.
+Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$.
 With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
 into a single equation.
+\small
 \begin{equation}
-\scriptstyle{
-Re\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2D_{drag}}{R_e}U\psi_y\right)
-}
+R\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2}{R}\frac{H^3}{\delta^3}U\psi_y\right)
 \end{equation}
-\begin{figure}[htb]
-  \includegraphics[scale=0.8]{fig1}
-\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
-\end{figure}
+\normalsize
 %In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
@@
 %flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
-
 %Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
@@
 %   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-where $R_{e}= \frac{\rho U_0 H}{\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is 
-non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, we get generalized orr-sommerfield equation 
-which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
+where $R= \frac{\rho U_0 H}{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
+we get generalized Orr-Sommerfield equation 
+which is an eigenvalue problem. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
 \small
@@
 \normalsize
-Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which there is no hydrodynamic instability (Fig(2)).
-Variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}
 \begin{figure}[htb]
@@
 \caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
+\label{Re_vs_delta}
 \end{figure}
+Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which the flow is stable 
+(Figure\ref{Re_vs_delta}). The solution also predicts a critical wavenumber $k$ with maximum growth rate.
+The variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}. This variation of critical 
+wavenumber with shear layer thickness can be divided into three distinct region. First, the region where critical wavelength does not increases linearly with the
+increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region identified as shear layer region, where variation in critical
+wavelength is proportional to the variation in shear layer thickness, the instability in this region can be understood to arise from a mechanism similar to Kelin-Helmholtz instability 
+observed in the free shear flow. Third, the region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
+submergence ratio and the shear layer thickness experienced by the canopy.
 \begin{figure}[htb]
@@
 \end{figure}
-Variation of critical wavenumber with shear layer thickness can be divided into three distinct region. First region, where critical wavelength does not increases linearly with the
-increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region named as shear layer region, where variation in critical
-wavelength is proportional to the variation in shear layer thickness, which can be understood to be arising from a mechanism similar to Kelin-Helmholtz instability observed in the free
-shear flow. Third region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
-submergence ratio and the shear layer thickness experienced by the canopy.
-\begin{figure}[]
-\includegraphics[]{Asymptotic_eig_set3.pdf}
-\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
-\label{Asymptotic_mode}
+\begin{figure}[htb]
+\includegraphics[]{Low_drag_k}
+\caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
+\label{Low_drag_k}
 \end{figure}
 \newline
-The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $ <5H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
-Canopy with low submergence ratio can also be understood to be close to the canopy in unbounded flow $h/H\sim0$ such as terrestrial canopy with atmospheric flow. 
+The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $5 <H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
+Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
 The linear variation of critical wavelength with the shear layer thickness in this regime is also consistent with the observed wavelength of honami for terrestrial canopies.
-However the deviation of critical wavelength from the prediction based on of instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
+However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
 regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
-comprising flow in shear layer and flow above the canopies. Whereas constant value of critical wavelength for all the shear layer smaller than a critical values shows the instability 
+comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
+variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.
+\newline
+The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
 in the region-3 is arising from the shear instability of flow above the canopy. The existence of a regime dominated by shear instability of flow above grass is further corroborated 
-by the collapse of eigen-modes for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the 
+by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
+Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
+below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
+instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). The regime of shear instability of flow above the 
 grass further predicts that for large drag $R_e \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
@@
 $R_e \sim (\frac{\delta}{H})^{-3/2}$.
+\begin{figure}[]
+\includegraphics[]{Asymptotic_eig_set3.pdf}
+\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
+\label{Asymptotic_mode}
+\end{figure}
 \begin{figure}[htb]
@@
 \newline
-Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monai. The occurrence of these three flow instability in a canopy
+Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monami. The occurrence of these three flow instability in a canopy
 depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to Kelvin-Helmholtz
@@
 might lead to better agreement between the observed and predicted frequency of monami. 
-\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
+%\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
 %\cite{Nepf02}\cite{Nepf04}\cite{Nepf08}\cite{Nepf08_2}\cite{Raupach11}\cite{Raupach94}

### 2014-04-10 Paper/grass.tex
"add new graphs"

@@
 (Figure\ref{Re_vs_delta}). The solution also predicts a critical wavenumber $k$ with maximum growth rate.
-The variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}. This variation of critical 
-wavenumber with shear layer thickness can be divided into three distinct region. First, the region where critical wavelength does not increases linearly with the
-increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region identified as shear layer region, where variation in critical
-wavelength is proportional to the variation in shear layer thickness, the instability in this region can be understood to arise from a mechanism similar to Kelin-Helmholtz instability 
-observed in the free shear flow. Third, the region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
-submergence ratio and the shear layer thickness experienced by the canopy.
+The variation of critical wave number with the variation in $\delta$ for different grass height is shown in Figure \ref{K_vs_shear}, where we can observe that above a 
+certain threshold $\delta$ the critical wavelength decreases with the decrease in $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow, for that
+reason this region is identified as shear layer region. We also observe that once the $\delta$ is decreased below the threshold value, there isn't any variation in the 
+critical wavelength with the decrease of shear layer thickness and $\lambda_{critical} \sim (H-h_{grass})$, we identify this region as shear instability of unvegetated region.
+The boundary of transition from instability of shear-layer near the tip of grass to the instability of unvegetated flow depends on the submergence ratio of grass. 
+%    This variation of critical 
+%wavenumber with shear layer thickness can be divided into three distinct region. First, the region where critical wavelength does not increases linearly with the
+%increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region identified as shear layer region, where variation in critical
+%wavelength is proportional to the variation in shear layer thickness, the instability in this region can be understood to arise from a mechanism similar to Kelin-Helmholtz instability 
+%observed in the free shear flow. Third, the region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
+%submergence ratio and the shear layer thickness experienced by the canopy.
 \begin{figure}[htb]

### 2014-08-11 Paper/grass.tex
Pages 1 and 2 half cleaned up

@@
 
+\newcommand{\shreyas}[1]{{\bf (#1)}}
+
 \begin{document}
@@
 Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
-transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87}\cite{Nepf99}. 
+transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
 The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
@@
 \newline
-Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56}\cite{Raupach96}, have also been observed.
+Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
@@
 \newline   
-Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02}\cite{Raupach96} (Ghisalberti 2002, Raupach 1996) due to
+Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
 different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
@@
 \newline  
-The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
-primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
-unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
-very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
-Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the 
-flow and the canopy. Drag force exerted by the grass is approximated via a 
-continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for realizing the shear layer structure of the flow.
+The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability.
+It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation).
+Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated (Py, et al 2006) that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
+
+% The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
+% primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
+% unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
+% very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
+% Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. 
+We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
+continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as
 \begin{equation}
@@
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity. Drag force itself is modeled 
+$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity.
+The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled
 as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
-normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field such as marine
-both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N \& N_g$ to be constants.
+normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass.
+We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis.
+In the field, both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
 %\begin{equation}
@@
 \newline
-The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   
+The steady uniform flow $U(y)$ resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   
 %We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
 constant pressure gradient satisfies
-$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. A sample solution is shown in figure (1) and 
-compared with mean flow observed in lab scale experiment. Flow consist of a region of approximately constant velocity
+$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $??<y<<??$ and $S(y)=0$ in the unvegetated layer $??
+< y<< ??$.
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=??$) and top ($y=??$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}.
+Flow consist of a region of approximately constant velocity
 \small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with-in grass and a parabolic velocity 
-profile in unvegetated region due to balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top 
-give rise to two boundary layer in the flow profile one near bottom and other near the tip of the grass. The boundary layer near the canopy top is identified 
-as a free shear layer in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
-In our model thickness of shear layer near top of grass is obtained by balance of viscous force 
+profile in unvegetated region due to balance of viscous force and pressure gradient. Continuity of flow speed and shear stress at canopy top 
+give rise to a boundary layer in the flow profile near the tip of the grass. The boundary layer near the canopy top may be identified 
+as a free shear layer\cite{Ghisal02} in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
+In our model, the thickness of shear layer near top of grass is obtained by balance of viscous force 
 with drag along with matching shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
-width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and
+thickness $\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and
 $H$ is half channel width.
@@
 \caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
+\label{fig:basicflow}
 \end{figure}
@@
 \begin{split}
-\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\
+\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu\\
 \frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0\\
@@
 \begin{equation}
-R\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2}{R}\frac{H^3}{\delta^3}U\psi_y\right)
+R \left[ (\del_t+\del_x U )\grad^2\psi - U_{yy}\psi_x \right] = \grad^4\psi-\frac{\del}{\del_y}\left({2}\frac{H^3}{\delta^3}S(y) U\psi_y\right)
 \end{equation}
@@
 %   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-where $R= \frac{\rho U_0 H}{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
-we get generalized Orr-Sommerfield equation 
-which is an eigenvalue problem. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
+where $R= {\rho U_0 H}/{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
+we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
 \small
@@
 \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
-&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}U\phi_y\right)
+&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}S(y) U\phi_y\right)
 \label{Orr-somerfield}
@@
 \end{figure}
-\begin{figure}[htb]
-\includegraphics[]{Low_drag_k}
-\caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
-\label{Low_drag_k}
-\end{figure}
+% \begin{figure}[htb]
+% \includegraphics[]{Low_drag_k}
+% \caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
+% \label{Low_drag_k}
+% \end{figure}
 \newline

### 2014-08-11 Paper/grass.tex
More changes to pages 1 and 2

@@
 \title{Waving and swaying}
-\author{Ravi Singh, Shreyas Mandre}
+\author{Ravi Singh}
 \affiliation{Brown University, Providence RI 02912 USA}
@@
 \affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
-%\author{Shreyas Mandre}
-%\affiliation{Brown University, Providence RI 02912 USA}
+\author{Shreyas Mandre}
+\affiliation{Brown University, Providence RI 02912 USA}
 
@@
 \newline
-While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
-several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
+While shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations,
+several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
 consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
-Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
-further investigation for better understanding of phenomenon.
-\newline 
-In this paper we extend the idea of Raupach et.al\cite{Raupach96} via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
-shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz 
-instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies and shear instability of flow comprising flow above the canopy 
-and flow in the shear layer. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory 
-with field and experimental observations.
-%\section{The model}
-%Monami is believed to be response to strong oscillations of flow in streamwise velocity associated with coherent vortices.
-\newline  
+Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$.
+Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
+These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.
+
+In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations.
+Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation.
+We call this new instability the unvegetated layer instability for brevity.
+We further show that inclusion of drag due to grass predicts a threshold condition for waving.
+We confirm our analysis by comparing prediction from theory with field and experimental observations.
+
 The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for realizing the shear layer structure of the flow.

### 2014-08-15 Paper/grass.tex

@@
 by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
-respectively, momentum and mass balance equation expanded in first order of perturbed variable yields.
-\small
+respectively, momentum and mass balance equation expanded in first order of perturbed variable yields
 \begin{equation}
 \begin{split}
-\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu\\
-\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0\\
+\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
+\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
  %\nabla\cdot \bu &= 0
@@
 \end{equation}
-\normalsize
 Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$.
-With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; momentum and mass balance can be combined 
-into a single equation.
-\small
-\begin{equation}
-R \left[ (\del_t+\del_x U )\grad^2\psi - U_{yy}\psi_x \right] = \grad^4\psi-\frac{\del}{\del_y}\left({2}\frac{H^3}{\delta^3}S(y) U\psi_y\right)
-\end{equation}
-\normalsize
-%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
-%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
-%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
-%of having been produced through equation of motion itself.\newline
-%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
-%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
-%Using half channel height $H$ of and associated advection time H/$U_{0}$, we use following scaling to non-dimensinlize equation 
-%of motion
-%   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
-where $R= {\rho U_0 H}/{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
-we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
-\small
+We are left with three non-dimensional parameters in the model, viz.
+the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, and the submergence ratio of the grass $h_g/H$.
+With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance;
+% momentum and mass balance can be combined  into a single equation.
+% \begin{equation}
+% R \left[ (\del_t+\del_x U )\grad^2\psi - U_{yy}\psi_x \right] = \grad^4\psi-\frac{\del}{\del_y}\left({2}\frac{H^3}{\delta^3}S(y) U\psi_y\right).
+% \end{equation}
+Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  we get modified Orr-Sommerfield equation.
+The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
+% \small
 \begin{equation}
@@
 \end{equation}
-\normalsize
+% \normalsize
+
 \begin{figure}[htb]
@@
 \end{figure}
-Solution of generalized Orr-Somerfield equation predicts a critical Reynolds number as function of grass number density, below which the flow is stable 
-(Figure\ref{Re_vs_delta}). The solution also predicts a critical wavenumber $k$ with maximum growth rate.
-The variation of critical wave number with the variation in $\delta$ for different grass height is shown in Figure \ref{K_vs_shear}, where we can observe that above a 
+Solution of \eqref{Orr-somerfield} predicts a critical $R$ as function of grass number density, below which the flow is stable (Re($\sigma$)$<0$) and above which it is unstable (Re($\sigma$)$>0$),
+as shown in Figure\ref{Re_vs_delta}. The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}. 
+\shreyas{Needs language clean up from here on. Describe the most important observation. The details are to be explained in the longer paper.}
+we can observe that above a 
 certain threshold $\delta$ the critical wavelength decreases with the decrease in $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow, for that
@@
 The boundary of transition from instability of shear-layer near the tip of grass to the instability of unvegetated flow depends on the submergence ratio of grass. 
-%    This variation of critical 
-%wavenumber with shear layer thickness can be divided into three distinct region. First, the region where critical wavelength does not increases linearly with the
-%increase of the shear layer thickness, this region occur for canopies with thick shear layer $H/\delta <10$. Second region identified as shear layer region, where variation in critical
-%wavelength is proportional to the variation in shear layer thickness, the instability in this region can be understood to arise from a mechanism similar to Kelin-Helmholtz instability 
-%observed in the free shear flow. Third, the region where critical wavelength does not vary with the variation of shear layer thickness. The boundary of these region depends on 
-%submergence ratio and the shear layer thickness experienced by the canopy.
 \begin{figure}[htb]

### 2014-08-15 Paper/grass.tex
Aug15

@@
 \usepackage{subfigure}
-\graphicspath{ {images/} }
+\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
 \newcommand{\ellcap}{{\ell_\text{c}}}
@@
 
-\section{Introduction}
+%\section{Introduction}
 Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
@@
 Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated (Py, et al 2006) that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated (Py, et al 2006) that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the
+instability of a shear layer.
 
@@
 We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
-continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as
+continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} \shreyas{Where did this model come from?
+Has such a model been used before?
+Citations to those needed.} as
 \begin{equation}
@@
 constant pressure gradient satisfies
-$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $??<y<<??$ and $S(y)=0$ in the unvegetated layer $??
-< y<< ??$.
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=??$) and top ($y=??$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}.
-Flow consist of a region of approximately constant velocity
+$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<h_{grass}$ and $S(y)=0$ in the unvegetated layer $h_{grass}< y< H$. 
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=H$) of the fluid layer, as can be seen from the comparison 
+between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}. Flow consist of a region of approximately constant velocity
 \small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with-in grass and a parabolic velocity 
@@
 \begin{figure}[htb]
-  \includegraphics[scale=0.8]{fig1}
-\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }
+  \includegraphics[scale=1]{Grass_Base_Vivoni_shear}
+\caption{Basic flow profile (left )from the experiment and model and the model of canopy (bottom) along with variation of shear layer (inset) }
 \label{fig:basicflow}
@@
 \shreyas{Needs language clean up from here on. Describe the most important observation. The details are to be explained in the longer paper.}
-we can observe that above a 
-certain threshold $\delta$ the critical wavelength decreases with the decrease in $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow, for that
-reason this region is identified as shear layer region. We also observe that once the $\delta$ is decreased below the threshold value, there isn't any variation in the 
+We observe that above a 
+certain threshold $\delta$ the critical wavenumber is inversely proportional to $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow,
+we identify this region as shear layer region. We also observe that once the $\delta$ is decreased below a critical value, there isn't any variation in the 
 critical wavelength with the decrease of shear layer thickness and $\lambda_{critical} \sim (H-h_{grass})$, we identify this region as shear instability of unvegetated region.
@@
 \end{figure}
-% \begin{figure}[htb]
-% \includegraphics[]{Low_drag_k}
-% \caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
-% \label{Low_drag_k}
-% \end{figure}
+\begin{figure*}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}
+
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens28_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens30_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens34_imgsc}}
+
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens40_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens42_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens46_imgsc}}
+
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens34_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens36_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens38_imgsc}}
+
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with contour of zero growth rate ($Re(\sigma)$=0). The figures in each 
+row corresponds to a particular grass height $h_{grass}/H = 0.8$ for (a-d), 1.0 for (e-h) and  1.2 for (i-l), these figures are arranged in order of increasing grass 
+number density. At low grass number density both the modes are indistinguishable, but as $N_g$ increases the distinction between the two modes become clear. For $N_g$ below
+a critical value, the K-H mode sets the onset of flow instability whereas above the critical $N_{g}$ onset of flow instability is dominated by shear instability of flow above the grass}
+\label{K_Re_sigma_set3}
+\end{figure*}
 \newline 
-The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $5 <H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
-Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
-The linear variation of critical wavelength with the shear layer thickness in this regime is also consistent with the observed wavelength of honami for terrestrial canopies.
-However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
-regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
-comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
-variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.
+The transition of dominant mode of flow instability from K-H mode to shear instability of flow above the grass can be further corroborated from Figure \ref{K_Re_sigma_set3}.
+We observe that the two different modes of instability are indistinguishable at low grass density. The distinction becomes apparent with the increase in grass number density, where we 
+can see that dominant mode of flow instability is 1st dominated by K-H mode before switching to shear instability of flow above the grass at relatively large grass density.  
+%\newline
+%The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness for canopy with low submergence ratio $h/H\leq 0.4$.
+%Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
+%The linear variation of critical wavelength with the shear layer thickness for terrestrial canopy ($h/H \sim 0$) is consistent with the observed wavelength of honami 
+%for terrestrial canopies.
+%However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
+%regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
+%comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
+%variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.
 \newline
-The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
-in the region-3 is arising from the shear instability of flow above the canopy. The existence of a regime dominated by shear instability of flow above grass is further corroborated 
-by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
-Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
-below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
-instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). The regime of shear instability of flow above the 
+%The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
+%in the region-3 is arising from the shear instability of flow above the canopy. 
+%The existence of a regime dominated by shear instability of flow above grass is further corroborated 
+%by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
+%Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
+%below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
+%instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). 
+The regime of shear instability of flow above the 
 grass further predicts that for large drag $R_e \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
@@
 \newline
-Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monami. The occurrence of these three flow instability in a canopy
-depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to Kelvin-Helmholtz
-instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer with flow above the grass.
-Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold Reynolds number is $O(100)$, which is 
-close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the grass filament
+Numerical investigation of flow in presence of canopy reveal two distinct modes of flow instability which might lead to monami. The occurrence of these modes for flow instability in a 
+canopy depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to 
+Kelvin-Helmholtz instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer 
+with flow above the grass. Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold
+ Reynolds number is $O(100)$, which is 
+close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the 
+grass filament
 might lead to better agreement between the observed and predicted frequency of monami.
+
 %\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}

### 2014-08-26 Paper/grass.tex
Added table of comparison with KH

@@
 \usepackage{graphicx}
+\usepackage{booktabs}
 %\usepackage{slashbox}
@@
 \usepackage{subfigure}
+\usepackage{tabularx}
+\usepackage{afterpage}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+
 \graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
@@
 \newcommand{\shreyas}[1]{{\bf (#1)}}
+\definecolor{tableShade}{gray}{0.8}
 
@@
 \newline
-While shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations,
-several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
-consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
-Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$.
+While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory.
+First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent theory.
+Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$.
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
@@
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity.
+$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity.
 The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled
@@
 
+\begin{table*}
+\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
+\renewcommand{\arraystretch}{1.4}
+ \begin{tabular}{l|l|l|l}
+			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
+ Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{l}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
+ Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{l}{$-1<y<1$} \\
+ Inflection point	& exists at $y=0$				& \multicolumn{2}{l}{$U''(y)$ discontinuous at $y=h_{grass}$} \\
+ Shear layer thickness	& $\delta$					& \multicolumn{2}{l}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
+ Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{l}{Equation \eqref{Orr-somerfield}} \\
+ Critical parameters	& none						& $R_e$, $\delta/H$ 	& $R_e \delta^{3/2}$ \\
+ Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
+ Mode localized?	& yes, near $y=0$				& yes, near $y=h_{grass}$			& no
+ \end{tabular}
+ \caption{Comparison between Kelvin-Helmholtz instability and the two modes presented here.}
+\end{table*}
+
 %\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}

### 2014-08-27 Paper/grass.tex
Added a figure
Argued that threshold condition is new

@@
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
-difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
-Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
+difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Ghisal02}.
+Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies. 
 \newline   
@@
 First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent theory.
-Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$.
+Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow.
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
@@
 Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated (Py, et al 2006) that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the
-instability of a shear layer.
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
 

### 2014-08-27 Paper/grass.tex
Explained using Two Modes instead of KH

@@
 Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98}. Another major
 difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Ghisal02}.
@@
 While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory.
-First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent theory.
-Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow.
+First, linear stability analysis\cite{Raupach96} does not take into account influence of drag due to grass hence is not a self consistent theory.
+Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
-These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.
+These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
 
-In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations.
-Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation.
+In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the threshold.
+The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed.
+In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability.
+In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+
+% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation.
 We call this new instability the unvegetated layer instability for brevity.

### 2014-08-28 Paper/grass.tex
figure label

@@
 is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
-inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
+inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed. We further show that in canopy with high submergence ratio
+Kelvin-Helmholtz insatbility is damped due to grass drag.
+
 \end{abstract}
@@
 While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
-First, linear stability analysis\cite{Raupach96} does not take into account influence of drag due to grass hence is not a self consistent theory.
-Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
-Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
+First, linear stability analysis\cite{Raupach96} does not take into account influence of drag due to grass hence is not a self consistent theory. Second, classical free shear flow 
+is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed. 
+Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
+ and therefore inconsistent with the assumptions of the free shear layer instability.
 These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
@@
 In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the threshold.
-The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed.
-In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability.
-In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+ The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. In the limit of a dense grass bed, two distinct unstable modes 
+can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. In the parameter regime typical of marine grass beds, 
+the two modes merge and are indistinguishable.
 
-% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation.
+% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only
+%through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the 
+%unvegetated flow coupled with flow deep within the vegetation.
 We call this new instability the unvegetated layer instability for brevity.
@@
 
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for realizing the shear layer structure of the flow.
-The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability.
-It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation).
-Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for 
+realizing the shear layer structure of the flow. The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from 
+the instability. It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation). 
+Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with 
+the surrounding fluid but does not deform. Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability 
+is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the 
+instability of a shear layer.
 
@@
 We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
-continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} \shreyas{Where did this model come from?
-Has such a model been used before?
-Citations to those needed.} as
+continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} 
+\shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as
 \begin{equation}
@@
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity.
-The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled
+$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to 
+be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 
 as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
-normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass.
-We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis.
-In the field, both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
+normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
+vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
 %\begin{equation}
@@
   \includegraphics[scale=1]{Grass_Base_Vivoni_shear}
-\caption{Basic flow profile (left )from the experiment and model and the model of canopy (bottom) along with variation of shear layer (inset) }
+\caption{Base flow profile (left )from the experiment of Vivoni\cite{Vivoni98} (Case-3A (Figure-3.6A) with 330 plants/$m^2$, 6 blades per plant, 
+plant height = 16.0$\pm0.2$ and blade width of 0.28cm)
+ and it's approximation with $U_0=31cm/s$ and $\delta = 0.2H$ with our model. The model of canopy is shown in the bottom and the variation of shear layer in the inset }
 \label{fig:basicflow}
@@
 Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
-We are left with three non-dimensional parameters in the model, viz.
-the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, and the submergence ratio of the grass $h_g/H$.
-With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance;
+We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, 
+and the submergence ratio of the grass $h_g/H$. With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; 
 % momentum and mass balance can be combined  into a single equation.
@@
 % \end{equation}
-Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  we get modified Orr-Sommerfield equation.
-The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
+Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  
+we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
 % \small

### 2014-08-28 Paper/grass.tex
Added spreadsheet with Ghisalberti's Threshold
The threshold Reynolds number with data from Ghisalberti and Nepf (2002) (heneforth GN2002) is calculated to be compared with our results. The data used from table 1 of GN2002 reports the flow rate, the grass height, and water column depth. This data can be converted in terms of our parameters into a scale for velocity U_0. Assuming a turbulent viscosity of 1e-4 m^2/s, gives a Reynolds number for each case. Reynolds numbers below 200 are reported to be stable, while the ones above 200 are reported to be unstable by GN2002. We can plot these Reynol

@@
 Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98}. Another major
-difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Ghisal02}.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. Another major
+difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Vivoni98,Ghisal02}.
 Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies. 
 \newline   
-Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
+Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
 different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
@@
 \newline
-While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
-First, linear stability analysis\cite{Raupach96} does not take into account influence of drag due to grass hence is not a self consistent theory. Second, classical free shear flow 
+While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
+First, linear stability analysis\cite{Raupach96} does not take into account influence of the vegetation and hence is not self consistent. Second, classical free shear flow 
 is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed. 

### 2014-08-28 Paper/grass.tex
figure recovery

@@
 \newcommand{\del}{\partial}
+\newcommand{\hg}{h_g}
 
@@
 We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
-continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} 
-\shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as
+continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}
@@
 vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
+Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
 %\begin{equation}
@@
 \newline
-The steady uniform flow $U(y)$ resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   
-%We perform linear stability analysis of a base state associated with the vegetated flow to gain insight about instability characteristics of flow in presence of grass.
-constant pressure gradient satisfies
-$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<h_{grass}$ and $S(y)=0$ in the unvegetated layer $h_{grass}< y< H$. 
+For consistency of analysis, we calculate the steady flow as a solution of \eqref{navier-stokes}, and examine its linear stability.
+The steady uniform flow $U(y)$ resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami.
+Steady state flow driven by constant pressure gradient satisfies
+$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< H$. 
 The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=H$) of the fluid layer, as can be seen from the comparison 
@@
 profile in unvegetated region due to balance of viscous force and pressure gradient. Continuity of flow speed and shear stress at canopy top 
-give rise to a boundary layer in the flow profile near the tip of the grass. The boundary layer near the canopy top may be identified 
-as a free shear layer\cite{Ghisal02} in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
-In our model, the thickness of shear layer near top of grass is obtained by balance of viscous force 
-with drag along with matching shear at grass tip exerted by flow in unvegetated region, which leads to shear layer 
-thickness $\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and
-$H$ is half channel width.
+give rise to a boundary layer in the flow profile near the tip of the grass.
+The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
+In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
+thickness $\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$ where $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity and
+$H$ is half channel width. Its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
 \begin{figure}[htb]
@@
 \newline
-Knowing the scale of shear layer we can systematically investigate its effect on flow instability
-by varying grass number density $N_g$ and pressure gradient $\frac{dP}{dx}$. Considering small perturbations $u, v, p$ about the base profile $U$ and $P$
-respectively, momentum and mass balance equation expanded in first order of perturbed variable yields
+Knowing the scale of shear layer we can systematically investigate its effect on flow instability by varying grass number density $N_g$ and pressure gradient ${dP}/{dx}$.
+Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in first order of perturbed variables yields
 \begin{equation}
@@
 Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
-We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, 
-and the submergence ratio of the grass $h_g/H$. With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; 
+We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
+and the submergence ratio of the grass $h_g/H$.
+With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation.
 % momentum and mass balance can be combined  into a single equation.
@@
 % \end{equation}
-Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  
-we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
+The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
 % \small
@@
 \end{equation}
+\shreyas{Boundary conditions?}
 % \normalsize
@@
 
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens28_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens30_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens34_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens38_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens40_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens42_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens46_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens34_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens36_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
 

### 2014-08-29 Paper/grass.tex
Explained threshold, and experimental comparison
Introduction of the Orr-Sommerfeld equation, and its solution. Existence of an instability, and a threshold in Reynolds number. Comparison of that threshold with lab experiments, and of the frequency of the fastest growing mode with experimental measurements.

@@
 \newcommand{\hg}{h_g}
+\newcommand{\Rey}{\text{R}}
 
@@
 in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper.
-\newline
+
 Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
@@
 Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies.
-\newline   
+
 Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
@@
 Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
-%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
-%are exhibited in form of large amplitude oscillations knows as mo-nami.
-\newline
+
 While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
 First, linear stability analysis\cite{Raupach96} does not take into account influence of the vegetation and hence is not self consistent. Second, classical free shear flow 
-is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed. 
+is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
@@
 
-% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only
-%through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the 
-%unvegetated flow coupled with flow deep within the vegetation.
-We call this new instability the unvegetated layer instability for brevity.
-We further show that inclusion of drag due to grass predicts a threshold condition for waving.
-We confirm our analysis by comparing prediction from theory with field and experimental observations.
-
 The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for 
@@
 
-% The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
-% primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
-% unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
-% very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
-% Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. 
 We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
@@
 Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
-%\begin{equation}
-% \mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
-%\end{equation}\
-%\begin{equation}
-%\begin{split}
-% \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
-% \frac{\del M}{\del s}&-N = 0\\
-% M &= B\kappa
-%\end{split}
-%\end{equation}
-%where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
-%\section{Stability Analysis}
-\newline
+
 For consistency of analysis, we calculate the steady flow as a solution of \eqref{navier-stokes}, and examine its linear stability.
@@
 \end{figure}
-\newline
+
 Knowing the scale of shear layer we can systematically investigate its effect on flow instability by varying grass number density $N_g$ and pressure gradient ${dP}/{dx}$.
@@
 \end{equation}
-Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
-We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
+Momentum and mass balance equation can be non-dimensionalized using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
+We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
 and the submergence ratio of the grass $h_g/H$.
 With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation.
-% momentum and mass balance can be combined  into a single equation.
-% \begin{equation}
-% R \left[ (\del_t+\del_x U )\grad^2\psi - U_{yy}\psi_x \right] = \grad^4\psi-\frac{\del}{\del_y}\left({2}\frac{H^3}{\delta^3}S(y) U\psi_y\right).
-% \end{equation}
 The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
@@
 \begin{split}
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi &= \frac{1}{R_{e}}\left[D^2 -k^{2} \right]^2\phi +ikU_{yy}\phi \\
-&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}S(y) U\phi_y\right)
+\left(\frac{\sigma}{ik}+U\right) \left(D^2-k^2\right)\phi &= \frac{1}{ik\Rey}\left[D^2 -k^{2} \right]^2\phi +U_{yy}\phi \\
+&-\frac{\del}{\del y}\left(\frac{2}{ik\Rey}\frac{H^3}{\delta^3}S(y) U\phi_y\right)
 \label{Orr-somerfield}
@@
 \end{figure}
-Solution of \eqref{Orr-somerfield} predicts a critical $R$ as function of grass number density, below which the flow is stable (Re($\sigma$)$<0$) and above which it is unstable (Re($\sigma$)$>0$),
-as shown in Figure\ref{Re_vs_delta}. The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}. 
-\shreyas{Needs language clean up from here on. Describe the most important observation. The details are to be explained in the longer paper.}
-We observe that above a 
-certain threshold $\delta$ the critical wavenumber is inversely proportional to $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow,
-we identify this region as shear layer region. We also observe that once the $\delta$ is decreased below a critical value, there isn't any variation in the 
-critical wavelength with the decrease of shear layer thickness and $\lambda_{critical} \sim (H-h_{grass})$, we identify this region as shear instability of unvegetated region.
-The boundary of transition from instability of shear-layer near the tip of grass to the instability of unvegetated flow depends on the submergence ratio of grass. 
+A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}.
+The dependence of this critical $\Rey$ on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.
+The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}.
+
+Comparison of frequency associated with mode of highest growth predicted by model with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in Figure \ref{Observed_calculated}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage.
+The deviation of predicted frequency from the observed can be attributed to the various simplification we have made in our model.
+In a real canopy, the drag coefficients are known to
+vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy.
+Although these variation
+are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
+observed frequency and the predicted frequency.
+
 \begin{figure}[htb]
@@
 \begin{figure*}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens28_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens30_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens34_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens40_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens42_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens46_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}}
-
-\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens32_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens34_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens36_imgsc}}
+\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens38_imgsc}}
 
 \caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with contour of zero growth rate ($Re(\sigma)$=0). The figures in each 
-row corresponds to a particular grass height $h_{grass}/H = 0.8$ for (a-d), 1.0 for (e-h) and  1.2 for (i-l), these figures are arranged in order of increasing grass 
+row corresponds to a particular grass height $\hg/H = 0.8$ for (a-d), 1.0 for (e-h) and  1.2 for (i-l), these figures are arranged in order of increasing grass 
 number density. At low grass number density both the modes are indistinguishable, but as $N_g$ increases the distinction between the two modes become clear. For $N_g$ below
@@
 \end{figure*}
-\newline 
+
 The transition of dominant mode of flow instability from K-H mode to shear instability of flow above the grass can be further corroborated from Figure \ref{K_Re_sigma_set3}.
@@
 can see that dominant mode of flow instability is 1st dominated by K-H mode before switching to shear instability of flow above the grass at relatively large grass density.
-%\newline
-%The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness for canopy with low submergence ratio $h/H\leq 0.4$.
-%Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
-%The linear variation of critical wavelength with the shear layer thickness for terrestrial canopy ($h/H \sim 0$) is consistent with the observed wavelength of honami 
-%for terrestrial canopies.
-%However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
-%regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
-%comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
-%variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.
-\newline
-%The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
-%in the region-3 is arising from the shear instability of flow above the canopy. 
-%The existence of a regime dominated by shear instability of flow above grass is further corroborated 
-%by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
-%Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
-%below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
-%instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). 
+
 The regime of shear instability of flow above the 
-grass further predicts that for large drag $R_e \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
-Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $R_{e} \gg 1$. So equation \eqref{Orr-somerfield}
-can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}{R_e} \ll 1$,
+grass further predicts that for large drag $\Rey \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $\Rey \gg 1$. So equation \eqref{Orr-somerfield}
+can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}\Rey \ll 1$,
 the equation within the grass simplifies into
-$\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
+$\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}\Rey}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
 $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. From the balance of coefficients of $\phi_{yy}$ inside the grass we can easily estimate that asymptotically 
-$R_e \sim (\frac{\delta}{H})^{-3/2}$.
+$\Rey \sim (\frac{\delta}{H})^{-3/2}$.
 \begin{figure}[]
@@
 \end{figure}
-\newline
-Comparison of frequency associated with mode of highest growth predicted by model with that of experimental observations of frequencies in the lab scale experiments is shown in
-Figure \ref{Observed_calculated}.
-The observed frequencies in the lab scale are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage. The agreement between the observed
-frequency and the predicted frequency along with shear layer width suggest that monami is associated with the instability of flow comprising shear layer flow and flow above the canopy.
-The deviation of predicted frequency from the observed one can be attributed to the various simplification we have made in our model. For real canopy the drag coefficients are known to
-decrease from bottom to tip of the grass blades. The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy. Although these variation
-are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
-observed frequency and the predicted frequency.   
-\newline
+
 Numerical investigation of flow in presence of canopy reveal two distinct modes of flow instability which might lead to monami. The occurrence of these modes for flow instability in a 
@@
  Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{l}{$-1<y<1$} \\
- Inflection point	& exists at $y=0$				& \multicolumn{2}{l}{$U''(y)$ discontinuous at $y=h_{grass}$} \\
+ Inflection point	& exists at $y=0$				& \multicolumn{2}{l}{$U''(y)$ discontinuous at $y=\hg$} \\
  Shear layer thickness	& $\delta$					& \multicolumn{2}{l}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
  Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{l}{Equation \eqref{Orr-somerfield}} \\
- Critical parameters	& none						& $R_e$, $\delta/H$ 	& $R_e \delta^{3/2}$ \\
+ Critical parameters	& none						& $\Rey$, $\delta/H$ 	& $\Rey \delta^{3/2}$ \\
  Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& yes, near $y=h_{grass}$			& no
+ Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no
  \end{tabular}
@@
 
-%\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
-%\cite{Nepf02}\cite{Nepf04}\cite{Nepf08}\cite{Nepf08_2}\cite{Raupach11}\cite{Raupach94}
-%\cite{Delangre06}
-%\begin{figure}[htb]Q
-%\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
-%\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
-%\caption{Comparison between observed and calculated velocity profile}
-%\end{figure}
-%\begin{figure}[htb]
-%\includegraphics[]{Vivoni_Fig3_7_hb_14match}
-%\caption{Comparison between observed and calculated velocity profile,hb=14}
-%\end{figure}
-%\begin{figure}[htb]
-%\includegraphics[]{Vivoni_Fig3_7_hb_17match}
-%\caption{Comparison between observed and calculated velocity profile,hb=17}
-%\end{figure}
 \bibliography{Grass}{}
 \bibliographystyle{plain}
-%\nocite{*}
 \end{document}

### 2014-08-29 Paper/grass.tex
Figure 4 and caption, and title

@@
 \begin{document}
-\title{Waving and swaying}
+\title{Linear instability underlying synchronous waving of marine grass in flow}
 \author{Ravi Singh}
@@
 \begin{figure*}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens28_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens30_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens34_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens38_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens40_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens42_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens46_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
+%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}
 
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens32_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens34_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens36_imgsc}}
-\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens38_imgsc}}
+\begin{tabular}{cccc}
+{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}} &
+{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}} &
+{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}} &
+{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}} &
+{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}} &
+{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
+\end{tabular}
 
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with contour of zero growth rate ($Re(\sigma)$=0). The figures in each 
-row corresponds to a particular grass height $\hg/H = 0.8$ for (a-d), 1.0 for (e-h) and  1.2 for (i-l), these figures are arranged in order of increasing grass 
-number density. At low grass number density both the modes are indistinguishable, but as $N_g$ increases the distinction between the two modes become clear. For $N_g$ below
-a critical value, the K-H mode sets the onset of flow instability whereas above the critical $N_{g}$ onset of flow instability is dominated by shear instability of flow above the grass}
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
+At low grass number density both the modes are indistinguishable, but as $N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
+a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $N_{g}$ onset is determined by Mode 2.}
 \label{K_Re_sigma_set3}

### 2014-08-29 Paper/grass.tex

@@
 
-Comparison of frequency associated with mode of highest growth predicted by model with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in Figure \ref{Observed_calculated}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage.
+Comparison of frequency (Im$(\sigma)$) of the fastest ggrowing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in Figure \ref{Observed_calculated}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \shreyas{cite someone}.
 The deviation of predicted frequency from the observed can be attributed to the various simplification we have made in our model.

### 2014-08-29 Paper/grass.tex
Introduced \epsilon = (\delta/H)^3
Need to systematically replace $\delta/H$ with \epsilon^{1/3} everywhere.

@@
 \begin{equation}
-\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
+\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
 \label{navier-stokes}
@@
 In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
-thickness $\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$ where $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity and
-$H$ is half channel width. Its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
+thickness $\delta = \epsilon^{1/3} H$, where $\epsilon = \left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]$, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and
+$H$ is half channel width.
+The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
 \begin{figure}[htb]
@@
 
-Knowing the scale of shear layer we can systematically investigate its effect on flow instability by varying grass number density $N_g$ and pressure gradient ${dP}/{dx}$.
-Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in first order of perturbed variables yields
+Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, equation \eqref{navier-stokes} expanded to linear order in the perturbations yields
 \begin{equation}
@@
 \end{equation}
-Momentum and mass balance equation can be non-dimensionalized using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
+These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
 We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
@@
 \begin{split}
-\left(\frac{\sigma}{ik}+U\right) \left(D^2-k^2\right)\phi &= \frac{1}{ik\Rey}\left[D^2 -k^{2} \right]^2\phi +U_{yy}\phi \\
-&-\frac{\del}{\del y}\left(\frac{2}{ik\Rey}\frac{H^3}{\delta^3}S(y) U\phi_y\right)
+\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
+&-D\left(2\epsilon^{-1} S(y) U D \phi\right)
 \label{Orr-somerfield}

### 2014-08-29 Paper/grass.tex
Redefined some symbols to help Fig 4
The labels in figure 4 need some redefinition of parameters to help including parameter values in the panels of the figure.

@@
 In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
-thickness $\delta = \epsilon^{1/3} H$, where $\epsilon = \left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]$, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and
-$H$ is half channel width.
+thickness $\delta = \epsilon^{1/3} H$, where $\epsilon = \left[\Rey (C_N d H N_g)\right]^{-1}$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity,
+$H$ is half channel width, and $\mu$ the turbulent viscosity.
 The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
@@
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
-We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
+We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey$, the vegetation drag parametrized by $C_N d H N_g = \Rey/\epsilon$, 
 and the submergence ratio of the grass $h_g/H$.

### 2014-08-29 Paper/grass.tex

@@
 In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
-thickness $\delta = \epsilon^{1/3} H$, where $\epsilon = \left[\Rey (C_N d H N_g)\right]^{-1}$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity,
+thickness $\delta = \Rey^{-1/3} \epsilon^{1/3} H$, where $\epsilon = \left[(C_N d H N_g)\right]^{-1}$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity,
 $H$ is half channel width, and $\mu$ the turbulent viscosity.
@@
 \left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&-D\left(2\epsilon^{-1} S(y) U D \phi\right)
+&-D\left(2R \epsilon^{-1} S(y) U D \phi\right)
 \label{Orr-somerfield}
@@
 \caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
-At low grass number density both the modes are indistinguishable, but as $N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
+As $\Re/\epsilon \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
 a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $N_{g}$ onset is determined by Mode 2.}
@@
 \end{figure*}
+
+To better understand the mechanism of waving, we consider the behavior of the instability as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
+$\epsilon \to 0$.
+\shreyas{comment about the transitions.} To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $kH$, for different $\hg/H$ and $\epsilon$.
+The lowest $\Rey$ on the neutral curve sets the threshold.
+We observe that as $\epsilon$ decreases, the unstable region splits into two.
+We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''.
 

### 2014-08-30 Paper/grass.tex
figure and its inclusion in .tex file
All main ideas are in the text!
Need to read, update figures, and check for consistency. We will not add any new ideas to the paper, unless absolutely necessary.

--- 2014-08-29 Paper/grass.tex
@@
 \newcommand{\Rey}{\text{R}}
+\newcommand{\Ndg}{\tilde{N}_g}
 
@@
 In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
-thickness $\delta = \epsilon^{1/3} H$, where $\epsilon = \left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]$, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and
-$H$ is half channel width.
+thickness $\delta = \Rey^{-1/3} \epsilon^{1/3} H$, where $\Ndg = \left[(C_N d H N_g)\right]$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity,
+$H$ is half channel width, and $\mu$ the turbulent viscosity.
 The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
@@
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
-We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey= {\rho U_0 H}/{\mu}$, the vegetation drag parametrized as $\delta/H$, 
+We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey$, the vegetation drag parametrized by $\Ndg$, 
 and the submergence ratio of the grass $h_g/H$.
+We also parametrize the grass density through the boundary layer thickness $\delta/H$ to help elucidate the mechanism of the instability.
 With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation.
@@
 \left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&-D\left(2\epsilon^{-1} S(y) U D \phi\right)
+&+D\left(2R \Ndg S(y) U D \phi\right),
 \label{Orr-somerfield}
@@
 \end{equation}
-\shreyas{Boundary conditions?}
-% \normalsize
-
+where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=\pm 1$. 
 \begin{figure}[htb]
-\includegraphics[]{Critical_Re_vs_delta}
+\includegraphics[]{Critical_Re_vs_delta_noshear}
 \caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
@@
 A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}.
-The dependence of this critical $\Rey$ on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.
-The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}.
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.
 
-Comparison of frequency (Im$(\sigma)$) of the fastest ggrowing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in Figure \ref{Observed_calculated}.
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
 The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \shreyas{cite someone}.
@@
 
-The transition of dominant mode of flow instability from K-H mode to shear instability of flow above the grass can be further corroborated from Figure \ref{K_Re_sigma_set3}.
-We observe that the two different modes of instability are indistinguishable at low grass density. The distinction becomes apparent with the increase in grass number density, where we 
-can see that dominant mode of flow instability is 1st dominated by K-H mode before switching to shear instability of flow above the grass at relatively large grass density.
+To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
+$\Ndg \gg 1$ and $\delta \ll H$.
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$.
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $kH$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}.
+The lowest $\Rey$ on the neutral curve sets the threshold.
+We observe that as $\Ndg$ increases, the unstable region splits into two.
+We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''.
+The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
 
-The regime of shear instability of flow above the 
-grass further predicts that for large drag $\Rey \sim (\frac{\delta}{H})^{-3/2}$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
-Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $\Rey \gg 1$. So equation \eqref{Orr-somerfield}
-can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}\Rey \ll 1$,
-the equation within the grass simplifies into
-$\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}\Rey}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
-$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. From the balance of coefficients of $\phi_{yy}$ inside the grass we can easily estimate that asymptotically 
-$\Rey \sim (\frac{\delta}{H})^{-3/2}$.
+The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability.
+Mode 2 has a threshold $\Rey \sim (\frac{\delta}{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large $\Ndg$, non-dimensional steady flow velocity $U \sim (\Rey \Ndg)^{-1/2}$ inside the grass and $\Rey \gg 1$. So equation \eqref{Orr-somerfield}
+can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass, the equation within the grass simplifies into
+$\sigma\left(D^2-k^2\right)\phi = -2{(\Rey\Ndg)^{-1/2}}D^2\phi$ while in the unvegetated region it simplifies to the Rayleigh equation 
+$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+The mode shape converges to a solution of these equations as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, as shown in figure \ref{Asymptotic_mode}.
+We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field.
+The only remaining parameter $\Rey \sim (\frac{\delta}{H})^{-3/2}$, therefore sets the threshold.
+
 \begin{figure}[]
@@
 
-Numerical investigation of flow in presence of canopy reveal two distinct modes of flow instability which might lead to monami. The occurrence of these modes for flow instability in a 
-canopy depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to 
-Kelvin-Helmholtz instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer 
-with flow above the grass. Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold
- Reynolds number is $O(100)$, which is 
-close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the 
-grass filament
-might lead to better agreement between the observed and predicted frequency of monami.
+On the other hand, Mode 1 localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold.
+This limit can be understood through asymptotically balancing the magnitude of the advection term ($\Rey/\delta^3 \sim \Rey^2 \Ndg $) with that of the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer ($\delta^{-4} \sim (\Rey \Ndg)^{4/3}$).
+The terms balance when $\Rey \sim \sqrt{\Ndg}$, explaining the asymptote.
+The resulting localization of the the mode structure to a length of $O(\delta)$, is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+
+To conclude, we note that numerical investigation of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami.
+The occurrence of these modes for flow instability in a
+canopy depends on its submergence ratio and shear layer thickness.
+Our analysis predict a threshold criteria characterized by a Reynolds number.
+The predicted value of threshold Reynolds number agrees well with experimentally observed parameters.
+Inclusion of variation of drag coefficient along the grass filament might lead to better agreement between the observed and predicted frequency of monami.
 

### 2014-08-30 Paper/grass.tex
Combined fig showing critical Re and k

@@
 \end{equation}
-%\begin{equation}
-% \mathbf{f}=-N\mathbf{f_{d}}
-%\end{equation}
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=\pm 1$.
+
 \begin{figure}[htb]
-\includegraphics[]{Critical_Re_vs_delta_noshear}
+\begin{center}
+\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{K_vs_shear_width_noshear}
+\end{center}
 \caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
@@
 \end{figure}
+
 A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}.
@@
 
-\begin{figure}[htb]
-\includegraphics[]{K_vs_shear_width_noshear}
-\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
-\label{K_vs_shear}
-\end{figure}
 \begin{figure*}
-
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
-%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}
-
 \begin{tabular}{cccc}
@@
 
-On the other hand, Mode 1 localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold.
+On the other hand, Mode 1 asymptotically localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold.
 This limit can be understood through asymptotically balancing the magnitude of the advection term ($\Rey/\delta^3 \sim \Rey^2 \Ndg $) with that of the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer ($\delta^{-4} \sim (\Rey \Ndg)^{4/3}$).

### 2014-08-30 Paper/grass.tex
Removed observed=predicted figure from text
combined graphs frequency

--- 2014-08-29 Paper/grass.tex
@@
 % \normalsize
-
-\begin{figure}[htb]
-\includegraphics[]{Critical_Re_vs_delta}
-\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
-\label{Re_vs_delta}
+\begin{figure}
+\begin{tabular}{c}
+%\begin{figure}[htb]
+{\centering{\includegraphics[]{Critical_Re_vs_delta_noshear}}}\\
+{\centering{\includegraphics[]{K_vs_shear_width_noshear}}}\\
+%\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
+%\label{Re_vs_delta}
+%\end{figure}
+\end{tabular}
 \end{figure}
@@
 
-\begin{figure}[htb]
-\includegraphics[]{K_vs_shear_width_noshear}
-\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
-\label{K_vs_shear}
-\end{figure}
+%\begin{figure}[htb]
+%\includegraphics[]{K_vs_shear_width_noshear}
+%\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
+%\label{K_vs_shear}
+%\end{figure}
+
 \begin{figure*}
-
 %\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
@@
 %\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}
-
 \begin{tabular}{cccc}
@@
 \end{figure}
-\begin{figure}[htb]
-\includegraphics[scale=0.34]{Observed_vs_calculated}
-\caption{Comparison between experimently oberved frequency and predicted frequency}
-\label{Observed_calculated}
-\end{figure}
+%\begin{figure}[htb]
+%\includegraphics[scale=0.34]{Observed_vs_calculated}
+%\caption{Comparison between experimently oberved frequency and predicted frequency}
+%\label{Observed_calculated}
+%\end{figure}
 

### 2014-08-30 Paper/grass.tex
Some conflicts resolved
Lean preamble

--- 2014-08-30 Paper/grass.tex
@@
-\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 %\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
-%\usepackage{fullpage}
 \usepackage{amsmath}
@@
 \usepackage{booktabs}
-%\usepackage{slashbox}
 \usepackage{color}
@@
 \usepackage{array}
-\usepackage{dashrule}
-\usepackage{dcolumn}% Align table columns on decimal point
-%\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
-\usepackage{ifthen}
-\usepackage{amsthm} % Theorem Formatting
-\usepackage{amssymb}	% Math symbols such as \mathbb
-\usepackage{calrsfs}
-\usepackage{subfigure}
-\usepackage{tabularx}
-\usepackage{afterpage}
+\usepackage{amssymb}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
@@
 \graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
-\newcommand{\ellcap}{{\ell_\text{c}}}
-\newcommand{\Fh}{{\boldsymbol{F}_h}}
-\newcommand{\Fv}{{\boldsymbol{F}_v}}
-\newcommand{\Tv}{{\boldsymbol{T}}}
-\newcommand{\bn}{{\boldsymbol{n}}}
-\newcommand{\bt}{{\boldsymbol{t}}}
-\newcommand{\bz}{{\boldsymbol{z}}}
-\newcommand{\bx}{{\boldsymbol{x}}}
-\newcommand{\by}{{\boldsymbol{y}}}
-\newcommand{\bT}{{\boldsymbol{T}}}
+\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
+\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
 \newcommand{\bu}{\mathbf{u}}
@@
 \newcommand{\Ndg}{\tilde{N}_g}
-
 \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 
-as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\hat{n}+C_{T}\rho u_{T}^{2}d\hat{t}$ where $d$ is average width of grass blade
+as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ where $d$ is average width of grass blade
 $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
-normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
+normal to grass while $\bt,\bn$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
 vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
@@
 \end{figure}
-\begin{figure}[htb]
-\includegraphics[scale=0.34]{Observed_vs_calculated}
-\caption{Comparison between experimently oberved frequency and predicted frequency}
-\label{Observed_calculated}
-\end{figure}
 
@@
 \renewcommand{\arraystretch}{1.4}
- \begin{tabular}{l|l|l|l}
+ \begin{tabular}{l|c|c|c}
 			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
- Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{l}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
- Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{l}{$-1<y<1$} \\
- Inflection point	& exists at $y=0$				& \multicolumn{2}{l}{$U''(y)$ discontinuous at $y=\hg$} \\
- Shear layer thickness	& $\delta$					& \multicolumn{2}{l}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
- Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{l}{Equation \eqref{Orr-somerfield}} \\
+ Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
+ Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
+ Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
+ Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
+ Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
  Critical parameters	& none						& $\Rey$, $\delta/H$ 	& $\Rey \delta^{3/2}$ \\
  Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no
+ Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
  \end{tabular}

### 2014-08-30 Paper/grass.tex

@@
 Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
-transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
-The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
-in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper.
+Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}.
+Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}.
+Such motion is considered to play a central role in the recruitment of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
+The hydrodynamic mechanism underlying monami is the focus of this paper.
 
 Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. Another major
-difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Vivoni98,Ghisal02}.
-Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
+Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
+Due to these differences between aquatic and marine flow, our analysis of flow in presence of vegetation in a bounded domain is applicable only to marine canopies.
 
-Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
-different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
-Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
+Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
+This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
+Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
 
 While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
-First, linear stability analysis\cite{Raupach96} does not take into account influence of the vegetation and hence is not self consistent. Second, classical free shear flow 
-is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
+First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account and hence this explanation is not self consistent.
+Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
@@
 
-In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the threshold.
-The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. In the limit of a dense grass bed, two distinct unstable modes 
-can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. In the parameter regime typical of marine grass beds, 
-the two modes merge and are indistinguishable.
+In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
+The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. 
+The threshold condition for the onset of the instability and the resulting waving frequency compares well with experimental observations.
+In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. 
+In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for 
-realizing the shear layer structure of the flow. The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from 
-the instability. It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation). 
-Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with 
-the surrounding fluid but does not deform. Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability 
-is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the 
-instability of a shear layer.
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is realizes the shear layer structure of the flow.
+The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability. 
+It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation). 
+Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
 
-We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
-continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
+We use a mean field model for the coupling between the flow and the canopy. 
+Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}
@@
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to 
-be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 
-as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ where $d$ is average width of grass blade
-$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively which are set to zero outside the grass; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
-normal to grass while $\bt,\bn$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
-vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.
+$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
+normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
+We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
+In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
 Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
 
-For consistency of analysis, we calculate the steady flow as a solution of \eqref{navier-stokes}, and examine its linear stability.
-The steady uniform flow $U(y)$ resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami.
-Steady state flow driven by constant pressure gradient satisfies
-$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< H$. 
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=H$) of the fluid layer, as can be seen from the comparison 
-between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}. Flow consist of a region of approximately constant velocity
-\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with-in grass and a parabolic velocity 
-profile in unvegetated region due to balance of viscous force and pressure gradient. Continuity of flow speed and shear stress at canopy top 
-give rise to a boundary layer in the flow profile near the tip of the grass.
+For consistency of analysis, we calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
+Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}.
+Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
+Continuity of flow speed and shear stress at canopy top give rise to a boundary layer in the flow profile near the tip of the grass. 
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
-In our model, this boundary layer thickness is obtained by balance of viscous force with drag, and matching the shear with that in unvegetated region. This leads to a shear layer 
-thickness $\delta = \Rey^{-1/3} \epsilon^{1/3} H$, where $\Ndg = \left[(C_N d H N_g)\right]$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity,
-$H$ is half channel width, and $\mu$ the turbulent viscosity.
+In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and $\mu$ the turbulent viscosity.
 The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
+
 \begin{figure}[htb]
@@
 \end{equation}
-These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$ . 
-We are left with three non-dimensional parameters in the model, viz. the Reynolds number $\Rey$, the vegetation drag parametrized by $\Ndg$, 
-and the submergence ratio of the grass $h_g/H$.
-We also parametrize the grass density through the boundary layer thickness $\delta/H$ to help elucidate the mechanism of the instability.
+These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
+We are left with three non-dimensional parameters in the model, viz. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
+We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
 With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation.
@@
 \end{equation}
-where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=\pm 1$.
+where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.
 
@@
 A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}.
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02} without any fitting parameters.
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02}.
 
@@
 The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \shreyas{cite someone}.
-The deviation of predicted frequency from the observed can be attributed to the various simplification we have made in our model.
+The deviation of predicted frequency from the observed may be attributed to the various simplification we have made in our model.
 In a real canopy, the drag coefficients are known to
@@
 To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
-$\Ndg \gg 1$ and $\delta \ll H$.
+$\Ndg \gg 1$ (or $\delta \ll H$).
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $kH$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}.
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$ \shreyas{Make sure to remone $H$ from $kH$ because $k$ is dimensionless.}, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}.
 The lowest $\Rey$ on the neutral curve sets the threshold. 
-We observe that as $\Ndg$ increases, the unstable region splits into two.
-We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''.
+We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
 The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
@@
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim (\frac{\delta}{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
-Noting that for large $\Ndg$, non-dimensional steady flow velocity $U \sim (\Rey \Ndg)^{-1/2}$ inside the grass and $\Rey \gg 1$. So equation \eqref{Orr-somerfield}
-can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass, the equation within the grass simplifies into
-$\sigma\left(D^2-k^2\right)\phi = -2{(\Rey\Ndg)^{-1/2}}D^2\phi$ while in the unvegetated region it simplifies to the Rayleigh equation 
-$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-The mode shape converges to a solution of these equations as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, as shown in figure \ref{Asymptotic_mode}.
-We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field.
-The only remaining parameter $\Rey \sim (\frac{\delta}{H})^{-3/2}$, therefore sets the threshold.
+Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = 2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}. 
+We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
+The only remaining parameter $\Rey/\Ndg$ sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
 
@@
 
-On the other hand, Mode 1 asymptotically localizes with a length scale $\delta$ around the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto \sqrt{\Ndg}$) at the threshold.
-This limit can be understood through asymptotically balancing the magnitude of the advection term ($\Rey/\delta^3 \sim \Rey^2 \Ndg $) with that of the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer ($\delta^{-4} \sim (\Rey \Ndg)^{4/3}$).
-The terms balance when $\Rey \sim \sqrt{\Ndg}$, explaining the asymptote.
-The resulting localization of the the mode structure to a length of $O(\delta)$, is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}.
+This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+
+From these asymptotic behaviors, Mode 1 may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
+Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability.
+Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
 
@@
  \end{tabular}
- \caption{Comparison between Kelvin-Helmholtz instability and the two modes presented here.}
+ \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
+ \label{tab:comparison}
 \end{table*}

### 2014-08-30 Paper/Base_flow_code/Grass_Nepf.m
Fig4 updates

@@ -1,0 +1,72 @@
+function [D_drag_vec,h_grass,U0rhs] = Grass_Nepf(H,Usurf,Uveg,hb,hp,Vivoni_datastruct)
+
+
+        D_drag_vec= [];
+        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+       
+%         D1_drag = (Usurf*5)^2/Uveg^2
+%         D2_drag = Usurf^2/(5*Uveg)^2
+        
+        N = 400;
+        h_grass = (-H+hb)/H;
+        %load Vivoni_Fig_3_6a1_25.dat;
+        U_data = Vivoni_datastruct(:,1);
+        z_data = Vivoni_datastruct(:,2);
+        
+%         U_data = U_data(1:end-8);
+%         z_data = z_data(1:end-8);
+        U_dz_cent = sum( U_data(1:end-1).*abs(z_data(2:end)-z_data(1:end-1)) );
+        U_dz_vec = [U_dz_cent-0.1*U_dz_cent,U_dz_cent,U_dz_cent+0.1*U_dz_cent];
+        
+        
+         for n_udz=2:2
+            U_dz = U_dz_vec(n_udz);
+            %Flow_rate = U_dz*38;
+            U0_upper = Usurf*50;
+            U0_lower = Usurf-Usurf*(49/50.);
+            
+            
+            
+            for j = 1:30
+                
+                U0_mid = (U0_upper+U0_lower)/2.;
+                U0rhs = (U0_mid+U0_upper)/2.;
+                U0lhs = (U0_mid+U0_lower)/2.;
+                Drhs_drag = (U0rhs)^2/Uveg^2;
+                Dlhs_drag = (U0lhs)^2/Uveg^2;
+                
+                
+                [Urhs,uy,uyy,z1] = grass_base3(N,Drhs_drag,h_grass);
+                [Ulhs,uy,uyy,z2] = grass_base3(N,Dlhs_drag,h_grass);
+                z1p = (z1+1)*(H);
+                z2p = (z2+1)*(H);
+                
+                z1p = z1p(z1p<=z_data(end));
+                z2p = z2p(z2p<=z_data(end));
+                Urhs = Urhs(z1p<=z_data(end));
+                Ulhs = Ulhs(z2p<=z_data(end));
+                
[…]

### 2014-08-30 Paper/Base_flow_code/grass_base3.m

@@ -1,0 +1,43 @@
+function [unew,uy,uyy,z] = grass_base3(N,D,h_grass)
+
+
+N = N-1;
+z = linspace(-1,1, N+1)';
+
+dz = z(2)-z(1);
+
+uguess = 0.001*z;
+error = 100000;
+ 
+%while(abs(error)>1.0e-12)
+for ii=1:250
+    
+       e = ones(N+1,1);
+       A = spdiags([e -2*e e], -1:1, N+1, N+1);
+       A(1,2)=2.;
+       Mat = A - spdiags(dz^2*D*abs(uguess).*(z<(h_grass+dz/2.)), 0, N+1,N+1);
+       Mat(N+1,N) = 2; 
+       rhs = 0*z-dz^2;
+       rhs(1) = 0; %rhs(N+1) = 1;
+        
+        unew = Mat\rhs;
+        %plot(z,uguess,z,unew);
+        uguess = uguess*0.5 + 0.5*unew;
+        error = sqrt(sum((uguess-unew).^2))/length(uguess);
+end
+%Mat(N+1,N+1)
+ uy = zeros(length(unew),1); 
+ uyy = zeros(length(unew),1);
+ 
+  for i = 2:length(unew)-1
+     uy(i)  = (unew(i+1)-unew(i-1))/(2*dz);
+     %uyy(i) = (unew(i+1)-2.*unew(i)+unew(i-1))/(dz^2);
+  end
+ 
+%  uy(2:end-1) = (unew(3:end)-unew(1:end-2))/(2*dz);
+%  uyy(2:end-1) = (unew(3:end)-2.*unew(2:end-1)+unew(1:end-2))/(dz^2);
+  uy  = interp1(z(2:end-2),uy(2:end-2),z,'cubic','extrap');
+%  uyy = interp1(z(2:end-2),uyy(2:end-2),z,'cubic','extrap');
+ 
+ uyy = D*unew.*unew.*(z<(h_grass+dz/2.))-1.;
+end

### 2014-08-30 Paper/Base_flow_code/Fit_vivoni_base.m
Fig4 updates

@@ -1,0 +1,375 @@
+D_drag_vec_collection = [];
+
+U0_vec_collection  = [];
+H_grass_vec_collection = [];
+Mut_vec_collection = [];
+Re_vec_collection = [];
+Ng_vec_collection = [];
+H_vec_collection = [];
+
+Uveg = 2.8;
+Usurf = 6.64;
+U_friction = 2.32;
+H = 10.;
+hb = 16.2; 
+hp = 16.0;
+load Vivoni_Fig_3_6a1_25.dat;
+U_data = Vivoni_Fig_3_6a1_25(:,1);
+z_data = Vivoni_Fig_3_6a1_25(:,2);
+
+[D_drag_vec,h_grass,U0] = Grass_vivoni(H,Usurf,Uveg,hb,hp,Vivoni_Fig_3_6a1_25);
+[U,Uy,Uyy,z] = grass_base3(500,D_drag_vec(2),h_grass);
+Num_den_non_dim = 0.76*(H/hp);
+Reynold_num = D_drag_vec(2)/Num_den_non_dim;
+
+D_drag_vec_collection(length(D_drag_vec_collection)+1) = D_drag_vec(2);
+U0_vec_collection(length(U0_vec_collection)+1) = U0;
+H_grass_vec_collection(length(H_grass_vec_collection)+1) = h_grass;
+Ng_vec_collection(length(Ng_vec_collection)+1) = Num_den_non_dim;
+Re_vec_collection(length(Re_vec_collection)+1) = Reynold_num;
+H_vec_collection(length(H_vec_collection)+1) = H;
+
+
+z1p = (z+1)*H/hp;
+lw = .8;
+ms = 1.5;
+
+
+% fileID = fopen('Vivoni_Fig3_6a_125_numerical.dat','w');
+% A = [U'*U0; (z+1)'*(H/hp)];
+% fprintf(fileID,'%6s %12s\n','U','z/hp');
+% fprintf(fileID,'%6.2f %12.8f\n',A);
+% fclose(fileID);
+
+figure(1)
+plot(U*U0,(z+1)*(H/hp),'--b',U_data,z_data,'-bo','LineWidth',lw,'MarkerSize',ms); hold on
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Uveg = 3.07;
+Usurf = 9.03;
+U_friction = 2.40;
+H = 12.;
[…]

### 2014-08-30 Paper/grass.tex
Included figure showing mode shape collapse
Fixed spreadsheet

@@
-\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,reprint,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \affiliation{Harvard University, Cambridge MA 02138 USA}
-\author{Mahesh Bandi\footnote{work performed while visiting Brown University}}
+\author{Mahesh Bandi\thanks{work performed while visiting Brown University}}
 \affiliation{Okinawa Institute of Science and Technology, Okinawa, Japan}
@@
 
+\begin{figure}
+\includegraphics[scale=1]{Grass_Base_Vivoni_shear}
+\caption{Base flow profile (left )from the experiment of Vivoni\cite{Vivoni98} (Case-3A (Figure-3.6A) with 330 plants/$m^2$, 6 blades per plant, 
+plant height = 16.0$\pm0.2$ and blade width of 0.28cm)
+ and it's approximation with $U_0=31cm/s$ and $\delta = 0.2H$ with our model. The model of canopy is shown in the bottom and the variation of shear layer in the inset }
+\label{fig:basicflow}
+\end{figure}
+
 While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{K_vs_shear_width_noshear}
+\end{center}
+\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
+\label{Re_vs_delta}
+\end{figure}
 We use a mean field model for the coupling between the flow and the canopy. 
@@
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
-In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and $\mu$ the turbulent viscosity.
+In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and $\mu$ the turbulent viscosity.
 The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
 
-\begin{figure}[htb]
-\includegraphics[scale=1]{Grass_Base_Vivoni_shear}
-\caption{Base flow profile (left )from the experiment of Vivoni\cite{Vivoni98} (Case-3A (Figure-3.6A) with 330 plants/$m^2$, 6 blades per plant, 
-plant height = 16.0$\pm0.2$ and blade width of 0.28cm)
- and it's approximation with $U_0=31cm/s$ and $\delta = 0.2H$ with our model. The model of canopy is shown in the bottom and the variation of shear layer in the inset }
-\label{fig:basicflow}
-\end{figure}
+\begin{figure*}
+\begin{tabular}{cccc}
+{\includegraphics[scale = 0.68]{Set4_dens28_imgsc}} &
+{\includegraphics[scale = 0.68]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = 0.68]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.68]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.68]{Set5_dens38_imgsc}} &
+{\includegraphics[scale = 0.68]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = 0.68]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.68]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = 0.68]{Set6_dens32_imgsc}} &
+{\includegraphics[scale = 0.68]{Set6_dens34_imgsc}} &
+{\includegraphics[scale = 0.68]{Set6_dens36_imgsc}} &
+{\includegraphics[scale = 0.68]{Set6_dens38_imgsc}}
+\end{tabular}
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
+As $\Re/\epsilon \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
+a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $N_{g}$ onset is determined by Mode 2.}
+\label{K_Re_sigma_set3}
+\end{figure*}
 
@@
 \begin{split}
-\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
-\frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
+% \frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
+% \frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
+ %\nabla\cdot \bu &= 0
+\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
+\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
  %\nabla\cdot \bu &= 0
@@
 \left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&+D\left(2R \Ndg S(y) U D \phi\right),
+&+D\left(2R \Ndg S U D \phi\right),
 \label{Orr-somerfield}
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.
-
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
-\end{center}
-\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
-\label{Re_vs_delta}
-\end{figure}
 
@@
 observed frequency and the predicted frequency.
-
-\begin{figure*}
-\begin{tabular}{cccc}
-{\includegraphics[scale = 0.72]{Set4_dens28_imgsc}} &
-{\includegraphics[scale = 0.72]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.72]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.72]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.72]{Set5_dens38_imgsc}} &
-{\includegraphics[scale = 0.72]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.72]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.72]{Set5_dens46_imgsc}} \\
-{\includegraphics[scale = 0.72]{Set6_dens32_imgsc}} &
-{\includegraphics[scale = 0.72]{Set6_dens34_imgsc}} &
-{\includegraphics[scale = 0.72]{Set6_dens36_imgsc}} &
-{\includegraphics[scale = 0.72]{Set6_dens38_imgsc}}
-\end{tabular}
-
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
-As $\Re/\epsilon \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
-a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $N_{g}$ onset is determined by Mode 2.}
-\label{K_Re_sigma_set3}
-\end{figure*}
 
@@
 
-\begin{figure}[]
-\includegraphics[]{Asymptotic_eig_set3.pdf}
+\begin{figure}
+\includegraphics[]{Asymptotic_noshear}
 \caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}

### 2014-08-30 Paper/Base_flow_code/grass_base3.m

@@ -10,7 +10,7 @@ uguess = 0.001*z;
 error = 100000;
  
 %while(abs(error)>1.0e-12)
-for ii=1:250
+for ii=1:100
     
        e = ones(N+1,1);
        A = spdiags([e -2*e e], -1:1, N+1, N+1);
@@ -36,7 +36,7 @@ end
  
 %  uy(2:end-1) = (unew(3:end)-unew(1:end-2))/(2*dz);
 %  uyy(2:end-1) = (unew(3:end)-2.*unew(2:end-1)+unew(1:end-2))/(dz^2);
-  uy  = interp1(z(2:end-2),uy(2:end-2),z,'cubic','extrap');
+  uy  = interp1(z(2:end-2),uy(2:end-2),z,'pchip','extrap');
 %  uyy = interp1(z(2:end-2),uyy(2:end-2),z,'cubic','extrap');
  
  uyy = D*unew.*unew.*(z<(h_grass+dz/2.))-1.;

### 2014-08-30 Paper/Base_flow_code/shreyas2.m

@@ -1,0 +1,27 @@
+Uveg = 1.0;
+Usurf = 9.4;
+
+H = 46.7/2;
+hb = 13.8; 
+hp = hb;
+
+data = load('Figure_2.png.dat');
+U_data = data(:,1);
+z_data = data(:,2);
+
+[D_drag_vec,h_grass,U0] = Grass_Nepf(H,Usurf,Uveg,hb,hp,data);
+[U,Uy,Uyy,z] = grass_base3(500,D_drag_vec(2),h_grass);
+Num_den_non_dim = 0.76*(H/hp);
+Reynold_num = D_drag_vec(2)/Num_den_non_dim;
+
+% D_drag_vec_collection(length(D_drag_vec_collection)+1) = D_drag_vec(2);
+% U0_vec_collection(length(U0_vec_collection)+1) = U0;
+% H_grass_vec_collection(length(H_grass_vec_collection)+1) = h_grass;
+% Ng_vec_collection(length(Ng_vec_collection)+1) = Num_den_non_dim;
+% Re_vec_collection(length(Re_vec_collection)+1) = Reynold_num;
+% H_vec_collection(length(H_vec_collection)+1) = H;
+
+
+z1p = (z+1)*H/hp;
+lw = .8;
+ms = 1.5;

### 2014-08-31 Paper/grass.tex
Finishing touches - captions

@@
 \includegraphics[scale=1]{Grass_Base_Vivoni_shear}
-\caption{Base flow profile (left )from the experiment of Vivoni\cite{Vivoni98} (Case-3A (Figure-3.6A) with 330 plants/$m^2$, 6 blades per plant, 
-plant height = 16.0$\pm0.2$ and blade width of 0.28cm)
- and it's approximation with $U_0=31cm/s$ and $\delta = 0.2H$ with our model. The model of canopy is shown in the bottom and the variation of shear layer in the inset }
+\caption{Steady flow profile from the experiment of XX\cite{XX} (Case-XXXX from Figure-XX with XX plants/$m^2$, XX blades per plant, 
+plant height = XX$\pm XX$ and blade width of XX cm)
+ and it's approximation with $U_0=XX$ cm/s and $\delta = 0.2H$ with our model.
+The canopy extends upto $y=h_g$ in the water column of depth $3H$.
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thicness $\delta$ near the canopy top.
+The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset.}
 \label{fig:basicflow}
@@
 While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
-First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account and hence this explanation is not self consistent.
+First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account, rendering this explanation inconsistent.
 Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
@@
 
-In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
+In this paper we present a mathematical\shreyas model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
 The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. 
@@
 
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is realizes the shear layer structure of the flow.
-The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability. 
-It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation). 
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it realizes the shear layer structure of the flow. 
+The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
+It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
 Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
@@
 \end{center}
-\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel.
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+Inset compares experimental observations of the dominant frequency $f_o$ (in Hz?) measured with the predictions from the solution of \eqref{Orr-somerfield}.
+The experimental data in the inset in obtained from the aricles by XX\cite{} and XX\cite{}.}
 \label{Re_vs_delta}
@@
 $P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
+The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
 normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
@@
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
-In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity, and $\mu$ the turbulent viscosity.
-The dependence of the boundary layer thickness on the grass density $N_g$ gives us a way to systematically control it and investigate its effect on the instability mechanism.
+In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
+This dependence of the boundary layer thickness on the grass density $N_g$, shown in figure \ref{Grass_Base_Vivoni_shear}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
 
@@
 \caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
-As $\Re/\epsilon \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $N_g$ below
-a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $N_{g}$ onset is determined by Mode 2.}
+As $\Ndg \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $\Ndg$ below
+a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
 \label{K_Re_sigma_set3}
@@
 \begin{split}
-% \frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu, \\
-% \frac{\del v}{\del  t}+ U\frac{\del v}{\del x} &= -\frac{1}{\rho}\frac{\del p}{\del y}+\frac{\mu}{\rho}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
- %\nabla\cdot \bu &= 0
 \rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
 \rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
- %\nabla\cdot \bu &= 0
 \end{split}
@@
 The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
-% \small
 \begin{equation}

### 2014-08-31 Paper/grass.tex
all fig except 4

@@
 \begin{figure}
-\includegraphics[scale=1]{Grass_Base_Vivoni_shear}
-\caption{Steady flow profile from the experiment of XX\cite{XX} (Case-XXXX from Figure-XX with XX plants/$m^2$, XX blades per plant, 
-plant height = XX$\pm XX$ and blade width of XX cm)
- and it's approximation with $U_0=XX$ cm/s and $\delta = 0.2H$ with our model.
-The canopy extends upto $y=h_g$ in the water column of depth $3H$.
+\includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\caption{Steady flow profile from the experiment of Ghisalberti\cite{Nepf04} (Case-I from Table-1 with 1250 plants/$m^2$, 
+plant height = 13.7$\pm .2$cm and blade width of 0.64 cm)
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02cm$ with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
 The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thicness $\delta$ near the canopy top.
@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel.
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
-Inset compares experimental observations of the dominant frequency $f_o$ (in Hz?) measured with the predictions from the solution of \eqref{Orr-somerfield}.
-The experimental data in the inset in obtained from the aricles by XX\cite{} and XX\cite{}.}
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density
+ parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
+also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the aricles by 
+Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
 \label{Re_vs_delta}

### 2014-09-01 Paper/grass.tex
All Figure captions are close to final version

--- 2014-08-31 Paper/grass.tex
@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density parametrized by the boundary layer thickness.
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratios as a function of vegetation density parametrized by the boundary layer thickness.
 Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel.
@@
 Inset compares experimental observations of the dominant frequency $f_o$ (in Hz?) measured with the predictions from the solution of \eqref{Orr-somerfield}.
-The experimental data in the inset in obtained from the aricles by XX\cite{} and XX\cite{}.}
+The experimental data in the inset in obtained from the articles by XX\cite{} and XX\cite{}.}
 \label{Re_vs_delta}
@@
 \end{tabular}
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
-As $\Ndg \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $\Ndg$ below
-a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified grass number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
+As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
+For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
 \label{K_Re_sigma_set3}
@@
 Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = 2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}. 
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}. 
 We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density.
+Mode 1 is shown in solid and Mode 2 is shown in dashed.
+The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$.
+The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached.
+Mode 2 shapes appear self-similar in shape as $\delta\to 0$, and are compared to each other on a rescaled coordinate in the inset.
+Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$.
+The modes approach a universal shape, indicating that an asymptotic limit has been reached.
+The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
 \label{Asymptotic_mode}

### 2014-09-01 Paper/grass.tex

@@
 \begin{figure}
-\includegraphics[scale=1]{Grass_Base_Vivoni_shear}
-\caption{Steady flow profile from the experiment of XX\cite{XX} (Case-XXXX from Figure-XX with XX plants/$m^2$, XX blades per plant, 
-plant height = XX$\pm XX$ and blade width of XX cm)
- and it's approximation with $U_0=XX$ cm/s and $\delta = 0.2H$ with our model.
-The canopy extends upto $y=h_g$ in the water column of depth $3H$.
+\includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\caption{Steady flow profile from the experiment of Ghisalberti\cite{Nepf04} (Case-I from Table-1 with 1250 plants/$m^2$, 
+plant height = 13.7$\pm .2$cm and blade width of 0.64 cm)
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
 The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thicness $\delta$ near the canopy top.
-The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset.}
+The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset and compared with our prediction.}
 \label{fig:basicflow}
@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratios as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel.
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
-Inset compares experimental observations of the dominant frequency $f_o$ (in Hz?) measured with the predictions from the solution of \eqref{Orr-somerfield}.
-The experimental data in the inset in obtained from the articles by XX\cite{} and XX\cite{}.}
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
+ parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
+also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
+Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
 \label{Re_vs_delta}

### 2014-09-01 Paper/grass.tex

@@
 plant height = 13.7$\pm .2$cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02cm$ with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
 The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thicness $\delta$ near the canopy top.
-The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset and compared with our prediction.}
+The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset.}
 \label{fig:basicflow}

### 2014-09-01 Paper/grass.tex
Removed comments from pdf, and added conclusion
The comments are still in the tex file but don't show in the pdf.
Conclusion now states application to other situations like coral reefs,etc.

--- 2014-09-01 Paper/grass.tex
@@
 \newcommand{\Ndg}{\tilde{N}_g}
-\newcommand{\shreyas}[1]{{\bf (#1)}}
+% \newcommand{\shreyas}[1]{{\bf (#1)}}
+\newcommand{\shreyas}[1]{}
 \definecolor{tableShade}{gray}{0.8}
@@
 In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
-This dependence of the boundary layer thickness on the grass density $N_g$, shown in figure \ref{Grass_Base_Vivoni_shear}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
+This dependence of the boundary layer thickness on the grass density $N_g$, shown in figure \ref{fig:basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
 
@@
 
-To conclude, we note that numerical investigation of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami.
-The occurrence of these modes for flow instability in a
-canopy depends on its submergence ratio and shear layer thickness.
+In conclusion we note that linear stability analysis of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami. 
+The occurrence of these modes for flow instability in a canopy depends on its submergence ratio and shear layer thickness.
 Our analysis predict a threshold criteria characterized by a Reynolds number.
@@
 Inclusion of variation of drag coefficient along the grass filament might lead to better agreement between the observed and predicted frequency of monami. 
-
+Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
 \begin{table*}

### 2014-09-01 Paper/grass.tex
all

--- 2014-09-01 Paper/grass.tex
@@
 The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset.}
-\label{fig:basicflow}
+\label{basicflow}
 \end{figure}
@@
 First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account, rendering this explanation inconsistent.
-Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed.
+Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold
+flow speed below which waving is not observed. 
 Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
@@
 
-In this paper we present a mathematical\shreyas model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
+In this paper we present a mathematical \shreyas model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
 The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. 
@@
 The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
-It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
+It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
 Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
+the flow instability is similar to the instability of a shear layer.
 
@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density
  parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
 also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the aricles by 
 Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
@@
 $P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
+The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as 
+$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
+where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
 normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
@@
 In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
-This dependence of the boundary layer thickness on the grass density $N_g$, shown in figure \ref{Grass_Base_Vivoni_shear}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
+This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
 
@@
 \end{tabular}
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified grass number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
-As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
-For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.  
+As $\Ndg \propto N_g$ increases the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. For $\Ndg$ below
+a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
 \label{K_Re_sigma_set3}
@@
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \shreyas{cite someone}.
-The deviation of predicted frequency from the observed may be attributed to the various simplification we have made in our model.
-In a real canopy, the drag coefficients are known to
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
+Figure \ref{Re_vs_delta}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
+The deviation of predicted frequency from the observed may be attributed to the various
+simplification we have made in our model. In a real canopy, the drag coefficients are known to
 vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy.
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
 Although these variation
@@
 Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = 2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}. 
+Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}. 
 We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density.
-Mode 1 is shown in solid and Mode 2 is shown in dashed.
-The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$.
-The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached.
-Mode 2 shapes appear self-similar in shape as $\delta\to 0$, and are compared to each other on a rescaled coordinate in the inset.
-Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$.
-The modes approach a universal shape, indicating that an asymptotic limit has been reached.
-The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+\caption{plot of $|\phi|$ for different shear layer thickness in limit of small shear layer thickness}
 \label{Asymptotic_mode}

### 2014-09-01 Paper/grass.tex
grass bbl

@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ration as a function of vegetation density
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
  parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
 also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the aricles by 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
 Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
@@
 This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}.
+
 We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.

### 2014-09-02 Paper/grass.tex
Abstract update

@@
 
-The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
-is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
-inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed. We further show that in canopy with high submergence ratio
-Kelvin-Helmholtz insatbility is damped due to grass drag.
+The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
+ Our analysis of flow through grass bed predicts a critical threshold condition for instability and associated waving frequency
+which compares well with experimental observations. We show that in the limit of a dense grass bed, two distinct unstable modes (Mode-1, Mode-2) can be asymptotically distinguished.
+while Mode-1 shares many characteristics with the Kelvin-Helmholtz instability where shear layer near canopy top plays a critical role, 
+Mode-2 can be thought of as instability of inviscid flow with dense grass bed for which shear 
+layer near grass tip doesn't play a critical role. We further show that in canopy with high submergence ratio
+Kelvin-Helmholtz insatbility is damped due to grass drag. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+
+% Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
+%inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
 

### 2014-09-03 Paper/grass.tex
abstact update september 3rd

@@
 
-The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
+%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
+We study the mechanism behind synchronous waving of marine grass, known as Mo-nami. We do linear stability analysis of a mathematical model which considers grass to exert a continuum drag on the flow.
 Our analysis of flow through grass bed predicts a critical threshold condition for instability and associated waving frequency
-which compares well with experimental observations. We show that in the limit of a dense grass bed, two distinct unstable modes (Mode-1, Mode-2) can be asymptotically distinguished.
-while Mode-1 shares many characteristics with the Kelvin-Helmholtz instability where shear layer near canopy top plays a critical role, 
-Mode-2 can be thought of as instability of inviscid flow with dense grass bed for which shear 
-layer near grass tip doesn't play a critical role. We further show that in canopy with high submergence ratio
-Kelvin-Helmholtz insatbility is damped due to grass drag. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+which compares well with experimental observations. Our analysis further predicts two distinct mode of instability which might lead to Mo-nami.
+One of these mode shares many characteristics with the Kelvin-Helmholtz instability, we termed this mode as ``Mode 1''. The other mode can be thought of as instability of 
+inviscid flow above grass coupled with the porous flow through grass bed, we refer to this mode as ``Mode 2''.  In the limit of dense grass bed these two modes can be asymptotically 
+distinguished. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
@@
 \maketitle
-
 %\section{Introduction}

### 2014-09-03 Paper/grass.tex
abstact update september 3rd

@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-We study the mechanism behind synchronous waving of marine grass, known as Mo-nami. We do linear stability analysis of a mathematical model which considers grass to exert a continuum drag on the flow.
-Our analysis of flow through grass bed predicts a critical threshold condition for instability and associated waving frequency
-which compares well with experimental observations. Our analysis further predicts two distinct mode of instability which might lead to Mo-nami.
-One of these mode shares many characteristics with the Kelvin-Helmholtz instability, we termed this mode as ``Mode 1''. The other mode can be thought of as instability of 
-inviscid flow above grass coupled with the porous flow through grass bed, we refer to this mode as ``Mode 2''.  In the limit of dense grass bed these two modes can be asymptotically 
-distinguished. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass. Flow through grass bed gives rise to a region of strong shear near tip of grass 
+known as shear layer, due to vertical discontinuity of grass drag. Linear stability analysis of a mathematical model incorporating grass drag as a continuum field has been 
+carried out. Our analysis predicts a critical threshold condition for flow instability leading to Mo-nami and associated waving frequency
+which compares well with experimental observations. Further analysis predicts existence of two distinct mode of instability.
+One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer plays a central role, we termed this as ``Mode 1''. 
+The other mode can be interpreted as instability of 
+inviscid flow through a grass bed where shear layer doesn't play a critical role, we refer this as ``Mode 2''. In the limit of dense grass bed these two modes can be
+distinguished. For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
@@
 Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}.
-Such motion is considered to play a central role in the recruitment of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
+Such motion is considered to play a central role in the recruitment
+of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
 The hydrodynamic mechanism underlying monami is the focus of this paper.

### 2014-09-03 Paper/grass.tex
update abstract

@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass. Flow through grass bed gives rise to a region of strong shear near tip of grass 
-known as shear layer, due to vertical discontinuity of grass drag. Linear stability analysis of a mathematical model incorporating grass drag as a continuum field has been 
+We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass.
+Linear stability analysis of a mathematical model incorporating grass drag as a continuum field has been
 carried out. Our analysis predicts a critical threshold condition for flow instability leading to Mo-nami and associated waving frequency
 which compares well with experimental observations. Further analysis predicts existence of two distinct mode of instability.
-One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer plays a central role, we termed this as ``Mode 1''. 
+One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role, we termed this as ``Mode 1''.
 The other mode can be interpreted as instability of 
-inviscid flow through a grass bed where shear layer doesn't play a critical role, we refer this as ``Mode 2''. In the limit of dense grass bed these two modes can be
+inviscid flow through a grass bed where shear layer doesn't play an important role, we refer this as ``Mode 2''. In the limit of dense grass bed these two modes can be
 distinguished. For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

### 2014-09-03 Paper/grass.tex
Update grass.tex

@@
 which compares well with experimental observations. Further analysis predicts existence of two distinct mode of instability.
-One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role, we termed this as ``Mode 1''.
+One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to drag discontinuity plays a critical role, we termed this as ``Mode 1''.
 The other mode can be interpreted as instability of 

### 2014-09-04 Paper/grass.tex
abstract update sep 4

@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass.
-Linear stability analysis of a mathematical model incorporating grass drag as a continuum field has been
+We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
+drag as a continuum field has been 
 carried out. Our analysis predicts a critical threshold condition for flow instability leading to Mo-nami and associated waving frequency
-which compares well with experimental observations. Further analysis predicts existence of two distinct mode of instability.
-One of these mode has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to drag discontinuity plays a critical role, we termed this as ``Mode 1''.
-The other mode can be interpreted as instability of 
-inviscid flow through a grass bed where shear layer doesn't play an important role, we refer this as ``Mode 2''. In the limit of dense grass bed these two modes can be
-distinguished. For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
+%Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
+%Mode 2 can be interpreted as instability of 
+%inviscid flow through a grass bed where shear layer doesn't play an important role. 
+The occurrence of these modes for flow instability depends on submergence ratio and the grass 
+number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
+scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
+For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 

### 2014-09-05 Paper/grass.tex
Moved comparison with terrestrial to discussion

@@
 The hydrodynamic mechanism underlying monami is the focus of this paper.
-
-Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
-Due to these differences between aquatic and marine flow, our analysis of flow in presence of vegetation in a bounded domain is applicable only to marine canopies.
 
@@
 
+Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
+Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
+Due to these differences between aquatic and marine flow, our analysis of flow in presence of vegetation in a bounded domain is applicable only to marine canopies.
+
 In conclusion we note that linear stability analysis of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami. 

### 2014-09-05 Paper/grass.tex
Comparison with terrestrial case made.
Also fixed missing figure reference.

@@
 Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}.
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}.
 Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
@@
 Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
-Due to these differences between aquatic and marine flow, our analysis of flow in presence of vegetation in a bounded domain is applicable only to marine canopies.
+Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
+We find that in this case, the transition from Mode 1 to Mode 2 appens at such a large grass density, so as to make Mode 2 irrelevant.
+We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.
 

### 2014-09-05 Paper/grass.tex
pull from shreyas
Added the emphasis on threshold condition in intro

@@
 
-In this paper we present a mathematical \shreyas model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
-The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. 
-The threshold condition for the onset of the instability and the resulting waving frequency compares well with experimental observations.
-In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. 
-In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
+The synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
+A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
+The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is a primary consideration that sets the threshold flow.
+We deduce this numerically and through an asymptotic analysis of the dense grass limit.
 
@@
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; departures from this scaling can be seen in the inset of figure \ref{basicflow}.
 
@@
 
+\begin{table*}
+\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
+\renewcommand{\arraystretch}{1.4}
+ \begin{tabular}{l|c|c|c}
+			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
+ Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
+ Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
+ Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
+ Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
+ Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
+ Critical parameters	& none						& $\Rey$, $\delta/H$ 	& $\Rey \delta^{3/2}$ \\
+ Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
+ Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
+ \end{tabular}
+ \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
+ \label{tab:comparison}
+\end{table*}
+
 To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
@@
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
-\begin{table*}
-\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
-\renewcommand{\arraystretch}{1.4}
- \begin{tabular}{l|c|c|c}
-			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
- Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
- Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
- Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
- Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
- Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
- Critical parameters	& none						& $\Rey$, $\delta/H$ 	& $\Rey \delta^{3/2}$ \\
- Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
- \end{tabular}
- \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
- \label{tab:comparison}
-\end{table*}
 

### 2014-09-06 Paper/grass.tex
The role of critical Reynolds

@@
 For consistency of analysis, we calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}.
-Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
+Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the 
+vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, 
+as can be seen from the comparison between sample solutions and mean flow observed in 
+laboratory experiment shown in Figure \ref{basicflow}. 
+Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of 
+drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient. 
 Continuity of flow speed and shear stress at canopy top give rise to a boundary layer in the flow profile near the tip of the grass. 
@@
 A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02}.
+The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
+due to grass drag. Indeed a calculation undertaken for terrestrial grass shows the suppression of growth rate of instability
+due to grass drag\cite{Delangre06}. It is important to notice the absence of a threshold condition in the previous analysis of 
+free shear flow with various assumptions about mean flow profile and the role of grass drag\cite{Raupach96,Delangre06}.
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
+on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
+threshold\cite{Ghisal02}.
 

### 2014-09-06 Paper/grass.tex
Tried to change the message of the paper
Focus on the threshold and the presence of grass drag in the mechanism.

--- 2014-09-05 Paper/grass.tex
@@
 and therefore inconsistent with the assumptions of the free shear layer instability. 
-These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
+% These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
 
@@
  parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
-also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
-Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
+also included in the top panel. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
+Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
 \label{Re_vs_delta}
@@
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
-where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
+where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
 normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
@@
 Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
-Continuity of flow speed and shear stress at canopy top give rise to a boundary layer in the flow profile near the tip of the grass. 
+Continuity of flow speed and shear stress at canopy top gives rise to a boundary layer in the flow profile near the tip of the grass. 
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
@@
 
-A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the threshold\cite{Ghisal02}.
+A threshold in $\Rey$ of O(100), above which  the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the solution of \eqref{Orr-somerfield}. 
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
+The threshold Reynolds number increases with the grass density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+We examine this competition more carefully using dominant balances in the limit of dense grass.
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
-Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
-The deviation of predicted frequency from the observed may be attributed to the various
-simplification we have made in our model. In a real canopy, the drag coefficients are known to
-vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
+The deviation of predicted frequency from the observed may be attributed to the various simplification we have made in our model. 
+In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
 The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
-Although these variation
-are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
-observed frequency and the predicted frequency.
+Although these variation are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the observed frequency and the predicted frequency.
 
@@
 
-To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
-$\Ndg \gg 1$ (or $\delta \ll H$).
+To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$ \shreyas{Make sure to remone $H$ from $kH$ because $k$ is dimensionless.}, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}.
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
 The lowest $\Rey$ on the neutral curve sets the threshold. 
@@
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
-Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}.
+Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield}.
+Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+Using this estimate, and $\Rey\gg 1$, simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}.
 
@@
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}.
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
 This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
@@
 Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
+The dominant role of vegetation drag in the mechanism for both the modes distinguishes them from the traditional Kelvin-Helmholtz instability.
 
@@
 
-In conclusion we note that linear stability analysis of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami. 
+In conclusion we note that linear stability analysis of flow in presence of canopy reveals the existence of a threshold flow for the onset of the instability.
+%two distinct modes of flow instability. 
 The occurrence of these modes for flow instability in a canopy depends on its submergence ratio and shear layer thickness.

### 2014-09-06 Paper/grass.tex
Reynolds section update

--- 2014-09-06 Paper/grass.tex
@@
 due to grass drag. Indeed a calculation undertaken for terrestrial grass shows the suppression of growth rate of instability
-due to grass drag\cite{Delangre06}. It is important to notice the absence of a threshold condition in the previous analysis of 
-free shear flow with various assumptions about mean flow profile and the role of grass drag\cite{Raupach96,Delangre06}.
+due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
+as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis \cite{Raupach96,Delangre06}.
 The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 

### 2014-09-06 Paper/grass.tex

--- 2014-09-06 Paper/grass.tex
@@
 Continuity of flow speed and shear stress at canopy top gives rise to a boundary layer in the flow profile near the tip of the grass.
+
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
@@
 The threshold Reynolds number increases with the grass density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-We examine this competition more carefully using dominant balances in the limit of dense grass.
+Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
+Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
 
@@
 Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
-We find that in this case, the transition from Mode 1 to Mode 2 appens at such a large grass density, so as to make Mode 2 irrelevant.
+We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
 We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.

### 2014-09-06 Paper/grass.tex
Improved discussion and conclusion
The comparison with frequency is dealt with first so as to combine the discussion of the threshold into one coherent section. The note about disagreement due to variation of grass density and eddy viscosity with depth is moved to a paragraph on limitations near the end of the paper. And the conclusion is improved to summarize the paper.

@@
 
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+
 A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the solution of \eqref{Orr-somerfield}. 
@@
 Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
-
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
-The deviation of predicted frequency from the observed may be attributed to the various simplification we have made in our model. 
-In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
-Although these variation are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the observed frequency and the predicted frequency.
 
@@
 
-In conclusion we note that linear stability analysis of flow in presence of canopy reveals the existence of a threshold flow for the onset of the instability.
-%two distinct modes of flow instability. 
-The occurrence of these modes for flow instability in a canopy depends on its submergence ratio and shear layer thickness.
-Our analysis predict a threshold criteria characterized by a Reynolds number.
-The predicted value of threshold Reynolds number agrees well with experimentally observed parameters. 
-Inclusion of variation of drag coefficient along the grass filament might lead to better agreement between the observed and predicted frequency of monami. 
+The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
+In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
+Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
+
+In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag.
+We show the origin of this threshold using a consistent continuum hydodynamic stabiity analysis for the shear flow that develops as a result of the drag.
+Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism.
+The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

### 2014-09-06 Paper/grass.tex
Globally changed monami, mo-nami, Mo-nami to \textit{monami}

@@
 \newcommand{\Ndg}{\tilde{N}_g}
+\newcommand{\monami}{\textit{monami}}
 % \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
+We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
 drag as a continuum field has been 
-carried out. Our analysis predicts a critical threshold condition for flow instability leading to Mo-nami and associated waving frequency
+carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
 which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
@@
 Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}.
-Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}.
+Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.
 Such motion is considered to play a central role in the recruitment
 of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
-The hydrodynamic mechanism underlying monami is the focus of this paper.
+The hydrodynamic mechanism underlying \monami is the focus of this paper.
 
-Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
+Existing explanation of \monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
 This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
@@
 
-While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
+While the shear layer model successfully appears to predict frequency of \monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
 First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account, rendering this explanation inconsistent.
@@
 
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it realizes the shear layer structure of the flow. 
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
 The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
@@
 
-The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
+The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
 In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
@@
 Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
 

### 2014-09-06 Paper/grass.tex
Abstract changed
Included consistent notation for "monami"

@@
 \newcommand{\Ndg}{\tilde{N}_g}
-\newcommand{\monami}{\textit{monami}}
+\newcommand{\monami}{\textit{monami }}
 % \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
-drag as a continuum field has been 
-carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
-which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
+%We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
+%drag as a continuum field has been 
+%carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
+%which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
 %Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
@@
 %inviscid flow through a grass bed where shear layer doesn't play an important role. 
-The occurrence of these modes for flow instability depends on submergence ratio and the grass 
-number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
-scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
-For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+%The occurrence of these modes for flow instability depends on submergence ratio and the grass 
+%number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
+%scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
+%For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
@@
 
+We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow.
+We show this by developing a self-consistent model of the flow past rigid vegetation.
+Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and as well as the origin of the mechanistic threshold.
+Asymptotic analysis of the model for dense grass reveals two unstable modes.
+One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it.
+The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations.
 \end{abstract}

### 2014-09-07 Paper/grass.tex
Abstract changed
more update on critical R

--- 2014-09-06 Paper/grass.tex
@@
 and therefore inconsistent with the assumptions of the free shear layer instability. 
-% These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
+These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
 
@@
  parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
-also included in the top panel. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
-Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+also included in the top panel. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from the articles by 
+Ghisalberti \cite{Ghisal02} and Vivoni \cite{Vivoni98}.}
 \label{Re_vs_delta}
@@
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
-where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
+where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and
 normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
@@
 For consistency of analysis, we calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}.
-Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
-Continuity of flow speed and shear stress at canopy top gives rise to a boundary layer in the flow profile near the tip of the grass.
-
+Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the 
+vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
+The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, 
+as can be seen from the comparison between sample solutions and mean flow observed in 
+laboratory experiment shown in Figure \ref{basicflow}. 
+Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of 
+drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient. 
+Continuity of flow speed and shear stress at canopy top give rise to a boundary layer in the flow profile near the tip of the grass. 
 The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
@@
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
+The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
+due to grass drag. Indeed a calculation undertaken for terrestrial canopy shows the suppression of growth rate of instability
+due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
+as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis 
+\cite{Raupach96,Delangre06}.
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
+on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
+threshold\cite{Ghisal02}. The good agreement between predicted and estimated threshold suggest that the grass drag plays a 
+crucial role in setting up the flow instability.
 
-A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
-The threshold Reynolds number increases with the grass density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
-Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
+Figure \ref{Re_vs_delta}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
+The deviation of predicted frequency from the observed may be attributed to the various
+simplification we have made in our model. In a real canopy, the drag coefficients are known to
+vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
+Although these variation
+are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
+observed frequency and the predicted frequency.
 
@@
 
-To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
+To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes asymptotically large, i.e.
+$\Ndg \gg 1$ (or $\delta \ll H$).
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$ \shreyas{Make sure to remone $H$ from $kH$ because $k$ is dimensionless.}, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}.
 The lowest $\Rey$ on the neutral curve sets the threshold. 
@@
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield}.
-Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-Using this estimate, and $\Rey\gg 1$, simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}.
+Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}.
 
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density.
-Mode 1 is shown in solid and Mode 2 is shown in dashed.
-The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$.
-The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached.
-Mode 2 shapes appear self-similar in shape as $\delta\to 0$, and are compared to each other on a rescaled coordinate in the inset.
-Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$.
-The modes approach a universal shape, indicating that an asymptotic limit has been reached.
-The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness 
+for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. 
+The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) 
+but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that 
+the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, 
+and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for 
+Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit 
+has been reached. The limit is not yet reached for the case $\delta/H = 0.05$ due to the influence of bottom boundary; 
+the grass height in this case is comparable to the boundary layer thickness.}
 \label{Asymptotic_mode}
@@
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}.
 This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
@@
 Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
-The dominant role of vegetation drag in the mechanism for both the modes distinguishes them from the traditional Kelvin-Helmholtz instability.
 
@@
 Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
-We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
+We find that in this case, the transition from Mode 1 to Mode 2 appens at such a large grass density, so as to make Mode 2 irrelevant.
 We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.
 
-The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
-In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
-Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-
-In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag.
-We show the origin of this threshold using a consistent continuum hydodynamic stabiity analysis for the shear flow that develops as a result of the drag.
-Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism.
-The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top.
+In conclusion we note that linear stability analysis of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami. 
+The occurrence of these modes for flow instability in a canopy depends on its submergence ratio and shear layer thickness.
+Our analysis predict a threshold criteria characterized by a Reynolds number.
+The predicted value of threshold Reynolds number agrees well with experimentally observed parameters. 
+Inclusion of variation of drag coefficient along the grass filament might lead to better agreement between the observed and predicted frequency of monami. 
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

### 2014-09-07 Paper/grass.tex

--- 2014-09-06 Paper/grass.tex
@@
 \newcommand{\Ndg}{\tilde{N}_g}
-\newcommand{\monami}{\textit{monami }}
+\newcommand{\monami}{\textit{monami}}
 % \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+%<<<<<<< HEAD
+% A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
+% The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
+% due to grass drag. Indeed a calculation undertaken for terrestrial canopy shows the suppression of growth rate of instability
+% due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
+% as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis 
+% \cite{Raupach96,Delangre06}.
+% The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
+% on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
+% threshold\cite{Ghisal02}. The good agreement between predicted and estimated threshold suggest that the grass drag plays a 
+% crucial role in setting up the flow instability.
+% 
+% Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
+% Figure \ref{Re_vs_delta}.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
+The deviation of predicted frequency from the observed may be attributed to the various
+% simplification we have made in our model. In a real canopy, the drag coefficients are known to
+% vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
+Although these variation
+% are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
+% observed frequency and the predicted frequency.
+% =======
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
+the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, 
+is also shown in Figure \ref{Re_vs_delta}. 
+The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of 
+vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
 
-A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
-The threshold Reynolds number increases with the grass density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
+A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the 
+solution of \eqref{Orr-somerfield}. 
+The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
+in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
+The threshold Reynolds number increases with the grass density, indicating a competition between the 
+destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition 
+either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying 
+assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
 Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
+%>>>>>>> bd11ad10d1dc4414f38f11c1e91e5e1d200223a2
 

### 2014-09-07 Paper/grass.tex
minor update from shreyas

@@
 
-We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow.
-We show this by developing a self-consistent model of the flow past rigid vegetation.
-Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and as well as the origin of the mechanistic threshold.
-Asymptotic analysis of the model for dense grass reveals two unstable modes.
-One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it.
-The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations.
+We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted 
+by the vegetation on the flow. We show this by developing a self-consistent model of the flow past rigid vegetation. 
+Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, 
+and as well as the origin of the mechanistic threshold. Asymptotic analysis of the model for dense grass reveals 
+two unstable modes. One of these modes shares many common features with the Kelvin-Helmholtz instability, 
+but is demonstrated to be different from it. The predictions of the model for threshold condition and the resulting 
+dominant frequency of waving agree with laboratory observations. 
 \end{abstract}
@@
 the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, 
-is also shown in Figure \ref{Re_vs_delta}. 
-The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of 
+is shown in Figure \ref{Re_vs_delta}. 
+The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
 vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.

### 2014-09-07 Paper/grass.tex

@@
 \begin{abstract}
-%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. 
-%We find that this inflection point is located inside the grass canopy but it does not lead to Kelvin Helmholtz instability which is 
-%normally observed in a free shear layer. We also show that the wavelength of coherent vortices scales with the unvegetated water depth, and not with 
-%the shear layer thickness as predicted by a theory based on Kelvin-Helmholtz instability. Based on these results, we propose a mechanism for waving 
-%marine grass based on the shear instability of the flow above the canopy.
-
-%The spontaneous waving of marine grass is though to be due to a Kelving-helmholtz instability resulting from strong shear near top of grass. On the contrary
-%we find that strong velocity shear resulting from vertical discontinuity of drag due to presence of grass does not lead to Kelvin-Helmholtz instability which is 
-%normally observed in free shear flows. We also show that wavelength of coherent vortices scales with the unvegetated water depth. Based on these results, 
-%we propose an alternative mechanism which attributes waving of marine grass on shear instability of flow above the canopy.
-
-%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-%We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
-%drag as a continuum field has been 
-%carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
-%which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
-%Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
-%Mode 2 can be interpreted as instability of 
-%inviscid flow through a grass bed where shear layer doesn't play an important role. 
-%The occurrence of these modes for flow instability depends on submergence ratio and the grass 
-%number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
-%scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
-%For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
-
-% Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
-%inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
-
-We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted 
-by the vegetation on the flow. We show this by developing a self-consistent model of the flow past rigid vegetation. 
-Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, 
-and as well as the origin of the mechanistic threshold. Asymptotic analysis of the model for dense grass reveals 
-two unstable modes. One of these modes shares many common features with the Kelvin-Helmholtz instability, 
-but is demonstrated to be different from it. The predictions of the model for threshold condition and the resulting 
-dominant frequency of waving agree with laboratory observations. 
+We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow. 
+We show this by developing a self-consistent model of the flow past rigid vegetation. 
+Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and explains its mechanistic origin. 
+Asymptotic analysis of the model for dense grass reveals two unstable modes. 
+One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
+The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 
 \end{abstract}

### 2014-09-07 Paper/grass.tex
pull from shreyas

--- 2014-09-07 Paper/grass.tex
@@
 %\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
+
 \usepackage{amsmath}
@@
 \newcommand{\Ndg}{\tilde{N}_g}
-\newcommand{\monami}{\textit{monami}}
 % \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 %The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
-%We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
-%drag as a continuum field has been 
-%carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
-%which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
+We study the mechanism behind Mo-nami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
+drag as a continuum field has been 
+carried out. Our analysis predicts a critical threshold condition for flow instability leading to Mo-nami and associated waving frequency
+which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
 %Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
@@
 %inviscid flow through a grass bed where shear layer doesn't play an important role. 
-%The occurrence of these modes for flow instability depends on submergence ratio and the grass 
-%number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
-%scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
-%For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
+The occurrence of these modes for flow instability depends on submergence ratio and the grass 
+number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
+scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
+For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
@@
 
-We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow.
-We show this by developing a self-consistent model of the flow past rigid vegetation.
-Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and as well as the origin of the mechanistic threshold.
-Asymptotic analysis of the model for dense grass reveals two unstable modes.
-One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it.
-The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations.
 \end{abstract}
@@
 Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}.
-Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.
+Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as mo-nami \cite{AckermanOkubo93}.
 Such motion is considered to play a central role in the recruitment
 of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
-The hydrodynamic mechanism underlying \monami is the focus of this paper.
+The hydrodynamic mechanism underlying monami is the focus of this paper.
 
-Existing explanation of \monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
+Existing explanation of monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
 This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
@@
 
-While the shear layer model successfully appears to predict frequency of \monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
+While the shear layer model successfully appears to predict frequency of monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
 First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account, rendering this explanation inconsistent.
@@
 and therefore inconsistent with the assumptions of the free shear layer instability. 
-% These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
+These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
 
 In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
-The synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
-A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
-The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is a primary consideration that sets the threshold flow.
-We deduce this numerically and through an asymptotic analysis of the dense grass limit.
+The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. 
+The threshold condition for the onset of the instability and the resulting waving frequency compares well with experimental observations.
+In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. 
+In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.
 
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it realizes the shear layer structure of the flow. 
 The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
@@
 Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
-Continuity of flow speed and shear stress at canopy top gives rise to a boundary layer in the flow profile near the tip of the grass.
-
-The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
+Continuity of flow speed and shear stress at canopy top give rise to a boundary layer in the flow profile near the tip of the grass. 
+The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of monami.
 In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; departures from this scaling can be seen in the inset of figure \ref{basicflow}.
 
@@
 We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
-With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation.
+With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
+the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
 The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
@@
 
-%<<<<<<< HEAD
-% A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
-% The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
-% due to grass drag. Indeed a calculation undertaken for terrestrial canopy shows the suppression of growth rate of instability
-% due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
-% as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis 
-% \cite{Raupach96,Delangre06}.
-% The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
-% on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
-% threshold\cite{Ghisal02}. The good agreement between predicted and estimated threshold suggest that the grass drag plays a 
-% crucial role in setting up the flow instability.
-% 
-% Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
-% Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
-The deviation of predicted frequency from the observed may be attributed to the various
-% simplification we have made in our model. In a real canopy, the drag coefficients are known to
-% vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
-Although these variation
-% are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
-% observed frequency and the predicted frequency.
-% =======
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, 
-is also shown in Figure \ref{Re_vs_delta}. 
-The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami and frequency of 
-vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}, which were not predicted in the previous calculation of 
+flow instability of free shear layer profile \cite{Raupach96}\cite{Delangre06}.
+The dependence of this critical $\Rey$ and the corresponding fastest growing wavenumber $k$, on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, 
+and compared with experimental observation of the threshold\cite{Ghisal02}. 
+The primary reason for existence of this threshold is the dissipation due to grass drag which has to be overcome by flow instability.
+The increase in the value of critical $Rey$ with the increase in grass number density can be attributed to the existence of increased dissipation  Dissipation due to grass drag increases with the increase in $\tilde{Ng}$; hence a higher $\Rey$ for flow instability to occur.
 
-A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the 
-solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
-in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
-The threshold Reynolds number increases with the grass density, indicating a competition between the 
-destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition 
-either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying 
-assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
-Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
-%>>>>>>> bd11ad10d1dc4414f38f11c1e91e5e1d200223a2
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the 
+vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in 
+Figure \ref{Re_vs_delta}. The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of 
+vortex passage \cite{Ghisal02}\shreyas{cite someone}. The deviation of predicted frequency from the observed may be attributed to the various 
+simplification we have made in our model. In a real canopy, the drag coefficients are known to
+vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. The turbulent viscosity is also known to vary between the flow of unvegetated region and 
+flow through the canopy\cite{Ghisal02}. Although these variation
+are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to 
+a better agreement between the
+observed frequency and the predicted frequency.
 
+To better understand the mechanism of waving, we consider the behavior of the instability, especially the fastest growing wavenumber, as a function of the vegetation drag, especially as it becomes 
+asymptotically large, i.e. $\Ndg \gg 1$ (or $\delta \ll H$).
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$ \shreyas{Make sure to remone $H$ from $kH$ because $k$ is dimensionless.}, 
+for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
+The lowest $\Rey$ on the neutral curve sets the threshold. 
+We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
+The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
+
+The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
+Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
+Noting that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+Using this estimate, and $\Rey\gg1$, \eqref{Orr-somerfield} within the grass simplifies into $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$ \shreyas{check sign}, while in the unvegetated region it simplifies to
+the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$. 
+This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with out numerical results shown in figure \ref{Asymptotic_mode}.
+
+We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
+The only remaining parameter $\Rey/\Ndg$
+sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
+
+\begin{figure}
+\includegraphics[]{Asymptotic_noshear}
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
+Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
+The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, 
+and are compared to each other on a rescaled coordinate in the inset.
+Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$.
+The modes approach a universal shape, indicating that an asymptotic limit has been reached.
+The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+\label{Asymptotic_mode}
+\end{figure}
+
+On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $,
+and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$),
+explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}.
+This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+
+From these asymptotic behaviors, Mode 1 may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column. 
+Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability.
+Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
+
+Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
+Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
+Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
+We find that in this case, the transition from Mode 1 to Mode 2 appens at such a large grass density, so as to make Mode 2 irrelevant.
+We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.
+
+In conclusion we note that linear stability analysis of flow in presence of canopy reveals two distinct modes of flow instability which might lead to monami. 
+The occurrence of these modes for flow instability in a canopy depends on its submergence ratio and shear layer thickness.
+Our analysis predict a threshold criteria characterized by a Reynolds number.
+The predicted value of threshold Reynolds number agrees well with experimentally observed parameters. 
+Inclusion of variation of drag coefficient along the grass filament might lead to better agreement between the observed and predicted frequency of monami. 
+Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
 \begin{table*}
@@
 
-To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
-The lowest $\Rey$ on the neutral curve sets the threshold. 
-We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
-The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
-
-The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield}.
-Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-Using this estimate, and $\Rey\gg 1$, simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}.
-
-We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
-The only remaining parameter $\Rey/\Ndg$ sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
-
-\begin{figure}
-\includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness 
-for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. 
-The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) 
-but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that 
-the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, 
-and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for 
-Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit 
-has been reached. The limit is not yet reached for the case $\delta/H = 0.05$ due to the influence of bottom boundary; 
-the grass height in this case is comparable to the boundary layer thickness.}
-\label{Asymptotic_mode}
-\end{figure}
-
-On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
-This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
-
-From these asymptotic behaviors, Mode 1 may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
-Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability.
-Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
-The dominant role of vegetation drag in the mechanism for both the modes distinguishes them from the traditional Kelvin-Helmholtz instability.
-
-Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
-Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
-We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
-We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.
-
-The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
-In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
-Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-
-In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag.
-We show the origin of this threshold using a consistent continuum hydodynamic stabiity analysis for the shear flow that develops as a result of the drag.
-Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism.
-The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top.
-Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
-
 \bibliography{Grass}{}

### 2014-09-07 Paper/grass.tex
Polishing of message and paper

--- 2014-09-07 Paper/grass.tex
@@
 Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}.
+Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
 Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.
@@
 Existing explanation of \monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
-This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
+This shear layer is interpreted to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
 Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
@@
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Steady flow profile from the experiment of Ghisalberti\cite{Nepf04} (Case-I from Table-1 with 1250 plants/$m^2$, 
-plant height = 13.7$\pm .2$cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02cm$ with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thicness $\delta$ near the canopy top.
-The depedence of the boundary layer on the vegetation density $\Ndg$ is shown in the inset.}
+\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top.
+The depedence of the boundary layer thickness on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
 \label{basicflow}
@@
 In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
-The synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
+Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
 A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
-The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is a primary consideration that sets the threshold flow.
-We deduce this numerically and through an asymptotic analysis of the dense grass limit.
+The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
+We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
 
@@
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; departures from this scaling can be seen in the inset of figure \ref{basicflow}.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
 
@@
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.68]{Set4_dens28_imgsc}} &
-{\includegraphics[scale = 0.68]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.68]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.68]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.68]{Set5_dens38_imgsc}} &
-{\includegraphics[scale = 0.68]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.68]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.68]{Set5_dens46_imgsc}} \\
-{\includegraphics[scale = 0.68]{Set6_dens32_imgsc}} &
-{\includegraphics[scale = 0.68]{Set6_dens34_imgsc}} &
-{\includegraphics[scale = 0.68]{Set6_dens36_imgsc}} &
-{\includegraphics[scale = 0.68]{Set6_dens38_imgsc}}
+{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
+{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
+{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
+{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
 \end{tabular}
@@
 
-%<<<<<<< HEAD
-% A threshold in $\Rey$ for the onset of the instability naturally emerges from the solution of \eqref{Orr-somerfield}. 
-% The primary reason for the existence of critical $\Rey$ is the presence of dissipation of flow instability 
-% due to grass drag. Indeed a calculation undertaken for terrestrial canopy shows the suppression of growth rate of instability
-% due to grass drag\cite{Delangre06}. Absence of threshold in the previous analysis can be attributed to the assumptions such 
-% as approximationn of mean flow with the free shear layer profile and the exclusion of grass drag in the stability analysis 
-% \cite{Raupach96,Delangre06}.
-% The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, 
-% on the $\delta/H$ and $\hg/H$ is shown in figure \ref{Re_vs_delta}, and compared with experimental observation of the 
-% threshold\cite{Ghisal02}. The good agreement between predicted and estimated threshold suggest that the grass drag plays a 
-% crucial role in setting up the flow instability.
-% 
-% Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in
-% Figure \ref{Re_vs_delta}.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of monami and frequency of vortex passage \cite{Ghisal02}\shreyas{cite someone}.
-The deviation of predicted frequency from the observed may be attributed to the various
-% simplification we have made in our model. In a real canopy, the drag coefficients are known to
-% vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}.
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}.
-Although these variation
-% are not central to the mechanism leading to the instability of flow, we believe an improved analysis of flow with these variation in the model might lead to a better agreement between the
-% observed frequency and the predicted frequency.
-% =======
 Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
@@
 assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
-Regardless of the assumptions, the introduction of vegetation drag damped the linear modes.
-%>>>>>>> bd11ad10d1dc4414f38f11c1e91e5e1d200223a2
+Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.
 
@@
  Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
- Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$} \\
+ Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
  Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
- Critical parameters	& none						& $\Rey$, $\delta/H$ 	& $\Rey \delta^{3/2}$ \\
+ Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
  Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
@@
 To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$. 
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
 To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
@@
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield}.
+Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but $\Rey/\Ndg \sim O(1)$,.
 Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-Using this estimate, and $\Rey\gg 1$, simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge as $\Rey\gg 1$ and $\Ndg \gg 1$ but $\Rey/\Ndg \sim O(1)$, in agreement with our numerical results shown in figure \ref{Asymptotic_mode}.
+This estimate simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
+This analysis predicts that the mode shape would converge in this limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}.
 
@@
 On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which has the same magnitude) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
 The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
@@
 Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
-The dominant role of vegetation drag in the mechanism for both the modes distinguishes them from the traditional Kelvin-Helmholtz instability.
+Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
+The appearance of vegetation drag parameter in the dominant balance of \eqref{Orr-somerfield} and the threshold criteria demonstrates its role in setting the threshold.
 
-Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
+Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
 Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
-Despite these differences, in the framework of this letter, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
+Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
-We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case, and explain the existence of a threshold flow speed.
+We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
 

### 2014-09-07 Paper/grass.tex

@@
 We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
-The only remaining parameter $\Rey/\Ndg$ sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
+The only remaining parameter $\Rey/\Ndg$
+sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
 
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness 
-for representative values of grass density. Mode 1 is shown in solid and Mode 2 is shown in dashed. 
-The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specifying in terms of $\delta/H$) 
-but $\Rey/\Ndg = O(1)$. The mode shapes approach each other for these small values of $\delta/H$ indicating that 
-the large grass density asymptote is reached. Mode 2 shapes appear self-similar in shape as $\delta\to 0$, 
-and are compared to each other on a rescaled coordinate in the inset. Inset shows $|\phi|$ for 
-Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit 
-has been reached. The limit is not yet reached for the case $\delta/H = 0.05$ due to the influence of bottom boundary; 
-the grass height in this case is comparable to the boundary layer thickness.}
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
+Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
+The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
+and are compared to each other on a rescaled coordinate in the inset. 
+Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
+The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
 \label{Asymptotic_mode}
@@
 On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.

### 2014-09-07 Paper/grass.tex
Final finishing touches
Changed caption for figure 4 to modify the value of delta/H to match the figure.
Explained how we obtained boundary layer thickness plotted in figure 1.

@@
 The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top.
-The depedence of the boundary layer thickness on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
+The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
 \label{basicflow}
@@
 Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
-The limit is not yet reached for the case $\delta/H = 0.01$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
 \label{Asymptotic_mode}

### 2014-09-09 Paper/grass.tex

@@
 In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
-We show the origin of this threshold using a consistent continuum hydodynamic stabiity analysis for the shear flow that develops as a result of the drag. 
+We show the origin of this threshold using a consistent continuum hydrodynamic stabiity analysis for the shear flow that develops as a result of the drag. 
 Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 

### 2014-09-19 Paper/grass.tex
Maha comment updated for introduction

@@
 %\section{Introduction}
-Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.
-Such motion is considered to play a central role in the recruitment
-of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
-The hydrodynamic mechanism underlying \monami is the focus of this paper.
+Sea grasses exhibit a rich set of dynamical behavier due to their
+ collective interaction with both steady and oscillatory flows. 
+%interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
+%Resulting 
+The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+The simplet response of these flexible grass canopies 
+%Flexible grass canopies respond 
+to steady currents 
+%in the form of 
+is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}. 
+%Such motion is considered to play a central role in the recruitment 
+%of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
+%The hydrodynamic mechanism underlying \monami is the focus of this paper. 
+In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
 
-Existing explanation of \monami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to different amounts of drag experienced by fluid in and above the canopy.
-This shear layer is interpreted to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz instability. 
-Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
+%Existing 
+Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability 
+%due to different amounts of drag experienced by fluid in and above the canopy. 
+%This shear layer is interpreted to become unstable to coherent vortices 
+through a mechanism similar to the Kelvin-Helmholtz instability. This is thought to lead to coherent eddies over the sea grass canopy which then drives 
+%Influence of these coherent eddies over sea grasses is manifested in their 
+large amplitude synchronous oscillations.
 
@@
 
-While the shear layer model successfully appears to predict frequency of \monami for number of experimental observations, several aspect of existing theory remain unsatisfactory. 
-First, while the free shear layer is established due to the vegetation drag, the linear stability analysis\cite{Raupach96} does not take this drag into account, rendering this explanation inconsistent.
-Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold
-flow speed below which waving is not observed. 
-Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
+While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
+(i) While the free shear layer is established due to the vegetation drag, linear stability analysis\cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
+(ii) Classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold
+flow speed below which \monami is not observed. 
+(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
 and therefore inconsistent with the assumptions of the free shear layer instability. 
@@
 
-In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the phenomenon.
-Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
-A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
-The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
-We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
-
+Here we present a self-consistent mathematical model that account correctly for these effects while also explaining 
+%exhibiting a rich behavior, which is not only consistent within itself, but also with 
+lab experiments and field observations.
+%Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
+%A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
+%The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
+%We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
+\newline
 The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
-The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
+The flexibility of the grass blades, on the other hand, serves merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
 It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
-Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform.
+Therefore, for simplicity, 
+%mathematical model for the linear instability, 
+we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
 Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
@@
 We use a mean field model for the coupling between the flow and the canopy. 
-Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
+The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}

### 2014-09-19 Paper/grass.tex
Maha comment updated for introduction and base profiles

@@
 We use a mean field model for the coupling between the flow and the canopy. 
-The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
+The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area. In the canopy the equations of continuity \& momentum balance can be written as 
+%in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}
-\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}, \quad \grad\cdot\bu = 0,
+\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
 \label{navier-stokes}
 \end{equation}
-where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
+%where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, 
+where $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
 $P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled as 
+The Reynolds number of the flow based on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by pressure and modeled as 
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
-where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
+where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are the normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
 normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
@@
 
-For consistency of analysis, we calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient satisfies $-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
-The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}.
-Flow consist of a region within grass of approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient and a parabolic velocity profile in unvegetated region due to balance of viscous force and pressure gradient.
-Continuity of flow speed and shear stress at canopy top gives rise to a boundary layer in the flow profile near the tip of the grass.
+To understand the onset of, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
+Such a steady state flow driven by constant pressure gradient satisfies 
+\begin{equation}
+ -{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
+\end{equation}
+where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$.
+In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the drag and hence the bottom surface ($y=0$) is approximated by zero shear.
+%The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, 
+%as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}. 
+A comparison of laboratory mean profile with the steady flow profile is shown in figure \ref{basicflow}.
+Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient.
+Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
+Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer 
+%in the flow profile 
+near the tip of the grass.
 
-The boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and may therefore be identified as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
-In our model, this leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
+This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
+In our model shear layer thickness can be estimated from the the balance between viscous forces and drag in the boundary layer near canopy top along with the continuity of shear stress at the grass tip.
+This leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.

### 2014-09-20 Paper/grass.tex
Compilation of an older version of the tex file to generate PDF
Changed Mahesh's name and institutional affiliation to preferred format

@@
-\documentclass[aps,prl,reprint,twocolumn,showpacs,superscriptaddress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,twocolumn,superscriptaddress,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \affiliation{Harvard University, Cambridge MA 02138 USA}
-\author{Mahesh Bandi\thanks{work performed while visiting Brown University}}
-\affiliation{Okinawa Institute of Science and Technology, Okinawa, Japan}
+\author{M. M. Bandi\thanks{work performed while visiting Brown University}}
+\affiliation{Collective Interactions Unit, OIST Graduate University, 1919-1 Tancha, Onna-son, Okinawa, Japan 904-0495}
 \author{Amala Mahadevan}

### 2014-09-21 Paper/grass.tex
more update with MAHA comments and figures

--- 2014-09-19 Paper/grass.tex
@@
  -{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
+\label{base_equ}
 \end{equation}
@@
 %as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}. 
-A comparison of laboratory mean profile with the steady flow profile is shown in figure \ref{basicflow}.
+A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
 Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient.
@@
 This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
-In our model shear layer thickness can be estimated from the the balance between viscous forces and drag in the boundary layer near canopy top along with the continuity of shear stress at the grass tip.
-This leads to a shear layer thickness $\delta = (\Rey\Ndg)^{-1/3} H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
+In our model the balance between viscous forces and the drag force in the shear layer 
+%$C_NdN_g\delta^2u/\nu \sim O(1)$ 
+together with the continuity of shear stress across the grass tip 
+%$u \sim \delta U_o/H$
+ leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
+% and velocity $u \sim \delta U_0$ in the shear layer. 
+Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity. 
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
@@
 {\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
-{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
-{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
-{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
-{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
+%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
 \end{tabular}
@@
 
-Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ respectively, equation \eqref{navier-stokes} expanded to linear order in the perturbations yields
+Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}.
+Substituting them in equation \eqref{navier-stokes} and expanding  to linear order in the perturbation variable we obtain.
 \begin{equation}
@@
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
-We are left with three non-dimensional parameters in the model, viz. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
+We are left with three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
 We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
 With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
-The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
+The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ can be shown to satisfy the eigenvalue problem.
+% provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
 \begin{equation}
@@
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, 
-is shown in Figure \ref{Re_vs_delta}. 
-The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
-vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+%We solve this eigenvalue problem 
+%and in fig\ref{Re_vs_delta} we plot the frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
+%the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field. 
+%The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
+%vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
 
-A threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), naturally emerges from the 
-solution of \eqref{Orr-somerfield}. 
-The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
+We solve this eigenvalue problem and a threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
+solution of \eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
 in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
@@
 destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-Previous linear stability calculations undertaken for terrestrial grass do not exhibit a threshold flow condition 
-either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of the over-simplifying 
-assumptions such as a tanh- or piece-wise linear mean velocity profile\cite{Raupach96,Delangre06}.
+Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
+either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of ad-hoc
+assumptions such as for the mean velocity profile\cite{Raupach96,Delangre06}.
+
+Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
+the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in the inset in figure \ref{Re_vs_delta}b. 
+The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
+vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
 Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.

### 2014-09-21 Paper/grass.tex
more update with MAHA comments and figures

@@
 %\section{Introduction}
-Sea grasses exhibit a rich set of dynamical behavier due to their
+Sea grasses exhibit a rich set of dynamical behavior due to their
  collective interaction with both steady and oscillatory flows. 
@@
 The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-The simplet response of these flexible grass canopies 
+The simplest response of these flexible grass canopies 
 %Flexible grass canopies respond 

### 2014-09-21 Paper/grass.tex
minor update

@@
 %\section{Introduction}
-Sea grasses exhibit a rich set of dynamical behavior due to their
- collective interaction with both steady and oscillatory flows. 
-%interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
-%Resulting 
+Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
 The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-The simplest response of these flexible grass canopies 
-%Flexible grass canopies respond 
-to steady currents 
-%in the form of 
-is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}. 
+The simplest response of these flexible grass canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.
 %Such motion is considered to play a central role in the recruitment 
@@
 In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
-
-%Existing 
 Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability 
@@
 and therefore inconsistent with the assumptions of the free shear layer instability.
-% These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.
-
-Here we present a self-consistent mathematical model that account correctly for these effects while also explaining 
-%exhibiting a rich behavior, which is not only consistent within itself, but also with 
-lab experiments and field observations.
+Here we present a self-consistent mathematical model that account correctly for these effects 
+while also explaining lab experiments and field observations.
 %Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
@@
 
-To understand the onset of, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
+To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
 Such a steady state flow driven by constant pressure gradient satisfies 

### 2014-09-21 Paper/grass.tex
Display equation in asymptotic analysis

@@
 The simplest response of these flexible grass canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.  
-%Such motion is considered to play a central role in the recruitment 
-%of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
-%The hydrodynamic mechanism underlying \monami is the focus of this paper. 
 In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
-Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability 
-%due to different amounts of drag experienced by fluid in and above the canopy. 
-%This shear layer is interpreted to become unstable to coherent vortices 
-through a mechanism similar to the Kelvin-Helmholtz instability. This is thought to lead to coherent eddies over the sea grass canopy which then drives 
-%Influence of these coherent eddies over sea grasses is manifested in their 
-large amplitude synchronous oscillations.
+Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability.
+This is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.
 
@@
 It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
-Therefore, for simplicity, 
-%mathematical model for the linear instability, 
-we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
+Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
 Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
@@
 where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
-In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the drag and hence the bottom surface ($y=0$) is approximated by zero shear.
-%The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=2H$) of the fluid layer, 
-%as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{basicflow}. 
+In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
 A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
@@
 Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
-Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer 
-%in the flow profile 
-near the tip of the grass.
+Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.
 
@@
 Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-This estimate simplifies \eqref{Orr-somerfield} within the grass to $\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi$, while in the unvegetated region it simplifies to the Rayleigh equation $ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
-This analysis predicts that the mode shape would converge in this limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}.
+This estimate simplifies \eqref{Orr-somerfield} to 
+\begin{align}
+\begin{split}
+\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,   &\text{ for $y<\hg$}  \\
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, &\text{ for $y>\hg$}.
+\end{split}
+\end{align}
+Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}.
 
-We interpret this as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
+We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
 The only remaining parameter $\Rey/\Ndg$
@@
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
-This analysis predicts that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), leading to a simplification of \eqref{Orr-somerfield} as
+\begin{align}
+ \begin{split}
+  \text{blah for $y<\hg$} \\
+  \text{blah for $y>\hg$.}
+ \end{split}
+\end{align}
+Since, $\Rey/\Ndg^{1/2}$ is the only remaining parameter in this model, explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}). (this language should be improved)
+This analysis concludes that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
 

### 2014-09-21 Paper/grass.tex
minor change

@@
 vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
-Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.
+%Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.
 
@@
 We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
-
+\newline
 The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
@@
 Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-
+\newline
 In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 

### 2014-09-21 Paper/grass.tex
changed the location of figures so that they appear in relevant para

@@
 %\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
-
 \usepackage{amsmath}
@@
 In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
+\newline
 Current explanations of \monami invoke the existence of a shear layer near the top of the canopy \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability.
-This is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.
-
-\begin{figure}
-\includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
-plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top.
-The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
-\label{basicflow}
-\end{figure}
-
+This instability is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.
+\newline
 While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
@@
 the flow instability is similar to the instability of a shear layer. 
-
-\begin{figure}
-\begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
-\end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
- parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
-also included in the top panel. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
-Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
-\label{Re_vs_delta}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+%\vspace{-6mm} \hspace{-3mm}
+%\includegraphics[]{K_vs_shear_width_noshear}
+%\end{center}
+%\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
+% parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
+%also included in the top panel. Inset compares experimental observations 
+%of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The %experimental data in the inset in obtained from publications by 
+%Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a %representative value of $\mu=0.1$ Pa~s was assumed.}
+%\label{Re_vs_delta}
+%\end{figure}
 We use a mean field model for the coupling between the flow and the canopy. 
@@
 Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
-
+\newline
 To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
@@
 \end{equation}
+\begin{figure}
+\includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top.
+The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
+\label{basicflow}
+\end{figure}
 where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
@@
 Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass. 
-
+\newline
 This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
 In our model the balance between viscous forces and the drag force in the shear layer 
-%$C_NdN_g\delta^2u/\nu \sim O(1)$ 
-together with the continuity of shear stress across the grass tip 
-%$u \sim \delta U_o/H$
- leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
+together with the continuity of shear stress across the grass tip leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
 % and velocity $u \sim \delta U_0$ in the shear layer. 
@@
 The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
-
-\begin{figure*}
-\begin{tabular}{cccc}
-{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
-%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
-%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
-%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
-%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
-\end{tabular}
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified grass number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
-As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
-For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
-\label{K_Re_sigma_set3}
-\end{figure*}
-
 Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}.
@@
 % provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
+\begin{figure}
+\begin{center}
+\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{K_vs_shear_width_noshear}
+\end{center}
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
+ parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
+also included in the top panel. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
+Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+\label{Re_vs_delta}
+\end{figure}
 \begin{equation}
@@
 
+\begin{figure*}
+\begin{tabular}{cccc}
+{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
+%{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens34_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens36_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
+\end{tabular}
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified grass number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
+As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
+For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
+\label{K_Re_sigma_set3}
+\end{figure*}
+
 Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
@@
 \end{table*}
-
 To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.

### 2014-09-21 Paper/grass.tex
changed the order of mode 1 mode 2 description and explained shear layer scaling

@@
 \end{equation}
+where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
+In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
+A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
+Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient. 
+Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
+Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.
 \begin{figure}
@@
 \end{figure}
-where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
-In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
-A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
-Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient. 
-Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
-Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass. 
-\newline
 This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
-In our model the balance between viscous forces and the drag force in the shear layer 
-together with the continuity of shear stress across the grass tip leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
-% and velocity $u \sim \delta U_0$ in the shear layer. 
-Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity. 
+In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_{sl}/\delta^2 \sim \rho C_N d N_g u_{sl}^2)$ and the continuity of shear stress across the grass tip $(u_{sl}/\delta \sim U_0/H)$ leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
+Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
 The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
+\newline
 Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}.
@@
 either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of ad-hoc
-assumptions such as for the mean velocity profile\cite{Raupach96,Delangre06}.
+assumptions for the mean velocity profile\cite{Raupach96,Delangre06}.
 
@@
 Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is shown in the inset in figure \ref{Re_vs_delta}b. 
+the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in figure \ref{Re_vs_delta}b. 
 The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
@@
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but $\Rey/\Ndg \sim O(1)$,.
-Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-This estimate simplifies \eqref{Orr-somerfield} to 
-\begin{align}
-\begin{split}
-\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,   &\text{ for $y<\hg$}  \\
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, &\text{ for $y>\hg$}.
-\end{split}
-\end{align}
-Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}.
-
-We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
-The only remaining parameter $\Rey/\Ndg$
-sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
-
-\begin{figure}
-\includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
-Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
-The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
-and are compared to each other on a rescaled coordinate in the inset. 
-Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
-The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
-\label{Asymptotic_mode}
-\end{figure}
-
-On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-
+Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
@@
 
+On the other hand Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but $\Rey/\Ndg \sim O(1)$,.
+Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+This estimate simplifies \eqref{Orr-somerfield} to 
+\begin{align}
+\begin{split}
+\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,   &\text{ for $y<\hg$}  \\
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, &\text{ for $y>\hg$}.
+\end{split}
+\end{align}
+\begin{figure}
+\includegraphics[]{Asymptotic_noshear}
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
+Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
+The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
+and are compared to each other on a rescaled coordinate in the inset. 
+Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
+The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+\label{Asymptotic_mode}
+\end{figure}
+Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}. 
+We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
+The only remaining parameter $\Rey/\Ndg$
+sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
+%On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+
+%This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+%The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), leading to a simplification of \eqref{Orr-somerfield} as
+%\begin{align}
+% \begin{split}
+%  \text{blah for $y<\hg$} \\
+%  \text{blah for $y>\hg$.}
+% \end{split}
+%\end{align}
+%Since, $\Rey/\Ndg^{1/2}$ is the only remaining parameter in this model, explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}). (this language should be improved)
+%This analysis concludes that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
 From these asymptotic behaviors, Mode 1 may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column. 
@@
 The appearance of vegetation drag parameter in the dominant balance of \eqref{Orr-somerfield} and the threshold criteria demonstrates its role in setting the threshold.
-
+\newline
 Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
@@
 Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-\newline
+\newline\newline
 In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
@@
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
-
 \bibliography{Grass}{}

### 2014-09-25 Paper/grass.tex
Explain Mode 1 asymptote better

@@
 This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
-In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_{sl}/\delta^2 \sim \rho C_N d N_g u_{sl}^2)$ and the continuity of shear stress across the grass tip $(u_{sl}/\delta \sim U_0/H)$ leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$.
-Where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is scale for velocity.
+In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the grass tip $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
 This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
@@
  Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
- Linearized dynamics	& Inviscid Rayleigh's equation			& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
+ Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
  Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
@@
 This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), leading to a simplification of \eqref{Orr-somerfield} as
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of \eqref{Orr-somerfield} as
 \begin{align}
- \begin{split}
-  \text{blah for $y<\hg$} \\
-  \text{blah for $y>\hg$.}
- \end{split}
+(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
 \end{align} 
-Since, $\Rey/\Ndg^{1/2}$ is the only remaining parameter in this model, explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}). (this language should be improved)
-This analysis concludes that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+in a region of O($\delta$) near $y=\hg$.
+Since the relative magnitude of the terms r.h.s.
+to the l.h.s.
+of this equation in the limit $\Rey \gg 1$ and $\Ndg \gg 1$ is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
+Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ explaining the numerically observation (see figure \ref{Re_vs_delta}).
+This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
 
-On the other hand Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but $\Rey/\Ndg \sim O(1)$,.
+On the other hand Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
@@
 \begin{split}
-\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,   &\text{ for $y<\hg$}  \\
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, &\text{ for $y>\hg$}.
+\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for $y<\hg$}  \\
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for $y>\hg$}.
 \end{split}

### 2014-09-25 Paper/grass.tex
Fix language

@@
 in a region of O($\delta$) near $y=\hg$.
-Since the relative magnitude of the terms r.h.s.
-to the l.h.s.
-of this equation in the limit $\Rey \gg 1$ and $\Ndg \gg 1$ is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
+Since the relative magnitude of the terms on the r.h.s.
+to those on the l.h.s.
+of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
 Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ explaining the numerically observation (see figure \ref{Re_vs_delta}).

### 2014-09-26 Paper/grass.tex
Minor language fix
Incorporate all of Maha's suggestions

@@
 \begin{document}
-\title{Linear instability underlying synchronous waving of marine grass in flow}
+\title{The onset of of hydrodynamically driven waving of marine grass in flow}
 \author{Ravi Singh}
@@
 The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
-The flexibility of the grass blades, on the other hand, serves merely aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
+The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
 It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
@@
 The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
-where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area. In the canopy the equations of continuity \& momentum balance can be written as 
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
+In the canopy the equations of continuity and momentum balance can be written as 
 %in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
@@
 \begin{equation}
- -{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
+ -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
 \label{base_equ}
@@
 In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
-A comparison of laboratory mean profile with the steady flow profile is shown in figure \eqref{basicflow}.
-Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient. 
-Flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient. 
+A comparison of the steady flow profile from the solution of \eqref{base_equ} with experimental measurements is shown in figure~\ref{basicflow}.
+Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
 Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.
@@
 In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the grass tip $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
-This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically control it and investigate its effect on the instability mechanism.
+This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
 The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
 \newline
-Considering small perturbations $u, v, p$ about the base profile $U$ and $P$ given by the solution of \eqref{base_equ}.
-Substituting them in equation \eqref{navier-stokes} and expanding  to linear order in the perturbation variable we obtain.
+To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in \eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
 \begin{equation}
@@
 the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
-The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ can be shown to satisfy the eigenvalue problem.
+The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ satisfies the eigenvalue problem.
 % provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
@@
 \end{table*}
-To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag.
+To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
@@
 We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
-The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the discontinuity in the most unstable mode.
+The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the most unstable mode to jump.
 
@@
 (D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
+\label{eqn:mode1asymp}
 \end{align} 
-in a region of O($\delta$) near $y=\hg$.
+in a region of thickness O($\delta$) near $y=\hg$.
 Since the relative magnitude of the terms on the r.h.s.
@@
 of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
-Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ explaining the numerically observation (see figure \ref{Re_vs_delta}).
+Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
 This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
 
-On the other hand Mode 2 has a threshold $\Rey \sim ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
@@
 \end{split}
+\label{eqn:mode2asymp}
 \end{align}
@@
 %This analysis concludes that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
-From these asymptotic behaviors, Mode 1 may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column. 
 Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+Because Mode 1 eigenfunction is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
 Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
 Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
-The appearance of vegetation drag parameter in the dominant balance of \eqref{Orr-somerfield} and the threshold criteria demonstrates its role in setting the threshold.
+The appearance of vegetation drag parameter in the dominant balances \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.
 \newline
@@
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
-We recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
+In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
 \newline

### 2014-09-26 Paper/grass.tex
Fix abstract and title, and ready to send out!

@@
 \begin{document}
-\title{The onset of of hydrodynamically driven waving of marine grass in flow}
+\title{The onset of hydrodynamically driven waving of marine grass in flow}
 \author{Ravi Singh}
@@
 \author{Amala Mahadevan}
-\affiliation{Woods Hold Institute of Oceanography, Woods Hole MA USA}
+\affiliation{Woods Hole Institute of Oceanography, Woods Hole MA USA}
 \author{Shreyas Mandre}
@@
 \begin{abstract}
-We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow. 
-We show this by developing a self-consistent model of the flow past rigid vegetation. 
-Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and explains its mechanistic origin. 
-Asymptotic analysis of the model for dense grass reveals two unstable modes. 
-One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
+We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field.
+To explain the onset of spontaneous waving, we perform a linear stability analysis of the steady flow resulting from this model.
+We show that a threshold flow condition for the spontaneous waving of marine grass arises naturally in this framework due to the drag exerted by the vegetation on the flow. 
+Beyond the threshold, we find two unstable modes; first is localized in a boundary layer near the canopy top, and second which spans the whole water column.
+The first mode shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. 
 The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 
 \end{abstract}
+% Asymptotic analysis of the model for dense grass reveals two unstable modes. 
+% We show this by  
+%  naturally leads to the threshold condition, and explains its mechanistic origin. 
 \maketitle

### 2014-09-26 Paper/grass.tex
Changes made by Mahesh
1) Added acknowledgments for myself.
2) Made few minor typo corrections (mostly singular/plural forms).
3) Changed symbol for Reynolds from R to Re
4) Replaced "figure" with "fig." for figure references and added "eq."
before all equation references.
5) Changed Order of magnitude symbol to math style rather than text.

@@
 \newcommand{\hg}{h_g}
-\newcommand{\Rey}{\text{R}}
+\newcommand{\Rey}{\text{Re}}
 \newcommand{\Ndg}{\tilde{N}_g}
@@
 \affiliation{Harvard University, Cambridge MA 02138 USA}
-\author{M. M. Bandi\thanks{work performed while visiting Brown University}}
+\author{M. M. Bandi}
 \affiliation{Collective Interactions Unit, OIST Graduate University, 1919-1 Tancha, Onna-son, Okinawa, Japan 904-0495}
 \author{Amala Mahadevan}
-\affiliation{Woods Hole Institute of Oceanography, Woods Hole MA USA}
+\affiliation{Woods Hole Oceanographic Institute, Woods Hole MA USA}
 \author{Shreyas Mandre}
@@
 \newline
-While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspect of existing theory remain unsatisfactory. 
-(i) While the free shear layer is established due to the vegetation drag, linear stability analysis\cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
-(ii) Classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations in the lab\cite{Ghisal02} and the field\cite{Grizzle96} indicate the existence of a threshold
+While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
+(i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
+(ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold
 flow speed below which \monami is not observed. 
@@
 and therefore inconsistent with the assumptions of the free shear layer instability.
-Here we present a self-consistent mathematical model that account correctly for these effects 
+Here we present a self-consistent mathematical model that accounts correctly for these effects 
 while also explaining lab experiments and field observations.
@@
 \newline
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
-The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability\cite{Nepf2012}. 
-It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels\cite{Ghisal02}. 
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism \cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
+The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability \cite{Nepf2012}. 
+It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels \cite{Ghisal02}. 
 Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated\cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
+Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
 the flow instability is similar to the instability of a shear layer. 
@@
 The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
-where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
 In the canopy the equations of continuity and momentum balance can be written as 
@@
 %where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, 
-where $\rho$ the fluid density, $\mathbf{u}$ the velocity, 
+where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
 $P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by pressure and modeled as 
+The Reynolds number (\Rey) of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
 where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are the normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
-normal to grass while $\bt,\bn$ being unit vector along and normal to grass. 
+normal to grass while $\bt,\bn$ are unit vectors along and normal to grass. 
 We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
-In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
-Based on previous experiments\cite{Vivoni98}, we take $C_N \approx 1$.
+In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variations to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
+Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.
 \newline
-To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of \eqref{navier-stokes}, and examine its linear stability. 
+To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 
 Such a steady state flow driven by constant pressure gradient satisfies 
@@
 In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
-A comparison of the steady flow profile from the solution of \eqref{base_equ} with experimental measurements is shown in figure~\ref{basicflow}.
-Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and with pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
+A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
+Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
 Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.
@@
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf\cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
 plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
@@
 \end{figure}
-This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer\cite{Ghisal02} in the previous explanation of \monami.
+This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
 In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the grass tip $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
-This dependence of the boundary layer thickness on the grass density $N_g$, shown in Figure \ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of figure \ref{basicflow}.
+This dependence of the boundary layer thickness on the grass density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
 \newline
-To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in \eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
+To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
 \begin{equation}
@@
 We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
-With these scaling, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
+With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
 the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
@@
 \caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
- parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
+ parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are 
 also included in the top panel. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of eq.~\eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
 Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
@@
 
-We solve this eigenvalue problem and a threshold in $\Rey$ of O(100), above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
-solution of \eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
-in figure \ref{Re_vs_delta}, and is found to compare well with experimental observations\cite{Ghisal02}.
+We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
+solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
+in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
 The threshold Reynolds number increases with the grass density, indicating a competition between the 
@@
 Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
-either due to the exclusion of the vegetation drag in their model\cite{Raupach96}, or because of ad-hoc
-assumptions for the mean velocity profile\cite{Raupach96,Delangre06}.
+either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
+assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.
 
@@
 Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in figure \ref{Re_vs_delta}b. 
+the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
 The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
@@
 To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see figure \ref{Re_vs_delta}). 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in figure \ref{K_Re_sigma_set3}. 
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
 The lowest $\Rey$ on the neutral curve sets the threshold. 
@@
 Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of \eqref{Orr-somerfield} as
+This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
 \begin{align}
@@
 of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
-Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}).
-This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
+Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}).
+This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in fig.~\ref{Asymptotic_mode}, supporting this argument.
 
-On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in figure \ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of \eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-This estimate simplifies \eqref{Orr-somerfield} to 
+This estimate simplifies eq.~\eqref{Orr-somerfield} to 
 \begin{align}
@@
 \end{figure}
-Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in figure \ref{Asymptotic_mode}. 
+Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in fig.~\ref{Asymptotic_mode}. 
 We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
@@
 Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
-The appearance of vegetation drag parameter in the dominant balances \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.
+The appearance of vegetation drag parameter in the dominant balances eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.
 \newline
-Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami}\cite{Inoue56,Raupach96}, have also been observed.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant\cite{Vivoni98,Ghisal02}.
-Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represents the hydrodynamic instability for the terrestrial case with stiff grass blades.
+Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami} \cite{Inoue56,Raupach96}, have also been observed.
+A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
+Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant \cite{Vivoni98,Ghisal02}.
+Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
@@
 \newline
-The deviation of our model predictions from the observed may be attributed to the various simplification we have made in our model. 
-In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades\cite{Vivoni98,Nepf00}. 
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy\cite{Ghisal02}. 
+The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
+The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy \cite{Ghisal02}. 
 Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
@@
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
+
+\acknowledgments
+MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
+
 \bibliography{Grass}{}

### 2014-09-26 Paper/grass.tex
Post Mahesh's corrections
- Fix real part vs Reynolds number conflict
- Remove introduction of Re for blade scale Reynolds number
- Minor cosmetic changes

@@
 where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number (\Rey) of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 
+$P$ the pressure, $\mu$ the eddy viscosity (the flow is usually turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
@@
 
-We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re($\sigma$)$>$0), emerges from the 
+We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (real part of $\sigma$ is positive), emerges from the 
 solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
@@
 \begin{align}
-(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \Rey \sigma (D^2-k^2)\phi,
+(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \sigma \Rey (D^2-k^2)\phi,
 \label{eqn:mode1asymp}
@@
 MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
+We thanks Heidi Nepf and Marco Ghisalberti for helpful discussions.
 

### 2014-09-26 Paper/grass.tex
Continue to improve language

@@
 \newcommand{\hg}{h_g}
-\newcommand{\Rey}{\text{Re}}
+\newcommand{\Rey}{\text{R}}
 \newcommand{\Ndg}{\tilde{N}_g}
-\newcommand{\monami}{\textit{monami }}
+\newcommand{\monami}{\textit{monami}}
 % \newcommand{\shreyas}[1]{{\bf (#1)}}
@@
 \begin{document}
-\title{The onset of hydrodynamically driven waving of marine grass in flow}
+\title{The onset of hydrodynamically driven waving of marine vegetation in flow}
 \author{Ravi Singh}
@@
 \author{M. M. Bandi}
-\affiliation{Collective Interactions Unit, OIST Graduate University, 1919-1 Tancha, Onna-son, Okinawa, Japan 904-0495}
+\affiliation{OIST Graduate University, Okinawa 904-0495, Japan}
 \author{Amala Mahadevan}
-\affiliation{Woods Hole Oceanographic Institute, Woods Hole MA USA}
+\affiliation{Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA}
 \author{Shreyas Mandre}
@@
 
-\begin{abstract}
-We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field.
+% \begin{abstract}
+% We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field.
 To explain the onset of spontaneous waving, we perform a linear stability analysis of the steady flow resulting from this model.
@@
 The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations.
+% \end{abstract}
+\begin{abstract}
+We develop a self-consistent mathematical model for the onset of synchronous waving of dense marine vegetation driven by a hydrodynamic instability.
+The vegetation, assumed rigid, only exerts a drag force on the flow.
+We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency, confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
+We also demonstrate that our instability is different from the Kelvin-Helmholtz instability.
 \end{abstract}
@@
 The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-The simplest response of these flexible grass canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}.  
+The simplest response of these flexible vegetation canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
 In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
 \newline
-Current explanations of \monami invoke the existence of a shear layer near the top of the canopy  \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability.
-This instability is thought to lead to coherent eddies over the sea grass canopy which then drives large amplitude synchronous oscillations.
+Current explanations of \monami ~invoke the existence of a shear layer due to vegetation drag  \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability.
+This instability is thought to lead to coherent eddies over the sea grass meadow, which then drive large amplitude synchronous oscillations.
 \newline
-While the shear layer model appears to predict the frequency of \monami consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
+While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
 (i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
-(ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold
-flow speed below which \monami is not observed. 
-(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness,
-and therefore inconsistent with the assumptions of the free shear layer instability.
-Here we present a self-consistent mathematical model that accounts correctly for these effects 
-while also explaining lab experiments and field observations.
+(ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
+(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the top of the meadow, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
+Here we present a self-consistent mathematical model for the hydrodynamic linear stability analysis that accounts correctly for these effects, while also explaining lab experiments and field observations.
 %Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
@@
 \newline
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism \cite{Ghisal02} of the instability underlying \monami because it realizes the shear layer structure of the flow. 
+The drag exerted by the grass bed on the flow is central to the hypothesized mechanism \cite{Ghisal02} of the instability underlying \monami ~because it realizes the shear layer structure of the flow.
 The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability \cite{Nepf2012}. 
@@
 Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
-Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed,
-the flow instability is similar to the instability of a shear layer. 
+Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
 %\begin{figure}
@@
 %\end{figure}
-We use a mean field model for the coupling between the flow and the canopy. 
-The force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
-where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
-In the canopy the equations of continuity and momentum balance can be written as 
+
+We use a mean field model for the coupling between the flow and the vegetation. 
+The force exerted by the vegetation is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
+where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
+In the vegetation, the equations of continuity and momentum balance can be written as 
 %in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}
-\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
+\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
 \label{navier-stokes}
@@
 where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
-$P$ the pressure, $\mu$ the eddy viscosity (the flow is usually turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of a grass blade is expected to be $O(10^2-10^3)$, and hence the drag force is dominated by pressure and modeled as 
-$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn + C_{T}\rho u_{T}^{2}d\bt$ \shreyas{citation needed} 
-where $d$ is average width of grass blade $C_{N}$ and $C_{T}$ are the normal and tangential drag coefficients respectively; $u_{T}$, $u_{N}$ are velocity components along and
-normal to grass while $\bt,\bn$ are unit vectors along and normal to grass. 
-We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. 
-In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variations to be central to the mechanism and therefore take $C_N$ and $N_g$ to be constants. 
+$p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow based on the scale of the grass blade is expected to be $O(10^2-10^3)$, implying that the form drag on the vegetation dominates over the skin-friction.
+Hence we model it as
+$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, \shreyas{citation needed} 
+where $d$ is average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass. 
+In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variations to be central to the instability mechanism and therefore take $C_N$ and $N_g$ to be constants.
 Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.
@@
 To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient satisfies 
+Such a steady state flow driven by constant pressure gradient $dP/dx$ satisfies 
 \begin{equation}
@@
 where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
-In the limit of large grass density, shear stress exerted by the bottom surface is estimated to be negligible in comparison with the vegetation drag and hence the bottom surface ($y=0$) is approximated by zero shear.
+In the limit of large vegetation density, shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, and hence the bottom surface ($y=0$) is approximated by zero shear.
 A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
-Flow within grass has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
-Continuity of flow speed and shear stresses at canopy top gives rise to a boundary layer near the tip of the grass.
+Flow within vegetation has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
+Continuity of flow speed and shear stresses at the top of the meadow gives rise to a boundary layer.
 \begin{figure}
@@
 plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm with our model. The canopy extends upto $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the grass bed, and (iii) a boundary layer of thickness $\delta$ near the canopy top.
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The meadow extends upto $y=h_g$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the top of the meadow.
 The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
@@
 \end{figure}
-This boundary layer near the canopy top results from purely local dynamics independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
-In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the grass tip $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
-This dependence of the boundary layer thickness on the grass density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
+This boundary layer results from purely local dynamics, independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the top of the vegetation $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
+This dependence of the boundary layer thickness on the vegetation density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
 The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
@@
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
-We are left with three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the grass $h_g/H$. 
-We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the grass density and help elucidate the mechanism of the instability. 
+We have three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$. 
+We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the mechanism of the instability. 
 With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
@@
 
-We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (real part of $\sigma$ is positive), emerges from the 
+We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
 solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
 in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
-The threshold Reynolds number increases with the grass density, indicating a competition between the 
+The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
 destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
@@
 \end{tabular}
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified grass number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
+\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
 As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
@@
 the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
-The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
+The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of 
 vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
@@
 			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
- Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$} \\
+ Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
  Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
@@
 To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical grass density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
 To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
@@
 We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
-The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the grass density increases, causing the most unstable mode to jump.
+The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
 
 The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
 This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
+The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
 \begin{align}
@@
 On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
-Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the grass is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
 This estimate simplifies eq.~\eqref{Orr-somerfield} to 
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of grass density. 
+\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 
 Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
-The mode shapes approach each other for these small values of $\delta/H$ indicating that the large grass density asymptote is reached. Mode 1 shapes appear self-similar in shape as $\delta\to 0$, 
+The mode shapes approach each other for these small values of $\delta/H$ indicating that the large vegetation density asymptote is reached.
+Mode 1 shapes appear self-similar in shape as $\delta\to 0$,
 and are compared to each other on a rescaled coordinate in the inset. 
 Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
-The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the grass height in this case is comparable to the boundary layer thickness.}
+The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.}
 \label{Asymptotic_mode}
@@
 Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in fig.~\ref{Asymptotic_mode}. 
-We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass tip plays no role.
+We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the top of the vegetation plays no role.
 The only remaining parameter $\Rey/\Ndg$
@@
 Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
-The appearance of vegetation drag parameter in the dominant balances eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the threshold criteria demonstrates its role in setting the threshold.
+The appearance of vegetation drag parameter in the dominant balances represented by eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
 \newline
@@
 Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
-We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large grass density, so as to make Mode 2 irrelevant.
+We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant.
 In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
@@
 The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-In a real canopy, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
-The turbulent viscosity is also known to vary between the flow of unvegetated region and flow through the canopy \cite{Ghisal02}. 
+In a real meadow, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
+The eddy viscosity is also known to vary between the flow of unvegetated region and flow through the meadow \cite{Ghisal02}. 
 Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
@@
 In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
-We show the origin of this threshold using a consistent continuum hydrodynamic stabiity analysis for the shear flow that develops as a result of the drag. 
+We show the origin of this threshold using a consistent continuum hydrodynamic stability analysis for the shear flow that develops as a result of the vegetation-induced drag. 
 Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
-The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the canopy top.
+The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

### 2014-10-01 Paper/grass.tex
minor change in abstract
Change title, abstract, and intro
- Change title and abstract as per discussions with Mahadevans
- Add time-scale separation argument to justify neglect of vegetation
  deformation

@@
 \newcommand{\monami}{\textit{monami}}
-% \newcommand{\shreyas}[1]{{\bf (#1)}}
-\newcommand{\shreyas}[1]{}
 \definecolor{tableShade}{gray}{0.8}
@@
 \begin{document}
-\title{The onset of hydrodynamically driven waving of marine vegetation in flow}
+\title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
 \author{Ravi Singh}
@@
 
-% \begin{abstract}
-% We develop a self-consistent model of the flow past rigid vegetation, where the vegetation is represented by a continuum drag field.
-To explain the onset of spontaneous waving, we perform a linear stability analysis of the steady flow resulting from this model.
-We show that a threshold flow condition for the spontaneous waving of marine grass arises naturally in this framework due to the drag exerted by the vegetation on the flow.
-Beyond the threshold, we find two unstable modes; first is localized in a boundary layer near the canopy top, and second which spans the whole water column.
-The first mode shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it.
-The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations.
-% \end{abstract}
 \begin{abstract}
-We develop a self-consistent mathematical model for the onset of synchronous waving of dense marine vegetation driven by a hydrodynamic instability.
-The vegetation, assumed rigid, only exerts a drag force on the flow.
-We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency, confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
-We also demonstrate that our instability is different from the Kelvin-Helmholtz instability.
+The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the steady flow.
+Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
+The damping from the vegetation drag leads to a finite threshold flow for the instability.
+This threshold agrees with experimental observations, and so does the predicted waving frequency.
+This role of vegetation drag differentiates our mechanism from the previous hypothesis that the Kelvin-Helmholz instability underlies \monami.
 \end{abstract}
-% Asymptotic analysis of the model for dense grass reveals two unstable modes. 
-% We show this by  
-%  naturally leads to the threshold condition, and explains its mechanistic origin. 
+
 \maketitle
-%\section{Introduction}
 Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
@@
 Here we present a self-consistent mathematical model for the hydrodynamic linear stability analysis that accounts correctly for these effects, while also explaining lab experiments and field observations.
-%Using this model, the synchronous waving may be understood to be arising from the linear hydrodynamic instability of a uniform flow over the grass bed, such that the principal role of the vegetation is to exert a drag on the flow.
-%A threshold flow speed arises naturally for the onset of the instability; this threshold condition and the resulting waving frequency compares well with experimental observations.
-%The vegetation drag introduces a damping mechanism in the flow, in addition to the viscous damping, and overcoming this damping is the primary consideration that sets the threshold flow.
-%We deduce this numerically and through an asymptotic scaling analysis in the limit of dense grass.
 \newline
-The drag exerted by the grass bed on the flow is central to the hypothesized mechanism \cite{Ghisal02} of the instability underlying \monami ~because it realizes the shear layer structure of the flow.
-The flexibility of the grass blades, on the other hand, serves merely to aid in visualizing the flow structure resulting from the instability \cite{Nepf2012}. 
-It has been demonstrated that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels \cite{Ghisal02}. 
-Therefore, for simplicity, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. 
-Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
-%\begin{figure}
-%\begin{center}
-%\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-%\vspace{-6mm} \hspace{-3mm}
-%\includegraphics[]{K_vs_shear_width_noshear}
-%\end{center}
-%\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
-% parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf\cite{Ghisal02} to exhibit or suppress synchronous waving are 
-%also included in the top panel. Inset compares experimental observations 
-%of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of \eqref{Orr-somerfield}. The %experimental data in the inset in obtained from publications by 
-%Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a %representative value of $\mu=0.1$ Pa~s was assumed.}
-%\label{Re_vs_delta}
-%\end{figure}
+The drag exerted by the grass bed on the flow is central to the hypothesized instability for \monami. 
+The instability and the resulting flow structures, though manifest in the waving of grasses, also persist if the grass blades are replaced by rigid dowels \cite{Ghisal02}. 
+Therefore, for simplicity, we consider the grass bed to exchange momentum with the surrounding fluid but not deform. 
+A criteria for the validity of this assumptions is presented at the end of this letter.
+% Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
 
@@
 In the vegetation, the equations of continuity and momentum balance can be written as 
-%in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} as
 \begin{equation}
@@
 \end{equation}
-%where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, $N_g$ the grass number density per unit area, 
 where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
@@
 Hence we model it as
-$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, \shreyas{citation needed} 
+$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
 where $d$ is average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass. 
@@
 The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ satisfies the eigenvalue problem.
-% provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$ through the solution of
 \begin{figure}
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.
-
-%We solve this eigenvalue problem 
-%and in fig\ref{Re_vs_delta} we plot the frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-%the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field. 
-%The observed frequencies are associated with the peaks in velocity spectra, frequency of $\monami$ and frequency of 
-%vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
 
@@
 vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
-%Regardless of the assumptions, the introduction of vegetation drag is observed to damp the linear modes\cite{Delangre06}.
 
@@
  Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
+ Dense grass limit &  & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
  Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
@@
 
-%This limit can be understood through asymptotically estimating the sizes of the terms in \eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-%The terms balance when $\Rey \sim {\Ndg}^{1/2}$ (or $\Rey \sim H/\delta$), leading to a simplification of \eqref{Orr-somerfield} as
-%\begin{align}
-% \begin{split}
-%  \text{blah for $y<\hg$} \\
-%  \text{blah for $y>\hg$.}
-% \end{split}
-%\end{align}
-%Since, $\Rey/\Ndg^{1/2}$ is the only remaining parameter in this model, explaining the numerically observed asymptote (see figure \ref{Re_vs_delta}). (this language should be improved)
-%This analysis concludes that the mode structure is self-similar over the length scale $\delta$; the verification of this expectation is shown in figure \ref{Asymptotic_mode}, supporting this argument.
 Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
@@
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid, whereas aquatic canopies are buoyant \cite{Vivoni98,Ghisal02}.
+Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tend to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}.
 Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
@@
 The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+We have neglected the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force of buoyancy to balance the fluid drag.
+This consideration can be cast using scaling analysis into a simple criteria that the buoyancy time scale $\sqrt{\rho H/\Delta \rho g}$, where $\Delta \rho$ is the difference in densities between the vegetation and the water and $g$ is the acceleration due to gravity, be much faster than the hydrodynamic time scale $H/U_0$.
+The former can be estimated using representative parameter values $\Delta \rho /\rho \approx 0.1$ and $H=1$ m to be about 1 s, and the latter assuming $U_0 \approx 0.1$ m/s to be about 10 s, which is about one order of magnitude slower.
+Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 
+We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
+Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.
+
 In a real meadow, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
@@
 MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
-We thanks Heidi Nepf and Marco Ghisalberti for helpful discussions.
+We thank Heidi Nepf and Marco Ghisalberti for helpful discussions.
 

### 2014-10-02 Paper/grass.tex
Cite White and Nepf, 2007

@@
 While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
-(i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent. 
+(i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96,White07} does not account for this, rendering this explanation inconsistent. 
 (ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
-(iii)The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the top of the meadow, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
-Here we present a self-consistent mathematical model for the hydrodynamic linear stability analysis that accounts correctly for these effects, while also explaining lab experiments and field observations.
+(iii) The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the top of the meadow, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
+Here we present a mathematical model for the hydrodynamic linear stability analysis that accounts correctly for these effects, while also explaining lab experiments and field observations.
 \newline
@@
 % Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
-
+\begin{figure}
+\includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
+ and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The meadow extends upto $y=h_g$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the top of the meadow.
+The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
+\label{basicflow}
+\end{figure}
+\begin{figure}
+\begin{center}
+\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{K_vs_shear_width_noshear}
+\end{center}
+\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
+ parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are 
+also included in the top panel. Inset compares experimental observations 
+of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of eq.~\eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
+Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+\label{Re_vs_delta}
+\end{figure}
 We use a mean field model for the coupling between the flow and the vegetation. 
@@
 \newline
-To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient $dP/dx$ satisfies 
-\begin{equation}
- -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
-\label{base_equ}
-\end{equation}
-where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
-In the limit of large vegetation density, shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, and hence the bottom surface ($y=0$) is approximated by zero shear.
-A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
-Flow within vegetation has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
-Continuity of flow speed and shear stresses at the top of the meadow gives rise to a boundary layer.
-\begin{figure}
-\includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
-plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The meadow extends upto $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the top of the meadow.
-The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
-\label{basicflow}
-\end{figure}
-This boundary layer results from purely local dynamics, independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
-In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the top of the vegetation $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
-This dependence of the boundary layer thickness on the vegetation density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
-\newline
-To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
-\begin{equation}
-\begin{split}
-\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
-\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
-\end{split}
-\end{equation}
-These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
-We have three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$. 
-We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the mechanism of the instability. 
-With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
-The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ satisfies the eigenvalue problem.
-\begin{figure}
-\begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
-\end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
- parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are 
-also included in the top panel. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of eq.~\eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
-Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
-\label{Re_vs_delta}
-\end{figure}
-\begin{equation}
-\begin{split}
-\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&+D\left(2R \Ndg S U D \phi\right),
-\label{Orr-somerfield}
-\end{split}
-\end{equation}
-where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.
-
-We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
-solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
-in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
-The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
-destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-Previous linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
-either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of ad-hoc
-assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.
-
 \begin{figure*}
@@
 \end{figure*}
+To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 
+Such a steady state flow driven by constant pressure gradient $dP/dx$ satisfies 
+\begin{equation}
+ -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
+\label{base_equ}
+\end{equation}
+where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
+In the limit of large vegetation density, shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, and hence the bottom surface ($y=0$) is approximated by zero shear.
+A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
+Flow within vegetation has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
+Continuity of flow speed and shear stresses at the top of the meadow gives rise to a boundary layer.
+In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the top of the vegetation $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
+This boundary layer results from purely local dynamics, independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+The dependence of the boundary layer thickness on the vegetation density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
+\newline
+To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
+\begin{equation}
+\begin{split}
+\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
+\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
+\end{split}
+\end{equation}
+These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
+We have three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$. 
+We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the mechanism of the instability. 
+With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
+The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ satisfies the eigenvalue problem
+\begin{equation}
+\begin{split}
+\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
+&+D\left(2R \Ndg S U D \phi\right),
+\label{Orr-somerfield}
+\end{split}
+\end{equation}
+where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$.
+
+We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
+solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
+in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
+The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
+destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+A similar conclusion was presented \cite{White07} in the case flow around emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
+Linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
+either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc}
+assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.
 

### 2014-10-02 Paper/grass.tex
three figure per row
pulling

--- 2014-09-26 Paper/grass.tex
@@
 The vegetation, assumed rigid, only exerts a drag force on the flow.
-We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency, confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
-We also demonstrate that our instability is different from the Kelvin-Helmholtz instability.
+We show that a linear instability of a steady pressure-driven flow explains the experimentally observed flow threshold for waving and the waving frequency,
+ confirming the hydrodynamic mechanism for the onset of waving, where the vegetation flexibility merely serves to visualize the flow.
+We demonstrate that flow instability mechanism leading to waving is different from the Kelvin-Helmholtz instability.
 \end{abstract}
@@
 assumptions for the mean velocity profile \cite{Raupach96,Delangre06}.
-
 \begin{figure*}
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = .95]{Set4_dens28_imgsc}} &
+%{\includegraphics[scale = 0.78]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = .95]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = .95]{Set4_dens34_imgsc}}\\
+{\includegraphics[scale = .95]{Set5_dens38_imgsc}} &
+%{\includegraphics[scale = 0.78]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = .95]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = .95]{Set5_dens46_imgsc}} \\
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &

### 2014-10-02 Paper/grass.tex
Change to H/delta in the table

--- 2014-10-02 Paper/grass.tex
@@
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.67]{Set4_dens28_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.67]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.67]{Set5_dens38_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.67]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = 0.95]{Set4_dens28_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
+{\includegraphics[scale = 0.95]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.95]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.95]{Set5_dens38_imgsc}} &
+%{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
+{\includegraphics[scale = 0.95]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.95]{Set5_dens46_imgsc}} \\
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
@@
  Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
- Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
+ Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
  Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no

### 2014-10-02 Paper/grass.tex
modify scale for figure 3 in tex file to be in margin
Change to accomodate fig 3 panels

@@
 Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
-The damping from the vegetation drag leads to a finite threshold flow for the instability.
-This threshold agrees with experimental observations, and so does the predicted waving frequency.
-This role of vegetation drag differentiates our mechanism from the previous hypothesis that the Kelvin-Helmholz instability underlies \monami.
+This damping leads to a finite threshold flow for the instability, which agrees with experimental observations.
+This role of vegetation drag differentiates our mechanism from the previous hypothesis that the Kelvin-Helmholtz instability underlies \monami.
 \end{abstract}
@@
 Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
-The resulting changes in hydrodynamic conditions can influence number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-The simplest response of these flexible vegetation canopies to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
+The hydrodynamic processes resulting from this behavior influence a number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+One such response of these grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
 In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
@@
 \newline
-While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, several aspects of existing theory remain unsatisfactory. 
-(i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96,White07} does not account for this, rendering this explanation inconsistent. 
-(ii) Classical free shear flow is known to be unstable for all Reynolds numbers \cite{drazin}, whereas observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
+% While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, s
+Although it predicts the \monami ~waving frequency, several of its aspects remain unsatisfactory: 
+(i) The profile of the shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
+    The role of drag in damping the perturbations to the shear profile is also sometimes ignored \cite{Raupach96}. 
+% (i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent (however, see ref.
+\cite{White07}).
+(ii) Observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
+No existing theory has explained this threshold condition.
 (iii) The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the top of the meadow, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
@@
 \newline
-The drag exerted by the grass bed on the flow is central to the hypothesized instability for \monami. 
-The instability and the resulting flow structures, though manifest in the waving of grasses, also persist if the grass blades are replaced by rigid dowels \cite{Ghisal02}. 
-Therefore, for simplicity, we consider the grass bed to exchange momentum with the surrounding fluid but not deform. 
-A criteria for the validity of this assumptions is presented at the end of this letter.
+Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
+The instability and the resulting flow structures persist even if the grass blades are replaced by rigid dowels \cite{Ghisal02}. 
+% Therefore, we consider the grass bed to be rigid to probe the essential features of the instability. 
+Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
+The criteria for the validity of this assumption is verified \textit{a posteriori} in this letter.
 % Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
@@
 \end{figure}
-We use a mean field model for the coupling between the flow and the vegetation. 
-The force exerted by the vegetation is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
+
+% We use a mean field model for the coupling between the flow and the vegetation. 
+The drag exerted by the vegetation on the flow is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
 where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
@@
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.95]{Set4_dens28_imgsc}} &
+% <<<<<<< HEAD
+{\includegraphics[scale = 0.85]{Set4_dens28_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.95]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.95]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.95]{Set5_dens38_imgsc}} &
+{\includegraphics[scale = 0.85]{Set4_dens32_imgsc}} &
+{\includegraphics[scale = 0.85]{Set4_dens34_imgsc}} \\
+{\includegraphics[scale = 0.85]{Set5_dens38_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.95]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.95]{Set5_dens46_imgsc}} \\
+{\includegraphics[scale = 0.85]{Set5_dens42_imgsc}} &
+{\includegraphics[scale = 0.85]{Set5_dens46_imgsc}} \\
+% =======
+% {\includegraphics[scale = 0.93]{Set4_dens28_imgsc}} &
+% %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
+% {\includegraphics[scale = 0.93]{Set4_dens32_imgsc}} &
+% {\includegraphics[scale = 0.93]{Set4_dens34_imgsc}} \\
+% {\includegraphics[scale = 0.93]{Set5_dens38_imgsc}} &
+% %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
+% {\includegraphics[scale = 0.93]{Set5_dens42_imgsc}} &
+% {\includegraphics[scale = 0.93]{Set5_dens46_imgsc}} \\
+% >>>>>>> origin/master
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
@@
  Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no
+ Mode localized?	& yes, near $y=0$				& ~~yes, near $y=\hg$~~			& ~~no, spans water column~~
  \end{tabular}
@@
 The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-We have neglected the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force of buoyancy to balance the fluid drag.
-This consideration can be cast using scaling analysis into a simple criteria that the buoyancy time scale $\sqrt{\rho H/\Delta \rho g}$, where $\Delta \rho$ is the difference in densities between the vegetation and the water and $g$ is the acceleration due to gravity, be much faster than the hydrodynamic time scale $H/U_0$.
-The former can be estimated using representative parameter values $\Delta \rho /\rho \approx 0.1$ and $H=1$ m to be about 1 s, and the latter assuming $U_0 \approx 0.1$ m/s to be about 10 s, which is about one order of magnitude slower.
+This consideration can be cast using scaling analysis into a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
+The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
+Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the time buoyancy time scale to be about 2 s.
+Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
 Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 

### 2014-10-03 Paper/grass.tex
Intermediate commit with language changes

@@
 \newcommand{\monami}{\textit{monami}}
+\newcommand{\ubl}{u_\text{bl}}
 \definecolor{tableShade}{gray}{0.8}
@@
 One such response of these grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
-In this letter we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
+In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
 \newline
-Current explanations of \monami ~invoke the existence of a shear layer due to vegetation drag  \cite{Ghisal02,Raupach96} and its instability through a mechanism similar to the Kelvin-Helmholtz instability.
-This instability is thought to lead to coherent eddies over the sea grass meadow, which then drive large amplitude synchronous oscillations.
+Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ghisal02,Raupach96}. 
+Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
+This model has been applied successfully to ...!
 \newline
 % While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, s
-Although it predicts the \monami ~waving frequency, several of its aspects remain unsatisfactory: 
-(i) The profile of the shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
-    The role of drag in damping the perturbations to the shear profile is also sometimes ignored \cite{Raupach96}. 
+However, several aspects of the shear layer model remain unsatisfactory: 
+(i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
+(ii)  The role of drag in damping the perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
 % (i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent (however, see ref.
 \cite{White07}).
-(ii) Observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
+(iii) Observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
 No existing theory has explained this threshold condition.
-(iii) The thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the top of the meadow, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability.
-Here we present a mathematical model for the hydrodynamic linear stability analysis that accounts correctly for these effects, while also explaining lab experiments and field observations.
+(iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the definition of the free shear layer.
+Here we present a mathematical model for the linear instability that accounts correctly for these effects, while also explaining lab experiments and field observations.
 \newline
 Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
-The instability and the resulting flow structures persist even if the grass blades are replaced by rigid dowels \cite{Ghisal02}. 
+The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02}. 
 % Therefore, we consider the grass bed to be rigid to probe the essential features of the instability. 
-Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
-The criteria for the validity of this assumption is verified \textit{a posteriori} in this letter.
+% Therefore, assuming that the restoring force of buoyancy is sufficiently strong, we neglect the motion of the grass blades in this analysis.
+% Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
+Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify it \textit{a posteriori}.
 % Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
@@
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Comparison of steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+\caption{Schematic setup and comparison of a steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
 plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
- and it's approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The meadow extends upto $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the top of the meadow.
-The dependence of the boundary layer thickness (estimated numerically as $|U/U_y|$ at $y=\hg$) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
+ and its approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The grass extends up to $y=h_g$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the grass top.
+The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
 \label{basicflow}
@@
 \end{center}
-\caption{Critical Reynolds number with flow (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density
- parametrized by the boundary layer thickness. Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are 
-also included in the top panel. Inset compares experimental observations 
-of the dominant frequency $f_o$ (in Hz) measured with the predictions from the solution of eq.~\eqref{Orr-somerfield}. The experimental data in the inset in obtained from publications by 
-Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+\caption{Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
+Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions from the solution of eq.~\eqref{Orr-somerfield}. 
+The experimental data in the inset is obtained from publications by Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. 
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
 \label{Re_vs_delta}
@@
 % We use a mean field model for the coupling between the flow and the vegetation. 
-The drag exerted by the vegetation on the flow is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  
-where $\mathbf{f_{d}}$ is drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}. 
-In the vegetation, the equations of continuity and momentum balance can be written as 
+The drag exerted by the vegetation on the flow is approximated by a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  where $\mathbf{f_{d}}$ is the drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit horizontal area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}.
+This drag enters the governing equations of continuity and momentum balance, written as 
 \begin{equation}
@@
 $p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is expected to be $O(10^2-10^3)$, implying that the form drag on the vegetation dominates over the skin-friction.
-Hence we model it as
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore the form drag on the vegetation dominates over the skin-friction. 
+We model it as 
 $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
-where $d$ is average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass. 
-In the field, both $C_N$ and $N_g$ vary with distance from the bottom but we do not expect these variations to be central to the instability mechanism and therefore take $C_N$ and $N_g$ to be constants.
+where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic).
+In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
 Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.
@@
 \begin{tabular}{cccc}
-% <<<<<<< HEAD
 {\includegraphics[scale = 0.85]{Set4_dens28_imgsc}} &
@@
 {\includegraphics[scale = 0.85]{Set5_dens46_imgsc}} \\
-% =======
-% {\includegraphics[scale = 0.93]{Set4_dens28_imgsc}} &
-% %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-% {\includegraphics[scale = 0.93]{Set4_dens32_imgsc}} &
-% {\includegraphics[scale = 0.93]{Set4_dens34_imgsc}} \\
-% {\includegraphics[scale = 0.93]{Set5_dens38_imgsc}} &
-% %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-% {\includegraphics[scale = 0.93]{Set5_dens42_imgsc}} &
-% {\includegraphics[scale = 0.93]{Set5_dens46_imgsc}} \\
-% >>>>>>> origin/master
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
@@
 \end{figure*}
-To understand the onset of \monami, we first calculate the steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ as a solution of eq.~\eqref{navier-stokes}, and examine its linear stability. 
-Such a steady state flow driven by constant pressure gradient $dP/dx$ satisfies 
+We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of eq.~\eqref{navier-stokes}. 
+Such a flow satisfies 
 \begin{equation}
@@
 \end{equation}
-where $S(y)=1$ in the vegetated layer $0<y<\hg$ and $S(y)=0$ in the unvegetated layer $\hg< y< 2H$. 
-In the limit of large vegetation density, shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, and hence the bottom surface ($y=0$) is approximated by zero shear.
-A comparison of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements is shown in fig.~\ref{basicflow}.
-Flow within vegetation has approximately constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from the balance between drag and pressure gradient, and the flow outside has a simple parabolic velocity profile due to the balance between viscous forces and the pressure gradient.
-Continuity of flow speed and shear stresses at the top of the meadow gives rise to a boundary layer.
-In our model the balance between viscous forces and the drag force in the shear layer $(\mu u_\text{sl}/\delta^2 \sim \rho C_N d N_g u_\text{sl}^2)$ and the continuity of shear stress across the top of the vegetation $(u_\text{sl}/\delta \sim U_0/H)$, leads to a shear layer thickness $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow, $u_{sl}$ is the velocity scale in shear layer and $U_0 = {(dP/dx)~H^2}/{\mu}$ is the velocity scale in the unvegetated region.
-This boundary layer results from purely local dynamics, independent of the influence of the boundaries, and we identify it as a free shear layer \cite{Ghisal02} in the previous explanation of \monami.
-The dependence of the boundary layer thickness on the vegetation density $N_g$, shown in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
+where $S(y)=1$ in the grass bed ($0<y<\hg$) and $S(y)=0$ above it ($\hg< y< 2H$). 
+Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
+The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag. 
+This expectations is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
+The profile $U(y)$ has three distinct regions.
+Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$, arising from the balance between drag and pressure gradient. 
+Outside the vegetation, the velocity has a simple parabolic profile due to the balance between viscous forces and the pressure gradient. 
+At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
+Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, and we identify to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+Denoting $\ubl$ to be the velocity scale in boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
+Solving for $\delta$ yields $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$.
+% ; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
 \newline
@@
 \end{equation}
-These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$. 
-We have three non-dimensional parameters in the model, ie. $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$. 
+These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$.
 We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the mechanism of the instability. 
 With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation. 
-The growth rate $\sigma$ for a given wave number $k=2\pi /\lambda$ satisfies the eigenvalue problem
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation 
 \begin{equation}
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+The growth rate $\sigma$ for a given wave number $k$ is the eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.
 
-We solve this eigenvalue problem and a threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$), emerges from the 
-solution of eq.~\eqref{Orr-somerfield}. The dependence of this critical $\Rey$, and the corresponding fastest growing wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
+A threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
+solution of eq.~\eqref{Orr-somerfield}. The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
 in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
@@
 destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented \cite{White07} in the case flow around emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
-Linear stability calculations for terrestrial grass do not exhibit a threshold flow condition 
+A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
+Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition 
 either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc}
@@
 The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of 
-vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies.
+vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. The experimentally observed \monami ~wavelengths are not available for comparison.
 
@@
 The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-This consideration can be cast using scaling analysis into a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
+The rigidity of the vegetation can be evaluated using a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
 The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
-Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the time buoyancy time scale to be about 2 s.
+Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the buoyancy time scale to be about 2 s.
 Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
-Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 
+% Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 
 We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
@@
 Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-\newline\newline
+\newline
 In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 

### 2014-10-03 Paper/grass.tex
pull

--- 2014-10-02 Paper/grass.tex
@@
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.95]{Set4_dens28_imgsc}} &
+{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.95]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.95]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.95]{Set5_dens38_imgsc}} &
+{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
+{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
+{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.95]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.95]{Set5_dens46_imgsc}} \\
+{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
+{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &
@@
  Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
- Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
+ Most unstable $k$ as $\delta \to 0$	& $\propto \delta^{-1}$		& $\propto \delta^{-1}$	& $O(1)$ \\
  Mode localized?	& yes, near $y=0$				& ~~~~yes, near $y=\hg$~~~~			& no

### 2014-10-03 Paper/grass.tex

--- 2014-10-03 Paper/grass.tex
@@
 \begin{tabular}{cccc}
-{\includegraphics[scale = 0.85]{Set4_dens28_imgsc}} &
+{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[scale = 0.85]{Set4_dens32_imgsc}} &
-{\includegraphics[scale = 0.85]{Set4_dens34_imgsc}} \\
-{\includegraphics[scale = 0.85]{Set5_dens38_imgsc}} &
+{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
+{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
+{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[scale = 0.85]{Set5_dens42_imgsc}} &
-{\includegraphics[scale = 0.85]{Set5_dens46_imgsc}} \\
+{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
+{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\
+
 %{\includegraphics[scale = 0.67]{Set6_dens32_imgsc}} &

### 2014-10-06 Paper/grass.tex
added figure with single colorbar and modified tex file accordingly

@@
 \begin{figure*}
-\begin{tabular}{cccc}
-{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
+ \includegraphics{SetAll_imgsc}
+%\end{figure}
+%\begin{figure*}
+%\begin{tabular}{cccc}
+%{\includegraphics[height = 4cm, width = 5.8cm]{Set4_dens28_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set4_dens30_imgsc}} &
-{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
-{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
-{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
+%{\includegraphics[height = 4cm,width = 5.8cm]{Set4_dens32_imgsc}} &
+%{\includegraphics[height = 4.15cm,width=6.5cm]{Set4_dens34_imgsc}} \\
+%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens38_imgsc}} &
 %{\includegraphics[scale = 0.67]{Set5_dens40_imgsc}} &
-{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
-{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\
+%{\includegraphics[height = 4cm, width = 5.8cm]{Set5_dens42_imgsc}} &
+%{\includegraphics[height = 4.15cm,width=6.5cm]{Set5_dens46_imgsc}} \\
 
@@
 %{\includegraphics[scale = 0.67]{Set6_dens38_imgsc}}
-\end{tabular}
+%\end{tabular}
 \caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.

### 2014-10-06 Paper/grass.tex
Almost finished language edit

--- 2014-10-03 Paper/grass.tex
@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \newcommand{\monami}{\textit{monami}}
-\newcommand{\ubl}{u_\text{bl}}
+\newcommand{\ubl}{U_\text{bl}}
 \definecolor{tableShade}{gray}{0.8}
@@
 \affiliation{Brown University, Providence RI 02912 USA}
-\author{L. Mahadevan}
-\affiliation{Harvard University, Cambridge MA 02138 USA}
+% \author{L. Mahadevan}
+% \affiliation{Harvard University, Cambridge MA 02138 USA}
 \author{M. M. Bandi}
@@
 % Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
-Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify it \textit{a posteriori}.
+Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify the assumption \textit{a posteriori}.
 % Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
@@
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag. 
+The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, whereas the latter is due to the free interface.
 This expectations is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
@@
 At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, and we identify to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
 Denoting $\ubl$ to be the velocity scale in boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
-Solving for $\delta$ yields $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\Rey\Ndg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
 This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$.
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. Using these variables, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$. 
 % ; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
 \newline
-To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
+Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$.
+These perturbations obey (dropping the tilde decoration)
 \begin{equation}
@@
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
-To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
+To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
 The lowest $\Rey$ on the neutral curve sets the threshold. 
@@
 Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
 The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
@@
 in a region of thickness O($\delta$) near $y=\hg$.
-Since the relative magnitude of the terms on the r.h.s.
-to those on the l.h.s.
-of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
+Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. 
 Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}). 
@@
 On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
-Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-This estimate simplifies eq.~\eqref{Orr-somerfield} to 
+% Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+In this limit eq.~\eqref{Orr-somerfield} simplifies to 
 \begin{align}
@@
 
-In a real meadow, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
+In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
 The eddy viscosity is also known to vary between the flow of unvegetated region and flow through the meadow \cite{Ghisal02}. 
@@
 Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
-The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
+% The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
@@
 \bibliography{Grass}{}
-\bibliographystyle{plain}
+% \bibliographystyle{plain}
+\bibliographystyle{aipauth4-1}
 \end{document}

### 2014-10-06 Paper/grass.tex
added figure with single colorbar and modified tex file accordingly
Language cleanup

@@
 \graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
+\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
 \newcommand{\bn}{{\boldsymbol{\hat{n}}}}
@@
 \newline
-% While the shear layer model appears to predict the frequency of \monami ~consistent with experimental observations, s
 However, several aspects of the shear layer model remain unsatisfactory: 
@@
 (ii)  The role of drag in damping the perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
-% (i) While the free shear layer is established due to the vegetation drag, linear stability analysis \cite{Raupach96} does not account for this, rendering this explanation inconsistent (however, see ref.
-\cite{White07}).
-(iii) Observations in the lab \cite{Ghisal02} and the field \cite{Grizzle96} indicate the existence of a threshold flow speed below which \monami ~is not observed.
-No existing theory has explained this threshold condition.
+(iii) No existing theory has explained the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
 (iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the definition of the free shear layer.
@@
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02}. 
-% Therefore, we consider the grass bed to be rigid to probe the essential features of the instability. 
-% Therefore, assuming that the restoring force of buoyancy is sufficiently strong, we neglect the motion of the grass blades in this analysis.
-% Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
-Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify the assumption \textit{a posteriori}.
-% Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented along the $y$ direction, and verify the assumption \textit{a posteriori}.
 \begin{figure}
@@
 % We use a mean field model for the coupling between the flow and the vegetation. 
-The drag exerted by the vegetation on the flow is approximated by a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  where $\mathbf{f_{d}}$ is the drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit horizontal area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}.
+The vegetation drag on the flow is approximated by a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  where $\mathbf{f_{d}}$ is the drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit horizontal area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}.
 This drag enters the governing equations of continuity and momentum balance, written as 
@@
 $p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore the form drag on the vegetation dominates over the skin-friction. 
-We model it as 
-$\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
-where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic).
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f_{d}}=C_N \rho u^{2}d\bx$, where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient.
+% We model it as 
+% $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
+% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic).
 In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
@@
 \rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
-\end{split}
+\end{split} \nonumber
 \end{equation}

### 2014-10-08 Paper/grass.tex
All changes made
- Waiting for the green to change to another color
- Waiting for the last sentence in paragraph two.

@@
 \newcommand{\hg}{h_g}
-\newcommand{\Rey}{\text{R}}
+\newcommand{\Rey}{{R}}
 \newcommand{\Ndg}{\tilde{N}_g}
@@
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied successfully to ...!
+This model has been applied successfully to predict the the observed frequency of \monami ...!
 \newline
@@
 (i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
-(ii)  The role of drag in damping the perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
+(ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
 (iii) No existing theory has explained the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
@@
 Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented along the $y$ direction, and verify the assumption \textit{a posteriori}.
-\begin{figure}
+We show that this drag results in an instability mode different from the Kelvin-Helmholtz instability.
+\begin{figure}[b]
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
@@
 Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
-Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions from the solution of eq.~\eqref{Orr-somerfield}. 
+Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p$ from the solution of eq.~\eqref{Orr-somerfield}. 
 The experimental data in the inset is obtained from publications by Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. 
@@
 
-% We use a mean field model for the coupling between the flow and the vegetation. 
-The vegetation drag on the flow is approximated by a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$,  where $\mathbf{f_{d}}$ is the drag force per unit length of the grass blade, and $N_g$ the grass blade number density per unit horizontal area \cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06}.
-This drag enters the governing equations of continuity and momentum balance, written as 
+The vegetation, assumed dense, exerts a drag modeled by a continuous body force $\mathbf{f}$.
+% We assume the vegetation to be dense and model the drag by a continuous body force $\mathbf{f}$.
+This drag enters the fluid mass and momentum balance equations as 
 \begin{equation}
@@
 $p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f_{d}}=C_N \rho u^{2}d\bx$, where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient.
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u^{2}d\bx$, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
 % We model it as 
@@
 In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
-Based on previous experiments \cite{Vivoni98}, we take $C_N \approx 1$.
+Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N \approx 1$.
 \newline
@@
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, whereas the latter is due to the free interface.
-This expectations is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
+The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter is due to the free interface.
+This expectation is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
 The profile $U(y)$ has three distinct regions.
@@
 At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
 Denoting $\ubl$ to be the velocity scale in boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
@@
 With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to  obtain a modified Orr-Sommerfield equation \cite{Drazin81} 
 \begin{equation}
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
-The growth rate $\sigma$ for a given wave number $k$ is the eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.
+The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.
 
-A threshold in $\Rey$ of $O(100)$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
-solution of eq.~\eqref{Orr-somerfield}. The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown 
-in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
-The threshold Reynolds number increases with the vegetation density, indicating a competition between the 
-destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
+solution of eq.~\eqref{Orr-somerfield}.
+The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
+The threshold Reynolds number increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
 A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
-Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition 
-either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc}
-assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.
+Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc} assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.
 
-Comparison of frequency (Im$(\sigma)$) of the fastest growing mode with that of experimental observations of frequencies in
-the lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b. 
-The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of 
-vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. The experimentally observed \monami ~wavelengths are not available for comparison.
+Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b.
+The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
+The experimentally observed \monami ~wavelengths are not available for comparison.
 
@@
  \end{tabular}
- \caption{Comparison between Kelvin-Helmholtz instability and the two modes resulting from solution of \ref{Orr-somerfield}.}
+ \caption{Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
  \label{tab:comparison}
@@
 To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
-The lowest $\Rey$ on the neutral curve sets the threshold. 
-We observe that as $\Ndg$ increases, the unstable region splits into two. We refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
+The lowest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
+We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
 The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
@@
 Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
-% Based on this separation of time scales, role of vegetation deformation appears negligible, or at least not central to the hydrodynamic instability mechanism. 
 We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
@@
 In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
-The eddy viscosity is also known to vary between the flow of unvegetated region and flow through the meadow \cite{Ghisal02}. 
-Although an improved analysis of flow with these variation in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
+The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
+Although these improvements in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
 \newline
@@
 Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
-% The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.

### 2014-10-08 Paper/grass.tex
added figures with different symbols and different color map

--- 2014-10-06 Paper/grass.tex
@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,twocolumn,superscriptaddress,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \newcommand{\monami}{\textit{monami}}
-\newcommand{\ubl}{U_\text{bl}}
+\newcommand{\ubl}{u_\text{bl}}
 \definecolor{tableShade}{gray}{0.8}
@@
 \affiliation{Brown University, Providence RI 02912 USA}
-% \author{L. Mahadevan}
-% \affiliation{Harvard University, Cambridge MA 02138 USA}
+\author{L. Mahadevan}
+\affiliation{Harvard University, Cambridge MA 02138 USA}
 \author{M. M. Bandi}
@@
 % Therefore we neglect the motion of the grass blades in this analysis, assuming that a small deviation from its equilibrium position generates sufficient restoring force to balance the fluid drag. 
-Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify the assumption \textit{a posteriori}.
+Therefore, to develop the essential mathematical model, we assume the vegetation to be rigid, and verify it \textit{a posteriori}.
 % Indeed, a more detailed calculation modeling terrestrial grass demonstrated \cite{Delangre06} that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.
@@
 \begin{figure*}
- \includegraphics{SetAll_imgsc}
+ \includegraphics{SetAll_imgsc_hsv}
 %\end{figure}
@@
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag, whereas the latter is due to the free interface.
+The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag. 
 This expectations is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
@@
 At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
+Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, and we identify to be analogous to the free shear layer \cite{Ghisal02} in the previous explanation of \monami.
 Denoting $\ubl$ to be the velocity scale in boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
-Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\Rey\Ndg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+Solving for $\delta$ yields $\delta=(\Rey\tilde{N_g})^{-1/3}H$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
 This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. Using these variables, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$. 
+The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$.
 % ; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
 \newline
-Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$.
-These perturbations obey (dropping the tilde decoration)
+To understand the evolution of small perturbations to the steady flow profile, we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to obtain (after dropping the tilde decoration)
 \begin{equation}
@@
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
-To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
+To aid in explaining the behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
 The lowest $\Rey$ on the neutral curve sets the threshold. 
@@
 Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim \delta^{-1}$, $\sigma \sim \delta^{-1}$, and $U\sim \delta$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
 The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
@@
 in a region of thickness O($\delta$) near $y=\hg$.
-Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. 
+Since the relative magnitude of the terms on the r.h.s.
+to those on the l.h.s.
+of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$.
 Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}).
@@
 On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
-% Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-In this limit eq.~\eqref{Orr-somerfield} simplifies to 
+Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
+This estimate simplifies eq.~\eqref{Orr-somerfield} to 
 \begin{align}
@@
 
-In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
+In a real meadow, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
 The eddy viscosity is also known to vary between the flow of unvegetated region and flow through the meadow \cite{Ghisal02}. 
@@
 Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
-% The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
+The extent of the structure of Mode 2 throughout the depth of the water column implies an efficient transport of suspended tracers compared to that due to the localized structure of Mode 1 near the top of the meadow.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
@@
 \bibliography{Grass}{}
-% \bibliographystyle{plain}
-\bibliographystyle{aipauth4-1}
+\bibliographystyle{plain}
 \end{document}

### 2014-10-08 Paper/grass.tex
Cite Ghisalberti and Nepf, 2006 as reference for rigid dowels

--- 2014-10-08 Paper/grass.tex
@@
 Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
-The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02}. 
+The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
 Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented along the $y$ direction, and verify the assumption \textit{a posteriori}.

### 2014-10-22 Paper/cover.tex
Add cover letter and complete the paper

\documentclass[10pt]{letter}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{palatino}
\name{Shreyas Mandre}
\date{\today}
% \address{
% Shreyas Mandre \\
% 29 Oxford Street \#328 \\
% Pierce Hall Harvard University \\
% Cambridge MA 02138 USA
% }

\telephone{\url{shreyasmandre@gmail.com} \quad +1 401 863 2602}
\signature{Shreyas Mandre}
\pagestyle{empty}
\begin{document}
\begin{letter}{
Editor \\
Physical Review Letters \\
1 Research Road, \\
Ridge, NY 11961-2701 
}
\opening{Dear Editor,}

Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Amala Mahadevan, and myself for publication in the Physical Review Letters. The subject of this manuscript is the synchronous coherent oscillations of marine grass, known as {\it monami} when subject to steady flow. These oscillations are similar to but distinct from the analogous phenomenon on terrestrial grass subject to strong wind. In the manuscript, we present a mathematical model for the onset of these oscillations.

Coherent motion of the grass blades enhances the transport of nutrients, gametes, larvae, and other microscopic life forms within the grass bed. This transport in turn dictates the seagrass ecosystem and regulates the biomass in it. Although seagrasses occupy less than 0.05\% of the ocean area, they contribute directly to about 15\% of the total biomass production rate in the ocean. Further, the seagrass meadows engineer a habitat supporting thousands of other species, sequester carbon dioxide, recycle nutrients, and stabilize sediments. Due to their effectiveness in performing these functions, seagrasses are considered some of the world's most valuable resources. An improved understanding of the processes performed very effectively by the seagrasses is critical in managing the planet's seagrass meadows. The synchronous motion of the meadow is one such process allowing the submerged seagrass meadow to ``breathe''. Our manuscript is one of the first theoretical treatments on the physical mechanism behind these oscillations.

Our analysis shows that the existing explanation for these oscillations is incorrect. The existing explanation invokes the existence of a shear layer near the top of the grass bed, which is established due to the drag exerted by the grass of the flow. This shear layer is assumed to be unstable to the Kelvin-Helmholtz instability; the vortices that form as a result of this instability then drive the synchronous oscillations of the grass blades. We test this explanation mathematically in our manuscript, using a self-consistent linear stability analysis of the flow. We demonstrate that indeed the steady flow is hydrodynamically unstable and the instability mechanism, although superficially similar to the Kelvin-Helmholtz instability, is in fact distinct from it. The results from our model are supported by qualitative observations in the field and quantitative lab experiments reported in the literature.

We believe that the topic will be of interest to a wide audience, not only because of the relation to the commonplace terrestrial analogy, but also because of the importance of the seagrass ecosystems. We have endeavored to write it in a form that is accessible to the general physicist.

Should you have any questions, please do not hesitate to contact me.

\closing{Thanking you, \\Yours Sincerely,}

\end{letter}
\end{document}

### 2014-10-22 Paper/grass.tex
Add cover letter and complete the paper

@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+% \documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \begin{abstract}
-The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the steady flow.
+The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the flow.
 Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
@@
 The hydrodynamic processes resulting from this behavior influence a number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-One such response of these grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
+One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
 In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
@@
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied successfully to predict the the observed frequency of \monami ...!
+This model has been applied successfully to predict the the observed frequency of \monami~\cite{Ghisal02} and to understand transport in the seagrass bed.
 \newline
@@
 (ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
-(iii) No existing theory has explained the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
+(iii) No existing theory has explained the threshold flow speed for instability, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
 (iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the definition of the free shear layer.
@@
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
-Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented along the $y$ direction, and verify the assumption \textit{a posteriori}.
-We show that this drag results in an instability mode different from the Kelvin-Helmholtz instability.
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically, and verify the assumption \textit{a posteriori}.
+We show that the drag of the grass results in an instability mode different from the Kelvin-Helmholtz instability.
 \begin{figure}[b]
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
-\caption{Schematic setup and comparison of a steady flow profile from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+\caption{Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
 plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
@@
 \begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
+% \includegraphics[]{Critical_Re_vs_delta_noshear} \\
+% \vspace{-6mm} \hspace{-3mm}
+% \includegraphics[]{K_vs_shear_width_noshear}
+\includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-5mm}
+\includegraphics[width=12cm]{K_vs_shear_width_noshear}
 \end{center}
@@
 Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
-Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p$ from the solution of eq.~\eqref{Orr-somerfield}. 
+Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p$ from the solution of ~\eqref{Orr-somerfield}. 
 The experimental data in the inset is obtained from publications by Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. 
@@
 where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
-$p$ the pressure, $\mu$ the eddy viscosity (the flow is turbulent) and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it based on the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u^{2}d\bx$, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+$p$ the pressure, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u^{2}d\bx$, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+The vorticity shed on the grass-blade scale imparts the turbulent flow within and above the grass with specific characteristics; in the interest of simplicity, we ignore these characteristics and model the turbulence using an eddy viscosity.
 % We model it as 
 % $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
-% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see fig.~\ref{basicflow} for a schematic).
+% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see Fig.~\ref{basicflow} for a schematic).
 In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
-Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N \approx 1$.
+Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
 \newline
 \begin{figure*}
- \includegraphics{SetAll_imgsc_hsv}
+ \includegraphics[width=\textwidth]{SetAll_imgsc}
 %\end{figure}
@@
 \end{figure*}
-We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of eq.~\eqref{navier-stokes}. 
+We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of ~\eqref{navier-stokes}. 
 Such a flow satisfies 
@@
 The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter is due to the free interface.
-This expectation is verified by a comparison, shown in fig.~\ref{basicflow}, of the steady flow profile from the solution of eq.~\eqref{base_equ} with experimental measurements.
+This expectation is verified by a comparison, shown in Fig.~\ref{basicflow}, of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements.
 The profile $U(y)$ has three distinct regions.
@@
 Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
-Denoting $\ubl$ to be the velocity scale in boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
+Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
 Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\Rey\Ndg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
-This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
+This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
 The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. Using these variables, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$. 
-% ; minor departures from this scaling can be seen in the inset of fig.~\ref{basicflow}.
-\newline
-Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in eq.~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$.
-These perturbations obey (dropping the tilde decoration)
+% ; minor departures from this scaling can be seen in the inset of Fig.~\ref{basicflow}.
+
+Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
 \begin{equation}
@@
 \rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
-\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0.
+\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0,
 \end{split} \nonumber
 \end{equation}
+where the tilde are dropped.
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$.
@@
 where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
-The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of eq. \eqref{Orr-somerfield}.
+The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
 
 A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
-solution of eq.~\eqref{Orr-somerfield}.
-The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
+solution of ~\eqref{Orr-somerfield}.
+The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
 The threshold Reynolds number increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile \textit{ad hoc}, and neglecting the viscous term.
+A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term.
 Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc} assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.
 
-Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in fig.~\ref{Re_vs_delta}b.
+Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b.
 The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
@@
 To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see fig.~\ref{Re_vs_delta}). 
-To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in fig.~\ref{K_Re_sigma_set3}. 
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
+To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in Fig.~\ref{K_Re_sigma_set3}. 
 The lowest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
@@
 Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in eq.~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of eq.~\eqref{Orr-somerfield} as
+This limit can be understood through asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of ~\eqref{Orr-somerfield} as
 \begin{align}
@@
 Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. 
-Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see fig.~\ref{Re_vs_delta}). 
-This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in fig.~\ref{Asymptotic_mode}, supporting this argument.
+Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
+This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in Fig.~\ref{Asymptotic_mode}, supporting this argument.
 
-On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of eq.~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of ~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-In this limit eq.~\eqref{Orr-somerfield} simplifies to 
+In this limit ~\eqref{Orr-somerfield} simplifies to 
 \begin{align}
@@
 \begin{figure}
-\includegraphics[]{Asymptotic_noshear}
+\includegraphics[width=15cm]{Asymptotic_noshear}
 \caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 
@@
 \end{figure}
-Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in fig.~\ref{Asymptotic_mode}. 
+Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}. 
 We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the top of the vegetation plays no role.
@@
 Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
-Because Mode 1 eigenfunction is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
+Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
 Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
 Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
-The appearance of vegetation drag parameter in the dominant balances represented by eq.~\eqref{eqn:mode1asymp} and eq.~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
-\newline
-Similar phenomenon of large amplitude coherent oscillation of terrestrial canopies in atmospheric flow known as \textit{honami} \cite{Inoue56,Raupach96}, have also been observed.
+The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
+
+Similarly observed large amplitude coherent oscillation of terrestrial canopies in atmospheric flow is known as \textit{honami} \cite{Inoue56,Raupach96}.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tend to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}.
+Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tends to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}.
 Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
@@
 In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
-\newline
-The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-The rigidity of the vegetation can be evaluated using a simple criteria that the buoyancy time scale  be much faster than the hydrodynamic time scale $H/U_0$.
+
+We now test the assumption of a rigid grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
 The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
 Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the buoyancy time scale to be about 2 s.
-Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore slower than the hydrodynamic time scale.
+Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
 We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
@@
 
+The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
 In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
@@
 \newline
-In conclusion, we note that the threshold flow condition observed in the field and in lab experiments arises due to the presence of vegetation drag. 
-We show the origin of this threshold using a consistent continuum hydrodynamic stability analysis for the shear flow that develops as a result of the vegetation-induced drag. 
-Such an analysis not only explains quantitative observations, but also provides an understanding of the hydrodynamic instability mechanism. 
-Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, etc. and therefore is expected to have a wider impact.
+In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
+The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
+While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
+The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips, while Mode 2 instability influences the whole water column.
+Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, flow through urban environments and therefore is expected to have a wider impact.
 
@@
 MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
-We thank Heidi Nepf and Marco Ghisalberti for helpful discussions.
+We thank Heidi Nepf, Marco Ghisalberti, and L.
+Mahadevan for helpful discussions.
 

### 2014-10-22 Paper/grass.tex
Revert back to two-column format

@@
-% \documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
-\documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+% \documentclass[aps,preprint,10pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied successfully to predict the the observed frequency of \monami~\cite{Ghisal02} and to understand transport in the seagrass bed.
+This model has been applied to predict the the observed frequency of \monami~\cite{Ghisal02} and to understand transport in the seagrass bed.
 \newline
@@
 We show that the drag of the grass results in an instability mode different from the Kelvin-Helmholtz instability.
-\begin{figure}[b]
+\begin{figure}
 \includegraphics[scale=1]{Grass_Base_Nepf_shear}
+% \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
 \caption{Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
@@
 \begin{center}
-% \includegraphics[]{Critical_Re_vs_delta_noshear} \\
-% \vspace{-6mm} \hspace{-3mm}
-% \includegraphics[]{K_vs_shear_width_noshear}
-\includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-5mm}
-\includegraphics[width=12cm]{K_vs_shear_width_noshear}
+\includegraphics[]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{K_vs_shear_width_noshear}
+% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
+% \vspace{-6mm} \hspace{-5mm}
+% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
 \end{center}
@@
 \begin{figure}
-\includegraphics[width=15cm]{Asymptotic_noshear}
+\includegraphics[]{Asymptotic_noshear}
 \caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 

### 2014-10-22 Paper/grass.tex
Finishing touches, and polish.

@@
 In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
-\newline
+
 Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ghisal02,Raupach96}. 
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied to predict the the observed frequency of \monami~\cite{Ghisal02} and to understand transport in the seagrass bed.
-\newline
+This model has been applied to predict the the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04}.
+
 However, several aspects of the shear layer model remain unsatisfactory: 
@@
 Here we present a mathematical model for the linear instability that accounts correctly for these effects, while also explaining lab experiments and field observations.
-\newline
+
 Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
@@
 Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
-\newline
+
 \begin{figure*}
@@
 The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
-Although these improvements in the model might lead to a better agreement between the observed and the predicted quantities, these variation are not central to the mechanism leading to the instability of flow.
-\newline
+Although these model improvements might lead to a better agreement between the observed and the predicted quantities, they are not central to the mechanism leading to the instability of flow.
+
 In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 

### 2014-10-22 Paper/cover.tex
Fixed spelling in cover letter

@@
 Our analysis shows that the existing explanation for these oscillations is incorrect.
-The existing explanation invokes the existence of a shear layer near the top of the grass bed, which is established due to the drag exerted by the grass of the flow.
+The existing explanation invokes the existence of a shear layer near the top of the grass bed, which is established due to the drag exerted by the grass on the flow.
 This shear layer is assumed to be unstable to the Kelvin-Helmholtz instability; the vortices that form as a result of this instability then drive the synchronous oscillations of the grass blades.

### 2014-10-24 Paper/cover.tex
Cover letter after Amala's changes

@@
 % Shreyas Mandre \\
-% 29 Oxford Street \#328 \\
-% Pierce Hall Harvard University \\
-% Cambridge MA 02138 USA
-% }
+%182 Hope Street \\
+%Box D \\
+%Providence RI 02912
+%}
 
-\telephone{\url{shreyasmandre@gmail.com} \quad +1 401 863 2602}
-\signature{Shreyas Mandre}
+\telephone{\url{shreyas_mandre@brown.edu} \quad +1 401 863 2602}
+\signature{Shreyas Mandre \\ Assistant Professor \\ School of Engineering, Brown University}
 \pagestyle{empty}
@@
 
-Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Amala Mahadevan, and myself for publication in the Physical Review Letters.
-The subject of this manuscript is the synchronous coherent oscillations of marine grass, known as {\it monami} when subject to steady flow.
-These oscillations are similar to but distinct from the analogous phenomenon on terrestrial grass subject to strong wind.
+Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Amala Mahadevan, and myself for publication in  Physical Review Letters.
+This manuscript addresses the synchronous coherent oscillations of aquatic vegetation in a steady flow, known as {\it monami}.
+The phenomenon is analogous to the more commonly observed synchronous waving of terrestrial grasses in the wind.
 In the manuscript, we present a mathematical model for the onset of these oscillations.
 
-Coherent motion of the grass blades enhances the transport of nutrients, gametes, larvae, and other microscopic life forms within the grass bed.
-This transport in turn dictates the seagrass ecosystem and regulates the biomass in it.
-Although seagrasses occupy less than 0.05\% of the ocean area, they contribute directly to about 15\% of the total biomass production rate in the ocean.
-Further, the seagrass meadows engineer a habitat supporting thousands of other species, sequester carbon dioxide, recycle nutrients, and stabilize sediments.
+Our work examines the instability at the interface of a submerged grass bed and overlying fluid in the presence of a unidirectional flow.
+The impedance of the flow by the grass sets up a shear layer above the grass.
+It has been previously postulated from experiments that this shear layer is susceptible to  Kelvin-Helmholtz instability, which drives the waving of the grass.
+Here, we challenge this notion by  showing that the presence of drag due to the grass generates a different instability mechanism, distinct from Kelvin-Helmholtz instability.
+This new instability has characteristics consistent with the experimental observations, and our theory makes verifiable predictions about the threshold velocity for its onset.
+
+Coherent motion of seagrass meadows enhances the transport of nutrients, gametes, larvae, and other microscopic life forms within the habitat they form.
+This transport in turn influences the seagrass ecosystem and regulates the biomass in it.
+Although seagrasses occupy less than 0.05\% of the ocean area, they contribute directly to about 15\% of the total biomass production rate in the ocean, provide habitat to thousands of other species, sequester carbon, recycle nutrients, and stabilize sediments.
 Due to their effectiveness in performing these functions, seagrasses are considered some of the world's most valuable resources.
 An improved understanding of the processes performed very effectively by the seagrasses is critical in managing the planet's seagrass meadows.
-The synchronous motion of the meadow is one such process allowing the submerged seagrass meadow to ``breathe''.
-Our manuscript is one of the first theoretical treatments on the physical mechanism behind these oscillations.
+The synchronous motion of the meadow is one such process allowing the submerged seagrass meadow to exchange matter with the overlying fluid.
+Our manuscript is one of the few theoretical treatments on the physical mechanism behind these oscillations.
 
-Our analysis shows that the existing explanation for these oscillations is incorrect.
-The existing explanation invokes the existence of a shear layer near the top of the grass bed, which is established due to the drag exerted by the grass on the flow.
+%Previous experiments showed the existence of a shear layer between the submerged grass and overlying fluid that was conjectured to be susceptible to  Kelvin-Helmholtz instability.
+%
+%Our analysis shows that the existing explanation for these oscillations is incorrect.
+The existing explanation invokes the existence of a shear layer near the top of the grass bed, which is established due to the drag exerted by the grass of the flow.
 This shear layer is assumed to be unstable to the Kelvin-Helmholtz instability; the vortices that form as a result of this instability then drive the synchronous oscillations of the grass blades.
@@
 
-We believe that the topic will be of interest to a wide audience, not only because of the relation to the commonplace terrestrial analogy, but also because of the importance of the seagrass ecosystems.
+We believe that the topic will be of interest because of its pertinence to the commonplace terrestrial analogy and the importance of the seagrass ecosystems.
 We have endeavored to write it in a form that is accessible to the general physicist.

### 2014-10-25 Paper/grass.tex
Add references and improve conclusions

@@
 
-Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ghisal02,Raupach96}. 
+Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}. 
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied to predict the the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04}.
+This model has been applied to predict the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
@@
 
-The vegetation, assumed dense, exerts a drag modeled by a continuous body force $\mathbf{f}$.
-% We assume the vegetation to be dense and model the drag by a continuous body force $\mathbf{f}$.
+The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
 This drag enters the fluid mass and momentum balance equations as 
@@
 $p$ the pressure, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u^{2}d\bx$, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
-The vorticity shed on the grass-blade scale imparts the turbulent flow within and above the grass with specific characteristics; in the interest of simplicity, we ignore these characteristics and model the turbulence using an eddy viscosity.
-% We model it as 
-% $\mathbf{f_{d}}=C_N \rho u_{N}^{2}d\bn$, 
-% where $d$ is the average width of grass blade projected in the flow direction, $C_{N}$ the drag coefficient, and $u_{N}$ and $\bn$ are the velocity component and unit vector normal to the grass (see Fig.~\ref{basicflow} for a schematic).
-In the field, both $C_N$ and $N_g$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take $C_N$ and $N_g$ to be constants. 
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+In the interest of simplicity, we model the turbulence using an eddy viscosity. 
+In the field, both $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
 Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
@@
 \begin{equation}
- -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0
+ -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU |U|=0
 \label{base_equ}
@@
 This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. Using these variables, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$. 
-% ; minor departures from this scaling can be seen in the inset of Fig.~\ref{basicflow}.
+The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$.
 
@@
 A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term.
-Previous linear stability calculations for terrestrial grass have not reported a threshold flow condition either due to the exclusion of the vegetation drag in their model \cite{Raupach96}, or because of the \textit{ad hoc} assumptions about the mean velocity profile \cite{Raupach96,Delangre06}.
+Previous calculations for terrestrial grass differ from our work as they either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile {\it ad hoc} \cite{Raupach96,Delangre06}.
+They have not reported a threshold flow condition.
 
@@
 Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood through asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
+This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
 The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of ~\eqref{Orr-somerfield} as
@@
 \end{figure}
-Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape would converge in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}. 
+Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape converges in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}. 
 We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the top of the vegetation plays no role.
@@
 Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
-Mode 1 appears to be closely related to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation. 
+Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
 Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
@@
 
-We now test the assumption of a rigid grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
+We now test the assumption of a undeformable grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
 The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
@@
 The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
-Although these model improvements might lead to a better agreement between the observed and the predicted quantities, they are not central to the mechanism leading to the instability of flow.
+Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
 
@@
 While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
-The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips, while Mode 2 instability influences the whole water column.
+The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, flow through urban environments and therefore is expected to have a wider impact.

### 2014-11-03 Paper/grass.tex
APS preprint style in 12 pts for uploading on the arXiv
Preprint format for revtex-4.1

@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
-% \documentclass[aps,preprint,10pt]{revtex4-1}  % for review and submission
+% \documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
-\author{Ravi Singh}
+\author{Ravi Singh}t
 \affiliation{Brown University, Providence RI 02912 USA}
@@
 \begin{figure}
-\includegraphics[scale=1]{Grass_Base_Nepf_shear}
-% \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
+% \includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\includegraphics[width=15cm]{Grass_Base_Nepf_shear}
 \caption{Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
@@
 \begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
-% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
-% \vspace{-6mm} \hspace{-5mm}
-% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
+% \includegraphics[]{Critical_Re_vs_delta_noshear} \\
+% \vspace{-6mm} \hspace{-3mm}
+% \includegraphics[]{K_vs_shear_width_noshear}
+\includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
+\vspace{-6mm} \hspace{-5mm}
+\includegraphics[width=12cm]{K_vs_shear_width_noshear}
 \end{center}
@@
 \begin{table*}
+\footnotesize
 \rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
-\renewcommand{\arraystretch}{1.4}
+\renewcommand{\arraystretch}{1.2}
  \begin{tabular}{l|c|c|c}
@@
  Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& ~~yes, near $y=\hg$~~			& ~~no, spans water column~~
+ Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no, spans water column
  \end{tabular}
@@
 \begin{figure}
-\includegraphics[]{Asymptotic_noshear}
+% \includegraphics[]{Asymptotic_noshear}
+\includegraphics[width=12cm]{Asymptotic_noshear}
 \caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 

### 2014-11-03 Paper/grass.tex
Add word count!

--- 2014-10-25 Paper/grass.tex
@@
 \documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
-% \documentclass[aps,preprint,10pt]{revtex4-1}  % for review and submission
+% \documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
 %\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
@@
 \newcommand{\ubl}{U_\text{bl}}
+\newcommand{\words}[1]{\textbf{(#1)~}}
+% \newcommand{\words}[1]{{}}
 \definecolor{tableShade}{gray}{0.8}
@@
 \maketitle
-Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
+\words{90} Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
 The hydrodynamic processes resulting from this behavior influence a number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
@@
 
-Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}. 
+\words{80} Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}.
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
@@
 
-However, several aspects of the shear layer model remain unsatisfactory: 
+\words{137} However, several aspects of the shear layer model remain unsatisfactory: 
 (i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
@@
 
-Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
+\words{95} Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
@@
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
-\caption{Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
+\caption{
+\words{129} Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
 plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
@@
 The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the grass top.
-The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.}
+The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.
+}
 \label{basicflow}
@@
 \end{center}
-\caption{Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+\caption{
+\words{110} Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
 Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
@@
 The experimental data in the inset is obtained from publications by Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. 
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+}
 \label{Re_vs_delta}
@@
 
-The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
+\words{28} The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
 This drag enters the fluid mass and momentum balance equations as 
@@
 \end{equation}
-where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
+\words{158} where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
 $p$ the pressure, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
@@
 %\end{tabular}
-\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
+\caption{
+\words{69} $Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
 As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
-For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.}
+For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.
+}
 \label{K_Re_sigma_set3}
 \end{figure*}
-We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of ~\eqref{navier-stokes}. 
+\words{32} We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of ~\eqref{navier-stokes}. 
 Such a flow satisfies 
@@
 \end{equation}
-where $S(y)=1$ in the grass bed ($0<y<\hg$) and $S(y)=0$ above it ($\hg< y< 2H$). 
+\words{377} where $S(y)=1$ in the grass bed ($0<y<\hg$) and $S(y)=0$ above it ($\hg< y< 2H$). 
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
@@
 
-Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
+\words{28} Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
 \begin{equation}
@@
 \end{equation}
-where the tilde are dropped.
+\words{124} where the tilde are dropped.
 These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$.
@@
 \end{equation}
-where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+\words{43} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
 The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
 
-A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
+\words{167} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
 solution of ~\eqref{Orr-somerfield}.
@@
 
-Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b.
+\words{86} Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b.
 The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
@@
 \begin{table*}
+% \footnotesize
 \rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
-\renewcommand{\arraystretch}{1.4}
+\renewcommand{\arraystretch}{1.2}
  \begin{tabular}{l|c|c|c}
@@
  Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& ~~yes, near $y=\hg$~~			& ~~no, spans water column~~
+ Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no, spans water column
  \end{tabular}
- \caption{Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
+ \caption{\words{17} Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
  \label{tab:comparison}
 \end{table*}
-To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
+\words{167} To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
@@
 
-The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
+\words{144} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
 Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
@@
 \end{align} 
-in a region of thickness O($\delta$) near $y=\hg$.
+\words{105} in a region of thickness O($\delta$) near $y=\hg$.
 Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. 
@@
 
-On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of ~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+\words{64} On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of ~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
@@
 \includegraphics[]{Asymptotic_noshear}
-\caption{Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 
+% \includegraphics[width=12cm]{Asymptotic_noshear}
+\caption{
+\words{158} Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 
 Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
@@
 \end{figure}
-Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape converges in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}. 
+\words{86} Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape converges in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}.
 We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the top of the vegetation plays no role.
@@
 
-Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+\words{142} Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
 Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
@@
 
-Similarly observed large amplitude coherent oscillation of terrestrial canopies in atmospheric flow is known as \textit{honami} \cite{Inoue56,Raupach96}.
+\words{144} Similarly observed large amplitude coherent oscillation of terrestrial canopies in atmospheric flow is known as \textit{honami} \cite{Inoue56,Raupach96}.
 A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
@@
 
-We now test the assumption of a undeformable grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
+\words{174} We now test the assumption of a undeformable grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
 The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
@@
 
-The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+\words{106} The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
 In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
@@
 
-In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
+\words{146} In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
 The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 

### 2014-11-05 Paper/grass.tex
Number of words reduced by 745!

@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+\documentclass[aps,prl,reprint,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
 % \documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
@@
 % \newcommand{\words}[1]{{}}
+
+\newcommand{\ReyNdg}{{\Rey\Ndg}}
+
 \definecolor{tableShade}{gray}{0.8}
@@
 \title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
-\author{Ravi Singh}t
+\author{Ravi Singh}
 \affiliation{Brown University, Providence RI 02912 USA}
@@
 \maketitle
-\words{90} Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with both steady and oscillatory flows.  
-The hydrodynamic processes resulting from this behavior influence a number of environmental processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+\words{80} Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
+The hydrodynamic processes resulting from this behavior influence many environmental processes such as sediment, contaminant, and dissolved oxygen transport, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
 One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
@@
 
-\words{80} Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}.
+\words{77} Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}.
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
@@
 
-\words{137} However, several aspects of the shear layer model remain unsatisfactory: 
+\words{127} However, several aspects of this model remain unsatisfactory: 
 (i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
 (ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
-(iii) No existing theory has explained the threshold flow speed for instability, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
-(iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the definition of the free shear layer.
+(iii) No existing theory has explained the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
+(iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated water depth, and therefore inconsistent with the definition of the free shear layer.
 Here we present a mathematical model for the linear instability that accounts correctly for these effects, while also explaining lab experiments and field observations.
 
-\words{95} Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
+\words{86} Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
-Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically, and verify the assumption \textit{a posteriori}.
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
 We show that the drag of the grass results in an instability mode different from the Kelvin-Helmholtz instability.
@@
 \caption{
-\words{129} Schematic setup and comparison of our steady flow profile with that from the experiments by Ghisalberti and Nepf \cite{Nepf04} (Case-I from Table-1 with 1250 plants/m$^2$, 
-plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
- and its approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The grass extends up to $y=h_g$ in the water column of depth $2H$. 
-The steady velocity profile can be decomposed into three regions, (i) A parabolic profile in the unvegetated region, (ii) a uniform profile deep within the vegetation, and (iii) a boundary layer of thickness $\delta$ near the grass top.
-The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\Rey \Ndg$ is shown in the inset.
+\words{111} Schematic setup and comparison of our steady flow profile with that from the experiments in ref.
+\cite{Nepf04} (Case I from Table 1)%  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
+ and its approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The grass extends up to $y=\hg$ in the water column of depth $2H$. 
+The steady velocity profile can be decomposed into a parabolic profile in the unvegetated region, a uniform profile deep within the vegetation, and a boundary layer of thickness $\delta$ near the grass top.
+The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\ReyNdg$ is shown in the inset.
 }
@@
 \caption{
-\words{110} Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters estimated from experiments reported by Ghisalberti and Nepf \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
-Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p$ from the solution of ~\eqref{Orr-somerfield}. 
-The experimental data in the inset is obtained from publications by Ghisalberti and Nepf \cite{Ghisal02} and Vivoni \cite{Vivoni98}. 
+\words{105} Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
+Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}. 
+The experimental data in the inset is obtained from refs. \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
 In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
@@
 
-\words{28} The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
+\words{27} The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
 This drag enters the fluid mass and momentum balance equations as 
 \begin{equation}
-\quad \grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
+\grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
 \label{navier-stokes}
 \end{equation}
-\words{158} where $\rho$ is the fluid density, $\mathbf{u}$ the velocity, 
-$p$ the pressure, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
+\words{138} where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
 The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
@@
 \caption{
-\words{69} $Re(\sigma)$ as function of wavenumber and Reynolds number for a specified vegetation number density $\Ndg$ along with the neutral curve ($Re(\sigma)$=0), for parameters shown in the corresponding panel.
-As $\Ndg \propto N_g$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
-For $\Ndg$ below a critical value, the Mode 1 sets the threshold $\Rey$ whereas above the critical $\Ndg$ onset is determined by Mode 2.
+\words{46} $Re(\sigma)$ and the neutral curve ($Re(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
+As $\Ndg$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
+For $\Ndg$ below (above) a critical value, Mode 1 (Mode 2) sets the threshold $\Rey$.
 }
@@
 \end{figure*}
-\words{32} We first calculate the fully developed steady flow $\bu = U(y)\boldsymbol{\hat{x}}$ driven by constant pressure gradient $dP/dx$ as a solution of ~\eqref{navier-stokes}. 
-Such a flow satisfies 
+\words{19} We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
 \begin{equation}
@@
 \end{equation}
-\words{377} where $S(y)=1$ in the grass bed ($0<y<\hg$) and $S(y)=0$ above it ($\hg< y< 2H$). 
+\words{298} where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises because, in the case of dense vegetation, the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter is due to the free interface.
-This expectation is verified by a comparison, shown in Fig.~\ref{basicflow}, of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements.
+The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter models the free interface.
+A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
 The profile $U(y)$ has three distinct regions.
-Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$, arising from the balance between drag and pressure gradient. 
-Outside the vegetation, the velocity has a simple parabolic profile due to the balance between viscous forces and the pressure gradient. 
+Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$, arising from balancing the drag with pressure gradient. 
+Outside the vegetation, the velocity has a simple parabolic profile. % due to the balance between viscous forces and the pressure gradient. 
 At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Because this boundary layer develops from purely local dynamics, independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
-Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance of the viscous forces and the vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$,
-Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\Rey\Ndg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
-This dependence of the boundary layer thickness on the vegetation density $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate its effect on the instability mechanism.
-The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\Rey \Ndg \gtrsim 100$. In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$.
+Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
+Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+Because this boundary layer develops independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
+This dependence of $\delta$ on $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
+The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$.
 
-\words{28} Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
+\words{27} Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
 \begin{equation}
@@
 \end{equation}
-\words{124} where the tilde are dropped.
-These equations can be non-dimensionalized using half channel height $H$, velocity $U_0$, and the associated advection time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the submergence ratio of the vegetation $h_g/H$.
-We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the mechanism of the instability. 
-With these scalings, and using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a wave solution of 
+\words{89} where the tilde are dropped.
+These equations are non-dimensionalized using half channel height $H$, velocity $U_0$, and time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the vegetation submergence ratio $\hg/H$.
+We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the instability mechanism. 
+Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
 the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfield equation \cite{Drazin81} 
@@
 \end{equation}
-\words{43} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+\words{39} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
 The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
 
-\words{167} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the 
-solution of ~\eqref{Orr-somerfield}.
+\words{159} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
 The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
-The threshold Reynolds number increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented \cite{White07} for flow around an emergent (i.e., $\hg>2H$) sea grass patch, but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term.
-Previous calculations for terrestrial grass differ from our work as they either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile {\it ad hoc} \cite{Raupach96,Delangre06}.
-They have not reported a threshold flow condition.
+The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+A similar conclusion was presented for a problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch) leading to an analogous model, but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \cite{White07}.
+Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \cite{Raupach96,Delangre06}.
+A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
 
-\words{86} Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b.
-The observed frequencies are associated with the peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}, and compare well with the predicted frequencies. 
+\words{72} Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies (peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}) in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b, and seen to agree well.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
 \end{table*}
-\words{167} To better understand the mechanism of waving and the origin of the threshold, we consider the behavior of the instability as a function of the vegetation drag characterized by $\Ndg$.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical vegetation density discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
+\words{140} To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical $\delta$ discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
 To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in Fig.~\ref{K_Re_sigma_set3}. 
-The lowest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
-We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the unstable region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
+The smallest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
+We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
 The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
 
-\words{144} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ allows us to understand the mechanism of the instability. 
-Mode 1 asymptotically localizes to the boundary layer near the top of the vegetation, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3 \sim \Rey^2 \Ndg $, and the viscous term (or the vegetation drag term, which is found to scale identically) in the boundary layer is $\delta^{-4} \sim (\Rey \Ndg)^{4/3}$.
-The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$) leading to a simplification of ~\eqref{Orr-somerfield} as
+\words{116} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
+Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim H/\delta$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3$  (or $\Rey^2 \Ndg$), and the viscous and the vegetation drag term in the boundary layer are $\delta^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$), simplifying ~\eqref{Orr-somerfield} to
 \begin{align}
@@
 \end{align} 
-\words{105} in a region of thickness O($\delta$) near $y=\hg$.
-Since the relative magnitude of the terms on the r.h.s. to those on the l.h.s. of this equation is $\Rey/\Ndg^{1/2}$, we expect identical asymptotic behavior for fixed $\Rey/\Ndg^{1/2}$. 
-Therefore the threshold obtained for Mode 1 is $\Rey \propto \Ndg^{1/2}$ (or $\Rey \propto H/\delta$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
-This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this expectation is shown in Fig.~\ref{Asymptotic_mode}, supporting this argument.
+\words{93} in a region of thickness O($\delta$) near $y=\hg$.
+We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s.
+to those on the l.h.s.
+of \eqref{eqn:mode1asymp}, which is $(\Rey H/\delta)^{3/2}$ (or $\Rey/\Ndg^{1/2}$).
+Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
+This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).
 
-\words{64} On the other hand, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by an asymptotic scaling analysis of ~\eqref{Orr-somerfield} by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+\words{45} Similarly, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
@@
 \caption{
-\words{158} Plot of mode shape $|\phi|$ for different shear layer thickness in limit of small shear layer thickness for representative values of vegetation density. 
-Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
-The mode shapes approach each other for these small values of $\delta/H$ indicating that the large vegetation density asymptote is reached.
-Mode 1 shapes appear self-similar in shape as $\delta\to 0$,
-and are compared to each other on a rescaled coordinate in the inset. 
-Inset shows $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$. The modes approach a universal shape, indicating that an asymptotic limit has been reached. 
-The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.}
+\words{75} Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
+% Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters for Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
+The approach of mode shapes to each other for these small values of $\delta/H$ indicates that the dense vegetation asymptote is reached. 
+Mode 1 shapes appear self-similar in shape as $\delta\to 0$.
+Inset shows rescaled $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$ approach a universal shape, indicating that an asymptotic limit has been reached. 
+% The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.
+}
 \label{Asymptotic_mode}
 \end{figure}
-\words{86} Since $\Rey/\Ndg$ is the only remaining parameter, the mode shape converges in the aforementioned limit, in agreement with our numerical results for Mode 2 shown in Fig.~\ref{Asymptotic_mode}.
-We interpret Mode 2 as the instability of an inviscid flow, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the top of the vegetation plays no role.
-The only remaining parameter $\Rey/\Ndg$
-sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$).
-%On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+\words{54} Since $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
+The parameter $\Rey/\Ndg$ also sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
+% We interpret Mode 2 as the instability of an inviscid fluid, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass top plays no role. 
+% On the other hand, Mode 1 asymptotically localizes to the boundary layer near the grass tip, and exhibits a different asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim (H/\delta)$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
 
-\words{142} Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+\words{122} Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
 Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
 Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
-Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability. 
-The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold.
+The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold, and in distinguishing them from the Kelvin-Helmholtz instability.
+% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability.
 
-\words{144} Similarly observed large amplitude coherent oscillation of terrestrial canopies in atmospheric flow is known as \textit{honami} \cite{Inoue56,Raupach96}.
-A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
-Another major difference between the two is the considerable difference of stiffness; terrestrial vegetation tends to be much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}.
-Despite these differences, in the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
+\words{96} The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
+% A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
+% Another difference is that terrestrial vegetation is much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
+In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
@@
 
-\words{174} We now test the assumption of a undeformable grass bed due to the restoring force of buoyancy, using a simple criteria that the buoyancy time scale  be much shorter than the hydrodynamic time scale $H/U_0$.
-The former can be estimated as $\sqrt{\rho H/V_f \Delta \rho g}$, where $\Delta \rho$ is the vegetation-water density different, $V_f$ is the vegetation volume fraction, and $g$ is the acceleration due to gravity.
-Using representative values for a common sea grass, \textit{Zostera Marina}, $\Delta \rho /\rho \approx 0.25$, $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98} yields the buoyancy time scale to be about 2 s.
-Similarly, the hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
-We have, however, neither accounted for the pre-factors appearing in the scaling argument, nor have we considered cases when the time-scale separation is not so evident.
-Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.
+\words{110} We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
+The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
+We have neither accounted for the pre-factors appearing in the scaling argument, nor considered cases when the time-scale separation is not so evident \cite{Delangre06}.
+% Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.
 
-\words{106} The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
-The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
-Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
+% \words{106} The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+% In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
+% The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
+% Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
 
-\words{146} In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
+\words{104} In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
 The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
 While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
-The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
+% The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
 Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, flow through urban environments and therefore is expected to have a wider impact.
@@
 Mahadevan for helpful discussions.
-
 \bibliography{Grass}{}

### 2014-11-05 Paper/grass.tex
changed intro one sentance

@@
 \words{80} Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
-The hydrodynamic processes resulting from this behavior influence many environmental processes such as sediment, contaminant, and dissolved oxygen transport, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
 One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  

### 2014-11-06 Paper/grass.tex
Proofread (word count 3500) and submit

--- 2014-11-05 Paper/grass.tex
@@
 
-\words{127} However, several aspects of this model remain unsatisfactory: 
+\words{126} However, several aspects of this explanation remain unsatisfactory: 
 (i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
 (ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
-(iii) No existing theory has explained the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
+(iii) No existing theory explains the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
 (iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated water depth, and therefore inconsistent with the definition of the free shear layer.
@@
 \words{111} Schematic setup and comparison of our steady flow profile with that from the experiments in ref.
-\cite{Nepf04} (Case I from Table 1)%  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
+\cite{Nepf04} (Case I from Table 1) %  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
  and its approximation with $U_0=7.28$ cm/s and $\delta = 5.02$ cm in our model. The grass extends up to $y=\hg$ in the water column of depth $2H$. 
@@
 \end{equation}
-\words{298} where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
+\words{318} where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
@@
 Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
 Because this boundary layer develops independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
 This dependence of $\delta$ on $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
-The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$.
+The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$.
+In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}).
 
@@
 Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \hat\phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfield equation \cite{Drazin81} 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfield equation \cite{Drazin81} 
 \begin{equation}
@@
 \end{equation}
-\words{39} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+\words{47} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
 The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
+We solve \eqref{Orr-somerfield} numerically for $\sigma$ and $\phi$.
 
-\words{159} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
+\words{155} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
 The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
 The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented for a problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch) leading to an analogous model, but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \cite{White07}.
+A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \cite{White07}.
 Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \cite{Raupach96,Delangre06}.
@@
 
-\words{72} Comparison of the frequency (Im$(\sigma)$) of the fastest growing mode with experimentally observed frequencies (peaks in velocity spectra, frequency of \monami ~and frequency of vortex passage \cite{Ghisal02}) in lab scale experiments, for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, is also shown in the inset in Fig.~\ref{Re_vs_delta}b, and seen to agree well.
+\words{65} The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \cite{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
 
-\words{116} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
-Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim O(H/\delta)$, and $\Rey \sim H/\delta$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey/\delta^3$  (or $\Rey^2 \Ndg$), and the viscous and the vegetation drag term in the boundary layer are $\delta^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+\words{117} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
+Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey \sim H/\delta$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
+This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^3$  (or $\Rey^2 \Ndg$), and the viscous and the vegetation drag term in the boundary layer are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
 The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$), simplifying ~\eqref{Orr-somerfield} to
@@
 to those on the l.h.s.
-of \eqref{eqn:mode1asymp}, which is $(\Rey H/\delta)^{3/2}$ (or $\Rey/\Ndg^{1/2}$).
+of \eqref{eqn:mode1asymp}, which is $(\Rey \delta/H)^{3/2}$ (or $\Rey/\Ndg^{1/2}$).
 Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
@@
 
-\words{96} The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
+\words{92} The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
 % A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
 % Another difference is that terrestrial vegetation is much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
-In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case with stiff grass blades.
+In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case.
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
@@
 
-\words{110} We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+\words{117} We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
 For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
 The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
-We have neither accounted for the pre-factors appearing in the scaling argument, nor considered cases when the time-scale separation is not so evident \cite{Delangre06}.
+We have neither accounted 
+Accounting for the pre-factors appearing in the scaling argument, or considering cases when the time-scale separation is not so evident can lead to further interesting behavior \cite{Delangre06}.
 % Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.

### 2015-03-03 Paper_JFM/grass.tex
Added A folder for JFM Paper

--- 2015-02-20 Paper/grass.tex
@@
-\documentclass[aps,prl,reprint,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
+%\documentclass[aps,prl,reprint,twocolumn,superscriptaddress,sort&compress,10pt]{revtex4-1}  % for review and submission
 % \documentclass[aps,preprint,12pt]{revtex4-1}  % for review and submission
@@
 %\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
+\documentclass{jfm}
 \usepackage{amsmath}
@@
 
-\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
+\graphicspath{ {../Paper/images/}{../Paper/images/Growth_Rate_Cont_Plots/} }
 \newcommand{\bx}{{\boldsymbol{\hat{x}}}}
@@
 \newcommand{\hg}{h_g}
-\newcommand{\Rey}{{R}}
+%\newcommand{\Rey}{{R}}
 \newcommand{\Ndg}{\tilde{N}_g}
@@
 \definecolor{tableShade}{gray}{0.8}
-
-\begin{document}
 \title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
@@
 \affiliation{Brown University, Providence RI 02912 USA}
+
+\begin{document}
+\maketitle
 
@@
 \begin{center}
-\includegraphics[]{Critical_Re_vs_delta_noshear} \\
-\vspace{-6mm} \hspace{-3mm}
+\includegraphics[]{new_graph_R_vs_delta} \\
+\vspace{-5mm} \hspace{-3mm}
 \includegraphics[]{K_vs_shear_width_noshear}
@@
  Dense grass limit &  & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
- Critical parameters	& none						& $\Rey \propto \Ndg^{1/2}$ 	& $\Rey \propto \Ndg$ \\
+ Critical parameters	& none						& $\Rey \propto \Ndg^{2}$ 	& $\Rey \propto \Ndg$ \\
  Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
@@
 \words{117} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
-Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey \sim H/\delta$ (or $\Rey \propto {\Ndg}^{1/2}$) at the threshold.
-This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim H/\delta$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^3$  (or $\Rey^2 \Ndg$), and the viscous and the vegetation drag term in the boundary layer are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
-The terms balance when $\Rey \sim H/\delta$ (or $\Rey \sim {\Ndg}^{1/2}$), simplifying ~\eqref{Orr-somerfield} to
-\begin{align}
-(D^2-k^2)^2 \phi- 2 \Rey \Ndg D( SUD\phi) = \sigma \Rey (D^2-k^2)\phi,
+Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey (\sim H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$) at the threshold.
+This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim O(1)$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous and the vegetation drag term in the boundary layer are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+The terms balance when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$). Using the boundary layer co-ordinate $\eta = y/\delta$, velocity scaling 
+$U(y) = (\delta/H)\bar{U}(\eta)$ and scaling of wavenumber $k = (H/\delta) \bar{k}$ in the boundary layer ~\eqref{Orr-somerfield} simplifies to
+\begin{equation}
+\begin{split}
+\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi \\
+&+\bar{D}\left(2S \bar{U} D \phi\right),
 \label{eqn:mode1asymp}
-\end{align} 
+\end{split}
+\end{equation}
+
 \words{93} in a region of thickness O($\delta$) near $y=\hg$.
-We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s.
+Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and soultion will converge in the limit of $R \sim \Ndg^2$, confirming our numerial finding.
+%We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s.
 to those on the l.h.s.
 of \eqref{eqn:mode1asymp}, which is $(\Rey \delta/H)^{3/2}$ (or $\Rey/\Ndg^{1/2}$).
-Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
-This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{1/2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).
+%Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
+This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).
 
@@
 \end{figure}
-\words{54} Since $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
+\words{54} Similarly here $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
 The parameter $\Rey/\Ndg$ also sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
@@
 Mahadevan for helpful discussions.
-\bibliography{Grass}{}
+
+% Note the spaces between the initials
+\bibliography{Grass.bib}{}
+\bibliographystyle{jfm}
+%\bibliography{Grass}{}
 % \bibliographystyle{plain}
-\bibliographystyle{aipauth4-1}
+%\bibliographystyle{aipauth4-1}
 \end{document}

### 2015-03-03 Paper_JFM/jfm-instructions.tex
Added A folder for JFM Paper

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%       release v3.0, 16th July 2014
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997, 2014 Cambridge University Press

\documentclass{jfm}
\usepackage{graphicx}
\usepackage{epstopdf, epsfig}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\shorttitle{Guidelines for authors}
\shortauthor{A. N. Other, H.-C. Smith and J. Q. Public}

\title{Guidelines for authors and submission template}

\author{Alan N. Other\aff{1}
  \corresp{\email{jfm@damtp.cam.ac.uk}},
  H. - C. Smith\aff{1}
 \and J. Q.  Public\aff{2}}

\affiliation{\aff{1}Department of Chemical Engineering, University of America,
Somewhere, IN 12345, USA
\aff{2}Department of Aerospace and Mechanical Engineering, University of
Camford, Academic Street, Camford CF3 5QL, UK}

\begin{document}

\maketitle

\begin{abstract}
This file contains instructions for authors planning to submit a paper to the {\it Journal of Fluid Mechanics}. These instructions were generated in {\LaTeX} using the JFM style, so the {\LaTeX} source file can be used as a template for submissions. The present paragraph appears in the \verb}abstract} environment. All papers should feature a single-paragraph abstract of no more than 250 words, which provides a summary of the main aims and results. 
\end{abstract}

\begin{keywords}
Authors should not enter keywords on the manuscript, as these must be chosen by the author during the online submission process and will then be added during the typesetting process (see http://journals.cambridge.org/data/\linebreak[3]relatedlink/jfm-\linebreak[3]keywords.pdf for the full list)
\end{keywords}

\section{How to submit to the \textbf{\textit{Journal of Fluid Mechanics}}}
Authors must submit using the online submission and peer review system, Scholar One Manuscripts (formerly Manuscript Central) at http://mc.manuscriptcentral.com/jfm. If visiting the site for the first time, users must create a new account by clicking on `register here'. Once logged in, authors should click on the `Corresponding Author Centre', from which point a new manuscript can be submitted, with step-by-step instructions provided. Authors must at this stage specify whether the submission is a standard paper, or a {\it JFM Rapids} paper (see \S\ref{sec:filetypes} for more details). In addition, authors must specify an editor to whom the paper will be submitted from the drop-down list provided. Note that all editors exclusively deal with either standard papers or {\it JFM Rapids} (clearly indicated on the list), so please ensure that you choose an editor accordingly. Once your submission is completed you will receive an email confirmation.  Book reviews should not be submitted via the online submission site and should instead be submitted by email to c.p.caulfield@damtp.cam.ac.uk.

\section{Rules of submission}\label{sec:rules_submission}
Submission of a paper implies a declaration by the author that the work has not previously been published, that it is not being considered for publication elsewhere and that it has not already been considered by a different editor of the Journal. Note that a report on a conference must be submitted within 3 months of the meeting.

\section{Types of paper}\label{sec:types_paper}
\subsection{Standard papers}
Regular submissions to JFM are termed `standard papers'. Note that such papers must contain original research, as review articles are not published in JFM. Papers should be written in a concise manner; though JFM has no page limit, each paper will be judged on its own merits, and those deemed excessive in length will be rejected or will require significant revision.  
\subsection{JFM Rapids}
{\it JFM Rapids} is devoted to the rapid publication of short, high-impact papers across the full range of fluid mechanics. Manuscripts submitted as {\it Rapids}  must be strictly 10 or fewer printed pages, and must be submitted in {\LaTeX} using the jfm.cls class file, so as to ensure that they meet the page limit and to expedite their production.  The principal and over-riding objective is to publish papers of the highest scientific quality.

Once a paper is submitted, referees are asked to provide reports with a short turnaround.  In order to be accepted for publication in {\it JFM Rapids}, such papers must require only minor revisions to improve clarity, usually without recourse to a second round of refereeing. In this case, and at the discretion of the editor, some additional pages may be allowed to address specific points raised by the referees, such as the addition of an extra figure or some explanatory text.

In cases where the editor, guided by the referees, judges that a paper has merit but requires substantial revision that will require significant refereeing, a decision of `revise and resubmit' will be given. On re-submission, such papers will be handled as standard JFM papers and if accepted will not subsequently appear as {\it JFM Rapids}.

{\it Rapids} will be published online within one month of final acceptance.  They will appear within a designated section on the Journal of Fluid Mechanics website.  Each {\it Rapid} will be cited and indexed as a JFM article but with a distinctive {\it Rapids} identifier, and will be assigned to a JFM volume. {\it Rapids} will not be included in the regular print versions of JFM, and accordingly they incur no colour figure charges.

\section{File types}\label{sec:filetypes}
Authors are strongly encouraged to compose their papers in {\LaTeX}, using the jfm.cls style file and supporting files provided at http://journals.cambridge.org/images/\linebreak[4]fileUpload/documents/jfm-ifc.zip, with the jfm-instructions.tex file serving as a template (note that this is mandatory for {\it JFM Rapids}). A PDF of the {\LaTeX} file should then be generated and submitted via the submission site. Please note that PDFs larger than 3MB are not acceptable for review. There is no need to submit the {\LaTeX} source file alongside the PDF, but upon provisional acceptance of the paper, the {\LaTeX} source file, along with individual figure files and a PDF of the final version, will need to be submitted for typesetting purposes. 
Authors may also compose standard papers in Word, though this will lead to the paper spending a longer period in production. If using Word, please note that equations must NOT be converted to picture format and the file must be saved with the option `make equation editable'. 
\section{Preparing your manuscript}
Authors should write their papers clearly and concisely in English, adhering to JFM's established style for notation, as provided in \S\ref{notstyle}. We encourage the submission of online supplementary material alongside the manuscript where appropriate (see section \ref{online}). Metric units should be used throughout and all abbreviations must be defined at first use, even those deemed to be well known to the readership. British spelling must be used, and should follow the \textit{Shorter Oxford English Dictionary}.

\subsection{Figures}
Figures should be as small as possible while displaying clearly all the information required, and with all lettering readable. Every effort should be taken to avoid figures that run over more than one page. Figures submitted in colour will appear online in colour but, with the exception of {\it JFM Rapids}, all figures will be printed in black and white unless authors specify during submission that figures should be printed in colour, for which there is a charge of \pounds200 per figure with a cap of \pounds1000 per article plus taxes where applicable (colour is free for {\it JFM Rapids}).  Note that separate figures for online and print will {\bf not} be accepted and it is the author's responsibility to ensure that if a figure is to appear in colour online only, that same figure will still be meaningful when printed in black and white (for example, do not rely upon colours to distinguish lines if those colours will just appear as similar shades of grey when printed). If using colour, authors should also ensure that some consistency is applied within the manuscript. For review purposes figures should be embedded within the manuscript. Upon final acceptance, however, individual figure files will be required for production. These should be submitted in EPS or high-resolution TIFF format (1200 dpi for lines, 300 dpi for halftone and colour in CMYK format, and 600 dpi for a mixture of lines and halftone). The minimum acceptable width of any line is 0.5pt. Each figure should be accompanied by a single caption, to appear beneath, and must be cited in the text. Figures should appear in the order in which they are first mentioned in the text and figure files must be named accordingly (`Figure 1.eps', `Figure 2a.tiff', etc) to assist the production process (and numbering of figures should continue through any appendices). The word \textit {figure} is only capitalized at the start of a sentence. For example see figures \ref{fig:ka} and \ref{fig:kd}. Failure to follow figure guidelines may result in a request for resupply and a subsequent delay in the production process. Note that {\em all} figures are relabelled by the typesetter, so please ensure all figure labels are carefully checked against your originals when you receive your proofs.

\begin{figure}
  \centerline{\includegraphics{trapped}}% Images in 100% size
  \caption{Trapped-mode wavenumbers, $kd$, plotted against $a/d$ for
    three ellipses:\protect\\
    ---$\!$---,
    $b/a=1$; $\cdots$\,$\cdots$, $b/a=1.5$.}
\label{fig:ka}
\end{figure}

\begin{figure}
  \centerline{\includegraphics{modes}}
  \caption{The features of the four possible modes corresponding to
  (\textit{a}) periodic\protect\\ and (\textit{b}) half-periodic solutions.}
\label{fig:kd}
\end{figure}

\subsection{Tables}
Tables, however small, must be numbered sequentially in the order in which they are mentioned in the text. The word \textit {table} is only capitalized at the start of a sentence. See table \ref{tab:kd} for an example.

\begin{table}
  \begin{center}
\def~{\hphantom{0}}
  \begin{tabular}{lccc}
      $a/d$  & $M=4$   &   $M=8$ & Callan \etal \\[3pt]
       0.1   & 1.56905 & ~~1.56~ & 1.56904\\
       0.3   & 1.50484 & ~~1.504 & 1.50484\\
       0.55  & 1.39128 & ~~1.391 & 1.39131\\
       0.7   & 1.32281 & ~10.322 & 1.32288\\
       0.913 & 1.34479 & 100.351 & 1.35185\\
  \end{tabular}
  \caption{Values of $kd$ at which trapped modes occur when $\rho(\theta)=a$}
  \label{tab:kd}
  \end{center}
\end{table}

\subsection{Online supplementary material}\label{online}
Relevant material which is not suitable for print production, such as movies or numerical simulations/animations, can be uploaded as part of the initial submission. Movies should be designated as `Movie' and each individual file must be accompanied by a separate caption and a suitable title (eg Movie 1). Accepted formats are .mov, .mpg, .mp4, and .avi, though they should be archived as a .zip or .tar file before uploading. Each movie should be no more than 10MB. Upon publication these materials will then be hosted online alongside the final published article. Likewise, should there be detailed mathematical relations, tables or figures which are likely to be useful only to a few specialists or take up excessive space in the printed journal, these can also be published online as supplementary material [designated as `Other supplementary material']. Note that supplementary material is published `as is', with no further production performed.

\section{Editorial decisions}
\subsection{Revision}
If a revision is requested, you should upload revised files following the same procedure as for submitting a new paper. You begin by clicking on `Manuscripts with decision' in your Corresponding Author Center, and then on `Create a revision'. (Note that if you abandon the process before completing the submission, to continue the submission, you must click on `Revised manuscripts in draft'.) There is a new first page showing the decision letter and a space for your reply to the referee's/editor's comments. You also have the opportunity at this stage to upload your reply to the comments as a separate file. All the values filled in on original submission are displayed again. The ID number of the paper will be appended `.R1'. Also note that if a manuscript is submitted as a {\it JFM Rapid}, but requires substantial revision, it will be re-designated as a standard paper, and the ID and paper type will be amended to reflect this.

\subsection{Provisional acceptance}
If the paper is accepted as suitable for publication you will be sent a provisional acceptance decision. This enables you to upload the final files required for production:
(1) the final PDF or word version of the paper, designated as a `main document';
(2) any source files (see section \ref{sec:filetypes}) which must be designated as `production (not for review)' and uploaded as a single .zip or .tar file.

\subsection{Acceptance}
On receipt of the production files you will be sent an email indicating completion of the acceptance process.

\section{Publication process}
Once a paper has been accepted for publication and the source files have been uploaded, the manuscript will be sent to Cambridge University Press for copyediting and typesetting, and will be assigned a digital object identifier (doi). When the proof is ready, authors will receive an email alert containing a link to the PDF of the proof, and instructions for its correction and return. It is imperative that authors check their proofs closely, particularly the equations and figures, which should be checked against the accepted file, as the production schedule does not allow for corrections at a later stage. Once ready for printing, papers will be published online on Cambridge Journals Online in the current `open' volume. {\it JFM Rapids} will also be assigned to the current open volume, but will receive an article number in place of traditional pagination, as they will not appear in the print version of that volume. Each volume will be kept open for approximately two weeks, at which point it will be considered `closed', and then printed and distributed. The following volume will be immediately opened. Note that the PDF published online is the Version of Record, matching the print version, and no further alterations/corrections to this document will be allowed. The corresponding author is emailed a link to the published article when it is first published online.

\section{Obtaining help}
Technical support for the online submission system is available by clicking on the `Get Help Now' link at the top-right corner of each page of the submission site. Any other questions relating to the submission or publication process should be directed to the JFM Editorial Assistant, Mrs Amanda Johns, at ajohns@cambridge.org.

\section{Notation and style}\label{notstyle}
Generally any queries concerning notation and journal style can be answered by viewing recent pages in the Journal. However, the following guide provides the key points to note. It is expected that Journal style will be followed, and authors should take care to define all variables or entities upon first use. Also note that footnotes are not normally accepted.

\subsection{Mathematical notation}
\subsubsection{Setting variables, functions, vectors, matrices etc}
{\bf Italic font} should be used for denoting variables, with multiple-letter symbols avoided except in the case of dimensionless numbers such as $\Rey$, $\Pran$ and $\Pen$ (Reynolds, Prandtl, and P\'eclet numbers respectively, which are defined as \verb}\Rey}, \verb}\Pran} and {\verb}\Pen} in the template).\\
{\bf Upright Roman font} (or upright Greek where appropriate) should be used for:\\
Operators: sin, log, d, $\Delta$, e etc.\\
Constants: i ($\sqrt{-1}$), $\upi$ (defined as \verb}\upi}), etc.\\
Functions: $\Ai$, $\Bi$ (Airy functions, defined as \verb}\Ai} and \verb}\Bi}), $\Real$ (real part, defined as \verb}\Real}), $\Imag$ (imaginary part, defined as \verb}\Imag}), etc.\\
Physical units: cm, s, etc\\
Abbreviations: c.c. (complex conjugate), h.o.t. (higher-order terms), DNS, etc.\\
{\bf Bold italic font} (or bold sloping Greek) should be used for:\\
Vectors (with the centred dot for a scalar product also in bold): $\boldsymbol{i \cdot j}$\\
{\bf Bold sloping sans serif font}, defined by the \verb}\mathsfbi} macro, should be used for:\\
Tensors and matrices: $\mathsfbi{D}$ \\
{\bf Script font} (for example $\mathcal{G}$, $\mathcal{R}$) can be used as an alternative to italic when the same letter denotes a different quantity (use \verb}\mathcal} in \LaTeX)\\
The product symbol ($\times$) should only be used to denote multiplication where an equation is broken over more than one line, to denote a cross product, or between numbers (the $\cdot$ symbol should not be used, except to denote a scalar product specifically).\\

\subsubsection{Other symbols}
A centred point should be used only for the scalar product of vectors.
Large numbers that are not scientific powers should not include commas, but have the
form 1600 or 16 000 or 160 000.
Use \textit{O} to denote `of the order of', not the \LaTeX\ $\mathcal{O}$.

\section{Citations and references}
All papers included in the References section must be cited in the article, and vice versa. Citations should be included as, for example ``It has been shown \citep{Rogallo81} that...'' (using the {\verb}\citep} command, part of the natbib package) ``recent work by \citet{Dennis85}...'' (using {\verb}\citet}).
The natbib package can be used to generate citation variations, as shown below.\\
\verb#\citet[pp. 2-4]{Hwang70}#:\\
\citet[pp. 2-4]{Hwang70} \\
\verb#\citep[p. 6]{Worster92}#:\\
\citep[p. 6]{Worster92}\\
\verb#\citep[see][]{Koch83, Lee71, Linton92}#:\\
\citep[see][]{Koch83, Lee71, Linton92}\\
\verb#\citep[see][p. 18]{Martin80}#:\\
\citep[see][p. 18]{Martin80}\\
\verb#\citep{Brownell04,Brownell07,Ursell50,Wijngaarden68,Miller91}#:\\
\citep{Brownell04,Brownell07,Ursell50,Wijngaarden68,Miller91}\\
The References section can either be built from individual \verb#\bibitem# commands, or can be built using BibTex. The BibTex files used to generate the references in this document can be found in the zip file at http://journals.cambridge.org/\linebreak[3]data/\linebreak[3]relatedlink/\linebreak[3]jfm-ifc.zip.\\

Acknowledgements should be included at the end of the paper, before the References section or any appendicies, and should be a separate paragraph without a heading. Several anonymous individuals are thanked for contributions to these instructions.

\appendix
\section{}\label{appA}
This appendix contains sample equations in the JFM style. Please refer to the {\LaTeX} source file for examples of how to display such equations in your manuscript.

\begin{equation}
  (\nabla^2+k^2)G_s=(\nabla^2+k^2)G_a=0
  \label{Helm}
\end{equation}

\begin{equation}
  \bnabla\bcdot\boldsymbol{v} = 0,\quad \nabla^{2}P=
    \bnabla\bcdot(\boldsymbol{v}\times \boldsymbol{w}).
\end{equation}

\begin{equation}
  G_s,G_a\sim 1 / (2\upi)\ln r
  \quad \mbox{as\ }\quad r\equiv|P-Q|\rightarrow 0,
  \label{singular}
\end{equation}

\begin{equation}
\left. \begin{array}{ll}  
\displaystyle\frac{\p G_s}{\p y}=0
  \quad \mbox{on\ }\quad y=0,\\[8pt]
\displaystyle  G_a=0
  \quad \mbox{on\ }\quad y=0,
 \end{array}\right\}
  \label{symbc}
\end{equation}

\begin{equation}
  -\frac{1}{2\upi} \int_0^{\infty} \gamma^{-1}[\mathrm exp(-k\gamma|y-\eta|)
   + \mathrm exp(-k\gamma(2d-y-\eta))] \cos k(x-\xi)t\:\mathrm{d} t,
   \qquad 0<y,\quad \eta<d,
\end{equation}

\begin{equation}
  \gamma(t) = \left\{
    \begin{array}{ll}
      -\mathrm{i}(1-t^2)^{1/2}, & t\le 1 \\[2pt]
      (t^2-1)^{1/2},         & t>1.
    \end{array} \right.
\end{equation}

\[
  -\frac{1}{2\upi}
   \pvi B(t)\frac{\cosh k\gamma(d-y)}{\gamma\sinh k\gamma d}
   \cos k(x-\xi)t\:\mathrm{d} t
\]

\begin{equation}
  G = -\frac{1}{4}\mathrm{i} (H_0(kr)+H_0(kr_1))
    - \frac{1}{\upi} \pvi\frac{\mathrm{e}^{-\kgd}}%
    {\gamma\sinh\kgd} \cosh k\gamma(d-y) \cosh k\gamma(d-\eta)
\end{equation}

Note that when equations are included in definitions, it may be suitable to render them in line, rather than in the equation environment: $\boldsymbol{n}_q=(-y^{\prime}(\theta),
x^{\prime}(\theta))/w(\theta)$.
Now $G_a=\squart Y_0(kr)+\Gat$ where
$r=\{[x(\theta)-x(\psi)]^2 + [y(\theta)-y(\psi)]^2\}^{1/2}$ and $\Gat$ is
regular as $kr\ttz$. However, any fractions displayed like this, other than $\thalf$ or $\squart$, must be written on the line, and not stacked (ie 1/3).

\begin{eqnarray}
  \ndq\left(\frac{1}{4} Y_0(kr)\right) & \sim &
    \frac{1}{4\upi w^3(\theta)}
    [x^{\prime\prime}(\theta)y^{\prime}(\theta)-
    y^{\prime\prime}(\theta)x^{\prime}(\theta)] \nonumber\\
  & = & \frac{1}{4\upi w^3(\theta)}
    [\rho^{\prime}(\theta)\rho^{\prime\prime}(\theta)
    - \rho^2(\theta)-2\rho^{\prime 2}(\theta)]
    \quad \mbox{as\ }\quad kr\ttz . \label{inteqpt}
\end{eqnarray}

\begin{equation}
  \frac{1}{2}\phi_i = \frac{\upi}{M} \sumjm\phi_j K_{ij}^a w_j,
  \qquad i=1,\,\ldots,\,M,
\end{equation}
where
\begin{equation}
  K_{ij}^a = \left\{
    \begin{array}{ll}
      \p G_a(\theta_i,\theta_j)/\p n_q, & i\neq j \\[2pt]
      \p\Gat(\theta_i,\theta_i)/\p n_q
      + [\rho_i^{\prime}\rho_i^{\prime\prime}-\rho_i^2-2\rho_i^{\prime 2}]
      / 4\upi w_i^3, & i=j.
  \end{array} \right.
\end{equation}

\refstepcounter{equation}
$$
  \rho_l = \lim_{\zeta \rightarrow Z^-_l(x)} \rho(x,\zeta), \quad
  \rho_{u} = \lim_{\zeta \rightarrow Z^{+}_u(x)} \rho(x,\zeta)
  \eqno{(\theequation{\mathit{a},\mathit{b}})}\label{eq35}
$$

\begin{equation}
  (\rho(x,\zeta),\phi_{\zeta\zeta}(x,\zeta))=(\rho_0,N_0)
  \quad \mbox{for}\quad Z_l(x) < \zeta < Z_u(x).
\end{equation}

\begin{subeqnarray}
  \tau_{ij} & = &
    (\overline{\overline{u}_i \overline{u}_j}
    - \overline{u}_i\overline{u}_j)
    + (\overline{\overline{u}_iu^{SGS}_j
    + u^{SGS}_i\overline{u}_j})
    + \overline{u^{SGS}_iu^{SGS}_j},\\[3pt]
  \tau^\theta_j & = &
    (\overline{\overline{u}_j\overline{\theta}}
    - \overline{u}_j \overline{\theta})
    + (\overline{\overline{u}_j\theta^{SGS}
    + u^{SGS}_j \overline{\theta}})
    + \overline{u^{SGS}_j\theta^{SGS}}.
\end{subeqnarray}

\begin{equation}
\setlength{\arraycolsep}{0pt}
\renewcommand{\arraystretch}{1.3}
\slsQ_C = \left[
\begin{array}{ccccc}
  -\omega^{-2}V'_w  &  -(\alpha^t\omega)^{-1}  &  0  &  0  &  0  \\
  \displaystyle
  \frac{\beta}{\alpha\omega^2}V'_w  &  0  &  0  &  0  &  \mathrm{i}\omega^{-1} \\
  \mathrm{i}\omega^{-1}  &  0  &  0  &  0  &  0  \\
  \displaystyle
  \mathrm{i} R^{-1}_{\delta}(\alpha^t+\omega^{-1}V''_w)  &  0
    & -(\mathrm{i}\alpha^tR_\delta)^{-1}  &  0  &  0  \\
  \displaystyle
  \frac{\mathrm{i}\beta}{\alpha\omega}R^{-1}_\delta V''_w  &  0  &  0
    &  0  & 0 \\
  (\mathrm{i}\alpha^t)^{-1}V'_w  &  (3R^{-1}_{\delta}+c^t(\mathrm{i}\alpha^t)^{-1})
    &  0  &  -(\alpha^t)^{-2}R^{-1}_{\delta}  &  0  \\
\end{array}  \right] .
\label{defQc}
\end{equation}

\begin{equation}
\etb^t = \skew2\hat{\etb}^t \exp [\mathrm{i} (\alpha^tx^t_1-\omega t)],
\end{equation}
where $\skew2\hat{\etb}^t=\boldsymbol{b}\exp (\mathrm{i}\gamma x^t_3)$. 
\begin{equation}
\mbox{Det}[\rho\omega^2\delta_{ps}-C^t_{pqrs}k^t_qk^t_r]=0,
\end{equation}

\begin{equation}
 \langle k^t_1,k^t_2,k^t_3\rangle = \langle
\alpha^t,0,\gamma\rangle  
\end{equation}

\begin{equation}
\boldsymbol{f}(\theta,\psi) = (g(\psi)\cos \theta,g(\psi) \sin \theta,f(\psi)).
\label{eq41}
\end{equation}

\begin{eqnarray}
f(\psi_1) = \frac{3b}{\upi[2(a+b \cos \psi_1)]^{{3}/{2}}}
  \int^{2\upi}_0 \frac{(\sin \psi_1 - \sin \psi)(a+b \cos \psi)^{1/2}}%
  {[1 - \cos (\psi_1 - \psi)](2+\alpha)^{1/2}}\mathrm{d}x,
\label{eq42}
\end{eqnarray}
\begin{eqnarray}
g(\psi_1) & = & \frac{3}{\upi[2(a+b \cos \psi_1)]^{{3}/{2}}}
  \int^{2\upi}_0 \left(\frac{a+b \cos \psi}{2+\alpha}\right)^{1/2}
  \left\{ \astrut f(\psi)[(\cos \psi_1 - b \beta_1)S + \beta_1P]
  \right. \nonumber\\
&& \mbox{}\times \frac{\sin \psi_1 - \sin \psi}{1-\cos(\psi_1 - \psi)}
  + g(\psi) \left[\left(2+\alpha - \frac{(\sin \psi_1 - \sin \psi)^2}
  {1- \cos (\psi - \psi_1)} - b^2 \gamma \right) S \right.\nonumber\\
&& \left.\left.\mbox{} + \left( b^2 \cos \psi_1\gamma -
  \frac{a}{b}\alpha \right) F(\frac{1}{2}\upi, \delta) - (2+\alpha)
  \cos\psi_1 E(\frac{1}{2}\upi, \delta)\right] \astrut\right\} \mathrm{d} \psi,
\label{eq43}
\end{eqnarray}
\begin{equation}
\alpha = \alpha(\psi,\psi_1) = \frac{b^2[1-\cos(\psi-\psi_1)]}%
  {(a+b\cos\psi) (a+b\cos\psi_1)},
  \quad
  \beta - \beta(\psi,\psi_1) = \frac{1-\cos(\psi-\psi_1)}{a+b\cos\psi}.
\end{equation}

\begin{equation}
\left. \begin{array}{l}
\displaystyle
H(0) = \frac{\epsilon \overline{C}_v}{\tilde{v}^{{1}/{2}}_T
(1- \beta)},\quad H'(0) = -1+\epsilon^{{2}/{3}} \overline{C}_u
+ \epsilon \skew5\hat{C}_u'; \\[16pt]
\displaystyle
H''(0) = \frac{\epsilon u^2_{\ast}}{\tilde{v}^{{1}/{2}}
_T u^2_P},\quad H' (\infty) = 0.
\end{array} \right\}
\end{equation}

\begin{lemma}
Let $f(z)$ be a trial \citet[][pp.~231--232]{Batchelor59} function defined on $[0,1]$.  Let $\varLambda_1$ denote
the ground-state eigenvalue for $-\mathrm{d}^2g/\mathrm{d} z^2=\varLambda g$,
where $g$ must satisfy $\pm\mathrm{d} g/\mathrm{d} z+\alpha g=0$ at $z=0,1$
for some non-negative constant~$\alpha$.  Then for any $f$ that is not
identically zero we have
\begin{equation}
\frac{\displaystyle
  \alpha(f^2(0)+f^2(1)) + \int_0^1 \left(
  \frac{\mathrm{d} f}{\mathrm{d} z} \right)^2 \mathrm{d} z}%
  {\displaystyle \int_0^1 f^2\mathrm{d} z}
\ge \varLambda_1 \ge
\left( \frac{-\alpha+(\alpha^2+8\upi^2\alpha)^{1/2}}{4\upi} \right)^2.
\end{equation}
\end{lemma}

\begin{corollary}
Any non-zero trial function $f$ which satisfies the boundary condition
$f(0)=f(1)=0$ always satisfies
\begin{equation}
  \int_0^1 \left( \frac{\mathrm{d} f}{\mathrm{d} z} \right)^2 \mathrm{d} z.
\end{equation}
\end{corollary}

\bibliographystyle{jfm}
% Note the spaces between the initials
\bibliography{jfm-instructions}

\end{document}

### 2015-06-18 Paper_JFM/grass.tex
Add section/subsection structure for JFM

@@
 \maketitle
-\words{80} Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
+\section{Introduction}
+Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
 The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
@@
 
-\words{77} Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}.
+Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}. 
 Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
@@
 
-\words{126} However, several aspects of this explanation remain unsatisfactory: 
+However, several aspects of this explanation remain unsatisfactory: 
 (i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
@@
 
-\words{86} Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
+\section{Mathematical model}
+Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
@@
 \begin{figure}
-\includegraphics[scale=1]{Grass_Base_Nepf_shear}
+\centerline{\includegraphics[scale=1]{Grass_Base_Nepf_shear}}
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
 \caption{
-\words{111} Schematic setup and comparison of our steady flow profile with that from the experiments in ref.
+Schematic setup and comparison of our steady flow profile with that from the experiments in ref.
 \cite{Nepf04} (Case I from Table 1) %  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
@@
 \begin{center}
-\includegraphics[]{new_graph_R_vs_delta} \\
-\vspace{-5mm} \hspace{-3mm}
-\includegraphics[]{K_vs_shear_width_noshear}
+\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
 % \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
@@
 \caption{
-\words{105} Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
 Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
@@
 
-\words{27} The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
+The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
 This drag enters the fluid mass and momentum balance equations as 
@@
 \end{equation}
-\words{138} where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
+where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
 The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
@@
 \caption{
-\words{46} $Re(\sigma)$ and the neutral curve ($Re(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
+$Re(\sigma)$ and the neutral curve ($Re(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
 As $\Ndg$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
@@
 \end{figure*}
-\words{19} We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
+
+\section{Linear stability analysis}
+We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
 \begin{equation}
@@
 \end{equation}
-\words{318} where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
+where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
@@
 
-\words{27} Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
+Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
 \begin{equation}
@@
 \end{equation}
-\words{89} where the tilde are dropped.
+where the tilde are dropped.
 These equations are non-dimensionalized using half channel height $H$, velocity $U_0$, and time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the vegetation submergence ratio $\hg/H$.
@@
 \end{equation}
-\words{47} where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
 The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
@@
 
-\words{155} A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
+\section{Results - unstable modes and critical parameters}
+A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
 The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
@@
 
-\words{65} The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \cite{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
+The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \cite{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
  \end{tabular}
- \caption{\words{17} Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
+ \caption{Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
  \label{tab:comparison}
 \end{table*}
-\words{140} To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
+To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
 The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical $\delta$ discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
@@
 
-\words{117} The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
+\section{Discussion -- comparison of unstable modes with Kelvin Helmholtz}
+The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
+\subsection{Mode 1}
 Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey (\sim H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$) at the threshold.
@@
 \end{equation}
-
-\words{93} in a region of thickness O($\delta$) near $y=\hg$.
+in a region of thickness O($\delta$) near $y=\hg$.
 Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and soultion will converge in the limit of $R \sim \Ndg^2$, confirming our numerial finding.
@@
 
-\words{45} Similarly, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+\subsection{Mode 2}
+Similarly, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
@@
 \caption{
-\words{75} Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
+Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
 % Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters for Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
@@
 \end{figure}
-\words{54} Similarly here $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
+Similarly here $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
 The parameter $\Rey/\Ndg$ also sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
@@
 
-\words{122} Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+\subsection{Comparison with Kelvin-Helmholtz}
+Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
 Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
@@
 
-\words{92} The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
+The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
 % A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
@@
 
-\words{117} We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+\section{Conclusion}
+We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
 For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
@@
 
-% \words{106} The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+% The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
 % In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
@@
 
-\words{104} In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
+In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
 The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
@@
 % Note the spaces between the initials
-\bibliography{Grass.bib}{}
+\bibliography{Grass}{}
 \bibliographystyle{jfm}

### 2015-06-18 Paper_JFM/grass.tex
Add discussion on Mode 2

@@
 \left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&+D\left(2R \Ndg S U D \phi\right),
+&+\Rey \Ndg D\left(2 S U D \phi\right),
 \label{Orr-somerfield}
@@
  Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
- Linearized dynamics	& $\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
- Dense grass limit &  & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
+ Linearized dynamics	& Equation \eqref{eqn:Rayleigh}		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
+ Dense grass limit &  no grass included & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
  Critical parameters	& none						& $\Rey \propto \Ndg^{2}$ 	& $\Rey \propto \Ndg$ \\
@@
 \subsection{Mode 2}
-Similarly, the threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
+The threshold condition for Mode 2 is numerically observed to be $\Rey \propto ({\delta}/{H})^{-3/2}$ (or $\Rey \propto \Ndg$) for $k\sim O(1)$, shown in Fig.~\ref{Re_vs_delta}, which can be understood by assuming $\Rey \gg 1$ but fixed $\Rey/\Ndg \sim O(1)$.
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-In this limit ~\eqref{Orr-somerfield} simplifies to 
+In this limit, the non-dimensional flow in the grass bed is $U\sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $\sigma + ikU \approx \sigma$ to leading order. 
+Furthermore, $U_{yy}$ decays to zero outside the boundary layer. 
+Outside the grass, the turbulent viscous stress term is negligible compared to the inertial term. Thus, \eqref{Orr-somerfield} simplifies to 
+\begin{subequations}
 \begin{align}
-\begin{split}
-\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for $y<\hg$}  \\
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for $y>\hg$}.
-\end{split}
+% \begin{split}
+\sigma\left(D^2-k^2\right)\phi = -2{(\Ndg/\Rey)^{1/2}}D^2\phi,  \quad &\text{ for } y<\hg  \label{eqn:mode2asympa} \\
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, \quad &\text{ for } y>\hg. \label{eqn:mode2asympb}
+% \end{split}
+\end{align}
 \label{eqn:mode2asymp}
-\end{align}
+\end{subequations}
+
 \begin{figure}
@@
 \end{figure}
-Similarly here $\Rey/\Ndg$ is the only remaining parameter in \eqref{eqn:mode2asymp}, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
-The parameter $\Rey/\Ndg$ also sets the threshold, leading to the asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
+The only remaining parameter in \eqref{eqn:mode2asymp} is $\Rey/\Ndg$. 
+For fixed $\Rey/\Ndg$, the mode shape converges in the aforementioned limit, in agreement with our numerical results shown in Fig.~\ref{Asymptotic_mode}.
+This convergence indicates that we have identified the correct asymptotic limit to investigate Mode 2.
+
+The parameter $\Rey/\Ndg$ therefore sets the threshold, justifying the numerically observed asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
+The structure of this mode in the aforementioned limit is such that $\phi$ is continuous at $y=h_g$, but $D\phi$ undergoes a rapid transition there on the scale of boundary layer thickness $\delta$.
+The eigenvalues and the mode shape are otherwise independent of $\delta$, and therefore we conclude that the boundary layer only plays a secondary role of regularizing the discontinuity in tangential velocity arising at $y=\hg$ in this instability mechanism.
 % We interpret Mode 2 as the instability of an inviscid fluid, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass top plays no role. 
@@
 In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
+
+Mode 2 is has characteristics distinct from Kelvin-Helmholtz. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation
+\begin{align}
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi. 
+\label{eqn:Rayleigh}
+\end{align}
+An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \cite{Rayleigh1879}. 
+However, for our $U(y)$ profiles, $U_{yy}(y) = -1$, and therefore does not change sign $y>\hg$. 
+Instead, the dynamics are coupled with the flow in the grass bed described by \eqref{eqn:mode2asympa} in $y< \hg$.
+The absence of $U_{yy}$ in \eqref{eqn:mode2asympa} indicates that $U_{yy}$ is approximated to be zero in $y<\hg$, and therefore the negative values of $U_{yy}$ that occurs in the boundary layer do not affect this mode of instability to leading order.
+Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
+Based on this analysis, we conclude that Mode 2 is distinct from the Kelvin-Helmholtz instability, and bears its existence purely because of the vegetation drag.
 

### 2015-06-18 Paper_JFM/grass.tex
Mode 1 discussion

--- 2015-06-18 Paper_JFM/grass.tex
@@
 \subsection{Mode 1}
-Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey (\sim H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$) at the threshold.
-This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim O(1)$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous and the vegetation drag term in the boundary layer are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
-The terms balance when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$). Using the boundary layer co-ordinate $\eta = y/\delta$, velocity scaling 
-$U(y) = (\delta/H)\bar{U}(\eta)$ and scaling of wavenumber $k = (H/\delta) \bar{k}$ in the boundary layer ~\eqref{Orr-somerfield} simplifies to
+Our numerical calculation show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a for a perturbation of  $k \sim H/\delta$.
+We also observe that critcal Reynolds number above which perturbation become unstable scales as  $\Rey (\sim H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$).
+The behavior of this critical Reynolds number can be understood by asymptotically estimating the size of various terms in boundary layer in ~\eqref{Orr-somerfield}.
+Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$), confirming our numerical finding.
+\newline
+Inorder to further understand the mechanism of Mode-1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/\delta$, 
+$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
 \begin{equation}
@@
 \end{equation}
+
 in a region of thickness O($\delta$) near $y=\hg$.
-Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and soultion will converge in the limit of $R \sim \Ndg^2$, confirming our numerial finding.
+Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and soulution will converge in the limit of $R \sim \Ndg^2$, confirming our numerial finding which is shown in Fig \ref{Asymptotic_mode}.
+\newline
+Since by increasing $\Rey$ in the above equation for a  fixed $\Ndg$ we recover the linearized dynamics of Kelvin-Helmholtz equation, signifying that Mode-1 is similar to Kelvin-Helmholtz instability.
+We can further corroborate this claim by observing that both Kelvin-Helmholtz and Mode-1 shows that (i) Critical wavenumber $k \sim O(H/\delta$, (ii) Strong shear in the boundary layer plays crucial role in the flow instability
+(iii) Dominant mode of instability is localized near the boundary layer.
+
 %We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s.
@@
 %Therefore the threshold obtained for Mode 1 is $\Rey \propto H/\delta$ (or $\Rey \propto \Ndg^{1/2}$) explaining the numerically observed asymptote (see Fig.~\ref{Re_vs_delta}). 
-This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).
+%This analysis also concludes that the mode structure is self-similar over the length scale $\delta$ for fixed $\Rey/\Ndg^{2}$; the verification of this idea is shown in Fig.~\ref{Asymptotic_mode} (inset).
 

### 2015-06-18 Paper_JFM/grass.tex
Initiate discussion on turbulence models

--- 2015-06-18 Paper_JFM/grass.tex
@@
 The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
-In the interest of simplicity, we model the turbulence using an eddy viscosity. 
+In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
 In the field, both $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
@@
 
+\subsection{Role of turbulence model}\label{subsec:turbulencemodel}
+We have modeled turbulence using a constant eddy viscosity. 
+While such a model is too simplistic and
+
 \section{Conclusion}

### 2015-06-18 Paper_JFM/grass.tex
Changes in Mode 2 and Comparison by Ravi

--- 2015-06-18 Paper_JFM/grass.tex
@@
 In this limit, the non-dimensional flow in the grass bed is $U\sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $\sigma + ikU \approx \sigma$ to leading order. 
-Furthermore, $U_{yy}$ decays to zero outside the boundary layer. 
+Furthermore, $U_{yy}$ decays to zero outside the boundary layer and within the grass. 
 Outside the grass, the turbulent viscous stress term is negligible compared to the inertial term. Thus, \eqref{Orr-somerfield} simplifies to 
@@
 
-Mode 2 is has characteristics distinct from Kelvin-Helmholtz. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation
+Mode 2 has characteristics distinct from Kelvin-Helmholtz. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation
 \begin{align}
@@
 An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \cite{Rayleigh1879}. 
-However, for our $U(y)$ profiles, $U_{yy}(y) = -1$, and therefore does not change sign $y>\hg$. 
+However, for our $U(y)$ profiles, $U_{yy}(y) = -1$ above the grass and therefore does not change sign for $y>\hg$. 
 Instead, the dynamics are coupled with the flow in the grass bed described by \eqref{eqn:mode2asympa} in $y< \hg$.
-The absence of $U_{yy}$ in \eqref{eqn:mode2asympa} indicates that $U_{yy}$ is approximated to be zero in $y<\hg$, and therefore the negative values of $U_{yy}$ that occurs in the boundary layer do not affect this mode of instability to leading order.
+The absence of $U_{yy}$ in \eqref{eqn:mode2asympa} indicates that $U_{yy}$ is approximated to be zero in $y<\hg$, and therefore the positive values of $U_{yy}$ that occurs in the boundary layer do not affect this mode of instability to leading order.
 Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
@@
 The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
-We have neither accounted 
-Accounting for the pre-factors appearing in the scaling argument, or considering cases when the time-scale separation is not so evident can lead to further interesting behavior \cite{Delangre06}.
+We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
+Accounting for these factors  can lead to further interesting behavior \cite{Delangre06}.
 % Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation.

### 2015-06-18 Paper_JFM/grass.tex
Mode 1 discussion modified and expanded

--- 2015-06-18 Paper_JFM/grass.tex
@@
 Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfield equation \cite{Drazin81} 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \cite{Drazin81} 
 \begin{equation}
@@
 \subsection{Mode 1}
-Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits an asymptotic behavior with $k \sim H/\delta$, and $\Rey (\sim H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$) at the threshold.
-This limit can be understood by asymptotically estimating the sizes of the terms in ~\eqref{Orr-somerfield} using $D\sim H/\delta$, $\sigma \sim O(1)$, and $\ubl \sim \delta/H$; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous and the vegetation drag term in the boundary layer are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
-The terms balance when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$). Using the boundary layer co-ordinate $\eta = y/\delta$, velocity scaling 
-$U(y) = (\delta/H)\bar{U}(\eta)$ and scaling of wavenumber $k = (H/\delta) \bar{k}$ in the boundary layer ~\eqref{Orr-somerfield} simplifies to
+We numerically observe that critical Reynolds number above which perturbation become unstable scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
+Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}).
+The behavior of this critical Reynolds number can be understood by considering the limit $\Rey \gg 1$ and $\Ndg \gg 1$.
+Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
+Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
+
+To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/\delta$, 
+$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
 \begin{equation}
@@
 \end{equation}
-in a region of thickness O($\delta$) near $y=\hg$.
-Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and soultion will converge in the limit of $R \sim \Ndg^2$, confirming our numerial finding.
+in a region of thickness O($\delta$) near $y=\hg$, where $\bar{D} = d/d\eta$. 
+Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
+Our numerical finding confirm this expectation; the critical $\Rey$ scales as $\Ndg^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
+
+Mode 1 shares many characteristics with the Kelvin-Helmholtz instability (see Table \ref{tab:comparison}). 
+The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to Kelvin-Helmholtz instability. 
+The extend of the unstable mode is also localized to the boundary layer region.
+Furthermore, the strong vorticity gradient $U_{yy}$ in the boundary layer plays a central role in destabilizing the flow and localizing the instability to that region.
+
+Our detailed description of Mode 1, given by \eqref{eqn:mode1asymp} also highlights key differences with formulations of Kelvin-Helmholtz. 
+Kelvin-Helmholtz is usually described using the inviscid Rayleigh's equation \eqref{eqn:Rayleigh}, and is therefore not parametrized by the Reynolds number.
+Describing the instability using the Orr-Sommerfeld equation introduces the Reynolds number as a parameter, but shear flows with tanh-profiles are unstable for all values of the parameter \cite{Drazin81}.
+Therefore, based on the inviscid and viscous formulations of Kelvin-Helmholtz instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.
+
+In our model, the vegetation, even if made very dense, may not simply be modeled by a no-slip condition.
+Instead the porous nature of the vegetation implies that the flow penetrates a thin boundary layer region and establishes a strong shear profile. 
+The (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1, in this formulation.
+The threshold flow condition arises from a competition between of the destabilizing role fluid inertia, which is very similar to the one played in Kelvin-Helmholtz, and the vegetation drag.
+The vegetation drag may not be neglected within this boundary layer, and therefore plays a central role in the Mode 1 instability mechanism.
+
 %We expect identical asymptotic behavior for a fixed relative magnitude of the terms on the r.h.s.

### 2015-06-18 Paper_JFM/grass.tex
Add almost complete role of turbulence model

@@
 The eigenvalues and the mode shape are otherwise independent of $\delta$, and therefore we conclude that the boundary layer only plays a secondary role of regularizing the discontinuity in tangential velocity arising at $y=\hg$ in this instability mechanism.
+The enhanced shear in the boundary layer plays no role for this mode of instability.
 % We interpret Mode 2 as the instability of an inviscid fluid, with the vegetation modeled by a continuum drag field, and for which the boundary layer near the grass top plays no role. 
@@
 Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
-Based on this analysis, we conclude that Mode 2 is distinct from the Kelvin-Helmholtz instability, and bears its existence purely because of the vegetation drag.
+Based on this analysis, we conclude that Mode 2 is distinct from the Kelvin-Helmholtz instability, and owes its existence to the vegetation drag.
 
@@
 We have modeled turbulence using a constant eddy viscosity. 
-While such a model is too simplistic and
+The simplicity of this model allows us to make progress and capture the essential features of the instability without accounting for the rich and detailed characteristics of the turbulence. 
+However, this simplicity in some cases only provides a qualitatively accurate description of the flow. 
+In this subsection, we present an account of the advantages and shortcomings of assuming a constant eddy viscosity to model the turbulence.
+
+As a consequence of the constant eddy viscosity, the boundary layer thickness $\delta$ scales as $\ReyNdg^{-1/3}$.
+Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$.
+While the precise thickness of the boundary layer is governed by the details of the turbulence model, the existence of this boundary layer for dense vegetation is a general feature of the flow. 
+We have captured one realization of this feature using constant eddy viscosity.
+Experiments have also shown that a mixing length $l$ model better approximates the turbulent characteristics of the flow with $l = \delta$; the boundary layer itself establishes eddies to transport momentum.
+The eddy viscosity corresponding to this model is $\mu = \rho U \delta$, and the leading order balance between turbulent momentum transport and vegetation drag is $\mu U/\delta^2 = \rho C_N d N_g U^2$.
+Substituting $\mu$ yields $\delta/H = \Ndg^{-1}$, in agreement with the experimental observations.
+This simple scaling analysis shows that the boundary layer thickness depends on the turbulence model, and indicates that turbulence models based on mixing lengths will yield more realistic scalings for boundary layer thickness.
+
+The Mode 1 instability is driven by the intense shear on the scale of the boundary layer.
+The driving mechanism for this instability is similar to that of Kelvin-Helmholtz, and relies only on the presence of this shear as presented in the $U_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}.
+Therefore, we expect Mode 1 instability to be exhibited independent of the turbulence model. 
+We further expect the fastest growing wavenumber to be proportional to $1/\delta$, and the mode to be localized to the boundary layer.
+The threshold parameters for Mode 1, however, may depend on the precise turbulence model used.
+
+For the Mode 2 instability, the turbulent momentum transport is found to be irrelevant to leading order.
+In the asymptotic limit of dense grass, \eqref{eqn:mode2asymp} shows that the instability is driven by the interaction of the unvegetated flow with the vegetation drag.
+The precise turbulence model used influences the regularization of the sharp transition in tangential velocity across the grass top.
+Therefore we expect Mode 2 to be present and its features to be preserved even if a different turbulence model is used.
 

### 2015-06-18 Paper_JFM/grass.tex
Almost all written

@@
 Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}. 
-Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed, and drive large amplitude synchronous oscillations.
-This model has been applied to predict the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed.
+The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
+Basic application of this model have been used to predict the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
@@
 (iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated water depth, and therefore inconsistent with the definition of the free shear layer.
-Here we present a mathematical model for the linear instability that accounts correctly for these effects, while also explaining lab experiments and field observations.
+
+Here we present a mathematical model for the linear instability that accounts correctly for these effects.
+The vegetation is modeled as a continuum through a drag on the fluid proportional to the grass blade density. 
+We solve for the steady fully-developed flow profile established due to an imposed pressure gradient in the presence of this drag, and perform linear stability analysis of this profile.
+We compare the results of our calculations with data available from lab experiments and field observations.
+To better understand the flow instability mechanisms and to compare them with the current explanations of \monami, we investigate the limit of asymptotically dense vegetation.
+
+We find two modes of instability labeled Mode 1 and Mode 2.  
+We find that Mode 1 resembles many characteristics of Kelvin-Helmholtz instability but with the inclusion of the vegetation drag. 
+Mode 2 is not related to the Kelvin-Helmholtz instability mechanism and arises purely due to the interaction of an inviscid layer of fluid with submerged rigid.
+The presence of vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
 
@@
 Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \cite{Drazin81} 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \cite{Drazin81,Chen97,Chu91} 
 \begin{equation}
@@
 
-\subsection{Comparison with Kelvin-Helmholtz}
-Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
-Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be understood as the instability on the scale of the water column.
-Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
-The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria demonstrates its role in setting the threshold, and in distinguishing them from the Kelvin-Helmholtz instability.
-% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability.
-
-The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
-% A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
-% Another difference is that terrestrial vegetation is much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
-In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case.
-We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
-In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
-
 Mode 2 has characteristics distinct from Kelvin-Helmholtz. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation
@@
 Based on this analysis, we conclude that Mode 2 is distinct from the Kelvin-Helmholtz instability, and owes its existence to the vegetation drag.
+
+% \subsection{Comparison with Kelvin-Helmholtz}
+% Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+% Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be  % understood as the instability on the scale of the water column.
+% Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
+% The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria % demonstrates its role in setting the threshold, and in distinguishing them from the Kelvin-Helmholtz instability.
+% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability.
 
@@
 
+\subsection{Comparison with previous models}
+The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
+% A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
+% Another difference is that terrestrial vegetation is much more rigid, whereas aquatic vegetation is buoyant \cite{Vivoni98,Ghisal02}. 
+In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case.
+We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
+In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
+
+Chen and Jirka (1997), White and Nepf (2007).
+What are their results and how do they relate to ours?
+These analyses have investigated instability of hyperbolic tangent profiles using modified versions of Orr Sommerfeld or Rayleigh's equations in the context of shallow water flow instabilities.
+They have represented the additional drag arising due to vegetation or bottom friction using a parametrization analogous to ours.
+
 \section{Conclusion}
+
 We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.

### 2015-06-19 Paper_JFM/grass.tex
Final finishing touches

@@
 \definecolor{tableShade}{gray}{0.8}
-\title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
+% \title{\textit{Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed}
+\title{Linear stability analysis for \textit{monami} in a submerged sea grass bed}
 \author{Ravi Singh}
@@
 Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
-This damping leads to a finite threshold flow for the instability, which agrees with experimental observations.
-This role of vegetation drag differentiates our mechanism from the previous hypothesis that the Kelvin-Helmholtz instability underlies \monami.
+We find two distinct modes of instability, which we label Mode 1 and Mode 2.
+Mode 1 mechanism is closely related to Kelvin-Helmholtz instability modified to account for vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction of the flow in the vegetated and unvegetated layers.
+The vegetation damping leads to a finite threshold flow for both these modes.
+The model predictions for the threshold flow criteria for onset and the resulting oscillation frequency compare well with experimental observations.
+The inclusion of vegetation drag differentiates our mechanism from the previous analyses of \monami.
 \end{abstract}
@@
 Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
-The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \cite{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\cite{AckermanOkubo93}.  
+The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \citep{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
 In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
 
-Current explanations of \monami ~invoke the existence of a free shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag  \cite{Ikeda96,Ghisal02,Raupach96}. 
-Its instability, through a mechanism similar to the Kelvin-Helmholtz instability, is thought to lead to coherent eddies over the grass bed.
+Current explanations of \monami~\citep{Ikeda96,Ghisal02,Raupach96} invoke the existence of a shear layer at the top of the grass bed (henceforth called grass top) due to vegetation drag. 
+Its instability, through a mechanism similar to the Kelvin-Helmholtz (KH) instability, is thought to lead to coherent eddies over the grass bed.
 The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
-Basic application of this model have been used to predict the observed frequency of \monami~and to understand transport in the seagrass bed \cite{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+Basic application of this model have been used to predict the observed frequency of \monami~and to understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
-However, several aspects of this explanation remain unsatisfactory: 
-(i)   The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \cite{Delangre06} or hyperbolic tangent \cite{Ghisal02,Raupach96}. 
-(ii)  The role of drag in damping the coherent perturbations to the shear profile is sometimes ignored \cite{Raupach96}. 
-(iii) No existing theory explains the threshold flow speed, observed in the lab \cite{Ghisal02} and the field \cite{Grizzle96}, below which \monami ~is not observed.
-(iv)  The thickness of the free shear layer is in many cases comparable to the unvegetated water depth, and therefore inconsistent with the definition of the free shear layer.
+However, several aspects of this explanation remain unsatisfactory. 
+The instability is modeled using the inviscid Rayleigh equation and the role of drag in damping the coherent perturbations to the shear profile is ignored \citep{Raupach96}. 
+It is so despite the shear layer itself being produced as a result of the vegetation drag.
+Furthermore, the shear layer is assumed to be not influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
+The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
+While these profiles qualitatively capture the features of the profile, the dependence of the shear layer thickness on the parameters of the problem remains unexplained.
+And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.
 
@@
 We find two modes of instability labeled Mode 1 and Mode 2.  
-We find that Mode 1 resembles many characteristics of Kelvin-Helmholtz instability but with the inclusion of the vegetation drag. 
-Mode 2 is not related to the Kelvin-Helmholtz instability mechanism and arises purely due to the interaction of an inviscid layer of fluid with submerged rigid.
+We find that Mode 1 resembles many characteristics of KH instability but is modified by the presence of the vegetation drag. 
+We also find another mode of instability, Mode 2, unrelated to the KH instability mechanism, arising purely due to the interaction of an inviscid layer of fluid with submerged rigid vegetation.
 The presence of vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
@@
 Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
-We show that the drag of the grass results in an instability mode different from the Kelvin-Helmholtz instability.
+We show that the drag of the grass results in an instability mode different from the KH instability.
 \begin{figure}
@@
 In the field, both $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
-Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
+% Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
+% Note that only the product $C_N N_g d$ appears together in our model.
 
@@
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\cite{Nepf00}, whereas the latter models the free interface.
+The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the latter models the free interface.
 A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
@@
 A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
-Because this boundary layer develops independent of the influence of the boundaries, we identify it to be analogous to the free shear layer \cite{Ghisal02,Nepf04} in the previous explanation of \monami.
-This dependence of $\delta$ on $N_g$, verified in Fig.~\ref{basicflow}, gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
-The figure also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$.
+We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
+This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
+Fig. \ref{basicflow} also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. 
 In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}).
@@
 Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \cite{Drazin81,Chen97,Chu91} 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \citep{Drazin81,Chen97,Chu91} 
 \begin{equation}
 \begin{split}
-\left(D^2 -k^{2} \right)^2\phi &= \Rey \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi \\
-&+\Rey \Ndg D\left(2 S U D \phi\right),
+\Rey^{-1}\left(D^2 -k^{2} \right)^2\phi &= \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi + \Ndg D\left(2 S U D \phi\right),
 \label{Orr-somerfield}
@@
 A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
-The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \cite{Ghisal02}.
+The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \citep{Ghisal02}.
 The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \cite{White07}.
-Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \cite{Raupach96,Delangre06}.
+A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \citep{White07}.
+Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
 A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
 
-The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \cite{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
+The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
  \begin{tabular}{l|c|c|c}
-			& Kelvin-Helmholtz 				& Mode 1 		& Mode 2 \\ \hline
+			& KH 				& Mode 1 		& Mode 2 \\ \hline
  Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
@@
  \end{tabular}
- \caption{Comparison between Kelvin-Helmholtz instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
+ \caption{Comparison between KH instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
  \label{tab:comparison}
@@
 \section{Discussion -- comparison of unstable modes with Kelvin Helmholtz}
-The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the Kelvin-Helmholtz instability. 
+The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the KH instability. 
 \subsection{Mode 1}
@@
 \begin{split}
-\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi \\
-&+\bar{D}\left(2S \bar{U} D \phi\right),
+\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi + \bar{D}\left(2S \bar{U} \bar{D} \phi\right),
 \label{eqn:mode1asymp}
@@
 
-Mode 1 shares many characteristics with the Kelvin-Helmholtz instability (see Table \ref{tab:comparison}). 
-The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to Kelvin-Helmholtz instability. 
+Mode 1 shares many characteristics with the KH instability (see Table \ref{tab:comparison}). 
+The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to KH instability. 
 The extend of the unstable mode is also localized to the boundary layer region.
@@
 
-Our detailed description of Mode 1, given by \eqref{eqn:mode1asymp} also highlights key differences with formulations of Kelvin-Helmholtz. 
-Kelvin-Helmholtz is usually described using the inviscid Rayleigh's equation \eqref{eqn:Rayleigh}, and is therefore not parametrized by the Reynolds number.
-Describing the instability using the Orr-Sommerfeld equation introduces the Reynolds number as a parameter, but shear flows with tanh-profiles are unstable for all values of the parameter \cite{Drazin81}.
-Therefore, based on the inviscid and viscous formulations of Kelvin-Helmholtz instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.
+Our detailed description of Mode 1, given by \eqref{eqn:mode1asymp} also highlights key differences with formulations of KH. 
+KH is usually described using the inviscid Rayleigh's equation, 
+\begin{align}
+\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi, 
+\label{eqn:Rayleigh}
+\end{align}
+and is therefore not parametrized by the Reynolds number.
+Describing the instability using the Orr-Sommerfeld equation introduces the Reynolds number as a parameter, but shear flows with tanh-profiles are unstable for all values of the parameter \citep{Drazin81}.
+Therefore, based on the inviscid and viscous formulations of KH instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.
 
@@
 The (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1, in this formulation.
-The threshold flow condition arises from a competition between of the destabilizing role fluid inertia, which is very similar to the one played in Kelvin-Helmholtz, and the vegetation drag.
+The threshold flow condition arises from a competition between of the destabilizing role of fluid inertia, which is very similar to the one played in KH, and the vegetation drag.
 The vegetation drag may not be neglected within this boundary layer, and therefore plays a central role in the Mode 1 instability mechanism.
@@
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-In this limit, the non-dimensional flow in the grass bed is $U\sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $\sigma + ikU \approx \sigma$ to leading order. 
-Furthermore, $U_{yy}$ decays to zero outside the boundary layer and within the grass. 
-Outside the grass, the turbulent viscous stress term is negligible compared to the inertial term. Thus, \eqref{Orr-somerfield} simplifies to 
+In this limit, the non-dimensional flow in the grass bed is $U_g/U_0 \sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $ikU \approx \sigma$ may be neglected in comparison to $\sigma$. 
+Furthermore, $U_{yy}$ decays to zero within the grass outside the boundary layer . 
+Outside the grass, the turbulent viscous stress term is negligible compared to the inertial term because $\Rey \gg 1$. 
+Thus, \eqref{Orr-somerfield} simplifies to 
 \begin{subequations}
@@
 
-Mode 2 has characteristics distinct from Kelvin-Helmholtz. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation
-\begin{align}
-\left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi. 
-\label{eqn:Rayleigh}
-\end{align}
-An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \cite{Rayleigh1879}. 
+Mode 2 has characteristics distinct from KH. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation \eqref{eqn:Rayleigh}.
+An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \citep{Rayleigh1879}. 
 However, for our $U(y)$ profiles, $U_{yy}(y) = -1$ above the grass and therefore does not change sign for $y>\hg$. 
@@
 Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
-Based on this analysis, we conclude that Mode 2 is distinct from the Kelvin-Helmholtz instability, and owes its existence to the vegetation drag.
+Based on this analysis, we conclude that Mode 2 is distinct from the KH instability, and owes its existence to the vegetation drag.
 
-% \subsection{Comparison with Kelvin-Helmholtz}
-% Table \ref{tab:comparison} compares the two modes to each other, and to the Kelvin-Helmholtz instability. 
+% \subsection{Comparison with KH}
+% Table \ref{tab:comparison} compares the two modes to each other, and to the KH instability. 
 % Because the eigenfunction of Mode 1 is localized over a length scale $\delta$, it may be interpreted as the instability of the flow in the boundary layer, whereas Mode 2 may be  % understood as the instability on the scale of the water column.
 % Mode 1 appears to be superficially similar to the Kelvin Helmholtz mechanism, whereas Mode 2 arises purely from the interaction between the unvegetated water column and the flow through the vegetation.
-% The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria % demonstrates its role in setting the threshold, and in distinguishing them from the Kelvin-Helmholtz instability.
-% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional Kelvin-Helmholtz instability.
+% The appearance of the vegetation drag parameter in the dominant balances represented by ~\eqref{eqn:mode1asymp} and ~\eqref{eqn:mode2asymp}, and the resulting threshold criteria % demonstrates its role in setting the threshold, and in distinguishing them from the KH instability.
+% Vegetation drag plays a dominant role in the mechanism for both the modes, which distinguishes our analysis from the traditional KH instability.
 
@@
 As a consequence of the constant eddy viscosity, the boundary layer thickness $\delta$ scales as $\ReyNdg^{-1/3}$.
-Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$.
+Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$ \citep{Nepf07}.
 While the precise thickness of the boundary layer is governed by the details of the turbulence model, the existence of this boundary layer for dense vegetation is a general feature of the flow. 
-We have captured one realization of this feature using constant eddy viscosity.
-Experiments have also shown that a mixing length $l$ model better approximates the turbulent characteristics of the flow with $l = \delta$; the boundary layer itself establishes eddies to transport momentum.
+We have captured one possible realization of this feature using constant eddy viscosity.
+Experiments have also shown that a model based on mixing length $l$ better approximates the turbulent characteristics of the flow with $l = \delta$; \textit{i.e.}, the boundary layer itself establishes eddies to transport momentum.
 The eddy viscosity corresponding to this model is $\mu = \rho U \delta$, and the leading order balance between turbulent momentum transport and vegetation drag is $\mu U/\delta^2 = \rho C_N d N_g U^2$.
 Substituting $\mu$ yields $\delta/H = \Ndg^{-1}$, in agreement with the experimental observations.
+Within our framework, the mixing length model implies a scale for the eddy viscosity $\mu = \rho U_g \delta$ at grass top, which corresponds to an effective $\Rey = U_0H/U_g \delta$.
+Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implied $U_g/U_0 \sim \delta/H$, and therefore $\Rey = (H/\delta)^2$.
+Substituting this relation in $\delta/H = (\Rey \Ndg)^{-1/3}$ and solving for $\delta$ yields $\delta/H = \Ndg^{-1}$, which is in agreement with experiments and the simple scaling analysis.
 This simple scaling analysis shows that the boundary layer thickness depends on the turbulence model, and indicates that turbulence models based on mixing lengths will yield more realistic scalings for boundary layer thickness.
+At the same time, the qualitative features of the instability may be predicted by re-interpreting results of our analysis.
 
 The Mode 1 instability is driven by the intense shear on the scale of the boundary layer.
-The driving mechanism for this instability is similar to that of Kelvin-Helmholtz, and relies only on the presence of this shear as presented in the $U_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}.
+The driving mechanism for this instability is similar to that of KH, and relies only on the presence of this shear as presented in the $U_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}. 
 Therefore, we expect Mode 1 instability to be exhibited independent of the turbulence model. 
-We further expect the fastest growing wavenumber to be proportional to $1/\delta$, and the mode to be localized to the boundary layer.
+We further expect the fastest growing wavenumber to be proportional to $1/\delta$, and the mode to be localized to the boundary layer because these results have a basis in dimensional analysis.
 The threshold parameters for Mode 1, however, may depend on the precise turbulence model used.
@@
 \subsection{Comparison with previous models}
-The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \cite{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
+
+% Chen and Jirka (1997), White and Nepf (2007).
+What are their results and how do they relate to ours?
+These analyses have investigated instability of hyperbolic tangent profiles using modified versions of Orr Sommerfeld or Rayleigh's equations in the context of shallow water flow instabilities.
+They have represented the additional drag arising due to vegetation or bottom friction using a parametrization analogous to ours.
+A modified version of the Orr-Sommerfeld equation was analyzed previously by \cite{Chu91}, \cite{Chen97}, and \cite{White07} in the context of instabilities in depth-averaged shallow water flows, where bottom friction replaces or augments vegetation drag.
+They assumed the steady profile to be a hyperbolic tangent, the drag to be isotropic, and the flow domain to be infinite in $y$.
+\cite{White07} also neglected the eddy viscosity in their stability analysis.
+While a detailed investigation needed to compare the consequence of the different assumptions is outside the scope of this paper, we discuss similarities and differences between their results and ours.
+These investigations only found one unstable mode.
+It is most likely so because they restricted their calculations to a parameter regime where the two modes have not yet been separated from each other, as is the case shown in Fig.
+\ref{K_Re_sigma_set3} for the lowest $\Ndg$.
+These investigations also found that increasing the drag could further destabilize the flow, which is consistent with our interpretation of Mode 2 instability mechanism.
+
+The analogous oscillation of terrestrial canopies in wind, known as \textit{honami} \citep{Inoue56,Raupach96}, is different because the atmospheric boundary layer is much larger than the vegetation height.
 % A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically \cite{Vivoni98,Nepf00}. 
@@
 We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
-In this manner, we recover the Kelvin-Helmholtz-like characteristics observed in the terrestrial case.
+In this manner, we recover the KH-like characteristics observed in the terrestrial case.
 
-Chen and Jirka (1997), White and Nepf (2007).
-What are their results and how do they relate to ours?
-These analyses have investigated instability of hyperbolic tangent profiles using modified versions of Orr Sommerfeld or Rayleigh's equations in the context of shallow water flow instabilities.
-They have represented the additional drag arising due to vegetation or bottom friction using a parametrization analogous to ours.
+% We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+% For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \citep{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
+% The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
+% We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
+% Accounting for these factors  can lead to further interesting behavior \citep{Delangre06}.
 
+While predictions of the threshold $\Rey$ for the onset of the instability and the frequency of oscillations are comparable to experimental observations, further improvement is possible.
+The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
+In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \citep{Vivoni98,Nepf00}. 
+The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \citep{Ghisal02, Nepf04}. 
+Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
 \section{Conclusion}
-
-We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
-For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
-The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
-We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
+% We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+% For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
+% The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
+% We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
 Accounting for these factors  can lead to further interesting behavior \cite{Delangre06}.
 % Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation. 
-
-% The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-% In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \cite{Vivoni98,Nepf00}. 
-% The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \cite{Ghisal02, Nepf04}. 
-% Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
-
-In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional Kelvin-Helmholtz due to the presence of the vegetation drag. 
+In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional KH due to the presence of the vegetation drag. 
 The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
@@
 MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
-We thank Heidi Nepf, Marco Ghisalberti, and L.
+We thank Heidi Nepf, Marco Ghisalberti, Mitul Luhar, and L.
 Mahadevan for helpful discussions.

### 2015-06-19 Paper_JFM/grass.tex
Fix authors list

@@
 \title{Linear stability analysis for \textit{monami} in a submerged sea grass bed}
-\author{Ravi Singh}
-\affiliation{Brown University, Providence RI 02912 USA}
-% \author{L. Mahadevan}
-% \affiliation{Harvard University, Cambridge MA 02138 USA}
-\author{M. M. Bandi}
-\affiliation{OIST Graduate University, Okinawa 904-0495, Japan}
-\author{Amala Mahadevan}
-\affiliation{Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA}
-\author{Shreyas Mandre}
-\affiliation{Brown University, Providence RI 02912 USA}
+\author{
+Ravi Singh\aff{1}, 
+M. M. Bandi\aff{2},
+Amala Mahadevan\aff{3}
+\and
+Shreyas Mandre\aff{4}}
+\affiliation{
+\aff{1}Department of Physics, Brown University, Providence RI 02912 USA
+\aff{2}OIST Graduate University, Okinawa 904-0495, Japan
+\aff{3}Physical Oceanography, Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA
+\aff{4}School of Engineering, Brown University, Providence RI 02912 USA
+}
 

### 2015-06-20 Paper_JFM/grass.tex
Changes Made by Ravi for frequency figure and few mother minor things

@@
 \section{Mathematical model}
-Although \monami ~is manifest in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
+Although \monami ~is manifested in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
 Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
-We show that the drag of the grass results in an instability mode different from the KH instability.
+%We show that the drag of the grass results in an instability mode different from the KH instability.
 \begin{figure}
-\centerline{\includegraphics[scale=1]{Grass_Base_Nepf_shear}}
+\centerline{\includegraphics[scale=.99]{Grass_Base_Nepf_shear} }
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
@@
 \end{figure}
+
+\begin{figure}
+\centerline{\includegraphics[]{new_graph_freq}}
+\caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
+The experimental data in the inset is obtained from refs. \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+\label{frequency_comparison}
+\end{figure}
+
 \begin{figure}
@@
 \caption{
-Critical Reynolds number (top) and the corresponding marginally stable wave number (bottom) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the top panel. 
-Inset compares experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}. 
-The experimental data in the inset is obtained from refs. \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+Critical Reynolds number (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
 }
@@
 where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
@@
 % Note that only the product $C_N N_g d$ appears together in our model.
+
+\section{Linear stability analysis}
+We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
+\begin{equation}
+ -\frac{dP}{dx}+\mu U''(y) -S(y) \rho C_N d N_gU |U|=0
+\label{base_equ}
+\end{equation}
+where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
+Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
+The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the latter models the free interface.
+A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
+The profile $U(y)$ has three distinct regions.
+Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{-dP/dx}{\rho C_N dN_g}}$, arising from balancing the drag with pressure gradient. 
+Outside the vegetation, the velocity has a simple parabolic profile. % due to the balance between viscous forces and the pressure gradient. 
+At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
+Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(-dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
+Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
+We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
+This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
+Fig. \ref{basicflow} also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. 
+In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}).
 
@@
 
-\section{Linear stability analysis}
-We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
-\begin{equation}
- -\frac{dP}{dx}+\mu U''(y) +S(y) \rho C_N d N_gU |U|=0
-\label{base_equ}
-\end{equation}
-where $S(y)=1$ for $0<y<\hg$ and $S(y)=0$ for $\hg< y< 2H$. 
-Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the latter models the free interface.
-A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
-The profile $U(y)$ has three distinct regions.
-Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$, arising from balancing the drag with pressure gradient. 
-Outside the vegetation, the velocity has a simple parabolic profile. % due to the balance between viscous forces and the pressure gradient. 
-At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
-Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
-A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
-We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
-This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
-Fig. \ref{basicflow} also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. 
-In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}).
-
 Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
@@
 A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \citep{White07}.
-Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
-A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
+%Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
+%A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
 
-The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in the inset in Fig.~\ref{Re_vs_delta}.
+Comparison of this predicted threshold against estimated Reynolds number for lab scale experiments are shown in  Fig \ref{Re_vs_delta}.
+We have also compared frequency (Im$(\sigma)$) of the fastest growing mode with the observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
 Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
-Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $\Rey (H/\delta)^2$  (or $\Rey^{5/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(\delta/H)^{-4}$ (or $(\Rey \Ndg)^{4/3})$.
+Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
 The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
 
-To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/\delta$, 
+To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/(\delta/H)$, 
 $U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
@@
 The (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1, in this formulation.
-The threshold flow condition arises from a competition between of the destabilizing role of fluid inertia, which is very similar to the one played in KH, and the vegetation drag.
+The threshold flow condition arises from a competition between the destabilizing role of fluid inertia, which is very similar to the one played in KH, and the vegetation drag.
 The vegetation drag may not be neglected within this boundary layer, and therefore plays a central role in the Mode 1 instability mechanism.
@@
 % Note that for large $\Ndg$, the non-dimensional steady flow velocity inside the vegetation is $U_g \sim (\Rey \Ndg)^{-1/2}$. 
-In this limit, the non-dimensional flow in the grass bed is $U_g/U_0 \sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $ikU \approx \sigma$ may be neglected in comparison to $\sigma$. 
+In this limit, the non-dimensional flow in the grass bed is $U_g/U_0 \sim (\Rey \Ndg)^{-1/2} \ll 1$, and therefore $ikU \ll \sigma$ may be neglected in comparison to $\sigma$. 
 Furthermore, $U_{yy}$ decays to zero within the grass outside the boundary layer . 
@@
 \begin{figure}
-\includegraphics[]{Asymptotic_noshear}
+\centerline{\includegraphics[]{Asymptotic_noshear}}
 % \includegraphics[width=12cm]{Asymptotic_noshear}

### 2015-06-20 Paper_JFM/grass.tex
Changes Made by Ravi, minor modification

@@
 
-Comparison of this predicted threshold against estimated Reynolds number for lab scale experiments are shown in  Fig \ref{Re_vs_delta}.
-We have also compared frequency (Im$(\sigma)$) of the fastest growing mode with the observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
+The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
 Describing the instability using the Orr-Sommerfeld equation introduces the Reynolds number as a parameter, but shear flows with tanh-profiles are unstable for all values of the parameter \citep{Drazin81}.
-Therefore, based on the inviscid and viscous formulations of KH instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.
+Therefore, based on the inviscid formulations of KH instability, the origin of the threshold flow conditions observed in experiments and the field is unclear.
 

### 2015-06-20 Paper_JFM/grass.tex

@@
 We also find another mode of instability, Mode 2, unrelated to the KH instability mechanism, arising purely due to the interaction of an inviscid layer of fluid with submerged rigid vegetation.
-The presence of vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
+The presence of dissipation due to vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
 
@@
 where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \cite{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ (\cite{Nepf99,Nepf00,Nepf04}), where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
@@
 \end{equation}
-where $D=d/dy$, and subject to the boundary conditions $D\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+where $D=d/dy$, and subject to the boundary conditions $\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
 The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
@@
 Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
-Our numerical finding confirm this expectation; the critical $\Rey$ scales as $\Ndg^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
+Our numerical finding confirm this expectation; the critical $\Rey$ scales as $(H/\delta)^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
 
@@
 Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
-Based on this analysis, we conclude that Mode 2 is distinct from the KH instability, and owes its existence to the vegetation drag.
+Based on this analysis, we conclude that Mode 2 is distinct from the KH instability as well, and owes its existence to the vegetation drag.
 
@@
 
-As a consequence of the constant eddy viscosity, the boundary layer thickness $\delta$ scales as $\ReyNdg^{-1/3}$.
+As a consequence of the constant eddy viscosity, the boundary layer thickness $\delta$ scales as $H(\ReyNdg)^{-1/3}$.
 Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$ \citep{Nepf07}.

### 2015-06-22 Paper_JFM/grass.tex
Final finishing touches

@@
 The vegetation damping leads to a finite threshold flow for both these modes.
-The model predictions for the threshold flow criteria for onset and the resulting oscillation frequency compare well with experimental observations.
-The inclusion of vegetation drag differentiates our mechanism from the previous analyses of \monami.
+Experimental observations for the onset and frequency of waving compare well with model predictions for the criteria for instability onset and the imaginary part of the complex growth rate respectively, but experiments appear to lie in a parameter regime where the two modes can not be distinguished.
+% The inclusion of vegetation drag differentiates our mechanism from the previous linear stability analyses of \monami.
 \end{abstract}
@@
 One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
-In this letter, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
+In this article, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
 
@@
 The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
-Basic application of this model have been used to predict the observed frequency of \monami~and to understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+Basic application of this model predict the observed frequency of \monami~and to understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
-However, several aspects of this explanation remain unsatisfactory. 
-The instability is modeled using the inviscid Rayleigh equation and the role of drag in damping the coherent perturbations to the shear profile is ignored \citep{Raupach96}. 
-It is so despite the shear layer itself being produced as a result of the vegetation drag.
-Furthermore, the shear layer is assumed to be not influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
-The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
-While these profiles qualitatively capture the features of the profile, the dependence of the shear layer thickness on the parameters of the problem remains unexplained.
-And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.
-
-Here we present a mathematical model for the linear instability that accounts correctly for these effects.
-The vegetation is modeled as a continuum through a drag on the fluid proportional to the grass blade density. 
-We solve for the steady fully-developed flow profile established due to an imposed pressure gradient in the presence of this drag, and perform linear stability analysis of this profile.
-We compare the results of our calculations with data available from lab experiments and field observations.
-To better understand the flow instability mechanisms and to compare them with the current explanations of \monami, we investigate the limit of asymptotically dense vegetation.
-
-We find two modes of instability labeled Mode 1 and Mode 2.  
-We find that Mode 1 resembles many characteristics of KH instability but is modified by the presence of the vegetation drag. 
-We also find another mode of instability, Mode 2, unrelated to the KH instability mechanism, arising purely due to the interaction of an inviscid layer of fluid with submerged rigid vegetation.
-The presence of dissipation due to vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
-
-\section{Mathematical model}
-Although \monami ~is manifested in the motion of the grass, the drag exerted by the grass bed on the flow is central to the hypothesized instability. 
-The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \cite{Ghisal02,Nepf06}. 
-Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
-%We show that the drag of the grass results in an instability mode different from the KH instability.
 \begin{figure}
@@
 \end{figure}
+However, several aspects of this explanation remain unsatisfactory. 
+The instability is modeled using the inviscid Rayleigh equation and despite the role of drag in producing the shear layer, its role in damping the perturbations is ignored \citep{Raupach96}. 
+Furthermore, the shear layer is assumed not to be influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
+The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
+Assuming the profiles \textit{ad hoc} leaves the dependence of the shear layer thickness on the parameters of the problem remains unexplained.
+And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.
 
+Here we present a mathematical model for the linear instability that accounts for these effects.
+Although \monami ~is manifested in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
+The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \citep{Ghisal02,Nepf06}. 
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
+The vegetation is modeled as a continuum drag on the fluid acting perpendicular to the blade orientation and proportional to the grass density.
+In the limit of dense vegetation, the steady profile established in the presence of the drag exhibits a localized region of enhanced shear gradient near the grass top and drives a flow instability.
+% provides an opportunity to compare the instability mechanism with KH.
+% The grass bed does not impose an effective no-slip boundary condition on the unvegetated flow; instead the flow penetrating the grass bed drives an instability.
+While some features of instability are similar to KH, we also find significant differences.
+This comparison is the focus of this article.
+
+% We solve for the steady fully-developed flow profile established due to an imposed pressure gradient in the presence of this drag, and perform linear stability analysis of this profile.
+% We compare the results of our calculations with data available from lab experiments and field observations.
+% To better understand the flow instability mechanisms and to compare them with the current explanations of \monami, we investigate the limit of asymptotically dense vegetation. 
+% 
+% We find two modes of instability labeled Mode 1 and Mode 2.  
+% We find that Mode 1 resembles many characteristics of KH instability but is modified by the presence of the vegetation drag. 
+% We also find another mode of instability, Mode 2, unrelated to the KH instability mechanism, arising purely due to the interaction of an inviscid layer of fluid with submerged rigid vegetation.
+% The presence of dissipation due to vegetation drag leads to a finite threshold flow condition for onset of the instability for both these modes.
+
+\section{Mathematical model}
+%We show that the drag of the grass results in an instability mode different from the KH instability.
 \begin{figure}
@@
 \caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
-The experimental data in the inset is obtained from refs. \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.}
+The experimental data in the inset is obtained from \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+}
 \label{frequency_comparison}
 \end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
-% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
-% \vspace{-6mm} \hspace{-5mm}
-% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
-\end{center}
-\caption{
-Critical Reynolds number (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters from experiments reported in ref. \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
-}
-\label{Re_vs_delta}
-\end{figure}
-
-The drag exerted by the vegetation, assumed dense, is modeled by a continuous body force $\mathbf{f}$.
-This drag enters the fluid mass and momentum balance equations as 
+The vegetation is assumed sufficiently dense so that the drag exerted by it may be modeled by a continuous body force $\mathbf{f}$ entering the equation governing the flow as.
 \begin{equation}
-\grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
+\grad\cdot\bu = 0,\hspace{3mm} \rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad p+\mu\grad^{2}\bu +\mathbf{f}, %+\rho\mathbf{g}
 \label{navier-stokes}
 \end{equation}
-where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and pressure respectively, $\mu$ the dynamic eddy viscosity and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ (\cite{Nepf99,Nepf00,Nepf04}), where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and the dynamic pressure respectively, and is $\mu$ the dynamic eddy viscosity. % and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
@@
 \section{Linear stability analysis}
-We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, satisfying
+We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, so that \eqref{navier-stokes} simplifies to
 \begin{equation}
- -\frac{dP}{dx}+\mu U''(y) -S(y) \rho C_N d N_gU |U|=0
+ -\frac{dP}{dx}+\mu U''(y) -S(y) \rho C_N d N_gU |U|=0,
 \label{base_equ}
@@
 Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(-dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
-Solving for $\delta$ and $\ubl$ yields $\delta/H = \ubl/U_0=(\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
-A numerical estimate of $\delta$ ($U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
+Solving for $\delta$ and $\ubl$ yields $\delta/H \sim \ubl/U_0 \sim (\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
+A numerical estimate of $\delta$ (estimated as $U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
 We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
@@
 In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}). 
+\begin{figure}
+\begin{center}
+\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
+% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
+% \vspace{-6mm} \hspace{-5mm}
+% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
+\end{center}
+\caption{
+Critical Reynolds number (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters from experiments reported by \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
+In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+}
+\label{Re_vs_delta}
+\end{figure}
 
+Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
+\begin{equation}
+\begin{split}
+\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
+\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0,
+\end{split} \nonumber
+\end{equation}
+where the tilde are dropped.
+These equations are non-dimensionalized using half channel height $H$, velocity $U_0$, and time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the vegetation submergence ratio $\hg/H$.
+We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the instability mechanism. 
+Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
+the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \citep{Drazin81,Chen97,Chu91} 
+\begin{equation}
+\begin{split}
+\Rey^{-1}\left(D^2 -k^{2} \right)^2\phi &= \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi + \Ndg D\left(2 S U D \phi\right),
+\label{Orr-somerfield}
+\end{split}
+\end{equation}
+where $D=d/dy$, and subject to the boundary conditions $\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
+The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
+We solve \eqref{Orr-somerfield} numerically for $\sigma$ and $\phi$.
+
+\section{Results - unstable modes and critical parameters}
+A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
+The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations by \cite{Ghisal02}.
+The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
+% A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \citep{White07}.
+%Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
+%A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
+The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
+The experimentally observed \monami ~wavelengths are not available for comparison.
+
+To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
+The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical $\delta$ discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
+To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in Fig.~\ref{K_Re_sigma_set3}. 
+The smallest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
+We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
+The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
+
+\section{Discussion -- comparison of unstable modes with Kelvin Helmholtz}
+The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguish them from each other and facilitate comparison with KH instability mechanism. 
+\subsection{Mode 1}
+We numerically observe that critical Reynolds number above which perturbation become unstable scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
+Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}).
+The behavior of this critical Reynolds number can be understood by considering the limit $\Rey \gg 1$ and $\Ndg \gg 1$.
+Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
+Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
+The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
+
+To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} near the grass-top using the the boundary layer scalings $\eta = y/(\delta/H)$, 
+$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
+\begin{equation}
+\begin{split}
+\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi + \bar{D}\left(2S \bar{U} \bar{D} \phi\right),
+\label{eqn:mode1asymp}
+\end{split}
+\end{equation}
+in a region of thickness O($\delta$) near $y=\hg$, where $\bar{D} = d/d\eta$. 
+Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
+Our numerical finding confirm this expectation; the critical $\Rey$ scales as $(H/\delta)^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
+
+Mode 1 shares many characteristics with the KH instability (see Table \ref{tab:comparison}). 
+The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to KH instability. 
+The extent of the unstable mode is also localized to the boundary layer region.
+The porous nature of the vegetation implies that a weak flow of magnitude $\ubl = U_0 \delta/H$ penetrates a thin boundary layer region $\delta$, and therefore the shear gradient $U_{yy} \sim U_0/\delta H$ is largest in this region.
+The strong shear gradient $U_{yy}$ in the boundary layer plays a central role in destabilizing the flow and localizing the instability to that region. 
 \begin{figure*}
@@
 
-Next we substitute $\bu = (U+\tilde{u}, \tilde{v})$, $p=P+\tilde{p}$ in ~\eqref{navier-stokes} and expand to linear order to investigate the evolution of small perturbations $(\tilde{u}, \tilde{v})$, which obey
-\begin{equation}
-\begin{split}
-\rho(u_t+U u_x+vU_y) &= -p_x+ {\mu}\nabla^2u-2S\rho C_{N}dN_{g}Uu, \\
-\rho(v_t+ Uv_x) &= -p_y+ {\mu}\nabla^2v, \hspace{0.3cm} \nabla\cdot\bu=0,
-\end{split} \nonumber
-\end{equation}
-where the tilde are dropped.
-These equations are non-dimensionalized using half channel height $H$, velocity $U_0$, and time $H/U_0$, leading to three non-dimensional parameters, \textit{viz.} $\Rey$, $\Ndg$, and the vegetation submergence ratio $\hg/H$.
-We also use $\delta/H$ in lieu of $\Ndg$ to parametrize the vegetation density and help elucidate the instability mechanism. 
-Using a stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance, we seek a solution of 
-the form $\left(u,v,\psi \right)= \left(\hat u(y), \hat v(y), \phi(y) \right)e^{ikx+\sigma t}$ to obtain a modified Orr-Sommerfeld equation \citep{Drazin81,Chen97,Chu91} 
-\begin{equation}
-\begin{split}
-\Rey^{-1}\left(D^2 -k^{2} \right)^2\phi &= \left[ \left({\sigma}+ikU\right) \left(D^2-k^2\right) -ikU_{yy}\right]\phi + \Ndg D\left(2 S U D \phi\right),
-\label{Orr-somerfield}
-\end{split}
-\end{equation}
-where $D=d/dy$, and subject to the boundary conditions $\phi = D^2\phi = 0$ at $y=0$ and $y=2$. 
-The growth rate $\sigma$ for a given wave number $k$ appears as an eigenvalue that allows a non-trivial solution $\phi$ of  \eqref{Orr-somerfield}.
-We solve \eqref{Orr-somerfield} numerically for $\sigma$ and $\phi$.
-
-\section{Results - unstable modes and critical parameters}
-A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
-The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations \citep{Ghisal02}.
-The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
-A similar conclusion was presented for an analogous problem (flow around an emergent (i.e., $\hg>2H$) sea grass patch), but by assuming $U(y)$ to be a tanh-profile, and neglecting the viscous term \citep{White07}.
-%Previous calculations for terrestrial grass either exclude the vegetation drag in their models \cite{Raupach96}, or assume the mean velocity profile \textit{ad hoc} \citep{Raupach96,Delangre06}.
-%A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
-
-The frequency (Im$(\sigma)$) of the fastest growing mode agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
-The experimentally observed \monami ~wavelengths are not available for comparison.
-
-\begin{table*}
-% \footnotesize
-\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
-\renewcommand{\arraystretch}{1.2}
- \begin{tabular}{l|c|c|c}
-			& KH 				& Mode 1 		& Mode 2 \\ \hline
- Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
- Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
- Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
- Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
- Linearized dynamics	& Equation \eqref{eqn:Rayleigh}		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
- Dense grass limit &  no grass included & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
- Critical parameters	& none						& $\Rey \propto \Ndg^{2}$ 	& $\Rey \propto \Ndg$ \\
- Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
- Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no, spans water column
- \end{tabular}
- \caption{Comparison between KH instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
- \label{tab:comparison}
-\end{table*}
-To better understand the instability mechanism, we consider the dependence of the fastest growing wavenumber on $\delta$.
-The fastest growing wavenumber first increases proportional to $H/\delta$, but at a critical $\delta$ discontinuously jumps and remains $O(1)$ (see Fig.~\ref{Re_vs_delta}). 
-To aid in explaining this behavior, we show heat maps of Re$(\sigma)$ as a function of $\Rey$ and $k$, for different $\hg/H$ and $\Ndg$ in Fig.~\ref{K_Re_sigma_set3}. 
-The smallest $\Rey$ on the neutral curve (Re$(\sigma)=0$) sets the threshold. 
-We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
-The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
-
-\section{Discussion -- comparison of unstable modes with Kelvin Helmholtz}
-The distinct asymptotic behavior of the two modes as $\Ndg \gg 1$ distinguishes it from the KH instability. 
-\subsection{Mode 1}
-We numerically observe that critical Reynolds number above which perturbation become unstable scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
-Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}).
-The behavior of this critical Reynolds number can be understood by considering the limit $\Rey \gg 1$ and $\Ndg \gg 1$.
-Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
-Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
-The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
-
-To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} in the the boundary layer using $\eta = y/(\delta/H)$, 
-$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
-\begin{equation}
-\begin{split}
-\left(\bar{D}^2 -\bar{k}^{2} \right)^2\phi &= (\Rey/\Ndg^2)^{1/3} \left[ \left({\sigma}+i\bar{k}\bar{U}\right) \left(\bar{D}^2-\bar{k}^2\right) -i\bar{k}\bar{U}_{\eta\eta}\right]\phi + \bar{D}\left(2S \bar{U} \bar{D} \phi\right),
-\label{eqn:mode1asymp}
-\end{split}
-\end{equation}
-in a region of thickness O($\delta$) near $y=\hg$, where $\bar{D} = d/d\eta$. 
-Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
-Our numerical finding confirm this expectation; the critical $\Rey$ scales as $(H/\delta)^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
-
-Mode 1 shares many characteristics with the KH instability (see Table \ref{tab:comparison}). 
-The fastest growing wavenumber at the critical $\Rey$ scales as $k \propto (H/\delta)$, similar to KH instability. 
-The extend of the unstable mode is also localized to the boundary layer region.
-Furthermore, the strong vorticity gradient $U_{yy}$ in the boundary layer plays a central role in destabilizing the flow and localizing the instability to that region.
-
 Our detailed description of Mode 1, given by \eqref{eqn:mode1asymp} also highlights key differences with formulations of KH. 
@@
 
-In our model, the vegetation, even if made very dense, may not simply be modeled by a no-slip condition.
-Instead the porous nature of the vegetation implies that the flow penetrates a thin boundary layer region and establishes a strong shear profile. 
-The (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1, in this formulation.
+In our model, (turbulent eddy) viscosity sets the scale of the boundary layer, and therefore for Mode 1.
+However, the boundary layer is established only in the vegetated region; the velocity profile does not saturate on the scale of $\delta$ in the unvegetated region.
 The threshold flow condition arises from a competition between the destabilizing role of fluid inertia, which is very similar to the one played in KH, and the vegetation drag.
@@
 \end{subequations}
-
-\begin{figure}
-\centerline{\includegraphics[]{Asymptotic_noshear}}
-% \includegraphics[width=12cm]{Asymptotic_noshear}
-\caption{
-Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
-% Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters for Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
-The approach of mode shapes to each other for these small values of $\delta/H$ indicates that the dense vegetation asymptote is reached. 
-Mode 1 shapes appear self-similar in shape as $\delta\to 0$.
-Inset shows rescaled $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$ approach a universal shape, indicating that an asymptotic limit has been reached. 
-% The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.
-}
-\label{Asymptotic_mode}
-\end{figure}
 The only remaining parameter in \eqref{eqn:mode2asymp} is $\Rey/\Ndg$. 
@@
 
+\begin{table*}
+% \footnotesize
+\rowcolors{3}{tableShade}{white}  %% start alternating shades from 3rd row
+\renewcommand{\arraystretch}{1.2}
+ \begin{tabular}{l|c|c|c}
+			& KH 				& Mode 1 		& Mode 2 \\ \hline
+ Base velocity profile 	& $U(y) = U_0 \tanh(y/\delta)$			& \multicolumn{2}{c}{Equation \eqref{base_equ}} \\
+ Domain 		& $-\infty < y < \infty$			& \multicolumn{2}{c}{$-1<y<1$} \\
+ Inflection point	& exists at $y=0$				& \multicolumn{2}{c}{$U''(y)$ discontinuous at $y=\hg$} \\
+ Shear layer thickness	& $\delta$					& \multicolumn{2}{c}{$\delta \sim  H\left(\Rey \Ndg \right)^{-1/3}$} \\
+ Linearized dynamics	& Equation \eqref{eqn:Rayleigh}		& \multicolumn{2}{c}{Equation \eqref{Orr-somerfield}} \\
+ Dense grass limit &  no grass included & Equation \eqref{eqn:mode1asymp} & Equation \eqref{eqn:mode2asymp}  \\
+ Critical parameters	& none						& $\Rey \propto \Ndg^{2}$ 	& $\Rey \propto \Ndg$ \\
+ Most unstable $k$ as $\delta \to 0$	& $\propto H/\delta$		& $\propto H/\delta$	& $O(1)$ \\
+ Mode localized?	& yes, near $y=0$				& yes, near $y=\hg$			& no, spans water column
+ \end{tabular}
+ \caption{Comparison between KH instability and the two unstable modes resulting from solution of \ref{Orr-somerfield}.}
+ \label{tab:comparison}
+\end{table*}
 \subsection{Role of turbulence model}\label{subsec:turbulencemodel}
 We have modeled turbulence using a constant eddy viscosity. 
-The simplicity of this model allows us to make progress and capture the essential features of the instability without accounting for the rich and detailed characteristics of the turbulence. 
+The simplicity of this model allows us to make progress and capture the essential features of the instability. 
+% without accounting for the rich and detailed characteristics of the turbulence. 
 However, this simplicity in some cases only provides a qualitatively accurate description of the flow. 
@@
 We have captured one possible realization of this feature using constant eddy viscosity.
-Experiments have also shown that a model based on mixing length $l$ better approximates the turbulent characteristics of the flow with $l = \delta$; \textit{i.e.}, the boundary layer itself establishes eddies to transport momentum.
-The eddy viscosity corresponding to this model is $\mu = \rho U \delta$, and the leading order balance between turbulent momentum transport and vegetation drag is $\mu U/\delta^2 = \rho C_N d N_g U^2$.
-Substituting $\mu$ yields $\delta/H = \Ndg^{-1}$, in agreement with the experimental observations.
-Within our framework, the mixing length model implies a scale for the eddy viscosity $\mu = \rho U_g \delta$ at grass top, which corresponds to an effective $\Rey = U_0H/U_g \delta$.
-Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implied $U_g/U_0 \sim \delta/H$, and therefore $\Rey = (H/\delta)^2$.
-Substituting this relation in $\delta/H = (\Rey \Ndg)^{-1/3}$ and solving for $\delta$ yields $\delta/H = \Ndg^{-1}$, which is in agreement with experiments and the simple scaling analysis.
+Experiments have also shown that a model based on mixing length $l$ better approximates the turbulent characteristics of the flow with $l \sim \delta$; \textit{i.e.}, the boundary layer itself establishes eddies to transport momentum.
+The eddy viscosity corresponding to this model is $\mu \sim \rho U \delta$, and the leading order balance between turbulent momentum transport and vegetation drag is $\mu U/\delta^2 \sim \rho C_N d N_g U^2$.
+Substituting $\mu$ yields $\delta/H \sim \Ndg^{-1}$, in agreement with the experimental observations.
+
+Replacing the constant eddy viscosity by the scale for it from the mixing length hypothesis also recovers the experimentally observed scaling for boundary layer thickness.
+Within our framework, the mixing length model implies a scale for the eddy viscosity $\mu \sim \rho \ubl \delta$ at grass top, which corresponds to an effective $\Rey \sim U_0H/\ubl \delta$.
+Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implied $\ubl/U_0 \sim \delta/H$, and therefore $\Rey \sim (H/\delta)^2$.
+Substituting this relation in $\delta/H \sim (\Rey \Ndg)^{-1/3}$ and solving for $\delta$ yields $\delta/H \sim \Ndg^{-1}$.
 This simple scaling analysis shows that the boundary layer thickness depends on the turbulence model, and indicates that turbulence models based on mixing lengths will yield more realistic scalings for boundary layer thickness.
@@
 
+\begin{figure}
+\centerline{\includegraphics[]{Asymptotic_noshear}}
+% \includegraphics[width=12cm]{Asymptotic_noshear}
+\caption{
+Plot of the neutral Mode 1 (solid) and Mode 2 (dashed) shape $|\phi|$ in the limit of small $\delta/H$ for $\hg/H=0.2$. 
+% Mode 1 is shown in solid and Mode 2 is shown in dashed. The parameters for Mode 2 shapes are chosen such that $\Rey \gg 1$, $\Ndg \gg 1$ (specified in terms of $\delta/H$) but $\Rey/\Ndg = O(1)$. 
+The approach of mode shapes to each other for these small values of $\delta/H$ indicates that the dense vegetation asymptote is reached. 
+Mode 1 shapes appear self-similar in shape as $\delta\to 0$.
+Inset shows rescaled $|\phi|$ for Mode 1 as a function of $(y-\hg)/\delta$ approach a universal shape, indicating that an asymptotic limit has been reached. 
+% The limit is not yet reached for the case $\delta/H = 0.005$ due to the influence of bottom boundary; the vegetation height in this case is comparable to the boundary layer thickness.
+}
+\label{Asymptotic_mode}
+\end{figure}
 While predictions of the threshold $\Rey$ for the onset of the instability and the frequency of oscillations are comparable to experimental observations, further improvement is possible.

### 2015-06-22 Paper_JFM/grass.tex
Add sentence on which mode is exhibited in experiments

@@
 The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
+All experimental data we have found corresponds to a vegetation density for which the two modes have not split from one another.
 
@@
 The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades \citep{Vivoni98,Nepf00}. 
+In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades due to variation in vegetation characteristics \citep{Vivoni98,Nepf00}. 
 The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \citep{Ghisal02, Nepf04}. 
 Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
+
 \section{Conclusion}

### 2015-06-22 Paper_JFM/grass.tex
Changed the caption for figure 3

@@
 \caption{
-Critical Reynolds number (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Critical Reynolds number, threshold Reynolds number for Mode 1 and Mode 2 (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
 Parameters from experiments reported by \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 

### 2015-07-02 Paper_JFM/grass.tex
Final modifications for reducing length to 10 pages

@@
 The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
-Basic application of this model predict the observed frequency of \monami~and to understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+Basic application of this model predicts the observed frequency of \monami~and helps understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
@@
 \section{Mathematical model}
-%We show that the drag of the grass results in an instability mode different from the KH instability.
 \begin{figure}
-\centerline{\includegraphics[]{new_graph_freq}}
-\caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
-The experimental data in the inset is obtained from \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
+\begin{center}
+\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
+% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
+% \vspace{-6mm} \hspace{-5mm}
+% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
+\end{center}
+\caption{
+Critical Reynolds number, threshold Reynolds number for Mode 1 and Mode 2 (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
+Parameters from experiments reported by \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
 In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
 }
-\label{frequency_comparison}
+\label{Re_vs_delta}
 \end{figure}
+%We show that the drag of the grass results in an instability mode different from the KH instability.
 The vegetation is assumed sufficiently dense so that the drag exerted by it may be modeled by a continuous body force $\mathbf{f}$ entering the equation governing the flow as.
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale = 0.95]{new_graph_R_vs_delta}\includegraphics[scale = 0.95]{new_graph_K_vs_delta}
-% \includegraphics[width=12cm]{Critical_Re_vs_delta_noshear} \\
-% \vspace{-6mm} \hspace{-5mm}
-% \includegraphics[width=12cm]{K_vs_shear_width_noshear}
-\end{center}
-\caption{
-Critical Reynolds number, threshold Reynolds number for Mode 1 and Mode 2 (left) and the corresponding marginally stable wave number (right) for different submergence ratio as a function of vegetation density parametrized by the boundary layer thickness.
-Parameters from experiments reported by \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
+\centerline{\includegraphics[]{new_graph_freq}}
+\caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
+The experimental data in the inset is obtained from \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
 In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
 }
-\label{Re_vs_delta}
+\label{frequency_comparison}
 \end{figure}
@@
 \begin{figure}
-\centerline{\includegraphics[]{Asymptotic_noshear}}
+\centerline{\includegraphics[width=9cm]{Asymptotic_noshear}}
 % \includegraphics[width=12cm]{Asymptotic_noshear}
@@
 The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \citep{Ghisal02, Nepf04}. 
-Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the dominance of the physical processes depicted in \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, are expected to remain.
+Although these model improvements might lead to a better agreement between the observed and the predicted quantities, the insight furnished by \eqref{eqn:mode1asymp} and \eqref{eqn:mode2asymp}, and therefore our main conclusions, remain useful.
 
@@
 \acknowledgments
-MMB was hosted by Brown University during this work and was supported by the OIST Graduate University with subsidy funding from the Cabinet Office, Japan.
+MMB acknowledges support from the OIST Graduate University with subsidy funding from the Cabinet Office, Japan, while visiting Brown University.
 We thank Heidi Nepf, Marco Ghisalberti, Mitul Luhar, and L.

### 2015-07-06 Paper_JFM/grass.tex
Incorporate corrections by Mahesh

@@
 The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the flow.
-Our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
+Unlike previous works, our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
 We find two distinct modes of instability, which we label Mode 1 and Mode 2.
-Mode 1 mechanism is closely related to Kelvin-Helmholtz instability modified to account for vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction of the flow in the vegetated and unvegetated layers.
-The vegetation damping leads to a finite threshold flow for both these modes.
-Experimental observations for the onset and frequency of waving compare well with model predictions for the criteria for instability onset and the imaginary part of the complex growth rate respectively, but experiments appear to lie in a parameter regime where the two modes can not be distinguished.
+Mode 1 mechanism is closely related to Kelvin-Helmholtz instability modified to account for vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction between the flow in the vegetated and unvegetated layers.
+Our model says that the vegetation damping leads to a finite threshold flow for both these modes.
+Experimental observations for the onset and frequency of waving compare well with model predictions for the instability onset criteria and the imaginary part of the complex growth rate respectively, but experiments lie in a parameter regime where the two modes can not be distinguished.
 % The inclusion of vegetation drag differentiates our mechanism from the previous linear stability analyses of \monami.
@@
 The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \citep{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-One such response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
-In this article, we provide a hydrodynamic mechanism for the onset of these coherent oscillations.
+One such behaviour of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
+In this article, we present a linear hydrodynamic instability underlying the onset of these coherent oscillations.
 
@@
 The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
-Assuming the profiles \textit{ad hoc} leaves the dependence of the shear layer thickness on the parameters of the problem remains unexplained.
+Assuming the profiles \textit{ad hoc} leaves the dependence of the shear layer thickness on the parameters of the problem unexplained.
 And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.
@@
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \citep{Ghisal02,Nepf06}. 
-Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, on average oriented vertically.
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, and oriented vertically on average.
 The vegetation is modeled as a continuum drag on the fluid acting perpendicular to the blade orientation and proportional to the grass density.
@@
 \end{equation}
-where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and the dynamic pressure respectively, and is $\mu$ the dynamic eddy viscosity. % and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting the skin friction, we model it as the form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the blade number density per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and the dynamic pressure respectively, and $\mu$ is the (dynamic) eddy viscosity. % and $\mathbf{g}$ the acceleration due to gravity. 
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting skin friction, we model it as form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the number of grass blades per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
-In the field, both $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
+In the field, $C_N$, $N_g$ and $\mu$ vary with position, but we do not expect these variations to be central to the instability mechanism, and therefore take them to be constants. 
 % Based on previous experiments \cite{Vivoni98,Nepf00}, we take $C_N = 1$.
@@
 \section{Linear stability analysis}
-We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, so that \eqref{navier-stokes} simplifies to
+We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, and use it to non-dimensionalize the mathematical model.
+The momentum balance \eqref{navier-stokes} for $U(y)$ simplifies to
 \begin{equation}
@@
 The profile $U(y)$ has three distinct regions.
-Within vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{-dP/dx}{\rho C_N dN_g}}$, arising from balancing the drag with pressure gradient. 
+Within the vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{-dP/dx}{\rho C_N dN_g}}$, arising from a balance of the drag with the pressure gradient. 
 Outside the vegetation, the velocity has a simple parabolic profile. % due to the balance between viscous forces and the pressure gradient. 
@@
 Fig. \ref{basicflow} also shows that the asymptotic regime of a thin boundary layer is expected to hold for $\ReyNdg \gtrsim 100$. 
-In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used for deriving \eqref{eqn:mode2asymp}). 
+In this notation, $U_g/U_0 = (\Rey \Ndg)^{-1/2}$ (used later in deriving \eqref{eqn:mode2asymp}). 
 \begin{figure}
@@
 We observe that as $\Ndg$ increases, the unstable region splits into two; we refer to the region with the higher $k$ as ``Mode 1'', and the one with the lower $k$ as ``Mode 2''. 
-The unstable region for Mode 1, depending on $\hg/H$ either recedes to higher $\Rey$ or shrinks to zero size, as the vegetation density increases, causing the most unstable mode to transition discontinuously.
-All experimental data we have found corresponds to a vegetation density for which the two modes have not split from one another.
+For $\hg/H\lesssim 0.9$, the unstable region for Mode 1 recedes to higher $\Rey$, and for $\hg/H \gtrsim 0.9$, the region shrinks to zero size.
+In either case, due to such behaviour the most unstable mode to transitions discontinuously.
+All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two.
 
@@
 \subsection{Mode 1}
-We numerically observe that critical Reynolds number above which perturbation become unstable scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
+We numerically observe that threshold Reynolds number for Mode 1 instability scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
 Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}).
@@
 Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
-The advection term, viscous term and vegetation drag terms balances when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
+The advection term, viscous term and vegetation drag terms balance when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
 
 To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} near the grass-top using the the boundary layer scalings $\eta = y/(\delta/H)$, 
-$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$, with these scalings ~\eqref{Orr-somerfield} simplifies to
+$U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$.
+With these scalings ~\eqref{Orr-somerfield} simplifies to
 \begin{equation}
@@
 Since $(\Rey/\Ndg^2)$ is the only remaining parameter in \eqref{eqn:mode1asymp}, the mode shape and solution are expected to converge in the limit $\Rey \gg 1$, $\Ndg \gg 1$, but $\Rey/\Ndg^2$ fixed.
-Our numerical finding confirm this expectation; the critical $\Rey$ scales as $(H/\delta)^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
+Our numerical findings confirm this expectation; the critical $\Rey$ scales as $(H/\delta)^2$ as shown in Fig.~\ref{Re_vs_delta} and the mode shapes are self-similar with length scale $\delta$, as shown in Fig \ref{Asymptotic_mode}.
 
@@
 \caption{
-$Re(\sigma)$ and the neutral curve ($Re(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
+$\text{Re}(\sigma)$ and the neutral curve ($\text{Re}(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
 As $\Ndg$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
@@
 The structure of this mode in the aforementioned limit is such that $\phi$ is continuous at $y=h_g$, but $D\phi$ undergoes a rapid transition there on the scale of boundary layer thickness $\delta$.
-The eigenvalues and the mode shape are otherwise independent of $\delta$, and therefore we conclude that the boundary layer only plays a secondary role of regularizing the discontinuity in tangential velocity arising at $y=\hg$ in this instability mechanism.
+The eigenvalues and the mode shape are otherwise independent of $\delta$.
+Therefore we conclude that the boundary layer only plays a secondary role of regularizing the discontinuity in tangential velocity arising at $y=\hg$ in this instability mechanism.
 The enhanced shear in the boundary layer plays no role for this mode of instability.
@@
 Furthermore, the presence of the critical parameter $\Rey/\Ndg$ in \eqref{eqn:mode2asympa} indicates the presence of alternative destabilizing dynamics, involving the interaction of flow in the unvegetated region governed by \eqref{eqn:Rayleigh} with the flow in the vegetated region incorporating the drag.
-Based on this analysis, we conclude that Mode 2 is distinct from the KH instability as well, and owes its existence to the vegetation drag.
+Therefore, we conclude that Mode 2 is distinct from the KH instability, and owes its existence to vegetation drag.
 
@@
 Experimental observations show that the boundary layer thickness scales instead as $\Ndg^{-1}$ \citep{Nepf07}.
-While the precise thickness of the boundary layer is governed by the details of the turbulence model, the existence of this boundary layer for dense vegetation is a general feature of the flow. 
+Whereas the precise boundary layer thickness is governed by the details of the turbulence model, the existence of this boundary layer for dense vegetation is independent of the turbulence model. 
 We have captured one possible realization of this feature using constant eddy viscosity.
@@
 The Mode 1 instability is driven by the intense shear on the scale of the boundary layer.
-The driving mechanism for this instability is similar to that of KH, and relies only on the presence of this shear as presented in the $U_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}. 
+The driving mechanism for this instability is similar to that of KH, and relies only on the presence of this shear as presented in the $\bar{U}_{\eta\eta}$ term in equation \eqref{eqn:mode1asymp}. 
 Therefore, we expect Mode 1 instability to be exhibited independent of the turbulence model. 
@@
 In the asymptotic limit of dense grass, \eqref{eqn:mode2asymp} shows that the instability is driven by the interaction of the unvegetated flow with the vegetation drag.
-The precise turbulence model used influences the regularization of the sharp transition in tangential velocity across the grass top.
+The influence of the turbulence model is limited to the regularization of the sharp transition in tangential velocity across the grass top.
 Therefore we expect Mode 2 to be present and its features to be preserved even if a different turbulence model is used.
@@
 \acknowledgments
-MMB acknowledges support from the OIST Graduate University with subsidy funding from the Cabinet Office, Japan, while visiting Brown University.
+MMB was supported by the Collective Interactions Unit, OIST Graduate University, while visiting Brown University.
 We thank Heidi Nepf, Marco Ghisalberti, Mitul Luhar, and L.

### 2015-09-02 Paper_JFM/grass.tex
Incorporate suggestions and typo corrections from Amala

@@
 \aff{2}OIST Graduate University, Okinawa 904-0495, Japan
-\aff{3}Physical Oceanography, Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA
+\aff{3}Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA
 \aff{4}School of Engineering, Brown University, Providence RI 02912 USA
@@
 The onset of \monami ~-- the synchronous waving of sea grass beds driven by a steady flow -- is modeled as a linear instability of the flow.
-Unlike previous works, our model treats the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
+Unlike previous works, our model considers the drag exerted by the grass in establishing the steady flow profile, and in damping out perturbations to it.
 We find two distinct modes of instability, which we label Mode 1 and Mode 2.
-Mode 1 mechanism is closely related to Kelvin-Helmholtz instability modified to account for vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction between the flow in the vegetated and unvegetated layers.
-Our model says that the vegetation damping leads to a finite threshold flow for both these modes.
+Mode 1 is closely related to Kelvin-Helmholtz instability modified by vegetation drag, whereas Mode 2 is unrelated to Kelvin-Helmholtz and arises from an interaction between the flow in the vegetated and unvegetated layers.
+The vegetation damping, according to our model, leads to a finite threshold flow for both these modes.
 Experimental observations for the onset and frequency of waving compare well with model predictions for the instability onset criteria and the imaginary part of the complex growth rate respectively, but experiments lie in a parameter regime where the two modes can not be distinguished.
@@
 \section{Introduction}
-Sea grasses exhibit a rich set of dynamical behavior due to their collective interaction with fluid flows.  
-The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transportation of dissolved oxygen, plant growth, and biomass production  \citep{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
-One such behaviour of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
+Sea grasses exhibit a rich set of dynamical behaviors due to their collective interaction with fluid flows.  
+The hydrodynamic processes resulting from this behavior influence many environmental processes such as sedimentation, transport of dissolved oxygen, plant growth, and biomass production  \citep{Fonseca87,Grizzle96,Nepf99,Nepf2012}.
+One response of the submerged grass beds to steady currents is the formation of coherent large amplitude oscillations, known as \monami ~\citep{AckermanOkubo93}.  
 In this article, we present a linear hydrodynamic instability underlying the onset of these coherent oscillations.
@@
 Its instability, through a mechanism similar to the Kelvin-Helmholtz (KH) instability, is thought to lead to coherent eddies over the grass bed.
-The grass responds to these eddies by deforming, which manifests itself as large amplitude synchronous oscillations.
-Basic application of this model predicts the observed frequency of \monami~and helps understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+The grass responds to these eddies by deforming, which leads to large amplitude synchronous oscillations of the grass blades.
+% Basic application of this model predicts the observed frequency of \monami~and helps understand transport in the seagrass bed \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
+This picture can be used to derive a simple scaling dependence of the \monami~frequency on the flow speed and the shear layer thickness, and understand transport in the seagrass bed  \citep{Nepf00,Ghisal02,Nepf04,Okamoto12}.
 
@@
 Furthermore, the shear layer is assumed not to be influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
-The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}. 
-Assuming the profiles \textit{ad hoc} leaves the dependence of the shear layer thickness on the parameters of the problem unexplained.
+The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}, with parameters fitted experimental observations.
+The origin of these profiles, the values of the fitted parameters, and their effect on \monami~remains unexplained.  
+% Assuming the profiles \textit{ad hoc} leaves the dependence of the shear layer thickness on the parameters of the problem unexplained.
 And finally, no existing theory explains the threshold flow speed, observed in the lab \citep{Ghisal02} and the field \citep{Grizzle96}, below which \monami ~is not observed.
@@
 Here we present a mathematical model for the linear instability that accounts for these effects.
-Although \monami ~is manifested in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
+Although \monami ~is manifest in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \citep{Ghisal02,Nepf06}. 
-Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, and oriented vertically on average.
+Therefore, to develop the essential mathematical model, we assume the grass blades to be rigid, and oriented vertically (along $y$ direction) on average.
 The vegetation is modeled as a continuum drag on the fluid acting perpendicular to the blade orientation and proportional to the grass density.
@@
 \end{equation}
-where $\rho$, $\bu$, and $p$ are the fluid density, velocity, and the dynamic pressure respectively, and $\mu$ is the (dynamic) eddy viscosity. % and $\mathbf{g}$ the acceleration due to gravity. 
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting skin friction, we model it as form drag on the vegetation, $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the number of grass blades per unit horizontal area, $C_{N}$ the drag coefficient, and $d$ the average blade width projected perpendicular to the flow.
+where $\rho$, $\bu=(u,v)$, and $p$ are the fluid density, velocity, and the dynamic pressure respectively, and $\mu$ is the (dynamic) eddy viscosity.
+% and $\mathbf{g}$ the acceleration due to gravity.
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting skin friction, we model the form drag on the vegetation as $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the number of grass blades per unit horizontal area, $C_{N}$ the drag coefficient for the flow normal to the grass, and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
@@
 \section{Linear stability analysis}
-We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$, and use it to non-dimensionalize the mathematical model.
+We first calculate the fully developed steady solution $\bu = U(y)\boldsymbol{\hat{x}}$ of ~\eqref{navier-stokes} driven by constant pressure gradient $dP/dx$ in a water column of depth $2H$ and vegetated depth $\hg$, and use it to non-dimensionalize the mathematical model.
 The momentum balance \eqref{navier-stokes} for $U(y)$ simplifies to
@@
 Eq. \eqref{base_equ} is solved subject to no shear at the boundaries, i.e., $U'(0) = U'(2H) = 0$.
-The former arises for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the latter models the free interface.
-A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements is shown in Fig.~\ref{basicflow}.
+The lower boundary conditions is appropriate for dense vegetation because the shear stress exerted by the bottom surface is expected to be negligible compared to the vegetation drag~\citep{Nepf00}, whereas the upper boundary condition models the free interface.
+A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements by \cite{Nepf04} is shown in Fig.~\ref{basicflow}.
 The profile $U(y)$ has three distinct regions.
@@
 At the grass top, continuity of shear stresses results in a boundary layer of thickness $\delta$. 
-Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(-dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $(\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2)$, and the continuity of shear stress across the grass top implies $(\ubl/\delta \sim U_0/H)$.
+Denoting $\ubl$ to be the velocity scale in the boundary layer, and $U_0 = {(-dP/dx)~H^2}/{\mu}$ the velocity scale in the unvegetated region, the balance between viscous forces and vegetation drag implies $\mu \ubl/\delta^2 \sim \rho C_N d N_g \ubl^2$, and the continuity of shear stress across the grass top implies $\ubl/\delta \sim U_0/H$.
 Solving for $\delta$ and $\ubl$ yields $\delta/H \sim \ubl/U_0 \sim (\ReyNdg)^{-1/3}$, where $\Ndg = \left(C_N d H N_g\right)$ is the vegetation frontal area per bed area, and $\Rey=\rho U_0 H/\mu$ is the Reynolds number of the flow.
@@
 \caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
-The experimental data in the inset is obtained from \cite{Ghisal02} (Ghisalberti) and \cite{Vivoni98} (Vivoni). 
+The experimental data in the inset is obtained from \cite{Ghisal02} and \cite{Vivoni98}. 
 In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
@@
 A threshold in $\Rey$, above which the flow is unstable (Re$(\sigma)>0$) for at least one $k$, emerges from the solution of ~\eqref{Orr-somerfield}. 
-The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations by \cite{Ghisal02}.
+The dependence of this threshold $\Rey$, and the corresponding marginally stable wavenumber $k$, on $\delta/H$ and $\hg/H$ is shown in Fig.~\ref{Re_vs_delta}, and is found to compare well with experimental observations of \cite{Ghisal02}.
 The threshold $\Rey$ increases with the vegetation density, indicating a competition between the destabilizing shear in the flow, and the stabilizing effect of damping due to vegetation drag.
@@
 %A threshold flow condition is not reported previously either for terrestrial or submerged marine meadows.
-The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed frequencies -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field, as shown in Fig.~\ref{frequency_comparison}.
+The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed behaviour -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field as shown in Fig.~\ref{frequency_comparison}.
 The experimentally observed \monami ~wavelengths are not available for comparison.
@@
 For $\hg/H\lesssim 0.9$, the unstable region for Mode 1 recedes to higher $\Rey$, and for $\hg/H \gtrsim 0.9$, the region shrinks to zero size.
-In either case, due to such behaviour the most unstable mode to transitions discontinuously.
+In either case, due to such behaviour the most unstable mode transitions discontinuously from Mode 1 to Mode 2.
 All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two.
@@
 \subsection{Mode 1}
-We numerically observe that threshold Reynolds number for Mode 1 instability scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
+We numerically observe that the threshold Reynolds number for Mode 1 instability scales as  $\Rey \sim (H/\delta)^2$ (or $\Rey \propto {\Ndg}^{2}$). 
 Our calculations also show that Mode 1 asymptotically localizes to the boundary layer near the grass top, and exhibits highest growth for a perturbation of  $k \sim H/\delta$ (see Fig.~\ref{Re_vs_delta}).
@@
 Estimating the size of various terms of ~\eqref{Orr-somerfield} within the boundary layer in this limit help us understand the behavior of the critical $\Rey$. 
-Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous as well as vegetation drag term are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
+Using $D\sim H/\delta$, $\sigma \sim O(1)$, and $U=\ubl \sim \delta/H$ in the boundary layer; the magnitude of the advection term is $ (H/\delta)^2$  (or $\Rey^{2/3} \Ndg^{2/3}$), and the viscous and vegetation drag terms are $(1/\Rey) (\delta/H)^{-4}$ (or $(\Rey^{1/3} \Ndg)^{4/3})$.
 The advection term, viscous term and vegetation drag terms balance when $\Rey \sim (H/\delta)^2$ (or $\Rey \sim {\Ndg}^{2}$).
@@
 \caption{
-$\text{Re}(\sigma)$ and the neutral curve ($\text{Re}(\sigma)$=0) as function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
+$\text{Re}(\sigma)$ and the neutral curve ($\text{Re}(\sigma)$=0) as a function of wavenumber and $\Rey$ for parameters shown in the corresponding panel.  
 As $\Ndg$ increases, the unstable region splits into two labeled as ``Mode 1'' and ``Mode 2''. 
@@
 This convergence indicates that we have identified the correct asymptotic limit to investigate Mode 2.
+The parameter $\Rey/\Ndg$ therefore sets the threshold, justifying the numerically observed asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
 
-The parameter $\Rey/\Ndg$ therefore sets the threshold, justifying the numerically observed asymptotic behavior $\Rey \propto \Ndg$ (or $\Rey \sim ({\delta}/{H})^{-3/2}$; see Fig.~\ref{Re_vs_delta} for comparison with numerical results).
-The structure of this mode in the aforementioned limit is such that $\phi$ is continuous at $y=h_g$, but $D\phi$ undergoes a rapid transition there on the scale of boundary layer thickness $\delta$.
+The structure of this mode in the aforementioned limit is such that $\phi$ is continuous at $y=h_g$, but $D\phi$ undergoes a rapid transition there, on the scale of boundary layer thickness $\delta$.
 The eigenvalues and the mode shape are otherwise independent of $\delta$.
@@
 Mode 2 has characteristics distinct from KH. Outside the grass, the unstable mode shape is governed by the inviscid Rayleigh's equation \eqref{eqn:Rayleigh}.
-An inflection point in $U(y)$ is necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \citep{Rayleigh1879}. 
+An inflection point in $U(y)$ is a necessary condition for instability arising from \eqref{eqn:Rayleigh} according to Rayleigh's criteria \citep{Rayleigh1879}. 
 However, for our $U(y)$ profiles, $U_{yy}(y) = -1$ above the grass and therefore does not change sign for $y>\hg$. 
@@
 
-Replacing the constant eddy viscosity by the scale for it from the mixing length hypothesis also recovers the experimentally observed scaling for boundary layer thickness.
+% Replacing the constant eddy viscosity by the scale for it from the mixing length hypothesis also recovers the experimentally observed scaling for boundary layer thickness.
 Within our framework, the mixing length model implies a scale for the eddy viscosity $\mu \sim \rho \ubl \delta$ at grass top, which corresponds to an effective $\Rey \sim U_0H/\ubl \delta$.
-Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implied $\ubl/U_0 \sim \delta/H$, and therefore $\Rey \sim (H/\delta)^2$.
+Furthermore, matching the slope of the velocity profile from the boundary layer to the unvegetated flow implies $\ubl/U_0 \sim \delta/H$, and therefore $\Rey \sim (H/\delta)^2$.
 Substituting this relation in $\delta/H \sim (\Rey \Ndg)^{-1/3}$ and solving for $\delta$ yields $\delta/H \sim \Ndg^{-1}$.
 This simple scaling analysis shows that the boundary layer thickness depends on the turbulence model, and indicates that turbulence models based on mixing lengths will yield more realistic scalings for boundary layer thickness.
-At the same time, the qualitative features of the instability may be predicted by re-interpreting results of our analysis.
+At the same time, the qualitative features of the instability are represented by our analysis.
 
@@
 The influence of the turbulence model is limited to the regularization of the sharp transition in tangential velocity across the grass top.
-Therefore we expect Mode 2 to be present and its features to be preserved even if a different turbulence model is used.
+Therefore we expect Mode 2 and its features to be preserved even if a different turbulence model is used.
 
@@
 These investigations only found one unstable mode.
-It is most likely so because they restricted their calculations to a parameter regime where the two modes have not yet been separated from each other, as is the case shown in Fig.
+It is most likely so because the calculations were restricted to a parameter regime where the two modes have not yet been separated from each other, as is the case shown in Fig.
 \ref{K_Re_sigma_set3} for the lowest $\Ndg$.
-These investigations also found that increasing the drag could further destabilize the flow, which is consistent with our interpretation of Mode 2 instability mechanism.
+These investigations also found that increasing the drag could further destabilize the flow, which is consistent with our interpretation of the Mode 2 instability mechanism.
 
@@
 In the framework of our model, the limit of $\hg/H \ll 1$ while $\delta/\hg$ = constant can be used to represent the hydrodynamic instability for the terrestrial case.
-We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, so as to make Mode 2 irrelevant. 
-In this manner, we recover the KH-like characteristics observed in the terrestrial case.
+We find that in this case, the transition from Mode 1 to Mode 2 happens at such a large vegetation density, that Mode 2 is irrelevant. 
+Hence, only the KH-like characteristics are observed in the terrestrial case.
 
@@
 \end{figure}
-While predictions of the threshold $\Rey$ for the onset of the instability and the frequency of oscillations are comparable to experimental observations, further improvement is possible.
-The deviation of our model predictions from the observed may be attributed to the various simplifications we have made in our model. 
-In real meadows, the drag coefficients are known to vary from bottom to tip of the grass blades due to variation in vegetation characteristics \citep{Vivoni98,Nepf00}. 
+While predictions of the threshold $\Rey$ for the onset of the instability and the frequency of oscillations are comparable to experimental observations, the deviation of our model predictions from the observed may be attributed to the various simplifications in our model.
+In real meadows, the grass is flexible, and the drag coefficients are known to vary from bottom to tip of the grass blades due to variation in vegetation characteristics \citep{Vivoni98,Nepf00}. 
 The turbulence model for the flow through the meadow can also be improved from one with constant eddy viscosity \citep{Ghisal02, Nepf04}. 
@@
 The threshold flow condition observed in the field and in lab experiments arises due to the presence of this drag. 
-While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
-% The spatial structure of the instability modes has direct implications for transport in the grass bed; Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
-Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, flow through urban environments and therefore is expected to have a wider impact.
+Furthermore, our linear stability analysis reveals two modes, namely Mode 1 with a mechanism similar to KH and Mode 2 with characteristics distinctly different from KH. 
+% While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.
+The spatial structure of the instability modes has direct implications for transport in the grass bed.
+% Mode 1 instability likely leads to enhanced transport near the grass tips as has been observed \cite{Nepf04,Okamoto12}, while Mode 2 instability influences the whole water column.
+Our analysis also informs flow structure formation in many other related scenarios, such as flow over coral reefs, permeable sediments, and flow through urban environments, and therefore is expected to have a wider impact.
 

### 2015-09-17 ArXiV/grass.tex
Update ArXiV copy

--- 2015-09-02 Paper_JFM/grass.tex
@@
 %\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
-%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
-\documentclass{jfm}
+\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
+% \documentclass{jfm}
 \usepackage{amsmath}
@@
 \newcommand{\hg}{h_g}
-%\newcommand{\Rey}{{R}}
+\newcommand{\Rey}{{R}}
 \newcommand{\Ndg}{\tilde{N}_g}
@@
 \title{Linear stability analysis for \textit{monami} in a submerged sea grass bed}
-\author{
-Ravi Singh\aff{1}, 
-M. M. Bandi\aff{2},
-Amala Mahadevan\aff{3}
-\and
-Shreyas Mandre\aff{4}}
-\affiliation{
-\aff{1}Department of Physics, Brown University, Providence RI 02912 USA
-\aff{2}OIST Graduate University, Okinawa 904-0495, Japan
-\aff{3}Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA
-\aff{4}School of Engineering, Brown University, Providence RI 02912 USA
-}
+% \author{
+% Ravi Singh\aff{1}, 
+% M. M. Bandi\aff{2},
+% Amala Mahadevan\aff{3}
+% \and
+% Shreyas Mandre\aff{4}}
+% \affiliation{
+% \aff{1}Department of Physics, Brown University, Providence RI 02912 USA
+% \aff{2}OIST Graduate University, Okinawa 904-0495, Japan
+% \aff{3}Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA
+% \aff{4}School of Engineering, Brown University, Providence RI 02912 USA
+% }
 
 \begin{document}
-\maketitle
+\author{Ravi Singh}
+\affiliation{Brown University, Providence RI 02912 USA}
+% \author{L. Mahadevan}
+% \affiliation{Harvard University, Cambridge MA 02138 USA}
+\author{M. M. Bandi}
+\affiliation{OIST Graduate University, Okinawa 904-0495, Japan}
+\author{Amala Mahadevan}
+\affiliation{Woods Hole Oceanographic Institution, Woods Hole MA 02543 USA}
+\author{Shreyas Mandre}
+\affiliation{Brown University, Providence RI 02912 USA}
 
@@
 They have represented the additional drag arising due to vegetation or bottom friction using a parametrization analogous to ours.
-A modified version of the Orr-Sommerfeld equation was analyzed previously by \cite{Chu91}, \cite{Chen97}, and \cite{White07} in the context of instabilities in depth-averaged shallow water flows, where bottom friction replaces or augments vegetation drag.
+A modified version of the Orr-Sommerfeld equation was analyzed previously \cite{Chu91,Chen97,White07} in the context of instabilities in depth-averaged shallow water flows, where bottom friction replaces or augments vegetation drag.
 They assumed the steady profile to be a hyperbolic tangent, the drag to be isotropic, and the flow domain to be infinite in $y$.
-\cite{White07} also neglected the eddy viscosity in their stability analysis.
+One study \cite{White07} also neglected the eddy viscosity in their stability analysis.
 While a detailed investigation needed to compare the consequence of the different assumptions is outside the scope of this paper, we discuss similarities and differences between their results and ours.

### 2015-10-14 Paper_JFM/Response.tex
Adding new figure and initial Response

\documentclass[a4paper,12pt]{article}
\pagenumbering {roman}
\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}

\usepackage[above]{placeins}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

\begin{document}
%\begin{abstract}
%aasfdakaldvgas;dfasdnk
%\end{abstract}

%\author{Ravi shanker singh }
%\begin{figure}
%\centerline{\includegraphics[scale=.5]{iitblogoHuge.pdf}}
%\end{figure}

%\centerline{\large{--------}}
%\vspace{1in}
%\centerline{\large {Department Of Physics}}
%\centerline{\Large{Brown University}}
%\centerline{\large{providence}}

We would like take this opportunity to thank both the referees for reviewing our work and providing their valuable feedback. We are pleased that the referee find 
our research work to be interesting.

In this letter we have tried to address the concerns and questions raised by the referees.

\textbf{One of question raised by a referee is about constant velocity profile $U(y)=const$ with in the grass.} 
\newline
The origin of approximately constant velocity with in the grass 
can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, 
with in vegetation the dominant balance is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass term as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
numerical simulation of 3.1). Using the dominant balance of pressure gradient and drag term 
give rise to a constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$, more in depth detailed discussion about such perturbation techniques can be found in the the book by Hinch titles as ``perturbation
methods''
\newline
\textbf{ About the Origin of boundary layer $\delta$}

\end{document}

### 2015-10-15 Paper_JFM/Response.tex
Adding new figure and initial Response

@@
 
-We would like take this opportunity to thank both the referees for reviewing our work and providing their valuable feedback. We are pleased that the referee find 
+We would like to take this opportunity to thank both the referees for reviewing our work and providing their valuable feedback. We are pleased that the referee find 
 our research work to be interesting.
@@
 The origin of approximately constant velocity with in the grass 
-can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, 
-with in vegetation the dominant balance is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little
-to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass term as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
+can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, the dominant balance
+with in vegetation is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
+to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
 numerical simulation of 3.1). Using the dominant balance of pressure gradient and drag term 

### 2015-10-20 Paper_JFM/Response.tex
comments and corrections for referee 1 incorporated

@@
 The origin of approximately constant velocity with in the grass 
-can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, the dominant balance
+can be understood by well known perturbation theory (Perturbation Methods by E.J.
+Hinch, Cambridge Texts in Applied Mathematics ).
+In case of sufficiently high grass density, the dominant balance of equation (3.1)
 with in vegetation is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
@@
 \textbf{ About the Origin of boundary layer $\delta$}
+\newline
+In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip.
+Again for in-depth discussion about such problem can be found in the book of Hinch.
+\newline
+\textbf{About the estimated value of $H/\delta$ from Figure-1}
+\newline
+I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
+Geophys.
+Res.
+107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ).
+In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$.
+We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
+\newline
+\textbf{Basis for eddy viscosity}
+\newline
+The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
+\textit{Water Resources Research 40 (7), 2004} )
 

### 2015-10-20 Paper_JFM/grass.tex
Adding new figure and initial Response
Minor commits before submit
comments and corrections for referee 1 incorporated

@@
 \newcommand{\hg}{h_g}
-%\newcommand{\Rey}{{R}}
+\renewcommand{\Rey}{{R}}
 \newcommand{\Ndg}{\tilde{N}_g}
@@
 Experimental observations for the onset and frequency of waving compare well with model predictions for the instability onset criteria and the imaginary part of the complex growth rate respectively, but experiments lie in a parameter regime where the two modes can not be distinguished.
-% The inclusion of vegetation drag differentiates our mechanism from the previous linear stability analyses of \monami.
+% The inclusion of vegetation drag
+differentiates our mechanism from the previous linear stability analyses of \monami.
 \end{abstract}
@@
 \begin{figure}
-\centerline{\includegraphics[scale=.99]{Grass_Base_Nepf_shear} }
+\centerline{\includegraphics[scale=0.95]{Grass_Base_Nepf_shear} \includegraphics[scale=0.95]{Grass_shear_scale}}
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
 \caption{
-Schematic setup and comparison of our steady flow profile with that from the experiments in ref.
+Schematic setup and comparison of our steady flow profile with that from the experiments (left) in ref.
 \cite{Nepf04} (Case I from Table 1) %  with 1250 plants/m$^2$, plant height = 13.7$\pm 0.2$ cm and blade width of 0.64 cm)
@@
 The steady velocity profile can be decomposed into a parabolic profile in the unvegetated region, a uniform profile deep within the vegetation, and a boundary layer of thickness $\delta$ near the grass top.
-The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\ReyNdg$ is shown in the inset.
+The dependence of the boundary layer thickness (estimated as $|U/U_y|$ at $y=\hg$ from the numerical solution of \eqref{base_equ}) on the vegetation density parameter $\ReyNdg$ is shown in the figure on right.
 }
@@
 Furthermore, the shear layer is assumed not to be influenced by the top and bottom boundary of the domain, although the experimentally measured thickness of the shear layer is in many cases comparable to the unvegetated water depth.
-The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}, with parameters fitted experimental observations.
+The velocity profile of the free shear layer is assumed \textit{ad hoc} to be piecewise linear \citep{Delangre06} or hyperbolic tangent \citep{Ghisal02,Raupach96}, with parameters fitted using the experimental observations.
 The origin of these profiles, the values of the fitted parameters, and their effect on \monami~remains unexplained.  
@@
 % and $\mathbf{g}$ the acceleration due to gravity.
-The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting skin friction, we model the form drag on the vegetation as $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the number of grass blades per unit horizontal area, $C_{N}$ the drag coefficient for the flow normal to the grass, and $d$ the average blade width projected perpendicular to the flow.
+The Reynolds number of the flow based on the scale of the grass blade is $O(10^2-10^3)$; therefore, neglecting skin friction, we model the form drag on the vegetation as $\mathbf{f}=-N_g C_N \rho u |u| d\bx$ \citep{Nepf99,Nepf00,Nepf04}, where  $N_g$ is the number of grass blades per unit horizontal area, $C_{N}$ the drag coefficient for the flow normal to the grass, $\bx$ is unit vector in the horizontal direction and $d$ the average blade width projected perpendicular to the flow.
 In the interest of simplicity, we model the turbulence using an eddy viscosity; see \S\ref{subsec:turbulencemodel} for a detailed discussion. 
@@
 A numerical estimate of $\delta$ (estimated as $U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
-We identify the boundary layer to be analogous to the shear layer \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
+We identify the boundary layer to be analogous to the shear layer discussed by Ghisalberti and Nepf \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
 This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
@@
 The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed behaviour -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field as shown in Fig.~\ref{frequency_comparison}.
-The experimentally observed \monami ~wavelengths are not available for comparison.
+Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.
 
@@
 In either case, due to such behaviour the most unstable mode transitions discontinuously from Mode 1 to Mode 2.
-All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two.
+All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to say if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode1 or Mode2.
 

### 2015-10-22 Paper_JFM/Response/Response.tex
Add structure to Response.tex

\documentclass[letterpaper,10pt]{article}
\pagenumbering {roman}
\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}

\usepackage[above]{placeins}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

\begin{document}

\centerline{Response to Referee 1}
We thank the referee for reviewing our work and providing their valuable feedback. 
We are pleased that the referee finds our contribution interesting, and the relevance of our work to fundamental and practical fluid mechanics certain.
Here we respond to the questions raised and the feedback provided.

\begin{enumerate}
\item The referee says:
\textit{
The claims of the paper are somehow contradictory, in terms of the application to sea grass beds. The title focuses on “monami in a sea grass bed” whereas motion of the sea grass is excluded from the analysis.  For the conclusions of the paper to be applicable to sea grass motion two things would be needed : (a) a much larger discussion on the possible effects of the unstable modes on the motion of the grass, (b) a discussion on the potential role of sea grass flexibility in the instability of the coupled flow-grass system. As it is, the  title of the paper should rather  be “linear stability of the flow over a submerged sea grass bed”. In other terms, I am not sure that the results of the present paper (except the existence of a threshold) have consequences on the motion of the sea grass.
}

While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of th edetailed interpretation. more explanation. We have now modified the manucript to clarify this issue. The new manuscript now reads: ``blah blah''.

\item The referee says:
\textit{
The practical existence of a threshold flow condition for monami (or rather,  for flow instability ) to occur is doubtless.  As stated page 6, the inviscid KH instability approach does not yield any threshold, of course. The finding of a threshold by including dissipation and other aspects is no surprise. The main contribution of the paper seems then to be the second mode. This brings back to the issue of the relevance of this new mode to the motion of the sea grass.
}

Do we agree or disagree with the referee? Why? What is our argument and interpretation? How did we revise the manuscript in response to this comment?

\item The referee says:
\textit{
The comparison with the case of wind-canopy interactions is interesting but does not seem consistent with the analysis found in the reference [1]  below, where the instability seems to be more dominantly KH type in water  (when flexibility is assumed).  This would need some discussion, at least to point out the differences.
}

Similar response.

\end{enumerate}

\newpage
\centerline{Response to Referee 2}
We thank the referee for reviewing our work and providing their valuable feedback. 
We are glad that we were able to engage the referee's attention with our presentation.
Here we respond in detail to his/her comments.
\begin{enumerate}
\item The referee says:
\textit{
XXX
}

We say

\item The referee says:
\textit{
blah
}

Do we agree or disagree with the referee? Why? What is our argument and interpretation? How did we revise the manuscript in response to this comment?

\item The referee says:
\textit{
blah blah
}

Similar response.

\end{enumerate}

\newpage
\centerline{\textbf{Matched asymptotic solution}}
Present the complete solution here. Refer to this solution in our response.

The origin of approximately constant velocity with in the grass 
can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, the dominant balance
with in vegetation is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
numerical simulation of 3.1). Using the dominant balance of pressure gradient and drag term 
give rise to a constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$, more in depth detailed discussion about such perturbation techniques can be found in the the book by Hinch titles as ``perturbation
methods''
\newline
\textbf{ About the Origin of boundary layer $\delta$}

\end{document}

### 2015-10-22 Paper_JFM/Response/Response.tex

@@
 
+\centerline{Ravi's new edits}
+
+In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip.
+Again for in-depth discussion about such problem can be found in the book of Hinch.
+\newline
+\textbf{About the estimated value of $H/\delta$ from Figure-1}
+\newline
+I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
+Geophys.
+Res.
+107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ).
+In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$.
+We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
+\newline
+\textbf{Basis for eddy viscosity}
+\newline
+The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
+\textit{Water Resources Research 40 (7), 2004} )
+
 \end{document}

### 2015-10-23 Paper_JFM/Response/Response.tex
Referee 2 comment included

@@
 \item The referee says:
-\textit{
-blah
+\textit{p.5 +10,11, Although experimentally observed wavelengths are not available, they do arise from the analysis of presumably, so some comment about the values and wheater they seem physically realistic (and/or how they compare with the scales of typical motions in the turbulent flow above the canopy which are believed to be linked with monami) would be instructive.
 }
 
-Do we agree or disagree with the referee? Why? What is our argument and interpretation? How did we revise the manuscript in response to this comment?
+We thank referee to point out for the usefulness to mention the wavelengths of the dominant mode.
+We have found that for the lab scale experiments, the wavelength of dominant mode is comparable to half channel width (H).
 
@@
 \textit{
-blah blah
+The final sentence, if I understand this correctly, implies that all the
+available data suggests that only mode 1 (i.e. only the left hand panels in
+fig.4) is actually relevant to real situations. On the other hand, the
+statement at the bottom on p.7 (mode 2 is distinct from KH), and the
+conclusions, seem together to imply that the observed monami are mode 2.
+I became confused at this point and would welcome greater clarity.
+And if mode 2 is not observed in experiments, the extended analysis of its
+character is perhaps not particularly relevant and should be shortened.
 }
 
-Similar response.
+It is true that all experimental data we have found corresponds to a vegetation density for which the unstable region in the R-k space has not split into two, so we are unable to say if flow instability in the lab scale experiments of Ghisalberti and Nepf are due to Mode1 or Mode 2.
+While currently Mode 2 is explicitly obseved we believe that its extended analysis would be crucial for someone who would like to test the prediction by designing suitable experiment.
+
+\item The referee says:
+\textit{
+P.4 +8:  
+It's not clear how a boundary layer thickness can be equated to a velocity
+ratio! This comment also appears in the caption of fig.1.
+Furthermore, the caption states that the profile is from Ghislaberti Nepf
+- case I. This case has H=12.3 cm and hg =9 cm and the caption says that $\delta$
+= 5.02 cm. This implies an $H/\delta$ rather lower than any of the experimental
+data shown in fig.2 (left), which seems inconsistent.
+}
+
+Since the shear stress $U_y$ is continuous across the grass tip we can equate the estimate of shear stress just above $(h_g^+$ the grass with that of just below the grass $h_g^{-}$. 
+Above the grass tip the base flow is a parabolic velocity profile and the estimate of $U_y$ at the grass tip is $U_0/H$, whereas below the grass tip the shear stress can be estimated to be $U_{bl}/\delta$, equating the shear stress above the grass and below the grass gives us $\delta/H = U_{bl}/U_0$.
+I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
+Geophys.
+Res.
+107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ).
+In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$.
+We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
+
+\item The referee says:
+\textit{  Fig 2 $\& 3$:
+In both captions a value for the eddy viscosity is given. Was there any
+physical basis for this value? If not, how was it chosen? And how critical
+are the comparisons between theory and experiment to its value?
+}
+
+The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
+\textit{Water Resources Research 40 (7), 2004} )
 

### 2015-10-23 Paper_JFM/Response/Response.tex
Referee 1 point 2 responded

@@
 
-Do we agree or disagree with the referee? Why? What is our argument and interpretation? How did we revise the manuscript in response to this comment?
+While here we agree with the referee that prediction of threshold flow condition and existence of a mode whose asymptotic behavior is very different from otherwise assumed behavior due to KH instability is one of the major finding of our work.
+We would also like to point out that without that our model successfully predicts threshold Reynolds number and  waving frequency  as well without making any adhoc assumption about velocity profile.
+
+Since all the current experimental data fall into a regime where unstable region in R-k space (Figure-4) has not split into two, we are unable to say if the flow instability is due to Mode1 or Mode2.
 
@@
 \item The referee says:
+
 \textit{  Fig 2 $\& 3$:

### 2015-10-23 Paper_JFM/Response/Response.tex
Referee point 3 responded

@@
 
-While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of th edetailed interpretation.
+While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of th detailed interpretation.
 more explanation.
-We have now modified the manucript to clarify this issue.
+We have now modified the manuscript to clarify this issue.
 The new manuscript now reads: ``blah blah''.
@@
 
-Similar response.
+It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of Gosselin $\&$ E.
+de Langre, where their model predicts enhancement of growth rate in the lock-in range due to less dissipation because of in phase movement of the flow and canopy.
+While in our present analysis we haven't taken into account the flexibility of plant as marine plants are flexible as compared to the terrestrial plants, we do expect deviation from the current prediction when the frequency of flow is similar to that of natural frequency of plant.
 

### 2015-10-23 Paper_JFM/Response/Response.tex
asymptotic solution below the grass added

@@
 \newcommand{\del}{\partial}
+
+%\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
+\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
+\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
+%\newcommand{\bu}{\mathbf{u}}
+%\newcommand{\grad}{\mathbf{\nabla}}
+%\newcommand{\del}{\partial}
+\newcommand{\hg}{h_g}
+%\renewcommand{\Rey}{{R}}
+\newcommand{\Ndg}{\tilde{N}_g}
+\newcommand{\monami}{\textit{monami}}
+\newcommand{\ubl}{U_\text{bl}}
+\newcommand{\words}[1]{\textbf{(#1)~}}
 
@@
 Present the complete solution here. Refer to this solution in our response.
+\textbf{Asymptotic Solution of 3.1 below the grass}
+
+Here we present asymptotic solution of equation 3.1, which should be helpful in clarifying many questions referee have related to the boundary layer near the grass tip.
+Using $ U_0 = (-dP/dx)H^2/\mu$ as velocity scale, $H$ as length scale, we rescale velocity as $U= U_0 \bar{U}$ and $y = H \bar{y}$, the equation 3.1 can be transformed into  
+\begin{equation}
+\begin{split}
+ \bar{U}_{\bar{y}\bar{y}}+1 - \frac{\rho U_0 H}{\mu} C_N d N_g \bar{U}^2 &=0 \\
+ \bar{U}_{\bar{y}\bar{y}}+1 - R \Ndg \bar{U}^2=0
+\end{split}
+\end{equation}
+We wil drop the overbar and for further analysis for simplicity and define drag parameter as $D_g = R\Ndg$
+\begin{equation}
+\begin{split}
+ {U}_{{y}{y}}+1 - D_g{U}^2=0
+\end{split}
+\end{equation}
+We observe that $U = \sqrt{(1/D_g)}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once will give us following equation
+\begin{equation}
+\begin{split}
+ \frac{1}{2} \left( \frac{dU}{dy} \right)^2 +U - \frac{1}{3} D_g U^3 + C = 0
+\end{split}
+\end{equation}
+Where C is a constant of integration, whose can be determined by observing that as $U$ approaches $1/sqrt{D_g}$, $U_y$ approaches zero.
+Applying this condition gives $C = -\frac{2}{3\sqrt{D_g}}$.
+Using the value of $C$, we can further write above equation as
+\begin{equation}
+\begin{split}
+ \frac{dU}{dy} = \sqrt{\frac{2}{3}D_g U^3+\frac{4}{3\sqrt{D_g}}-2U }
+\end{split}
+\end{equation}
+We further use the scaling $U=\frac{1}{\sqrt{D_g}} u $, in the above equation, which simplified into 
+\begin{equation}
+\begin{split}
+ \frac{1}{\sqrt{D_g}}\frac{du}{dy} &= \frac{1}{D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
+ \frac{du}{dy} &= {D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
+ \frac{du}{(u-1)\sqrt{u+2} } &= {D_g^{1/4}}\sqrt{\frac{2}{3}} dy
+\end{split}
+\end{equation}
+Above equation can be further integrated and provides solution in the region of grass
+\begin{equation}
+\begin{split}
+u &= 3 \coth \left(\frac{C-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2 \\
+U &= \frac{1}{\sqrt{D_g}} \left( 3 \coth \left(\frac{C-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2    \right)
+\end{split}
+\end{equation}
+Where $C$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+\textbf{Solution above the grass}
+
+We use similar scaling for flow above the grass to non-dimensionlize the equation 3.1 
+\begin{equation}
+
+\end{equation}
 

### 2015-10-23 Paper_JFM/Response/Response.tex
Asymptic matching done

@@
 \newcommand{\words}[1]{\textbf{(#1)~}}
+\DeclareMathOperator{\sech}{sech}
+\DeclareMathOperator{\csch}{csch}
 
@@
 \centerline{\textbf{Matched asymptotic solution}}
-Present the complete solution here. Refer to this solution in our response.
-\textbf{Asymptotic Solution of 3.1 below the grass}
 
 Here we present asymptotic solution of equation 3.1, which should be helpful in clarifying many questions referee have related to the boundary layer near the grass tip.
+For simplicity of calculation in this particular calculation we assume grass extend from $y=0$ to $y=H$ ie submergance ratio is 0.5.
 Using $ U_0 = (-dP/dx)H^2/\mu$ as velocity scale, $H$ as length scale, we rescale velocity as $U= U_0 \bar{U}$ and $y = H \bar{y}$, the equation 3.1 can be transformed into  
@@
 \end{equation}
-We wil drop the overbar and for further analysis for simplicity and define drag parameter as $D_g = R\Ndg$
+We will drop the overbar and for further analysis for simplicity and define drag parameter as $D_g = R\Ndg$
 \begin{equation}
 \begin{split}
- {U}_{{y}{y}}+1 - D_g{U}^2=0
+ {U}_{{y}{y}}+1 - D_g{U}^2 &=0
 \end{split}
 \end{equation}
+where $D_g = 0$ for $0<y<=1$, grass tip is at $y=0$ and the bottom surface is at $y=-1$, we use the zero shear boundary condition for on both the surface as mentioned in the manuscript.
+
+\textbf{Asymptotic Solution of 3.1 below the grass}
+
 We observe that $U = \sqrt{(1/D_g)}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once will give us following equation
@@
 \end{equation}
-Above equation can be further integrated and provides solution in the region of grass
+Above equation can be integrated which provides solution in the region of grass
 \begin{equation}
 \begin{split}
-u &= 3 \coth \left(\frac{C-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2 \\
-U &= \frac{1}{\sqrt{D_g}} \left( 3 \coth \left(\frac{C-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2    \right)
+u &= 3 \coth \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2 \\
+U &= \frac{1}{\sqrt{D_g}} \left( 3 \coth^2 \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)-2    \right) \hspace{1cm} \text{for $-1<=y<=0$}
+\label{under_grass_sol}
 \end{split}
 \end{equation}
-Where $C$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+Where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+
 \textbf{Solution above the grass}
 
-We use similar scaling for flow above the grass to non-dimensionlize the equation 3.1 
+Above the grass tip the non-dimenlize equation form of equation 3.1 translates in to.
 \begin{equation}
+ U_{yy}+1=0
+\end{equation}
+which can be integrated together with the zero shear boundary condition at top to provided
+\begin{equation}
+ U(y) = y-y^2/2+C_2 \hspace{1cm} \text{for $0<y<1$}
+ \label{above_grass_sol}
+\end{equation}
+Where $C_2$ is constant of integration.
 
+\textbf{Matching the solution obtained by \eqref{under_grass_sol} and \eqref{above_grass_sol} }
+
+The value of $C_1$ and $C_2$ can be determined by matching the solution \eqref{under_grass_sol} and \eqref{above_grass_sol} at the grass tip.
+Matching stress stress at the grass tip results into.
+\begin{equation}
+\frac{3\sqrt{2}}{D_g^{1/4}} \coth(C_1/\sqrt{2}) \csch^2(C_1/\sqrt{2}) = 1;
+\label{C1_eq}
 \end{equation}
-
-The origin of approximately constant velocity with in the grass 
-can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, the dominant balance
-with in vegetation is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
-to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
-numerical simulation of 3.1). Using the dominant balance of pressure gradient and drag term 
-give rise to a constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$, more in depth detailed discussion about such perturbation techniques can be found in the the book by Hinch titles as ``perturbation
-methods''
+Above equation can be solved through some numerical method like bisection etc, here we will do an asymptotic expansion of terms in \eqref{C1_eq} to get an estimate of dependence of $C_1$ on the drag parameter.
+\begin{equation}
+\begin{split}
+\frac{3\sqrt{2}}{D_g^{1/4}} \frac{\sqrt{2}}{C_1} \frac{2}{C_1^2} &= 1 \\
+C_1 = \frac{(12)^{1/3}}{D_g^{1/12}}
+\end{split}
+\end{equation}
+Through constant $C_1$ a natural length scale $\delta$ arises when $\delta D_g^{1.4}$ balances $C_1$ in equation \eqref{under_grass_sol}, which gives $\delta \sim D_g^{-1/3}$ or $\delta \sim (R\Ndg)^{-1/3}$.
+We can also obtain scale of velocity in the boundary layer by substituting $y=0$ in the \eqref{under_grass_sol} and using asymptotic expansion of $\coth(x)$, doing this expansion leads to
+\begin{equation}
+\begin{split}
+U(y)-U_{tip} &\sim \frac{1}{\sqrt{D_g}}\frac{1}{C_1^2}\\
+U(y)-U_{tip} &\sim D_g^{-1/3}\\
+U(y)-U_{tip} &\sim (R\Ndg)^{-1/3}
+\end{split}
+\end{equation}
+near the grass tip. More detailed analysis of similar kind can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ).
 \newline

### 2015-10-23 Paper_JFM/grass.tex
added figure for swaped legend of figure 3

@@
 The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed behaviour -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field as shown in Fig.~\ref{frequency_comparison}.
-Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.
+\textbf{Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.}
 
@@
 In either case, due to such behaviour the most unstable mode transitions discontinuously from Mode 1 to Mode 2.
-All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to say if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode1 or Mode2.
+\textbf{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to say if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode1 or Mode2.}
 

### 2015-10-26 Paper_JFM/Response/Response.tex
Response to referees complete (version 1)

--- 2015-10-23 Paper_JFM/Response/Response.tex
@@
 \documentclass[letterpaper,10pt]{article}
+% \documentclass{jfm}
 \pagenumbering {roman}
@@
 
+\usepackage{natbib}
 \usepackage{graphicx}% Include figure files
 \usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
-\usepackage{ifthen}
-\usepackage{amsthm} % Theorem Formatting
-\usepackage{amssymb}	% Math symbols such as \mathbb
+% \usepackage{bm}% bold math
+% \usepackage{amssymb}	% Math symbols such as \mathbb
 \usepackage{calrsfs}
+\usepackage{fullpage}
 
@@
 \newcommand{\del}{\partial}
+
+%\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
+\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
+\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
+%\newcommand{\bu}{\mathbf{u}}
+%\newcommand{\grad}{\mathbf{\nabla}}
+%\newcommand{\del}{\partial}
+\newcommand{\hg}{h_g}
+\newcommand{\Rey}{{R}}
+\newcommand{\Ndg}{\tilde{N}_g}
+\newcommand{\monami}{\textit{monami}}
+\newcommand{\ubl}{U_\text{bl}}
+\newcommand{\words}[1]{\textbf{(#1)~}}
+\DeclareMathOperator{\sech}{sech}
+\DeclareMathOperator{\csch}{csch}
 
@@
 \textit{
+The authors claim that their approach yields not only the classical Kelvin-Helmholtz instability modes know[sic] for these systems, but also a new type of instability mode.
+The later is thought to be present in practical cases.
+}
+
+The referee did understand the gist of our conclusion, but we would like to take this opportunity to clarify further our position on this matter.
+In summary, we have refrained from making a determination about which mode is observed in experiments and the field.
+It is so because all experimental and field data available to us exists in a parameter regime in which the two modes cannot be distinguished.
+
+We remind the referee that the two modes bifurcate above a critical vegetation density from what appears to be a single mode.
+Below this value, the physical processes underlying the two unstable mode are both active, we have difficulty distinguishing between them.
+
+Our approach could be stated in terms of two questions: (i) do experiments support our mathematical formulation, and (ii) of the two modes, which one is observed in experiments/field?
+The answer to the first question is a likely yes.
+The threshold condition and the waving frequency both are in good agreement with the results of our analysis.
+The question of which mode is physically observed is more subtle, because it is a question of fluid dynamic mechanism.
+We think there is a danger of using superficial flow features to deduce the underlying mechanism.
+Therefore, the approach we took was to distinguish between the modes and the underlying fluid dynamics using the dense grass asymptotic limit.
+A concise and complete presentation of the difference is presented in Table 1 of the manuscript.
+One conclusion that can be drawn from our analysis is a challenge to the traditional view that the observed instability is Kelvin-Helmholtz.
+Instead, we conclude that that there are two modes of instability, and which one is predominantly active in the experiments and the field is still an open question and should be subject of future investigations.
+
+To clarify this discussion in the manuscript, we now say after \S 4:
+
+\noindent
+\textbf{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}
+
+and in the conclusion:
+
+\noindent
+\textbf{
+We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation.
+Since the two modes merge for the experimental parameters, KH may not be assumed to underlie \monami.
+}
+
+\item The referee says:
+\textit{
 The claims of the paper are somehow contradictory, in terms of the application to sea grass beds.
@@
 
-While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of th detailed interpretation.
-more explanation.
-We have now modified the manuscript to clarify this issue.
-The new manuscript now reads: ``blah blah''.
+While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of the detailed interpretation. 
+Experimental observations have revealed that the flow structures underlying \monami persist when the flexible grass mimics are replaced by rigid woden dowels.
+These observations are interpreted as support for the origin of the \monami being a hydrodynamic in nature.
+In this simple picture, the flexible grass simply deforms in response to flow structures.
+In reality, the flexibility of the grass does modify the details of the instability.
+Previous investigations have identified this modification to be in the form of frequency-locking of the hydro-elastic oscillations with the hydrodynamic instability frequency, with different strengths depending on the canopy being terrestrial \citep{Delangre04,Delangre06} or aquatic \cite{Gosselin2009}.
+However, the motion of the vegetation is not \textit{essential} to the existence of the instability, and therefore a minimal model is justified in neglecting it.
+This approach in simplification is similar in spirit to, for example, the one employed by \cite{Gosselin2009}, in which they determine the frequency of the coupled fluid-vegetation oscillations in the absence of the flow, and then use the frequency in a dynamic model in the presence of the flow but assuming the mode shape \textit{ad hoc} instead of solving for it self-consistently.
+Since the precise mode shape is not an essential feature for the phenomenon of frequency locking, such a substitution yields simplification without reducing the utility of the analysis.
+
+However, we also recognize the referee's concern with the neglect of the vegetation flexibility.
+In particular, we have presented a mathematical model for a rigid vegetation bed, and therefore it will be useful to determine boundaries in the parameter space where our results will be quantitatively applicable, instead of merely qualitatively useful.
+To alleviate this concern, we present a back-of-the-envelope analysis determining the criteria for neglecting the vegetation flexibility.
+
+We have now added the following analysis to the discussion:
+
+\textbf{
+We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+For a common seagrass, \textit{Zostera Marina}, the relative density difference is $\Delta \rho /\rho \approx 0.25$, the volume fraction is $V_f \approx 0.1$ and $H=1$ m \citep{Fonseca98}, yielding the buoyancy time scale to be $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
+The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the buoyancy time scale.
+We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. 
+Accounting for these factors  can lead to further interesting behavior \citep{Delangre06,Gosselin2009}.
+Indeed, the case where these time-scales are comparable can lead to interesting behavior \citep{Delangre06}, and motivates further investigation. 
+}
 
@@
 The practical existence of a threshold flow condition for monami (or rather,  for flow instability ) to occur is doubtless.
-As stated page 6, the inviscid KH instability approach does not yield any threshold, of course.
+As stated [on] page 6, the inviscid KH instability approach does not yield any threshold, of course.
 The finding of a threshold by including dissipation and other aspects is no surprise.
@@
 
-While here we agree with the referee that prediction of threshold flow condition and existence of a mode whose asymptotic behavior is very different from otherwise assumed behavior due to KH instability is one of the major finding of our work.
-We would also like to point out that without that our model successfully predicts threshold Reynolds number and  waving frequency  as well without making any adhoc assumption about velocity profile.
+We agree partly with the referee. The existence of a threshold, in our opinion also, is doubtless. 
+But it is not obvious to deduce that including some aspect of dissipation would yield that threshold. 
+We are sure the referee knows that sometimes dissipative effects can be destabilizing \citep{Krechetnikov2007}. 
+In our case, adding viscous (turbulent) dissipation alone does not lead to a threshold in Reynolds number for the shear instability of a tanh profile. 
+Furthermore, adding vegetation-induced drag adds another parameter to the problem, and its interaction with the turbulent dissipation in establishing a threshold is not trivial. 
+In fact, we find very non-trivial dependencies of the threshold criteria on both the turbulent dissipation parameter ($\Rey$) and the vegetation drag parameter ($\Ndg$). 
+Therefore, following our analysis, we disagree with the referee that finding the threshold  by including dissipation or other effects is expected. 
+Only an analysis such as ours can conclusively establish the threshold.
 
-Since all the current experimental data fall into a regime where unstable region in R-k space (Figure-4) has not split into two, we are unable to say if the flow instability is due to Mode1 or Mode2.
+We also agree with the referee that a significant contribution of our analysis is the Mode 2 of instability.
+As explained in item 1 of this response, it is not clear which mode of instability underlies \monami.
+Since the existence of the two modes was unknown before our analysis, Mode 2 as a possible candidate for \monami was never considered. 
+We are unable to determine which of the two modes predominantly underlies \monami.
+Therefore, Mode 2 remains a possible candidate for the synchronous waving of aquatic vegetation.
 
@@
 
-It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of Gosselin $\&$ E.
-de Langre, where their model predicts enhancement of growth rate in the lock-in range due to less dissipation because of in phase movement of the flow and canopy.
+It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of \cite{Gosselin2009}, where their model correlates enhancement of growth rate in the frequency lock-in range.
 While in our present analysis we haven't taken into account the flexibility of plant as marine plants are flexible as compared to the terrestrial plants, we do expect deviation from the current prediction when the frequency of flow is similar to that of natural frequency of plant.
+We address the flexibility of the grass in item 2 of this response.
+
+In addition, we would like to point out that the instability mode found by \cite{Gosselin2009} is assumed to be KH, but not carefully demonstrated to be so.
+Thus, we interpret the referee's comment as saying that the flexibility enhances the instability of whichever mode is observed by \cite{Gosselin2009}. 
+This mode could very well be Mode 2 that we found. 
+Further investigation is needed before this question can be conclusively answered.
+Especially, cleverly designed experiments with observations to distinguish between the two modes are needed in the future.
 
@@
 \textit{
-XXX
+In this paper the authors consider a simple model for a vegetated channel flow and analyse the resulting velocity profile on the basis of a modified Orr-Sommerfeld equation containing both a viscous (Reynolds number) term and a term representing the drag of the vegetated layer at the bottom of the channel.
+They compare some experimental data from an earlier paper from Nepf’s group with the implications of their theory and, via a detailed analysis of the instability, conclude (perhaps not surprisingly) that it differs from traditional KH instability because of the presence of this vegetated drag force.
 }
 
-We say
+We would like to take this opportunity to clarify our position, as well as our understanding of the literature.
+Existing literature on synchronous waving of terrestrial and aquatic vegetation repeatedly implicated KH instability as the underlying mechanism, even in the presence of the vegetation. 
+Therefore, we would invoke the flexibility in the referee's comment to argue that, while it is not surprising \textit{a posteriori}, this possibility was has not been recognized before in the existing literature.
 
 \item The referee says:
-\textit{p.5 +10,11, Although experimentally observed wavelengths are not available, they do arise from the analysis of presumably, so some comment about the values and wheater they seem physically realistic (and/or how they compare with the scales of typical motions in the turbulent flow above the canopy which are believed to be linked with monami) would be instructive.
+\textit{p.1 Fig.1 requires attention. The legend needs to be repositioned for greater
+clarity (perhaps to the left of the velocity profile).}
+
+We have accepted the referee's suggestion. The legend is repositioned. We have also made other changes to the figure to improve its clarity.
+
+\item The referee says:
+\textit{
+p.2, +6 ‘fitted using the experimental observations.
+}
+
+We have corrected the error.
+
+\item The referee says:
+\textit{
+p.3, +4 I think x (hat) needs to be defined here (the unit vector?).
+}
+
+We now introduce $\hat{\bx}$.
+
+\item The referee says:
+\textit{
+p.3, \S 3 It’s not quite clear to me how (3.1) leads to $U(y)$=const. below $y=h_g$. I note (from what follows on p.4) that the shear stress (i.e. $U’(y)$ presumably) is
+matched from upper and lower layers, but this seems insufficient. A little more explanation here would be helpful.
+}
+
+This follows from a matched asymptotic analysis of (3.1).
+For the referee's perusal, we have included a version of it at the end of this response.
+However, while the method and the solution is not unique (because different solutions may differ in accounting for higher orders), it is a standard procedure in perturbation methods with standard textbooks on the topic.
+We now cite one such textbook \citep{Hinch1991} for readers who may be unfamiliar with the topic.
+
+\item The referee says:
+\textit{
+p.4, +1,2 On a related point, why does ‘continuity of shear stresses result in a boundary layer of thickness $\delta$’?
+No natural scale ($\delta$) arises from the solution of (3.1) does it?
+I cannot see qualitatively what distinguishes this `boundary layer' from the entire `shear layer' (indeed, the authors suggest that they are in fact analogous).
+The layer is (looking at fig.
+1) much less well defined than, say, the strong shear layer just above an urban canopy, where the velocity gradient is clearly very much greater than in the boundary layer above or the canopy below.
+}
+
+We refer the referee to the solution of (3.1) we present at the end of this response.
+Many of the features, we assume, will be familiar to readers of the \textit{Journal of Fluid Mechanics}.
+There is a natural boundary layer thickness that arises from the a competition between turbulent viscous stress and vegetation drag.
+It is this thickness we call $\delta$.
+The boundary layer on this thickness scale exists inside the vegetation, but not outside (the velocity profile outside is simply parabolic). 
+The shear established in the boundary layer is, indeed, equal to the shear in the unvegetated region, as a consequence of matching the shear stress.
+However, the shear gradient ($U_{yy}(y)$) in the vegetated region is $O((1/\delta)$ larger than that in the unvegetated region.
+It is this shear gradient that underlies the traditional picture of KH instability in a shear flow.
+Furthermore, the presence of the inflection point in the velocity profile, which is often invoked to suggest the presence of an instability, is absent in our profile.
+Instead the shear gradient switches sign at $y=h_g$.
+
+To clarify the statement in the manuscript about the matching of shear stress determining the boundary layer thickness, we are sure the referee realizes that there are two conditions imposed at $y=h_g$, the continuity of velocity and of shear.
+The continuity of velocity can be imposed by equating the velocity of the parabolic profile outside with the
+
+\item The referee says:
+\textit{
+p.4, +9: It would be better to say ‘the shear layer discussed by Ghisalberti and Nepf (2002, 2004)’.
+}
+
+We thank the referee for pointing this out. We prefer the term ``invoked by Ghisalberti and Nepf (2002, 2004)'', and assume that this revision is in-line with the referee's suggestion.
+
+\item The referee says:
+\textit{p.5 +10,11, Although experimentally observed wavelengths are not available, they do arise from the analysis of presumably, so some comment about the values and whether they seem physically realistic (and/or how they compare with the scales of typical motions in the turbulent flow above the canopy which are believed to be linked with monami) would be instructive.
 }
@@
 \textit{
-The final sentence, if I understand this correctly, implies that all the
-available data suggests that only mode 1 (i.e. only the left hand panels in
-fig.4) is actually relevant to real situations. On the other hand, the
-statement at the bottom on p.7 (mode 2 is distinct from KH), and the
-conclusions, seem together to imply that the observed monami are mode 2.
+The final sentence, if I understand this correctly, implies that all the available data suggests that only mode 1 (i.e.
+only the left hand panels in fig.
+4) is actually relevant to real situations.
+On the other hand, the statement at the bottom on p.7 (mode 2 is distinct from KH), and the conclusions, seem together to imply that the observed monami are mode 2.
 I became confused at this point and would welcome greater clarity.
-And if mode 2 is not observed in experiments, the extended analysis of its
-character is perhaps not particularly relevant and should be shortened.
+And if mode 2 is not observed in experiments, the extended analysis of its character is perhaps not particularly relevant and should be shortened.
 }
 
-It is true that all experimental data we have found corresponds to a vegetation density for which the unstable region in the R-k space has not split into two, so we are unable to say if flow instability in the lab scale experiments of Ghisalberti and Nepf are due to Mode1 or Mode 2.
-While currently Mode 2 is explicitly obseved we believe that its extended analysis would be crucial for someone who would like to test the prediction by designing suitable experiment.
+We understand that our presentation was not perfect, and have now clarified the confusion.
+The fact of the matter is that it is taken for granted based on superficial similarities with the KH instability, that \monami is based on KH instability.
+However, a careful analysis is needed to determine if it is really the case.
+We have presented such an analysis, and find that the picture is not simple.
+We find that there are two kinds of unstable modes for large vegetation density, and they merge together for smaller vegetation density.
+Both mechanisms of instability are presumably active when the two modes appear merged.
+All experiments lie in this regime where the vegetation density is small.
+Therefore, we are unable to distinguish based on the experimental observations and our analysis, if it is Mode 1 or Mode 2 that dominates in region where the two modes merge.
+Future experiments may be cleverly designed to resolve this ambiguity.
+
+To clarify this picture, we now say in the manuscript:
+
+\noindent
+\textbf{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}
+
+and in the conclusion:
+
+\noindent
+\textbf{
+We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation.
+Since the two modes merge for the experimental parameters, KH may not be assumed to underlie \monami.
+}
 
@@
 \textit{
-P.4 +8:  
-It's not clear how a boundary layer thickness can be equated to a velocity
-ratio! This comment also appears in the caption of fig.1.
-Furthermore, the caption states that the profile is from Ghislaberti Nepf
-- case I. This case has H=12.3 cm and hg =9 cm and the caption says that $\delta$
+P.4 +8: It's not clear how a boundary layer thickness can be equated to a velocity ratio!
+This comment also appears in the caption of fig.1.
+Furthermore, the caption states that the profile is from Ghislaberti and Nepf - case I.
+This case has H=12.3 cm and hg =9 cm and the caption says that $\delta$
 = 5.02 cm. This implies an $H/\delta$ rather lower than any of the experimental
@@
 
-Since the shear stress $U_y$ is continuous across the grass tip we can equate the estimate of shear stress just above $(h_g^+$ the grass with that of just below the grass $h_g^{-}$. 
+Since the shear stress $U_y$ is continuous across the grass tip we can equate the scaling estimate of shear stress just above the grass  ($h_g^+$) with that of just below the grass ($h_g^{-}$). 
 Above the grass tip the base flow is a parabolic velocity profile and the estimate of $U_y$ at the grass tip is $U_0/H$, whereas below the grass tip the shear stress can be estimated to be $U_{bl}/\delta$, equating the shear stress above the grass and below the grass gives us $\delta/H = U_{bl}/U_0$.
-I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
+
+We think referee has understood the case-I mentioned for Figure 1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
+Geophys.
+Res.
+107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ).
+In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$.
+We would also like to point out that we have used a logarithmic scale on y-axis in Figure 2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure 2.
+
+\item The referee says:
+\textit{
+p.5, -9: `boundary layer, the' (i.e. singular).
+}
+
+We have accepted the referee's suggestion.
+
+\item The referee says:
+
+\textit{  Fig 2 $\& 3$:
+In both captions a value for the eddy viscosity is given.
+Was there any physical basis for this value?
+If not, how was it chosen?
+And how critical are the comparisons between theory and experiment to its value?
+}
+
+The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure 5 for lab scale experiments of Ghisalberti and Nepf (2004).
+As the referee realized, the exact value of the eddy viscosity varies with depth in the water column.
+But overall, the eddy viscosity arises from the eddies the vegetation generates.
+The scale of velocity for these eddies is the free stream speed ($U \sim 0.1$ m/s), and the size of the eddies is the width of the grass blade ($d \sim 0.01$ m).
+According to Ghisalberti and Nepf (2004), the eddy viscosity is approximately $\mu \approx 0.1 \rho U d = 0.1$ Pa s).
+The Reynolds number $\Rey$ is the only parameter that depends on the eddy viscosity.
+Any uncertainty in the eddy viscosity translates into an uncertainty in the critical Reynolds number for the onset of the instability.
+{\bf Any changes to the manuscript.}
+% (The limited growth of vegetated shear layers. \textit{Water Resources Research 40 (7), 2004} )
+
+We have now revised the manuscript to include this discussion. The manuscript now reads:
+
+\textbf{
+The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.
+}
+
+and we have removed the statement of the value of eddy viscosity from the captions of figures 2 and 3.
+
+\end{enumerate}
+
+\newpage
+\centerline{\textbf{Matched asymptotic solution}}
+
+Here we present an asymptotic solution of equation (3.1) for $N_g \gg 1$, which should be helpful in clarifying many questions the referee has related to the boundary layer near the grass top.
+For simplicity of calculation in this particular calculation we assume the grass to extend from $y=0$ to $y=H$, i.
+e.
+a submergence ratio is 0.5.
+Using $ U_0 = (-dP/dx)H^2/\mu$ as velocity scale, $H$ as length scale, we rescale velocity as $U= U_0 \bar{U}$ and $y = H \bar{y}$. Equation (3.1) is then transformed to  
+\begin{equation}
+\begin{split}
+ \bar{U}_{\bar{y}\bar{y}}+1 - \frac{\rho U_0 H}{\mu} C_N d N_g \bar{U}^2 =0, \quad &\text{ for } y<0\\
+ \bar{U}_{\bar{y}\bar{y}}+1 - R \Ndg \bar{U}^2=0, \quad &\text{ for } y>0.
+\end{split}
+\end{equation}
+For notational convenience, we will drop the overbar and define a drag parameter as $D_g = R\Ndg$ for $-1<y\le 0$, $D_g = 0$ for $0<y\le 1$, grass top is at $y=0$ and the bottom surface is at $y=-1$.
+Equation (3.1) now reads
+\begin{equation}
+\begin{split}
+ {U}_{{y}{y}}+1 - D_g{U}^2 &=0.
+\end{split}
+\end{equation}
+We use the zero shear boundary condition on both the boundaries ($y=-1,1$) as mentioned in the manuscript.
+
+\vspace{2mm}
+\noindent
+\textbf{Asymptotic Solution of 3.1 below the grass}
+
+\noindent
+We observe that $U = \sqrt{(1/D_g)}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once, we get
+\begin{equation}
+\begin{split}
+ \frac{1}{2} \left( \frac{dU}{dy} \right)^2 +U - \frac{1}{3} D_g U^3 + C = 0
+\end{split}
+\label{eqn:above}
+\end{equation}
+Where C is a constant of integration, which can be determined by observing that as $U$ approaches $1/\sqrt{D_g}$, $U_y$ approaches zero.
+Applying this condition gives $C = -\dfrac{2}{3\sqrt{D_g}}$.
+Using the value of $C$, \eqref{eqn:above} becomes
+\begin{equation}
+\begin{split}
+ \frac{dU}{dy} = \sqrt{\frac{2}{3}D_g U^3+\frac{4}{3\sqrt{D_g}}-2U }
+\end{split}
+\end{equation}
+We further use the substitution $U=\dfrac{1}{\sqrt{D_g}} u $ in the above equation, which simplifies it to
+\begin{equation}
+\begin{split}
+%  \frac{1}{\sqrt{D_g}}\frac{du}{dy} &= \frac{1}{D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
+%  \frac{du}{dy} &= {D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
+ \frac{du}{(u-1)\sqrt{u+2} } &= {D_g^{1/4}}\sqrt{\frac{2}{3}} dy
+\end{split}
+\end{equation}
+Above equation can be integrated to obtain the solution in the vegetated region as
+\begin{equation}
+\begin{split}
+u &= 3 \coth \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2 \\
+U &= \frac{1}{\sqrt{D_g}} \left( 3 \coth^2 \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)-2    \right)
+\label{under_grass_sol}
+\end{split}
+ \hspace{1cm} \text{for $-1\le y\le 0$,}
+\end{equation}
+where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+
+\vspace{2mm}
+\noindent
+\textbf{Solution above the grass}
+
+\noindent
+Above the grass, the dimensionless form of equation (3.1) translates to
+\begin{equation}
+ U_{yy}+1=0
+\end{equation}
+which can be integrated together with the zero shear boundary condition at top to provided
+\begin{equation}
+ U(y) = y-y^2/2+C_2 \hspace{1cm} \text{for $0<y<1$}
+ \label{above_grass_sol}
+\end{equation}
+Where $C_2$ is constant of integration.
+
+\vspace{2mm}
+\noindent
+\textbf{Matching the solution obtained by \eqref{under_grass_sol} and \eqref{above_grass_sol} }
+
+\noindent
+The value of $C_1$ and $C_2$ can be determined by matching the solution \eqref{under_grass_sol} and \eqref{above_grass_sol} at the grass tip.
+Matching the shear stress at the grass top results in
+\begin{equation}
+\frac{3\sqrt{2}}{D_g^{1/4}} \coth\left(\dfrac{C_1}{\sqrt{2}}\right) \csch^2\left(\dfrac{C_1}{\sqrt{2}}\right) = 1;
+\label{C1_eq}
+\end{equation}
+Above equation can be solved numerically, but here we will solve it asymptotically for $D_g \gg 1$.
+We expect $C_1 \ll 1$ in this limit, and the terms in \eqref{C1_eq} may be expanded in this limit to get
+\begin{equation}
+\begin{split}
+\frac{3\sqrt{2}}{D_g^{1/4}} \left( \frac{\sqrt{2}}{C_1} \right) \left( \frac{2}{C_1^2} \right) &= 1 \\
+C_1 = \frac{(12)^{1/3}}{D_g^{1/12}}
+\end{split}
+\end{equation}
+The boundary layer length scale may be derived from examining the scale of $y$ for which the argument of \eqref{under_grass_sol} changes asymptotic order.
+This examination implies $C_1 \sim O(y D_g^{1/4})$ or $y = O(D_g^{-1/3})$.
+The scale for velocity in the boundary layer is the value of \eqref{under_grass_sol} at $y=0$.
+\begin{align}
+ U_\text{top} &=  \frac{1}{\sqrt{D_g}} \left( 3 \coth^2 \left(\frac{C_1}{\sqrt{2}}  \right)-2    \right) \\
+              &\approx  \frac{1}{\sqrt{D_g}} \left( \frac{6}{C_1^2} -2    \right) \\
+              &\approx \frac{1}{\sqrt{D_g}} \frac{6D_g^{1/6}}{12^{2/3}} \\
+              &\approx \left( \dfrac{3}{2} \right)^{1/3} D_g^{-1/3}.
+\end{align}
+
+% 
+% 
+% Through constant $C_1$ a natural length scale $\delta$ arises when $\delta D_g^{1/4}$ balances $C_1$ in equation \eqref{under_grass_sol}, which gives $\delta \sim D_g^{-1/3}$ or $\delta \sim (R\Ndg)^{-1/3}$.
+We can also obtain scale of velocity in the boundary layer by substituting $y=0$ in the \eqref{under_grass_sol} and using asymptotic expansion of $\coth(x)$, doing this expansion leads to
+% \begin{equation}
+% \begin{split}
+% U(y)-U_{tip} &\sim \frac{1}{\sqrt{D_g}}\frac{1}{C_1^2}\\
+% U(y)-U_{tip} &\sim D_g^{-1/3}\\
+% U(y)-U_{tip} &\sim (R\Ndg)^{-1/3}
+% \end{split}
+% \end{equation}
+% near the grass tip. 
+We have employed one of the many possible methods to understand the structure of the solution of (3.1); other methods can also be used to reach the same conclusions.
+In th manuscript we have presented the outline of an essentially identical method.
+For a general treatment of the topic, we refer the referee to the book by \cite{Hinch1991}. % (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ).
+
+\newpage
+\bibliography{Grass}{}
+\bibliographystyle{jfm}
+% 
+% \newpage
+% \textbf{ About the Origin of boundary layer $\delta$}
+% 
+% \centerline{Ravi's new edits}
+% 
+% In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip.
+Again for in-depth discussion about such problem can be found in the book of Hinch.
+% 
+% \textbf{About the estimated value of $H/\delta$ from Figure-1}
+% 
+% I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
 Geophys.
@@
 We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
-
-\item The referee says:
-
-\textit{  Fig 2 $\& 3$:
-In both captions a value for the eddy viscosity is given. Was there any
-physical basis for this value? If not, how was it chosen? And how critical
-are the comparisons between theory and experiment to its value?
-}
-
-The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
-\textit{Water Resources Research 40 (7), 2004} )
-
-\end{enumerate}
-
-\newpage
-\centerline{\textbf{Matched asymptotic solution}}
-Present the complete solution here. Refer to this solution in our response.
-
-The origin of approximately constant velocity with in the grass 
-can be understood by well known perturbation theory (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ). In case of sufficiently high grass density, the dominant balance
-with in vegetation is between pressure gradient term $-\frac{dP}{dx}$ and the drag term experienced by the flow $-\rho C_N dNg U^2$ whereas the viscous term contribute very little 
-to (3.1), so we can treat the viscous term $\mu U_{yy}$ with in the grass as a small perturbation term in equation (3.1) ( $U(y)$ is approximately constant as can be seen from experimental data as well from
-numerical simulation of 3.1). Using the dominant balance of pressure gradient and drag term 
-give rise to a constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$, more in depth detailed discussion about such perturbation techniques can be found in the the book by Hinch titles as ``perturbation
-methods''
-\newline
-\textbf{ About the Origin of boundary layer $\delta$}
-
-\centerline{Ravi's new edits}
-
-In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip.
-Again for in-depth discussion about such problem can be found in the book of Hinch.
-\newline
-\textbf{About the estimated value of $H/\delta$ from Figure-1}
-\newline
-I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J.
-Geophys.
-Res.
-107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ).
-In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$.
-We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
-\newline
-\textbf{Basis for eddy viscosity}
-\newline
-The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
+% 
+% \textbf{Basis for eddy viscosity}
+% 
+% The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers.
 \textit{Water Resources Research 40 (7), 2004} )

### 2015-10-26 Paper_JFM/grass.tex
Response to referees complete (version 1)

--- 2015-10-20 Paper_JFM/grass.tex
@@
 \newcommand{\ReyNdg}{{\Rey\Ndg}}
+\newcommand{\revise}[1]{{\color{blue} {#1}}}
 
@@
 Experimental observations for the onset and frequency of waving compare well with model predictions for the instability onset criteria and the imaginary part of the complex growth rate respectively, but experiments lie in a parameter regime where the two modes can not be distinguished.
-% The inclusion of vegetation drag
-differentiates our mechanism from the previous linear stability analyses of \monami.
+% The inclusion of vegetation drag differentiates our mechanism from the previous linear stability analyses of \monami.
 \end{abstract}
@@
 \begin{figure}
-\centerline{\includegraphics[scale=0.95]{Grass_Base_Nepf_shear} \includegraphics[scale=0.95]{Grass_shear_scale}}
+\centerline{\includegraphics[scale=0.8]{Grass_Base_Nepf_shear} \includegraphics[scale=0.95]{Grass_shear_scale}}
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
@@
 Parameters from experiments reported by \cite{Ghisal02} to exhibit or suppress synchronous waving are also included in the left panel. 
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+% In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
 }
@@
 A comparison of the steady flow profile from the solution of ~\eqref{base_equ} with experimental measurements by \cite{Nepf04} is shown in Fig.~\ref{basicflow}.
-The profile $U(y)$ has three distinct regions.
+The profile $U(y)$ has three distinct regions\revise{, as determined by a matched asymtptotic analysis \citep{Hinch1991}}.
 Within the vegetation, it is approximately uniform with $ U(y) \approx U_g = \sqrt{\frac{-dP/dx}{\rho C_N dN_g}}$, arising from a balance of the drag with the pressure gradient. 
@@
 A numerical estimate of $\delta$ (estimated as $U/U_y$ at $y=\hg$) is compared with this prediction in Fig. \ref{basicflow} (inset).
-We identify the boundary layer to be analogous to the shear layer discussed by Ghisalberti and Nepf \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
+We identify the boundary layer to be analogous to the shear layer \revise{invoked} by Ghisalberti and Nepf \citep{Ghisal02,Nepf04} in the previous explanation of \monami.
 This dependence of $\delta$ on $N_g$ gives us a way to systematically investigate the effect of the shear layer thickness on the instability mechanism.
@@
 \begin{figure}
-\centerline{\includegraphics[]{new_graph_freq}}
+\centerline{\includegraphics[width=6cm]{new_graph_freq}}
 \caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.
 The experimental data in the inset is obtained from \cite{Ghisal02} and \cite{Vivoni98}. 
-In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
+% In order to estimate the $\Rey$ for these experiments, a representative value of $\mu=0.1$ Pa~s was assumed.
 }
@@
 The frequency (Im$(\sigma)$) of the fastest growing mode also agrees well with observed behaviour -- frequency of \monami, maxima in the velocity spectra, and frequency of vortex passage in lab scale experiments \citep{Ghisal02} -- for cases where the vegetation was sufficiently dense to be modeled by a continuum drag field as shown in Fig.~\ref{frequency_comparison}.
-Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.
+\revise{ The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.}
+\revise{Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.}
 
@@
 In either case, due to such behaviour the most unstable mode transitions discontinuously from Mode 1 to Mode 2.
-All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to say if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode1 or Mode2.
+\revise{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}
 
@@
 
-To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} near the grass-top using the the boundary layer scalings $\eta = y/(\delta/H)$, 
+To further understand the mechanism of Mode 1, we rescale \eqref{Orr-somerfield} near the grass-top using the boundary layer scalings $\eta = y/(\delta/H)$, 
 $U(y) = (\delta/H)\bar{U}(\eta)$ and $k = (H/\delta) \bar{k}$.
@@
 
+\revise{
+We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
+For a common seagrass, \textit{Zostera Marina}, the relative density difference is $\Delta \rho /\rho \approx 0.25$, the volume fraction is $V_f \approx 0.1$ and $H=1$ m \citep{Fonseca98}, yielding the buoyancy time scale to be $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
+The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the buoyancy time scale.
+We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. 
+Accounting for these factors  can lead to further interesting behavior \citep{Delangre06,Gosselin2009}.
+Indeed, the case where these time-scales are comparable can lead to interesting behavior \citep{Delangre06}, and motivates further investigation.
+}
+
 \section{Conclusion}
-% We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
-% For a common seagrass, \textit{Zostera Marina}, the relative density difference $\Delta \rho /\rho \approx 0.25$, the volume fraction $V_f \approx 0.1$ and $H=1$ m \cite{Fonseca98}, yielding the buoyancy time scale $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
-% The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the hydrodynamic time scale.
-% We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident.
-Accounting for these factors  can lead to further interesting behavior \cite{Delangre06}.
-% Indeed, the case where these time-scales are comparable can lead to interesting behavior\cite{Delangre06}, and motivates further investigation. 
 In conclusion, we show that the hydrodynamic instability underlying \monami ~differs from the traditional KH due to the presence of the vegetation drag. 
@@
 Furthermore, our linear stability analysis reveals two modes, namely Mode 1 with a mechanism similar to KH and Mode 2 with characteristics distinctly different from KH. 
+\revise{
+We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation.
+Since the two modes merge for the experimental parameters, KH may not be assumed to underlie \monami.
+}
 % While further investigation is needed to understand the sensitivity of the results to the various simplifying assumptions made in our model, the agreement with experiments is encouraging.

### 2015-10-27 Paper_JFM/Response/Response.tex
My response added with red colored

@@
 \usepackage{fullpage}
-
+\usepackage{color}
 \usepackage[above]{placeins}
@@
 One conclusion that can be drawn from our analysis is a challenge to the traditional view that the observed instability is Kelvin-Helmholtz.
-Instead, we conclude that that there are two modes of instability, and which one is predominantly active in the experiments and the field is still an open question and should be subject of future investigations.
+Instead, we conclude that that there are two modes of instability, and which one
+is predominantly active in the experiments and the field is still an open question and should be subject of future investigations.
 
@@
 It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of \cite{Gosselin2009}, where their model correlates enhancement of growth rate in the frequency lock-in range.
-While in our present analysis we haven't taken into account the flexibility of plant as marine plants are flexible as compared to the terrestrial plants, we do expect deviation from the current prediction when the frequency of flow is similar to that of natural frequency of plant.
+While in our present analysis we haven't taken into account the flexibility of plant as marine plants are flexible as compared to the terrestrial plants, we do expect deviation from the current prediction when the frequency of flow {\color{red}instability} is similar to that of natural frequency of plant.
 We address the flexibility of the grass in item 2 of this response.
@@
 To clarify the statement in the manuscript about the matching of shear stress determining the boundary layer thickness, we are sure the referee realizes that there are two conditions imposed at $y=h_g$, the continuity of velocity and of shear.
-The continuity of velocity can be imposed by equating the velocity of the parabolic profile outside with the
+{\color{red} \textbf{We can remove this sentence I guess.}
+The continuity of velocity can be imposed by equating the velocity of the parabolic profile outside with the }
 
@@
 \end{equation}
-where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass. {\color{red} Deep inside grass the 
+solution saturates to $U(y)=\frac{1}{\sqrt{D_g}}$ independent of the value of $C_1$ due to the dominant balance between drag and pressure gradient.}
 

### 2015-10-27 Paper_JFM/grass.tex
changed a word in manuscript (bold letter) and adjusted figure's aspect ratios

@@
 \begin{figure}
-\centerline{\includegraphics[scale=0.8]{Grass_Base_Nepf_shear} \includegraphics[scale=0.95]{Grass_shear_scale}}
+\centerline{\includegraphics[scale=0.9]{Grass_Base_Nepf_shear} \includegraphics[scale=0.87]{Grass_shear_scale}}
 % \includegraphics[width=15cm]{Grass_Base_Nepf_shear}
@@
 Here we present a mathematical model for the linear instability that accounts for these effects.
-Although \monami ~is manifest in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
+Although \monami ~is \textbf{manifested} in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \citep{Ghisal02,Nepf06}. 
@@
 \begin{figure}
-\centerline{\includegraphics[width=6cm]{new_graph_freq}}
+%\centerline{\includegraphics[width=6cm]{new_graph_freq}}
+\centerline{\includegraphics[scale=0.9]{new_graph_freq}}
 \caption{Comparison of experimental observations of the experimentally measured dominant frequency $f_o$ (in Hz) with the predictions $f_p=\text{Im}(\sigma)$ from the solution of ~\eqref{Orr-somerfield}.

### 2015-10-28 Paper_JFM/Response/Response.tex
Just changed th to the
Revisions and Response complete

--- 2015-10-26 Paper_JFM/Response/Response.tex
@@
 \DeclareMathOperator{\csch}{csch}
+\newcommand{\revise}[1]{{\color{blue} {#1}}}
 
@@
 \noindent
-\textbf{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}
+\revise{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}
 
@@
 \noindent
-\textbf{
+\revise{
 We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation.
@@
 While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of the detailed interpretation. 
-Experimental observations have revealed that the flow structures underlying \monami persist when the flexible grass mimics are replaced by rigid woden dowels.
-These observations are interpreted as support for the origin of the \monami being a hydrodynamic in nature.
+Experimental observations have revealed that the flow structures underlying \monami~ persist when the flexible grass mimics are replaced by rigid woden dowels.
+These observations are interpreted as support for the origin of the \monami~ being a hydrodynamic in nature.
 In this simple picture, the flexible grass simply deforms in response to flow structures.
@@
 However, we also recognize the referee's concern with the neglect of the vegetation flexibility.
-In particular, we have presented a mathematical model for a rigid vegetation bed, and therefore it will be useful to determine boundaries in the parameter space where our results will be quantitatively applicable, instead of merely qualitatively useful.
+In particular, we have presented a mathematical model for a rigid vegetation bed, and therefore it will be useful to determine boundaries in the parameter space where our results will be quantitatively applicable, instead of merely being qualitatively useful.
 To alleviate this concern, we present a back-of-the-envelope analysis determining the criteria for neglecting the vegetation flexibility.
@@
 
-\textbf{
+\revise{
 We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
@@
 We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. 
-Accounting for these factors  can lead to further interesting behavior \citep{Delangre06,Gosselin2009}.
-Indeed, the case where these time-scales are comparable can lead to interesting behavior \citep{Delangre06}, and motivates further investigation. 
+Accounting for these factors can lead to further interesting behavior \citep{Delangre06,Gosselin2009}, and motivates further investigation. 
 }
@@
 Furthermore, adding vegetation-induced drag adds another parameter to the problem, and its interaction with the turbulent dissipation in establishing a threshold is not trivial. 
-In fact, we find very non-trivial dependencies of the threshold criteria on both the turbulent dissipation parameter ($\Rey$) and the vegetation drag parameter ($\Ndg$). 
+In fact, we find very non-trivial dependency of the threshold criteria on both the turbulent dissipation parameter ($\Rey$) and the vegetation drag parameter ($\Ndg$). 
 Therefore, following our analysis, we disagree with the referee that finding the threshold  by including dissipation or other effects is expected. 
@@
 As explained in item 1 of this response, it is not clear which mode of instability underlies \monami.
-Since the existence of the two modes was unknown before our analysis, Mode 2 as a possible candidate for \monami was never considered. 
+Since the existence of the two modes was unknown before our analysis, Mode 2 as a possible candidate for \monami~ was never considered. 
 We are unable to determine which of the two modes predominantly underlies \monami.
@@
 It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of \cite{Gosselin2009}, where their model correlates enhancement of growth rate in the frequency lock-in range.
-While in our present analysis we haven't taken into account the flexibility of plant as marine plants are flexible as compared to the terrestrial plants, we do expect deviation from the current prediction when the frequency of flow is similar to that of natural frequency of plant.
-We address the flexibility of the grass in item 2 of this response.
+While in our present analysis we have not taken into account the flexibility of plant, marine plants are flexible as compared to the terrestrial plants.
+Buoyancy as a restoring force for marine vegetation dominates over the restoring action of the vegetation bending stiffness.
+For this reason, we do expect deviation from the current prediction when the frequency of flow instability is comparable to that of natural (buoyancy) frequency of grass bed.  
+We address this issue in item 2 of this response.
 
@@
 Existing literature on synchronous waving of terrestrial and aquatic vegetation repeatedly implicated KH instability as the underlying mechanism, even in the presence of the vegetation. 
-Therefore, we would invoke the flexibility in the referee's comment to argue that, while it is not surprising \textit{a posteriori}, this possibility was has not been recognized before in the existing literature.
+Therefore, we would invoke the flexibility in the referee's comment to argue that, while it is not surprising \textit{a posteriori}, this possibility has not been recognized before in the existing literature.
 
@@
 We refer the referee to the solution of (3.1) we present at the end of this response.
-Many of the features, we assume, will be familiar to readers of the \textit{Journal of Fluid Mechanics}.
+Many features of this solution, we assume, will be familiar to readers of the \textit{Journal of Fluid Mechanics}.
 There is a natural boundary layer thickness that arises from the a competition between turbulent viscous stress and vegetation drag.
@@
 The shear established in the boundary layer is, indeed, equal to the shear in the unvegetated region, as a consequence of matching the shear stress.
-However, the shear gradient ($U_{yy}(y)$) in the vegetated region is $O((1/\delta)$ larger than that in the unvegetated region.
+However, the shear gradient ($U_{yy}(y)$) in the vegetated region is $O(1/\delta)$ larger than that in the unvegetated region.
 It is this shear gradient that underlies the traditional picture of KH instability in a shear flow.
@@
 
-To clarify the statement in the manuscript about the matching of shear stress determining the boundary layer thickness, we are sure the referee realizes that there are two conditions imposed at $y=h_g$, the continuity of velocity and of shear.
-The continuity of velocity can be imposed by equating the velocity of the parabolic profile outside with the
+To clarify the statement in the manuscript about the matching of shear stress determining the boundary layer thickness, we are sure the referee realizes that there are two conditions imposed at $y=h_g$, the continuity of velocity and the shear stress.
+However, only the continuity of shear stress influences the boundary layer thickness, as can be seen from the solution we present at the end of this response.
 
@@
 
-We thank referee to point out for the usefulness to mention the wavelengths of the dominant mode.
-We have found that for the lab scale experiments, the wavelength of dominant mode is comparable to half channel width (H).
+We thank the referee for pointing out for the usefulness mentioning the wavelengths of the dominant mode. 
+For parameters corresponding to the lab scale experiments ($4<H/\delta<10$ in figure 2 of the manuscript), the dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$).
+We have revied the manuscript to say:
+
+\revise{Figure 4 also shows that the predicted dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$)}
 
@@
 
-\textbf{
+\revise{
 The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.
@@
 \newpage
-\centerline{\textbf{Matched asymptotic solution}}
+\centerline{\large \textbf{Matched asymptotic solution}}
+~\vspace{1mm}
 
 Here we present an asymptotic solution of equation (3.1) for $N_g \gg 1$, which should be helpful in clarifying many questions the referee has related to the boundary layer near the grass top.
-For simplicity of calculation in this particular calculation we assume the grass to extend from $y=0$ to $y=H$, i.
+For clarity of presentation in this particular calculation we assume the grass to extend from $y=0$ to $y=H$, i.
 e.
-a submergence ratio is 0.5.
-Using $ U_0 = (-dP/dx)H^2/\mu$ as velocity scale, $H$ as length scale, we rescale velocity as $U= U_0 \bar{U}$ and $y = H \bar{y}$. Equation (3.1) is then transformed to  
+the submergence ratio is 0.5.
+Using $H$ as length scale, and $U_0 = (-dP/dx)H^2/\mu$ as velocity scale, we rescale $U= U_0 \bar{U}$ and $y-H = H \bar{y}$. 
+Equation (3.1) is then transformed to  
 \begin{equation}
 \begin{split}
- \bar{U}_{\bar{y}\bar{y}}+1 - \frac{\rho U_0 H}{\mu} C_N d N_g \bar{U}^2 =0, \quad &\text{ for } y<0\\
- \bar{U}_{\bar{y}\bar{y}}+1 - R \Ndg \bar{U}^2=0, \quad &\text{ for } y>0.
+ \bar{U}_{\bar{y}\bar{y}}+1 - \Rey \Ndg \bar{U}^2 =0, \quad &\text{ for } -1<\bar{y}<0\\
+ \bar{U}_{\bar{y}\bar{y}}+1 =0, \quad &\text{ for } 1>\bar{y}>0.
 \end{split}
@@
 \textbf{Asymptotic Solution of 3.1 below the grass}
+\vspace{1mm}
 
 \noindent
-We observe that $U = \sqrt{(1/D_g)}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once, we get
+We observe that $U = \sqrt{1/D_g}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once, we get
 \begin{equation}
 \begin{split}
- \frac{1}{2} \left( \frac{dU}{dy} \right)^2 +U - \frac{1}{3} D_g U^3 + C = 0
+ \frac{1}{2} \left( \frac{dU}{dy} \right)^2 +U - \frac{1}{3} D_g U^3 + C = 0,
 \end{split}
@@
 \end{equation}
-Where C is a constant of integration, which can be determined by observing that as $U$ approaches $1/\sqrt{D_g}$, $U_y$ approaches zero.
+where $C$ is a constant of integration, which can be determined by observing that as $U$ approaches $1/\sqrt{D_g}$, $U_y$ approaches zero.
 Applying this condition gives $C = -\dfrac{2}{3\sqrt{D_g}}$.
@@
 where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass.
+Deeper inside grass ($h_g-y \gg \delta$) the solution saturates to $U(y)=\frac{1}{\sqrt{D_g}}$ independent of the value of $C_1$ due to the dominant balance between drag and pressure gradient.
 
@@
 \textbf{Solution above the grass}
+\vspace{1mm}
 
@@
 \textbf{Matching the solution obtained by \eqref{under_grass_sol} and \eqref{above_grass_sol} }
+\vspace{1mm}
 
@@
 \frac{3\sqrt{2}}{D_g^{1/4}} \left( \frac{\sqrt{2}}{C_1} \right) \left( \frac{2}{C_1^2} \right) &= 1 \\
-C_1 = \frac{(12)^{1/3}}{D_g^{1/12}}
+C_1 = \frac{(12)^{1/3}}{D_g^{1/12}},
 \end{split}
 \end{equation}
+consistent with our expectation.
 The boundary layer length scale may be derived from examining the scale of $y$ for which the argument of \eqref{under_grass_sol} changes asymptotic order.
@@
 We have employed one of the many possible methods to understand the structure of the solution of (3.1); other methods can also be used to reach the same conclusions.
-In th manuscript we have presented the outline of an essentially identical method.
+In the manuscript we have presented the outline of an essentially identical method.
 For a general treatment of the topic, we refer the referee to the book by \cite{Hinch1991}. % (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ).

### 2015-10-28 Paper_JFM/grass-revise.tex
Revisions and Response complete

--- 2015-10-27 Paper_JFM/grass.tex
@@
 Here we present a mathematical model for the linear instability that accounts for these effects.
-Although \monami ~is \textbf{manifested} in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
+Although \monami ~is manifested in the motion of the grass, the drag exerted by the vegetation on the flow is central to the hypothesized instability. 
 The instability and the resulting flow structures persist in lab experiments even when flexible grass mimics are replaced by rigid dowels \citep{Ghisal02,Nepf06}. 
@@
 \revise{ The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.}
-\revise{Although the experimentally observed \monami ~wavelengths are not available for comparison, the predicted wavelength for lab scale experiments \citep{Ghisal02} are found to be $O(H)$.}
+\revise{Figure 4 also shows that the predicted dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$)}
 
@@
 We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. 
-Accounting for these factors  can lead to further interesting behavior \citep{Delangre06,Gosselin2009}.
-Indeed, the case where these time-scales are comparable can lead to interesting behavior \citep{Delangre06}, and motivates further investigation.
+Accounting for these factors  can lead to further interesting behavior \citep{Delangre06,Gosselin2009}, and motivates further investigation.
 }

### 2015-10-28 JFM_Revision/grass.tex
Revisions sent to JFM

--- 2015-10-28 Paper_JFM/grass-revise.tex
@@
 \newcommand{\ReyNdg}{{\Rey\Ndg}}
-\newcommand{\revise}[1]{{\color{blue} {#1}}}
+% \newcommand{\revise}[1]{{\color{blue} {#1}}}
+\newcommand{\revise}[1]{{{#1}}}
 
@@
 \revise{ The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.}
-\revise{Figure 4 also shows that the predicted dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$)}
+\revise{Figure 2 also shows that the predicted dimensionless wavenumber of the dominant mode for parameters corresponding to the lab scale experiments ($4<H/\delta<10$) is very close to unity, and therefore the dimensional wavelength is approximately $2\pi H$.}
 
@@
 % Note the spaces between the initials
-\bibliography{Grass}{}
-\bibliographystyle{jfm}
-%\bibliography{Grass}{}
-% \bibliographystyle{plain}
-%\bibliographystyle{aipauth4-1}
+\input{grass.bbl}
+% \bibliography{Grass}{}
+% \bibliographystyle{jfm}
 \end{document}

### 2015-10-28 Paper_JFM/Response/Response1.tex
Revisions sent to JFM

\documentclass[letterpaper,10pt]{article}
% \documentclass{jfm}
\pagenumbering {roman}
\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{natbib}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
% \usepackage{bm}% bold math
% \usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\usepackage{fullpage}
\usepackage{color}
\usepackage[above]{placeins}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

%\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
%\newcommand{\bu}{\mathbf{u}}
%\newcommand{\grad}{\mathbf{\nabla}}
%\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
\newcommand{\ubl}{U_\text{bl}}
\newcommand{\words}[1]{\textbf{(#1)~}}
\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\csch}{csch}
\newcommand{\revise}[1]{{\color{blue} {#1}}}

\begin{document}

\centerline{Response to Referee 1}
We thank the referee for reviewing our work and providing their valuable feedback. 
We are pleased that the referee finds our contribution interesting, and the relevance of our work to fundamental and practical fluid mechanics certain.
Here we respond to the questions raised and the feedback provided.

\begin{enumerate}
\item The referee says:
\textit{
The authors claim that their approach yields not only the classical Kelvin-Helmholtz instability modes know[sic] for these systems, but also a new type of instability mode. The later is thought to be present in practical cases.
}

The referee did understand the gist of our conclusion, but we would like to take this opportunity to clarify further our position on this matter. In summary, we have refrained from making a determination about which mode is observed in experiments and the field. It is so because all experimental and field data available to us exists in a parameter regime in which the two modes cannot be distinguished.

We remind the referee that the two modes bifurcate above a critical vegetation density from what appears to be a single mode. Below this value, the physical processes underlying the two unstable mode are both active, we have difficulty distinguishing between them.

Our approach could be stated in terms of two questions: (i) do experiments support our mathematical formulation, and (ii) of the two modes, which one is observed in experiments/field? The answer to the first question is a likely yes. The threshold condition and the waving frequency both are in good agreement with the results of our analysis. The question of which mode is physically observed is more subtle, because it is a question of fluid dynamic mechanism. We think there is a danger of using superficial flow features to deduce the underlying mechanism. Therefore, the approach we took was to distinguish between the modes and the underlying fluid dynamics using the dense grass asymptotic limit. A concise and complete presentation of the difference is presented in Table 1 of the manuscript. One conclusion that can be drawn from our analysis is a challenge to the traditional view that the observed instability is Kelvin-Helmholtz. Instead, we conclude that that there are two modes of instability, and which one 
is predominantly active in the experiments and the field is still an open question and should be subject of future investigations.

To clarify this discussion in the manuscript, we now say after \S 4:

\noindent
\revise{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}

and in the conclusion:

\noindent
\revise{
We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation. Since the two modes merge for the experimental parameters, KH may not be assumed to underlie \monami.
}

\item The referee says:
\textit{
The claims of the paper are somehow contradictory, in terms of the application to sea grass beds. The title focuses on “monami in a sea grass bed” whereas motion of the sea grass is excluded from the analysis.  For the conclusions of the paper to be applicable to sea grass motion two things would be needed : (a) a much larger discussion on the possible effects of the unstable modes on the motion of the grass, (b) a discussion on the potential role of sea grass flexibility in the instability of the coupled flow-grass system. As it is, the  title of the paper should rather  be “linear stability of the flow over a submerged sea grass bed”. In other terms, I am not sure that the results of the present paper (except the existence of a threshold) have consequences on the motion of the sea grass.
}

While we agree whole-heartedly with the spirit of the referee's comment, we disagree with some of the detailed interpretation. 
Experimental observations have revealed that the flow structures underlying \monami~ persist when the flexible grass mimics are replaced by rigid woden dowels.
These observations are interpreted as support for the origin of the \monami~ being a hydrodynamic in nature.
In this simple picture, the flexible grass simply deforms in response to flow structures.
In reality, the flexibility of the grass does modify the details of the instability.
Previous investigations have identified this modification to be in the form of frequency-locking of the hydro-elastic oscillations with the hydrodynamic instability frequency, with different strengths depending on the canopy being terrestrial \citep{Delangre04,Delangre06} or aquatic \cite{Gosselin2009}.
However, the motion of the vegetation is not \textit{essential} to the existence of the instability, and therefore a minimal model is justified in neglecting it.
This approach in simplification is similar in spirit to, for example, the one employed by \cite{Gosselin2009}, in which they determine the frequency of the coupled fluid-vegetation oscillations in the absence of the flow, and then use the frequency in a dynamic model in the presence of the flow but assuming the mode shape \textit{ad hoc} instead of solving for it self-consistently. 
Since the precise mode shape is not an essential feature for the phenomenon of frequency locking, such a substitution yields simplification without reducing the utility of the analysis.

However, we also recognize the referee's concern with the neglect of the vegetation flexibility.
In particular, we have presented a mathematical model for a rigid vegetation bed, and therefore it will be useful to determine boundaries in the parameter space where our results will be quantitatively applicable, instead of merely being qualitatively useful.
To alleviate this concern, we present a back-of-the-envelope analysis determining the criteria for neglecting the vegetation flexibility.

We have now added the following analysis to the discussion:

\revise{
We now test the assumption of an undeformable grass bed due to the dominant restoring force of buoyancy, using the criteria that the buoyancy time scale be much shorter than the hydrodynamic time scale $H/U_0$.
For a common seagrass, \textit{Zostera Marina}, the relative density difference is $\Delta \rho /\rho \approx 0.25$, the volume fraction is $V_f \approx 0.1$ and $H=1$ m \citep{Fonseca98}, yielding the buoyancy time scale to be $\sqrt{\rho H/V_f \Delta \rho g} \approx 2$ s.
The hydrodynamic time scale assuming $U_0 \approx 0.1$ m/s is 10 s, and therefore longer than the buoyancy time scale.
We have neither accounted for the pre-factors appearing in the scaling argument, or considered cases when the time-scale separation is not so evident. 
Accounting for these factors can lead to further interesting behavior \citep{Delangre06,Gosselin2009}, and motivates further investigation. 
}

\item The referee says:
\textit{
The practical existence of a threshold flow condition for monami (or rather,  for flow instability ) to occur is doubtless.  As stated [on] page 6, the inviscid KH instability approach does not yield any threshold, of course. The finding of a threshold by including dissipation and other aspects is no surprise. The main contribution of the paper seems then to be the second mode. This brings back to the issue of the relevance of this new mode to the motion of the sea grass.
}

We agree partly with the referee. The existence of a threshold, in our opinion also, is doubtless. 
But it is not obvious to deduce that including some aspect of dissipation would yield that threshold. 
We are sure the referee knows that sometimes dissipative effects can be destabilizing \citep{Krechetnikov2007}. 
In our case, adding viscous (turbulent) dissipation alone does not lead to a threshold in Reynolds number for the shear instability of a tanh profile. 
Furthermore, adding vegetation-induced drag adds another parameter to the problem, and its interaction with the turbulent dissipation in establishing a threshold is not trivial. 
In fact, we find very non-trivial dependency of the threshold criteria on both the turbulent dissipation parameter ($\Rey$) and the vegetation drag parameter ($\Ndg$). 
Therefore, following our analysis, we disagree with the referee that finding the threshold  by including dissipation or other effects is expected. 
Only an analysis such as ours can conclusively establish the threshold.

We also agree with the referee that a significant contribution of our analysis is the Mode 2 of instability.
As explained in item 1 of this response, it is not clear which mode of instability underlies \monami.
Since the existence of the two modes was unknown before our analysis, Mode 2 as a possible candidate for \monami~ was never considered. 
We are unable to determine which of the two modes predominantly underlies \monami.
Therefore, Mode 2 remains a possible candidate for the synchronous waving of aquatic vegetation.

\item The referee says:
\textit{
The comparison with the case of wind-canopy interactions is interesting but does not seem consistent with the analysis found in the reference [1]  below, where the instability seems to be more dominantly KH type in water  (when flexibility is assumed).  This would need some discussion, at least to point out the differences.
}

It is true that inclusion of flexibility can give interesting results and we thank referee for drawing our attention to the interesting work of \cite{Gosselin2009}, where their model correlates enhancement of growth rate in the frequency lock-in range. 
While in our present analysis we have not taken into account the flexibility of plant, marine plants are flexible as compared to the terrestrial plants.
Buoyancy as a restoring force for marine vegetation dominates over the restoring action of the vegetation bending stiffness.
For this reason, we do expect deviation from the current prediction when the frequency of flow instability is comparable to that of natural (buoyancy) frequency of grass bed.  
We address this issue in item 2 of this response.

In addition, we would like to point out that the instability mode found by \cite{Gosselin2009} is assumed to be KH, but not carefully demonstrated to be so.
Thus, we interpret the referee's comment as saying that the flexibility enhances the instability of whichever mode is observed by \cite{Gosselin2009}. 
This mode could very well be Mode 2 that we found. 
Further investigation is needed before this question can be conclusively answered.
Especially, cleverly designed experiments with observations to distinguish between the two modes are needed in the future.

\end{enumerate}

\bibliography{Grass}{}
\bibliographystyle{jfm}
% 
% \newpage
% \textbf{ About the Origin of boundary layer $\delta$}
% 
% \centerline{Ravi's new edits}
% 
% In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip. Again for in-depth discussion about such problem can be found in the book of Hinch.   
% 
% \textbf{About the estimated value of $H/\delta$ from Figure-1}
% 
% I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J. Geophys. Res. 107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ). In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$. We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
% 
% \textbf{Basis for eddy viscosity}
% 
% The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers. \textit{Water Resources Research 40 (7), 2004} )

\end{document}

### 2015-10-28 Paper_JFM/Response/Response2.tex
Revisions sent to JFM

\documentclass[letterpaper,10pt]{article}
% \documentclass{jfm}
\pagenumbering {roman}
\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{natbib}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
% \usepackage{bm}% bold math
% \usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\usepackage{fullpage}
\usepackage{color}
\usepackage[above]{placeins}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

%\newcommand{\bx}{{\boldsymbol{\hat{x}}}}
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
%\newcommand{\bu}{\mathbf{u}}
%\newcommand{\grad}{\mathbf{\nabla}}
%\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
\newcommand{\ubl}{U_\text{bl}}
\newcommand{\words}[1]{\textbf{(#1)~}}
\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\csch}{csch}
\newcommand{\revise}[1]{{\color{blue} {#1}}}

\begin{document}

\centerline{Response to Referee 2}
We thank the referee for reviewing our work and providing their valuable feedback. 
We are glad that we were able to engage the referee's attention with our presentation.
Here we respond in detail to his/her comments.
\begin{enumerate}
\item The referee says:
\textit{
In this paper the authors consider a simple model for a vegetated channel flow and analyse the resulting velocity profile on the basis of a modified Orr-Sommerfeld equation containing both a viscous (Reynolds number) term and a term representing the drag of the vegetated layer at the bottom of the channel. They compare some experimental data from an earlier paper from Nepf’s group with the implications of their theory and, via a detailed analysis of the instability, conclude (perhaps not surprisingly) that it differs from traditional KH instability because of the presence of this vegetated drag force.
}

We would like to take this opportunity to clarify our position, as well as our understanding of the literature.
Existing literature on synchronous waving of terrestrial and aquatic vegetation repeatedly implicated KH instability as the underlying mechanism, even in the presence of the vegetation. 
Therefore, we would invoke the flexibility in the referee's comment to argue that, while it is not surprising \textit{a posteriori}, this possibility has not been recognized before in the existing literature.

\item The referee says:
\textit{p.1 Fig.1 requires attention. The legend needs to be repositioned for greater
clarity (perhaps to the left of the velocity profile).}

We have accepted the referee's suggestion. The legend is repositioned. We have also made other changes to the figure to improve its clarity.

\item The referee says:
\textit{
p.2, +6 ‘fitted using the experimental observations.
}

We have corrected the error.

\item The referee says:
\textit{
p.3, +4 I think x (hat) needs to be defined here (the unit vector?).
}

We now introduce $\hat{\bx}$.

\item The referee says:
\textit{
p.3, \S 3 It’s not quite clear to me how (3.1) leads to $U(y)$=const. below $y=h_g$. I note (from what follows on p.4) that the shear stress (i.e. $U’(y)$ presumably) is
matched from upper and lower layers, but this seems insufficient. A little more explanation here would be helpful.
}

This follows from a matched asymptotic analysis of (3.1). For the referee's perusal, we have included a version of it at the end of this response. However, while the method and the solution is not unique (because different solutions may differ in accounting for higher orders), it is a standard procedure in perturbation methods with standard textbooks on the topic. We now cite one such textbook \citep{Hinch1991} for readers who may be unfamiliar with the topic.

\item The referee says:
\textit{
p.4, +1,2 On a related point, why does ‘continuity of shear stresses result in a boundary layer of thickness $\delta$’? No natural scale ($\delta$) arises from the solution of (3.1) does it? I cannot see qualitatively what distinguishes this `boundary layer' from the entire `shear layer' (indeed, the authors suggest that they are in fact analogous). The layer is (looking at fig. 1) much less well defined than, say, the strong shear layer just above an urban canopy, where the velocity gradient is clearly very much greater than in the boundary layer above or the canopy below.
}

We refer the referee to the solution of (3.1) we present at the end of this response.
Many features of this solution, we assume, will be familiar to readers of the \textit{Journal of Fluid Mechanics}.
There is a natural boundary layer thickness that arises from the a competition between turbulent viscous stress and vegetation drag.
It is this thickness we call $\delta$.
The boundary layer on this thickness scale exists inside the vegetation, but not outside (the velocity profile outside is simply parabolic). 
The shear established in the boundary layer is, indeed, equal to the shear in the unvegetated region, as a consequence of matching the shear stress.
However, the shear gradient ($U_{yy}(y)$) in the vegetated region is $O(1/\delta)$ larger than that in the unvegetated region.
It is this shear gradient that underlies the traditional picture of KH instability in a shear flow.
Furthermore, the presence of the inflection point in the velocity profile, which is often invoked to suggest the presence of an instability, is absent in our profile.
Instead the shear gradient switches sign at $y=h_g$.

To clarify the statement in the manuscript about the matching of shear stress determining the boundary layer thickness, we are sure the referee realizes that there are two conditions imposed at $y=h_g$, the continuity of velocity and the shear stress.
However, only the continuity of shear stress influences the boundary layer thickness, as can be seen from the solution we present at the end of this response.

\item The referee says:
\textit{
p.4, +9: It would be better to say ‘the shear layer discussed by Ghisalberti and Nepf (2002, 2004)’.
}

We thank the referee for pointing this out. We prefer the term ``invoked by Ghisalberti and Nepf (2002, 2004)'', and assume that this revision is in-line with the referee's suggestion.

\item The referee says:
\textit{p.5 +10,11, Although experimentally observed wavelengths are not available, they do arise from the analysis of presumably, so some comment about the values and whether they seem physically realistic (and/or how they compare with the scales of typical motions in the turbulent flow above the canopy which are believed to be linked with monami) would be instructive.
}

We thank the referee for pointing out for the usefulness mentioning the wavelengths of the dominant mode. 
For parameters corresponding to the lab scale experiments ($4<H/\delta<10$ in figure 2 of the manuscript), the dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$). We have revied the manuscript to say:

\revise{Figure 2 also shows that the predicted dimensionless wavenumber of the dominant mode for parameters corresponding to the lab scale experiments ($4<H/\delta<10$) is very close to unity, and therefore the dimensional wavelength is approximately $2\pi H$.}

\item The referee says:
\textit{
The final sentence, if I understand this correctly, implies that all the available data suggests that only mode 1 (i.e. only the left hand panels in fig. 4) is actually relevant to real situations. On the other hand, the statement at the bottom on p.7 (mode 2 is distinct from KH), and the conclusions, seem together to imply that the observed monami are mode 2. I became confused at this point and would welcome greater clarity. And if mode 2 is not observed in experiments, the extended analysis of its character is perhaps not particularly relevant and should be shortened.
}

We understand that our presentation was not perfect, and have now clarified the confusion. The fact of the matter is that it is taken for granted based on superficial similarities with the KH instability, that \monami is based on KH instability. However, a careful analysis is needed to determine if it is really the case. We have presented such an analysis, and find that the picture is not simple. We find that there are two kinds of unstable modes for large vegetation density, and they merge together for smaller vegetation density. Both mechanisms of instability are presumably active when the two modes appear merged. All experiments lie in this regime where the vegetation density is small. Therefore, we are unable to distinguish based on the experimental observations and our analysis, if it is Mode 1 or Mode 2 that dominates in region where the two modes merge. Future experiments may be cleverly designed to resolve this ambiguity.

To clarify this picture, we now say in the manuscript:

\noindent
\revise{All experimental data we have found corresponds to a vegetation density for which the unstable region in the $\Rey-k$ space has not split into two, so we are unable to determine if flow instability in the lab scale experiments \citep{Ghisal02} are due to Mode 1 or Mode 2.}

and in the conclusion:

\noindent
\revise{
We are unable to determine based on observations, and therefore have refrained from identifying, which mode is observed in experiments and it still remains a subtle question and subject of future investigation. Since the two modes merge for the experimental parameters, KH may not be assumed to underlie \monami.
}

\item The referee says:
\textit{
P.4 +8: It's not clear how a boundary layer thickness can be equated to a velocity ratio! This comment also appears in the caption of fig.1. Furthermore, the caption states that the profile is from Ghislaberti and Nepf - case I. This case has H=12.3 cm and hg =9 cm and the caption says that $\delta$
= 5.02 cm. This implies an $H/\delta$ rather lower than any of the experimental
data shown in fig.2 (left), which seems inconsistent.
}

Since the shear stress $U_y$ is continuous across the grass tip we can equate the scaling estimate of shear stress just above the grass  ($h_g^+$) with that of just below the grass ($h_g^{-}$). 
Above the grass tip the base flow is a parabolic velocity profile and the estimate of $U_y$ at the grass tip is $U_0/H$, whereas below the grass tip the shear stress can be estimated to be $U_{bl}/\delta$, equating the shear stress above the grass and below the grass gives us $\delta/H = U_{bl}/U_0$.

We think referee has understood the case-I mentioned for Figure 1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J. Geophys. Res. 107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ). In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$. We would also like to point out that we have used a logarithmic scale on y-axis in Figure 2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure 2.

\item The referee says:
\textit{
p.5, -9: `boundary layer, the' (i.e. singular).
}

We have accepted the referee's suggestion.

\item The referee says:

\textit{  Fig 2 $\& 3$:
In both captions a value for the eddy viscosity is given. Was there any physical basis for this value? If not, how was it chosen? And how critical are the comparisons between theory and experiment to its value?
}

The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure 5 for lab scale experiments of Ghisalberti and Nepf (2004). As the referee realized, the exact value of the eddy viscosity varies with depth in the water column. But overall, the eddy viscosity arises from the eddies the vegetation generates. The scale of velocity for these eddies is the free stream speed ($U \sim 0.1$ m/s), and the size of the eddies is the width of the grass blade ($d \sim 0.01$ m). According to Ghisalberti and Nepf (2004), the eddy viscosity is approximately $\mu \approx 0.1 \rho U d = 0.1$ Pa s). The Reynolds number $\Rey$ is the only parameter that depends on the eddy viscosity. Any uncertainty in the eddy viscosity translates into an uncertainty in the critical Reynolds number for the onset of the instability. {\bf Any changes to the manuscript.}
% (The limited growth of vegetated shear layers. \textit{Water Resources Research 40 (7), 2004} )

We have now revised the manuscript to include this discussion. The manuscript now reads:

\revise{
The measured values of eddy viscosity vary with depth in the water column \citep{Nepf04}; we use $\mu= 0.1$ Pa s as a representative value from this range for comparison with experiments.
}

and we have removed the statement of the value of eddy viscosity from the captions of figures 2 and 3.

\end{enumerate}

\newpage
\centerline{\large \textbf{Matched asymptotic solution}}
~\vspace{1mm}

Here we present an asymptotic solution of equation (3.1) for $N_g \gg 1$, which should be helpful in clarifying many questions the referee has related to the boundary layer near the grass top. For clarity of presentation in this particular calculation we assume the grass to extend from $y=0$ to $y=H$, i. e. the submergence ratio is 0.5. 
Using $H$ as length scale, and $U_0 = (-dP/dx)H^2/\mu$ as velocity scale, we rescale $U= U_0 \bar{U}$ and $y-H = H \bar{y}$. 
Equation (3.1) is then transformed to  
\begin{equation}
\begin{split}
 \bar{U}_{\bar{y}\bar{y}}+1 - \Rey \Ndg \bar{U}^2 =0, \quad &\text{ for } -1<\bar{y}<0\\
 \bar{U}_{\bar{y}\bar{y}}+1 =0, \quad &\text{ for } 1>\bar{y}>0.
\end{split}
\end{equation}
For notational convenience, we will drop the overbar and define a drag parameter as $D_g = R\Ndg$ for $-1<y\le 0$, $D_g = 0$ for $0<y\le 1$, grass top is at $y=0$ and the bottom surface is at $y=-1$. Equation (3.1) now reads
\begin{equation}
\begin{split}
 {U}_{{y}{y}}+1 - D_g{U}^2 &=0.
\end{split}
\end{equation}
We use the zero shear boundary condition on both the boundaries ($y=-1,1$) as mentioned in the manuscript.

\vspace{2mm}
\noindent
\textbf{Asymptotic Solution of 3.1 below the grass}
\vspace{1mm}

\noindent
We observe that $U = \sqrt{1/D_g}$ is a stationary point of above equation, where both $U_{yy}=0$ and $U_y =0 $. By multiplying with $U_y$ and integrating once, we get
\begin{equation}
\begin{split}
 \frac{1}{2} \left( \frac{dU}{dy} \right)^2 +U - \frac{1}{3} D_g U^3 + C = 0,
\end{split}
\label{eqn:above}
\end{equation}
where $C$ is a constant of integration, which can be determined by observing that as $U$ approaches $1/\sqrt{D_g}$, $U_y$ approaches zero. Applying this condition gives $C = -\dfrac{2}{3\sqrt{D_g}}$. Using the value of $C$, \eqref{eqn:above} becomes 
\begin{equation}
\begin{split}
 \frac{dU}{dy} = \sqrt{\frac{2}{3}D_g U^3+\frac{4}{3\sqrt{D_g}}-2U }
\end{split}
\end{equation}
We further use the substitution $U=\dfrac{1}{\sqrt{D_g}} u $ in the above equation, which simplifies it to
\begin{equation}
\begin{split}
%  \frac{1}{\sqrt{D_g}}\frac{du}{dy} &= \frac{1}{D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
%  \frac{du}{dy} &= {D_g^{1/4}}\sqrt{\frac{2}{3}} \sqrt{ u^3+2-3u } \\
 \frac{du}{(u-1)\sqrt{u+2} } &= {D_g^{1/4}}\sqrt{\frac{2}{3}} dy
\end{split}
\end{equation}
Above equation can be integrated to obtain the solution in the vegetated region as
\begin{equation}
\begin{split}
u &= 3 \coth \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)^2-2 \\
U &= \frac{1}{\sqrt{D_g}} \left( 3 \coth^2 \left(\frac{C_1-y D_g^{1/4}}{\sqrt{2}}  \right)-2    \right)
\label{under_grass_sol}
\end{split}
 \hspace{1cm} \text{for $-1\le y\le 0$,}
\end{equation}
where $C_1$ is constant of integration, which can be found by matching the shear stress applied by the flow above the grass at the tip of the grass. Deeper inside grass ($h_g-y \gg \delta$) the solution saturates to $U(y)=\frac{1}{\sqrt{D_g}}$ independent of the value of $C_1$ due to the dominant balance between drag and pressure gradient.

\vspace{2mm}
\noindent
\textbf{Solution above the grass}
\vspace{1mm}

\noindent
Above the grass, the dimensionless form of equation (3.1) translates to
\begin{equation}
 U_{yy}+1=0
\end{equation}
which can be integrated together with the zero shear boundary condition at top to provided
\begin{equation}
 U(y) = y-y^2/2+C_2 \hspace{1cm} \text{for $0<y<1$}
 \label{above_grass_sol}
\end{equation}
Where $C_2$ is constant of integration.

\vspace{2mm}
\noindent
\textbf{Matching the solution obtained by \eqref{under_grass_sol} and \eqref{above_grass_sol} }
\vspace{1mm}

\noindent
The value of $C_1$ and $C_2$ can be determined by matching the solution \eqref{under_grass_sol} and \eqref{above_grass_sol} at the grass tip.
Matching the shear stress at the grass top results in
\begin{equation}
\frac{3\sqrt{2}}{D_g^{1/4}} \coth\left(\dfrac{C_1}{\sqrt{2}}\right) \csch^2\left(\dfrac{C_1}{\sqrt{2}}\right) = 1;
\label{C1_eq}
\end{equation}
Above equation can be solved numerically, but here we will solve it asymptotically for $D_g \gg 1$. We expect $C_1 \ll 1$ in this limit, and the terms in \eqref{C1_eq} may be expanded in this limit to get
\begin{equation}
\begin{split}
\frac{3\sqrt{2}}{D_g^{1/4}} \left( \frac{\sqrt{2}}{C_1} \right) \left( \frac{2}{C_1^2} \right) &= 1 \\
C_1 = \frac{(12)^{1/3}}{D_g^{1/12}},
\end{split}
\end{equation}
consistent with our expectation.
The boundary layer length scale may be derived from examining the scale of $y$ for which the argument of \eqref{under_grass_sol} changes asymptotic order. This examination implies $C_1 \sim O(y D_g^{1/4})$ or $y = O(D_g^{-1/3})$. The scale for velocity in the boundary layer is the value of \eqref{under_grass_sol} at $y=0$.
\begin{align}
 U_\text{top} &=  \frac{1}{\sqrt{D_g}} \left( 3 \coth^2 \left(\frac{C_1}{\sqrt{2}}  \right)-2    \right) \\
              &\approx  \frac{1}{\sqrt{D_g}} \left( \frac{6}{C_1^2} -2    \right) \\
              &\approx \frac{1}{\sqrt{D_g}} \frac{6D_g^{1/6}}{12^{2/3}} \\
              &\approx \left( \dfrac{3}{2} \right)^{1/3} D_g^{-1/3}.
\end{align}

% 
% 
% Through constant $C_1$ a natural length scale $\delta$ arises when $\delta D_g^{1/4}$ balances $C_1$ in equation \eqref{under_grass_sol}, which gives $\delta \sim D_g^{-1/3}$ or $\delta \sim (R\Ndg)^{-1/3}$. We can also obtain scale of velocity in the boundary layer by substituting $y=0$ in the \eqref{under_grass_sol} and using asymptotic expansion of $\coth(x)$, doing this expansion leads to 
% \begin{equation}
% \begin{split}
% U(y)-U_{tip} &\sim \frac{1}{\sqrt{D_g}}\frac{1}{C_1^2}\\
% U(y)-U_{tip} &\sim D_g^{-1/3}\\
% U(y)-U_{tip} &\sim (R\Ndg)^{-1/3}
% \end{split}
% \end{equation}
% near the grass tip. 
We have employed one of the many possible methods to understand the structure of the solution of (3.1); other methods can also be used to reach the same conclusions.
In the manuscript we have presented the outline of an essentially identical method.
For a general treatment of the topic, we refer the referee to the book by \cite{Hinch1991}. % (Perturbation Methods by E.J. Hinch, Cambridge Texts in Applied Mathematics ).

\newpage
\bibliography{Grass}{}
\bibliographystyle{jfm}
% 
% \newpage
% \textbf{ About the Origin of boundary layer $\delta$}
% 
% \centerline{Ravi's new edits}
% 
% In order to match shear stress at the grass tip applied by the fluid above the grass, the flow with in the grass increases from an approximately constant velocity $U_g = \sqrt{\frac{-dp/dx}{\rho C_N d N_g}}$ in a region $\delta$ to a value so that both velocity and shear stress are continuous across the grass tip. Again for in-depth discussion about such problem can be found in the book of Hinch.   
% 
% \textbf{About the estimated value of $H/\delta$ from Figure-1}
% 
% I think reviewer have understood the case-I mentioned for Figure-1 as the one from 2002 paper ( Mixing layer and coherent structures in vegetated aquatic flows, \textit{J. Geophys. Res. 107} ), whereas we are referring to the case-I from the 2004 paper ( The limited growth of vegetated shear layers, \textit{Water Resource Research 40(7)} ). In this case the channel width is $41cm$, implying $H=20.5 cm$ hence $H/\delta=4.02$. We would also like to point out that we have used log scale on y-axis in Figure-2, and indeed $H/\delta \approx 5$ for all the experimental observation shown in Figure-2.
% 
% \textbf{Basis for eddy viscosity}
% 
% The constant eddy viscosity of 0.1 Pa s is chosen based on eddy viscosity profile shown in Figure-5 for lab scale experiments of Ghisalberti and Nepf (The limited growth of vegetated shear layers. \textit{Water Resources Research 40 (7), 2004} )

\end{document}

### 2015-10-28 Paper_JFM/Response/Response.tex

@@
 
-\revise{Figure 4 also shows that the predicted dimensionless wavenumber of dominant mode is very close to unity, and therefore the dimensional wavelength is approximately half channel width ($2\pi H$)}
+\revise{Figure 2 also shows that the predicted dimensionless wavenumber of the dominant mode for parameters corresponding to the lab scale experiments ($4<H/\delta<10$) is very close to unity, and therefore the dimensional wavelength is approximately $2\pi H$.}
 

### 2017-03-06 Paper/cover.tex
Add logo to cover letter
Archiving this folder

@@
 \usepackage{palatino}
+\usepackage{graphicx}
+\usepackage{fancyhdr}
+\usepackage{geometry}
+
+\geometry{headheight = 0.7in,tmargin=1in}
+\fancypagestyle{firstpage}{\fancyhf{}\fancyhead[L]{\includegraphics[height=0.9in, keepaspectratio=true]{logo}}}
+\fancypagestyle{plain}{\fancyhf{}\fancyhead[L]{\includegraphics[height=0.9in, keepaspectratio=true]{logo}}}
+\pagestyle{plain}
+\renewcommand{\headrulewidth}{0pt}
 \name{Shreyas Mandre}
@@
 
-Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Amala Mahadevan, and myself for publication in  Physical Review Letters.
+Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Mahesh Bandi, Amala Mahadevan, and myself for publication in  Physical Review Letters.
 This manuscript addresses the synchronous coherent oscillations of aquatic vegetation in a steady flow, known as {\it monami}.

### 2017-03-06 Paper_JFM/cover.tex
Archiving this folder

--- 2017-03-06 Paper/cover.tex
@@
 \begin{letter}{
-Editor \\
-Physical Review Letters \\
-1 Research Road, \\
-Ridge, NY 11961-2701 
+Prof. Detlef Lohse, \\
+Editor, Journal of Fluid Mechanics, \\
+Cambridge University Press, \\
+Cambridge CB2 8RU UK
 }
-\opening{Dear Editor,}
+\opening{Dear Prof. Lohse,}
 
-Please consider a manuscript entitled, ``{\it Monami} as an oscillatory hydrodynamic instability in a submerged sea grass bed'' by Ravi Singh, Mahesh Bandi, Amala Mahadevan, and myself for publication in  Physical Review Letters.
+Please consider a manuscript entitled, ``Linear stability analysis for \textit{monami} in a submerged sea grass bed'' by Ravi Singh, Mahesh Bandi, Amala Mahadevan, and myself for publication in Journal of Fluid Mechanics.
 This manuscript addresses the synchronous coherent oscillations of aquatic vegetation in a steady flow, known as {\it monami}.
 The phenomenon is analogous to the more commonly observed synchronous waving of terrestrial grasses in the wind.
-In the manuscript, we present a mathematical model for the onset of these oscillations.
+In the manuscript, we present a mathematical model for the onset of these oscillations using linear stability analysis.
+Our manuscript is 10 pages long, therefore we wish to be considered under the \textit{Rapids} section of the journal.
 
@@
 It has been previously postulated from experiments that this shear layer is susceptible to  Kelvin-Helmholtz instability, which drives the waving of the grass.
-Here, we challenge this notion by showing that the presence of drag due to the grass generates a different instability mechanism, distinct from Kelvin-Helmholtz instability.
-This new instability has characteristics consistent with the experimental observations, and our theory makes verifiable predictions about the threshold velocity for its onset.
+However, a careful treatment of this instability is missing in the literature.
+Here, we provide such a treatment, and show that the presence of drag due to the grass also generates an instability mechanism, in addition to and distinct from a modified Kelvin-Helmholtz instability.
 
@@
 Our manuscript is one of the few theoretical treatments on the physical mechanism behind these oscillations.
+We believe that the topic will be of interest because of its pertinence to the commonplace terrestrial analogy and the importance of the seagrass ecosystems.
+We have endeavored to write it in a form that is accessible to the general physicist.
 
@@
 
-We believe that the topic will be of interest because of its pertinence to the commonplace terrestrial analogy and the importance of the seagrass ecosystems.
-We have endeavored to write it in a form that is accessible to the general physicist.
-
 Should you have any questions, please do not hesitate to contact me.

### 2017-03-06 Paper_JFM/jfm-template.tex
Archiving this folder

--- 2015-03-03 Paper_JFM/jfm-instructions.tex
@@
-% This is file JFM2esam.tex
-% first release v1.0, 20th October 1996
-%       release v1.01, 29th October 1996
-%       release v1.1, 25th June 1997
-%       release v2.0, 27th July 2004
-%       release v3.0, 16th July 2014
-%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
-% Copyright (C) 1996, 1997, 2014 Cambridge University Press
+\documentclass{jfm}
 
-\documentclass{jfm}
-\usepackage{graphicx}
-\usepackage{epstopdf, epsfig}
+\begin{document}
 
@@
 
-\shorttitle{Guidelines for authors}
-\shortauthor{A. N. Other, H.-C. Smith and J. Q. Public}
+\shorttitle{Guidelines for authors} %for header on odd pages
+\shortauthor{A. N. Other et al} %for header on even pages
 
@@
 
-\author{Alan N. Other\aff{1}
+\author
+ {
+ Alan N. Other\aff{1}
   \corresp{\email{jfm@damtp.cam.ac.uk}},
-  H. - C. Smith\aff{1}
- \and J. Q.  Public\aff{2}}
+  H. - C. Smith\aff{1, 2}
+  \and 
+  J. Q.  Public\aff{2}
+  }
 
-\affiliation{\aff{1}Department of Chemical Engineering, University of America,
-Somewhere, IN 12345, USA
-\aff{2}Department of Aerospace and Mechanical Engineering, University of
-Camford, Academic Street, Camford CF3 5QL, UK}
-
-\begin{document}
+\affiliation
+{
+\aff{1}
+Department of Chemical Engineering, University of America, Somewhere, IN 12345, USA
+\aff{2}
+Department of Aerospace and Mechanical Engineering, University of Camford, Academic Street, Camford CF3 5QL, UK
+}
 
@@
 \end{abstract}
-
-\begin{keywords}
-Authors should not enter keywords on the manuscript, as these must be chosen by the author during the online submission process and will then be added during the typesetting process (see http://journals.cambridge.org/data/\linebreak[3]relatedlink/jfm-\linebreak[3]keywords.pdf for the full list)
-\end{keywords}
 
@@
 \subsection{JFM Rapids}
-{\it JFM Rapids} is devoted to the rapid publication of short, high-impact papers across the full range of fluid mechanics.
-Manuscripts submitted as {\it Rapids}  must be strictly 10 or fewer printed pages, and must be submitted in {\LaTeX} using the jfm.cls class file, so as to ensure that they meet the page limit and to expedite their production.
+{\it JFM Rapids} is an evolution of the Fast Track scheme, and is devoted to the rapid publication of short, high-impact papers across the full range of fluid mechanics.
+{\it Rapids}  must occupy 10 or fewer printed pages, and must be submitted in {\LaTeX} using the jfm.cls class file, so as to ensure that they meet the page limit and to expedite their production.
 The principal and over-riding objective is to publish papers of the highest scientific quality.
-
 Once a paper is submitted, referees are asked to provide reports with a short turnaround.
 In order to be accepted for publication in {\it JFM Rapids}, such papers must require only minor revisions to improve clarity, usually without recourse to a second round of refereeing.
-In this case, and at the discretion of the editor, some additional pages may be allowed to address specific points raised by the referees, such as the addition of an extra figure or some explanatory text.
-
-In cases where the editor, guided by the referees, judges that a paper has merit but requires substantial revision that will require significant refereeing, a decision of `revise and resubmit' will be given.
-On re-submission, such papers will be handled as standard JFM papers and if accepted will not subsequently appear as {\it JFM Rapids}.
+Where the editor, guided by the referees, judges that a paper has merit but requires substantial revision that will require significant refereeing, a decision of `revise and resubmit' will be given.
+On re-submission, such papers will be handled as standard JFM papers and will not subsequently appear as {\it JFM Rapids}.
 
 {\it Rapids} will be published online within one month of final acceptance.
-They will appear within a designated section on the Journal of Fluid Mechanics website.
-Each {\it Rapid} will be cited and indexed as a JFM article but with a distinctive {\it Rapids} identifier, and will be assigned to a JFM volume.
-{\it Rapids} will not be included in the regular print versions of JFM, and accordingly they incur no colour figure charges.
+They will appear within a new devoted section on the Journal of Fluid Mechanics website, which has been re-designed as part of this development.
+Each {\it Rapid} will still be cited and indexed as a JFM article but with a distinctive {\it Rapids} identifier, and will be assigned to a JFM volume.
+{\it Rapids} will not be included in the regular print versions of JFM, but a collection of all {\it Rapids} will be available to order in print at the end of each year.
+{\it JFM Rapids} will have free colour and for the first year of operation will be freely accessible to all.
 
@@
 Every effort should be taken to avoid figures that run over more than one page.
-Figures submitted in colour will appear online in colour but, with the exception of {\it JFM Rapids}, all figures will be printed in black and white unless authors specify during submission that figures should be printed in colour, for which there is a charge of \pounds200 per figure with a cap of \pounds1000 per article plus taxes where applicable (colour is free for {\it JFM Rapids}).
+Figures submitted in colour will appear online in colour but, with the exception of {\it JFM Rapids}, all figures will be printed in black and white unless authors specify during submission that figures should be printed in colour, for which there is a charge of \pounds300 for the first page and \pounds150 for each subsequent page that uses colour (colour is free for {\it JFM Rapids}).
 Note that separate figures for online and print will {\bf not} be accepted and it is the author's responsibility to ensure that if a figure is to appear in colour online only, that same figure will still be meaningful when printed in black and white (for example, do not rely upon colours to distinguish lines if those colours will just appear as similar shades of grey when printed).
@@
 \begin{figure}
-  \centerline{\includegraphics{trapped}}% Images in 100% size
-  \caption{Trapped-mode wavenumbers, $kd$, plotted against $a/d$ for
-    three ellipses:\protect\\
-    ---$\!$---,
-    $b/a=1$; $\cdots$\,$\cdots$, $b/a=1.5$.}
-\label{fig:ka}
-\end{figure}
-
-\begin{figure}
-  \centerline{\includegraphics{modes}}
-  \caption{The features of the four possible modes corresponding to
-  (\textit{a}) periodic\protect\\ and (\textit{b}) half-periodic solutions.}
-\label{fig:kd}
+  \centerline{\includegraphics{trapped}}
+  \caption{\label{}}
 \end{figure}
@@
  \begin{center}
-\def~{\hphantom{0}}
   \begin{tabular}{lccc}
@@
   \end{tabular}
-  \caption{Values of $kd$ at which trapped modes occur when $\rho(\theta)=a$}
-  \label{tab:kd}
+  \caption{Values of $kd$ at which trapped modes occur when $\rho(\theta)=a$}{\label{tab:kd}}
  \end{center}
@@
 \begin{lemma}
+
 Let $f(z)$ be a trial \citet[][pp.~231--232]{Batchelor59} function defined on $[0,1]$.  Let $\varLambda_1$ denote
@@
 
+\bibliography{jfm-references}
 \bibliographystyle{jfm}
-% Note the spaces between the initials
-\bibliography{jfm-instructions}
 
